### 2024-07-19 images/tex/SciModel_simp1.tex
section:preliminaries,theory; process:write,images

\documentclass[tikz, border=1mm]{standalone}
\usepackage{tikz} 
\usetikzlibrary{arrows.meta}
\usepackage{pgfplots}

\begin{document}

\begin{tikzpicture}

    % main graph
    % nodes
    \node at (-3,1) {\tiny{\shortstack{characteristic \\ unit}}};
    \node at (-3,0) {\tiny{\shortstack{error \\ unit}}};
    \node at (-1.5,3) {\tiny{\shortstack{characteristic \\ judge}}};
    \node at (-1.5,2) {\tiny{\shortstack{error \\ judge}}};
    \node at (-1.5,1) {\tiny{\shortstack{characteristic \\ sub-unit}}};
    \node at (-1.5,0.1) {\tiny{\shortstack{trait \\ unit}}};
    \node at (-1.5,-0.85) {\tiny{\shortstack{error \\ sub-unit}}};
    \node at (0,2.1) {\tiny{\shortstack{discrimination \\ judge}}};
    \node at (0,0.2) {\tiny{\shortstack{trait \\ sub-unit}}};
	\node at (1.6,1.15) {\tiny{\shortstack{perception \\ difference}}};
    \node at (3,1.15) {\tiny{\shortstack{dichotomous \\ outcome}}};

	% paths
    \draw[{Circle}-{latex}](-3,0.7) to (-1.6,-0.28); % X_i -> theta_i
    \draw[{Circle[open]}-{latex}](-3,-0.3) to (-1.6,-0.32); % e_i -> theta_i
    \draw[{Circle}-{latex}](-1.5,2.7) to (-0.1,1.72); % Z_k -> gamma_k
    \draw[{Circle[open]}-{latex}](-1.5,1.7) to (-0.1,1.68); % e_k -> gamma_k
    \draw[{Circle}-{latex}](-1.5,0.7) to (-0.1,-0.28); % Y_r -> theta_ir
    \draw[{Circle[open]}-{latex}](-1.5,-0.3) to (-0.1,-0.3); % theta_i -> theta_ir
    \draw[{Circle[open]}-{latex}](-1.5,-1.3) to (-0.1,-0.32); % e_ir -> theta_ir
    \draw[{Circle[open]}-{latex}](0,1.7) to (1.4,0.7); % gamma_k -> delta_kijr
    \draw[{Circle[open]}-{latex}](0,-0.3) to (1.4,0.7); % theta_ir -> delta_kijr
    \draw[{Circle[open]}-{latex}{Circle}](1.5,0.7) to (3,0.7); % delta_kijr -> O_kijr

	% extras
	% nodes
    \node at (-4.5,0.1) {\tiny{\shortstack{population \\ unit}}}; % population of units
    \node at (-3,2.1) {\tiny{\shortstack{population \\ judges}}}; % population of judges
    \node at (-3,-0.9) {\tiny{\shortstack{population \\ sub-unit}}}; % population of objects

    % paths
    \draw[{Circle}-{latex}](-4.5,-0.3) to (-3.1,-0.28); % I -> e_i
    \draw[{Circle}-{latex}](-3,1.7) to (-1.6,1.7); % K -> e_k
    \draw[{Circle}-{latex}](-3,-1.27) to (-1.6,-1.27); % R_i -> e_r(i)

    % X_i -> R_i
    %\draw[-{latex}](-3,0.65) to[out=180,in=85] (-5,-0.3) to[out=-95,in=180] (-3.1,-1.27);

\end{tikzpicture}

\end{document}