### 2024-12-02 index.qmd
section:issues,CaseV; process:write

---
title: "Let's talk about Thurstone & Co.: An information-theoretical model for comparative judgments, and its statistical translation"
author:
  - name: 
      given: Jose Manuel
      family: Rivera Espejo
    orcid: 0000-0002-3088-2783
    url: https://www.uantwerpen.be/en/staff/jose-manuel-rivera-espejo_23166/
    email: JoseManuel.RiveraEspejo@uantwerpen.be
    corresponding: true
    affiliation:
      - name: University of Antwerp
        department: Training and education sciences
        group: Edubron
  - name: 
      given: Tine
      family: van Daal
      non-dropping-particle: van
    orcid: https://orcid.org/0000-0001-9398-9775
    url: https://www.uantwerpen.be/en/staff/tine-vandaal/
    email: tine.vandaal@uantwerpen.be
    corresponding: false
    affiliation:
      - name: University of Antwerp
        department: Training and education sciences
        group: Edubron
  - name: 
      given: Sven
      family: De Maeyer
      non-dropping-particle: De
    orcid: 0000-0003-2888-1631
    url: https://www.uantwerpen.be/en/staff/sven-demaeyer/
    email: sven.demaeyer@uantwerpen.be
    corresponding: false
    affiliation:
      - name: University of Antwerp
        department: Training and education sciences
        group: Edubron
  - name: 
      given: Steven
      family: Gillis
    orcid: 
    url: https://www.uantwerpen.be/nl/personeel/steven-gillis/
    email: steven.gillis@uantwerpen.be
    corresponding: false
    affiliation:
      - name: University of Antwerp
        department: Linguistics
        group: Centre for computational linguistics, psycholinguistics, and sociolinguistics (CLiPS)
funding: 
  statement: "The project was founded through the Research Fund of the University of Antwerp (BOF)."
keywords:
  - Probability
  - Directed Acyclic Graphs
  - Bayesian methods
  - Thurstonian model
  - Comparative judgement
  - Structural Causal Models
  - Statistical modeling
abstract: |
  (to do)
key-points:
  - (to do)
date: last-modified
bibliography: references.bib
notebook-links: global
lightbox: true
citation: true
  # type: article-journal
  # container-title: "Psychometrika"
  # doi: ""
  # # url: https://example.com/summarizing-output
execute: 
  cache: true
  # eval: true
  echo: false
  # output: true
  # include: true
  warning: false
  error: false
  message: false
---

<!-- abstract -->
<!-- recording: Sven and Tine 24.11.25; time: 00:58:20 - 00:59:20 -->

# Introduction {#sec-introduction}

<!-- recording: Sven 24.10.04; time: 00:00:00 - 00:09:10 -->
<!-- recording: Sven 24.10.04; time: 00:14:30 - 00:21:00 -->
<!-- recording: Sven and Tine 24.10.18; time: 00:04:05 - 00:08:30 -->
<!-- recording: Sven and Tine 24.10.25; time: 00:00:00 - 00:17:00 -->

<!-- 1. Overview of CJ method in assessing traits -->
In *comparative judgment* (CJ) studies, judges assess a specific trait or attribute across various stimuli by performing pairwise comparisons [@Thurstone_1927a; @Thurstone_1927b]. Each comparison produces a dichotomous outcome, indicating which stimulus is perceived to exhibit a higher trait level. For example, when assessing text quality, judges compare pairs of written texts (the stimuli) to determine the relative quality each text exhibit (the trait) [@Laming_2004; @Pollitt_2012b; @Whitehouse_2012; @vanDaal_et_al_2016; @Lesterhuis_2018_thesis; @Coertjens_et_al_2017; @Goossens_et_al_2018; @Bouwer_et_al_2023].

<!-- and spoken language [@Boonen_et_al_2020]. Additionally, it has been applied to evaluate various competencies, including mathematical problem-solving skills [@Jones_et_al_2015], engineering design skills [@Bartholomew_et_al_2018], build real-time web-based portfolios of performance [@@Kimbell_2012], conceptual understanding in algebra [@Jones_et_al_2019], statistical and English knowledge [@Marshall_et_al_2020], and STEM knowledge and skills [@Bartholomew_et_al_2020]. -->

<!-- 2. What is the effectiveness of CJ -->
Numerous studies have documented the effectiveness of CJ in assessing traits and competencies over the past decade. These studies have emphasized three aspects of the method's effectiveness: its reliability, validity, and practical applicability. Research on reliability indicates that CJ requires a relatively small number of pairwise comparisons [@Verhavert_et_al_2019; @Crompvoets_et_al_2022] to produce trait scores that are as precise and consistent as those generated by other assessment methods [@Coertjens_et_al_2017; @Goossens_et_al_2018; @Bouwer_et_al_2023]. Furthermore, evidence suggests that the reliability and time efficiency of CJ are comparable, if not superior, to those of other assessment methods when employing adaptive comparison algorithms [@Pollitt_2012b; @Verhavert_et_al_2022; @Mikhailiuk_et_al_2021]. Meanwhile, research on validity suggests that scores generated by CJ can accurately represent the traits under measurement [@Whitehouse_2012; @vanDaal_et_al_2016; @Lesterhuis_2018_thesis; @Bartholomew_et_al_2018; @Bouwer_et_al_2023], while research on practical applicability highlights the method's versatility across both educational and non-educational contexts [@Kimbell_2012; @Jones_et_al_2015; @Bartholomew_et_al_2018; @Jones_et_al_2019; @Marshall_et_al_2020; @Bartholomew_et_al_2020; @Boonen_et_al_2020].

<!-- rubrics [@Coertjens_et_al_2017; @Goossens_et_al_2018]  -->
<!-- holistic benchmark ratings [@Bouwer_et_al_2023] -->

<!-- Adaptive comparison algorithms dynamically present stimuli to judges, based on the results of previous comparisons. They aim to enhance the informativeness of comparisons while optimizing judges' time [@Bramley_2015].  -->

<!-- 3. But what critical issues remain in CJ Research?, what is the organization of the study? -->
Nevertheless, despite the increasing number of CJ studies, unsystematic and fragmented research approaches have left several critical issues unaddressed. The present study primarily focuses on two: the over-reliance on the assumptions of Thurstone's Case V in the statistical analysis of CJ data, and the apparent disconnect between CJ's trait measurement and hypothesis testing. The following sections begin with a brief overview of Thurstone's theory and a detailed discussion of these issues. Subsequently, the study introduces a theoretical model for CJ that builds upon Thurstone's theory, alongside its statistical translation, designed to address the two concerns simultaneously.

# Thurstone's theory {#sec-thurstone_theory}
<!-- recording: Sven and Tine 24.11.18; time: 00:04:30 - 00:14:50 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:19:04 - 00:21:37 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:00:35 - 00:06:55 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:30:35 - 00:35:10 -->

<!-- 1. Thurstone's Theory of CJ: discriminal process -->
In its most general form, Thurstone's theory deals with pairwise comparisons of stimuli made by a single judge [@Thurstone_1927b, pp. 267]. The theory proposes that two key factors determine the dichotomous outcome of these comparisons: the discriminal process of each stimulus and their discriminal difference. The *discriminal process* represents the psychological impact each stimulus has on judges or, more simply, their underlying perception of the stimulus' trait level. According to the theory, the discriminal process for each stimulus follows a Normal distribution, where its mode (or the mean), referred to as the *modal discriminal process*, indicates the stimulus' position on the trait continuum, and its dispersion, known as the *discriminal dispersion*, reflects the variability in the perceived trait level of the stimulus.

<!-- 2. Example of discriminal processes -->
For instance, @fig-discriminal_process illustrates the discriminal process distributions along a quality trait continuum for two written texts. The figure shows that these processes follow a Normal distribution. Moreover, it depicts differences in the texts' positions along the quality trait continuum, where text B is positioned further along the continuum than text A, as indicated by their modal discriminal processes ($S_{B}$ and $S_{A}$). Finally, it highlights differences in the texts' discriminal dispersions ($\sigma_{B}$ and $\sigma_{A}$), showing that text B exhibits a greater variability in its perceived quality than text A, as reflected by its wider distribution.

![Example distributions of the discriminal processes for two written texts](/images/figures/discriminal_process.png){#fig-discriminal_process width=70%}

<!-- 3. Thurstone's Theory of CJ: discriminal difference -->
However, because the discriminal process of a single stimulus is not directly observable, the theory introduces the *law of comparative judgment*. This law posits that in pairwise comparisons, a judge perceives the stimulus positioned further along the trait continuum as having a higher level of that trait. This principle highlights that the outcome of a pairwise comparison likely depends on the relative distance between stimuli rather than their absolute positions on the trait continuum.

Indeed, the theory assumes that the observed dichotomous outcome arises from the distribution of the difference between the underlying discriminal processes of the stimuli, known as the *discriminal difference*. Since the individual discriminal processes follow a Normal distribution, their difference also follows a Normal distribution [@Andrich_1978]. The mode (or the mean) of this distribution, representing the (average) relative separation, is given by the difference between the modal discriminal processes of the stimuli $S_{BA} = S_{B} - S_{A}$. Meanwhile, the dispersion of the distribution, reflecting the variability in the relative separation, is calculated as $\sigma_{BA} = \sqrt{\sigma_{B}^{2} + \sigma_{A}^{2} - \rho\sigma_{B}\sigma_{A}}$. Here, $\sigma_{B}$ and $\sigma_{A}$ are the previously defined discriminal dispersions, while $\rho$ measures the correlation between the discriminal processes of the stimuli. This correlation quantifies the dependence of the judge's perception of the trait in one stimulus on his perception of the same trait in the other.

<!-- 4. Example of discriminal difference -->
@fig-discriminal_difference shows the distribution of the discriminal difference for the texts depicted in @fig-discriminal_process, assuming a correlation of $\rho = 0.6$. The figure reveals that, under these conditions, the judge perceives text B as having significantly higher quality than text A, as indicated by the shaded gray area under the curve $P(B > A)$. As a result, the dichotomous outcome of this comparison would mostly favor text B over text A.

![Example distribution of the discriminal difference for the two texts shown in @fig-discriminal_process, assuming a correlation of 0.6](/images/figures/discriminal_difference.png){#fig-discriminal_difference width=70%}

<!-- 5. The effects of correlation on the discriminal difference -->
Notably, the correlation between the discriminal processes, $\rho$, plays a pivotal role in determining comparison outcomes by shaping the distribution of the discriminal difference between the stimuli. Specifically, as the correlation increases, reflecting a stronger dependence of the judge's perception of quality in one stimulus on his perception of the other, the distribution of the discriminal difference narrows. This narrowing ultimately impacts the area under the curve that determines the comparison outcome and, consequently, the conclusions drawn from this outcome.

<!-- 6. example of the effects of correlation on the discriminal difference -->
@fig-correlation illustrates how varying correlations influence the distribution of the discriminal difference for the two texts depicted in @fig-discriminal_process. Since the texts differ in quality, higher correlations increase the likelihood that the discriminal difference distinctly favors text B over text A. This is evident from the larger proportion of the area under the curve, $P(B > A)$, that lies above zero (shaded gray area in @fig-correlation). Conversely, it is not hard to infer that if the texts had similar or identical quality levels, higher correlations would likely reduce the chance that the discriminal difference distinctly favors one text over the other. This probability reduction occurs because the distribution of the discriminal difference would become more narrowly centered around zero (not illustrated).

![The effect of correlation on the distribution of the discriminal difference of the same two written text](/images/figures/correlation.png){#fig-correlation width=70%}

# Three critical issues in CJ literature {#sec-theory-issues}

## The Case V and the statistical analysis of CJ data {#sec-theory-issue1}

<!-- recording: Sven and Tine 24.10.25; time: 00:31:10 - 00:52:20 -->
<!-- recording: Sven and Tine 24.11.12; time: 00:00:00 - 00:00:00 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:00:25 - 00:04:30 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:14:50 - 00:19:04 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:21:37 - 00:47:20 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:06:55 - 00:24:50 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:35:10 - 00:49:30 -->

<!-- 1. Thurstone's Cases and their simplifying assumptions  -->
The previous section outlines the general form of Thurstone's theory, which applies to a CJ design where a single judge evaluates multiple stimuli. For the practical application of the theory, Thurstone developed four additional cases derived from this general form, where each successive case incorporates additional simplifying assumptions. Case I represents the general form of the theory. Case II extends this by allowing multiple judges to make comparisons rather than restricting them to a single judge. Case III introduces the assumption of zero correlation between stimuli. Case IV builds on this by assuming that the stimuli have similar dispersions. Finally, Case V replaces this assumption with the condition that the stimuli have equal discriminal dispersions. @tbl-thurstone_cases summarizes these cases and their assumptions. For a detailed discussion of this progression, refer to @Thurstone_1927b and @Bramley_2008 [pp. 248-253].

<!-- , suggesting all stimuli have the same variability in their perceived trait levels -->

![Thurstones cases and asumptions](/images/tables/thurstone_cases.png){#tbl-thurstone_cases width=90%}

<!-- 2. The Case V and its simpler statistical representation: the BTL model -->
Despite its reliance on the largest number of simplifying assumptions [@Bramley_2008, pp. 253; @Kelly_et_al_2022, pp. 677], Case V remains the most widely used case in the CJ literature. This popularity stems mainly from its simplified statistical representation in the Bradley-Terry-Luce (BTL) model [@Bradley_et_al_1952; @Luce_1959]. The BTL model mirrors the assumptions of Case V, with one key difference: while Case V assumes a Normal distribution for the discriminal processes of the stimuli, the BTL model uses the more mathematically tractable Logistic distribution [@Andrich_1978; @Bramley_2008, pp. 254] (see @tbl-thurstone_cases). This substitution has little impact on the model's estimation or interpretation, as the Normal and Logistic distributions share similar statistical properties, differing only by a scaling factor of approximately $1.7$ [@vanderLinden_et_al_2017_I, pp. 16] (see @fig-logistic_vs_normal).

<!-- the mathematical function linking dichotomous outcomes to trait scores -->

::: {#fig-logistic_vs_normal layout-nrow=2}

![Probability density](/images/figures/density.png){#fig-density width=70%}

![Cummulative probability](/images/figures/cummulative.png){#fig-cumulative width=70%}

Probability density and cumulative probability of the logistic and Normal distributions. Extracted from @Bramley_2008 [pp. 254-255].
:::

<!-- 3. But what are limitations of Case V (and the BTL model)? -->
However, Thurstone originally developed Case V to provide a "rather coarse scaling" of traits [@Thurstone_1927b, pp. 269], prioritizing statistical simplicity over precision in trait measurement [@Kelly_et_al_2022, pp. 677]. As a result, its assumptions may not be suitable for applications beyond the psycho-physical contexts for which it was created. Thurstone himself cautioned that its use "should not be made without (an) experimental test" [@Thurstone_1927b, pp. 270], acknowledging that some assumptions could prove problematic in the presence of complex traits or heterogeneous stimuli, such as handwriting or English compositions [@Thurstone_1927a, pp. 374]. Consequently, given that modern CJ applications frequently involve these traits and stimuli, two main assumptions of Case V may not consistently hold in theory or practice: the zero correlation and equal dispersion between stimuli.

<!-- Adding assumptions simplifies the formula and improves its tractability, but one cannot place as much confidence in the outputs of the formula if those assumptions are not met [@Kelly_et_al_2022, pp. 677]. -->

<!-- 4. What does the zero correlation between stimuli imply? and does it hold? -->
As outlined in @sec-thurstone_theory, the assumption of *zero correlation between stimuli* implies that, during a pairwise comparison, a judge's perception of quality in one text does not influence his perception of the same trait in another text. Thurstone attributed this independence to the cancellation of potential judges' biases, driven by two opposing and equally weighted effects occurring during the pairwise comparisons [@Thurstone_1927b, pp. 268]. @Andrich_1978 mathematically demonstrated this cancellation using the BTL model under the assumption of discriminal processes with additive biases. However, it is easy to imagine at least two scenarios where the zero correlation assumption almost certainly does not hold: when the pairwise comparison involves multidimensional, complex traits with heterogeneous stimuli and when an additional hierarchical structure is relevant to the stimuli.

<!-- "It is a safe assumption that when the stimulus series is very homogeneous with no distracting attributes, the correlation between discriminal deviations is low and possible even zero" [@Thurstone_1927b, pp. 268]. -->

<!-- 5. The first scenario where the assumption does not hold -->
In the first scenario, the intricate aspects of multidimensional, complex traits may introduce dependencies between the stimuli due to certain judges' biases that resist cancellation. Research on text quality suggests that when judges evaluating these traits, they often rely on various intricate characteristics of the stimuli [@vanDaal_et_al_2016; @Lesterhuis_2018; @Chambers_et_al_2022]. These additional relevant characteristics, which are unlikely to be equally weighted or opposing, can unevenly influence judges' perceptions, creating biases in their judgments and, consequently, introducing dependencies between stimuli [@vanderLinden_et_al_2017_II, pp. 346]. For example, this could occur when a judge assessing the argumentative quality of a text places disproportionate emphasis on grammatical accuracy, ultimately favoring texts with fewer errors but weaker arguments. While direct evidence for this specific scenario is lacking, studies such as @Pollitt_et_al_2003 demonstrate the presence of such biases, supporting the idea that the factors influencing pairwise comparisons may not always cancel out.

<!-- If the test taker’s position on one latent trait is fixed, the assumption of local stochastic independence requires the association between the responses to the items to vanish. Suppose the assumption is violated and more than one dimension is necessary to describe the test takers’ position in the latent space. The association between the responses to the items given one proficiency parameter does then not vanish. [@vanderLinden_et_al_2017_II, pp. 346]. -->

<!-- research shows that, in the presence of these traits and stimuli, the accuracy of the judges' trait perception is influenced by the rank-order distance of the stimuli [@vanDaal_et_al_2017; @Gijsen_et_al_2021], a coarse measure of the relative positioning of the stimuli within the trait continuum. -->

<!-- 6. The second scenario where the assumption does not hold -->
In the second scenario, the shared context or inherent connections created by additional hierarchical structures may further introduce dependencies between stimuli, a statistical phenomenon commonly known as clustering [@Everitt_et_al_2010]. Although the CJ literature acknowledges the presence of such hierarchical structures in CJ data, the statistical handling of this extra source of dependency has been inadequate. For example, when CJ data includes multiple samples of stimuli from the same individuals, researchers often rely on (average) estimated BTL scores to conduct subsequent analyses and tests at the individual hierarchical level [@Bramley_et_al_2019; @Boonen_et_al_2020; @Bouwer_et_al_2023; @vanDaal_et_al_2017; @Jones_et_al_2019; @Gijsen_et_al_2021]. This approach, however, has the significant limitation of ignoring the uncertainty associated with the BTL scores, which generates additional statistical and measurement issues, as discussed in section @sec-theory-issue2.

<!-- 7. But what are the challenges when these assumptions are violated? and how to solve them? -->
In any case, the psychometric and statistical literature underscores the importance of incorporating additional relevant traits and clustering (grouping) structures into the modeling of CJ data to address potential dependencies between stimuli. Oversights in these areas can lead to misleading conclusions about the trait's reliability and ultimately compromise its validity [@Perron_et_al_2015, pp. 2]. Neglecting additional traits relevant to the stimuli or the trait of interest, such as judges' biases, introduces a dimensional mismatch in the modeling process of the CJ data. This mismatch often results in an overestimation of the trait's reliability [@Hoyle_et_al_2023, pp. 341] or, worse, introduces biases into the estimated trait [@Ackerman_1989]. Therefore, these traits must be incorporated into the modeling of CJ data, as demonstrated by @Andrich_1978 and @Wainer_et_al_1978. Similarly, ignoring hierarchical (grouping) structures reduces the accuracy of model parameter estimates and further exacerbates the overestimation of reliability [@Hoyle_et_al_2023, pp. 482]. To address this, any additional hierarchical structure relevant to the stimuli should also be included in the model, as in Multilevel Structural Equation models (MSEM) [@Hoyle_et_al_2023, chap. 26].

<!-- This further implies that the shape of the discriminal difference distribution that determines the comparison outcome and, consequently, the conclusions drawn from this outcome (see @sec-thurstone_theory, @fig-correlation). -->

<!-- To show the practical consequences, in Figure 18.1, we present the test information for the two models; clearly, the information is positively biased when the multidimensional data are forced into a unidimensional IRT model. Moreover, in this same figure, we show the test response curves (how expected scores change as a function of trait level) for the two models; again, they, clearly, are not overlapping. The estimated implication is that if one applied a unidimensional IRT model, psychometric information is inflated due to the multidimensionality and examinee standard errors, which are the inverse square root of information, are too small. [@Hoyle_et_al_2023, pp. 340-341] -->

<!-- Accounting for measurement error in defining latent variables improves the accuracy of the model’s estimated structural parameters between the latent constructs, which in turn can provide a stronger test of a proposed theoretical model. [@Hoyle_et_al_2023, pp. 482] -->

<!-- Reliability is a necessary but not sufficient condition for validity. Reliability can exist without validity but validity cannot exist without reliability [@Perron_et_al_2015, pp. 2]. If your scores are not reliable they cannot be valid. People have been employing a model with assumptions that they do not know if those assumptions apply for their cases. And when you do that, the scores you produce most likely will not be reliable, which, in turn, renders them invalid. -->

<!-- Brennen generalizability theory, generalize conclusions over groups. To generalize over judges, you need to estimate them. Akin to Frisch-Wough-Lovel theorem. -->

<!-- Remember: the "mitigate" paragraph have the purpose of being the design principles that will guide the development of the theoretical and statistical model. -->

<!-- 9. What does the equal dispersion between stimuli imply? and does it hold? -->
<!-- In contrast, the assumption of *equal dispersion between stimuli* suggests that the variability in the perceived trait level of the stimulus is the same across all stimuli. While Thurstone acknowledged that this assumption may be violated when "dealing with less conspicuous attributes or with less homogeneous stimuli" [@Thurstone_1927a, pp. 374], ... -->

<!-- It is a ridiculous assumption, because multidimensional, complex traits -->
<!-- The dispersion is not the same for every text, a weak writer will produce consistently weak quality texts, in all the dimensions, this implies that every judge will agree on its lower quality, thereby the dispersion of the texts is going to be narrow. In contrast, when the texts move further along the quality trait continuum, implying that some areas of the texts are improved while some other remains, then there is a higher potential that there are discrepancies between the judges, as they focus in different aspects of the stimuli. So it seem that the extreme of the discriminal process distribution of the stimuli have higher reliability than the center? -->

<!-- Some text will provoke more variability in the judgments, and that is because of the text. There is something in the text that cause doubts in the judges. -->

<!-- Example of creativity from Tine. Misfiting text might be texts with larger dispersion, because there is a lot of confusion about them. When assessing creativity, for instance, student can do an assignment in really creative but an unexpected way, but because the judges did not expected such a way, they do not what to do with it, and that creates more dispersion in the judgment. Then the discriminal difference is all over the place. -->

<!-- Of these, the most questionable is the one that the discriminal dispersions of all the objects are equal. Thurstone clearly did not expect this to hold for any but the most simple stimuli. The scripts used in cross-moderation exercises are obviously far more complex than any used by Thurstone, so it seems naive to expect this assumption to hold here [@Bramley_2008]. -->

<!-- 10. But what are the challenges when these assumptions are violated? -->
<!-- But the statistical literature warns against mis-specifying the trait’s measurement model. When mis-specifying the measurement model by assuming, for example, equal discriminal dispersions across stimuli, the model can overlook meaningful differences in the reliability of the trait. Taken together, these mis-specifications and oversights can lead to misleading conclusions about the trait measurement, ultimately casting doubts about the validity of the trait, as validity cannot exist without reliability [@Perron_et_al_2015, pp. 2]. -->

<!-- Reliability is a necessary but not sufficient condition for validity. Reliability can exist without validity but validity cannot exist without reliability [@Perron_et_al_2015, pp. 2] -->

<!-- If your scores are not reliable they cannot be valid. People have been employing a model with assumptions that they do not know if those assumptions apply for their cases. And when you do that, the scores you produce most likely will not be reliable, which, in turn, renders them invalid. -->

<!-- 11. So, how do we mitigate these risks? -->
<!-- ["to mitigate" paragraph]{style="color:red;"} -->

<!-- This has been considered in the literature as a (see https://doi.org/10.1177/1471082X15571817) or a multilevel BTL model (find sources) -->

## The disconnect between trait measurement and hypothesis testing {#sec-theory-issue2}

<!-- recording: Sven 24.10.04; time: 00:14:30 - 00:21:00 -->
<!-- recording: Sven and Tine 24.10.18; time: 00:08:30 - 00:14:55 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:49:30 - 00:53:30 -->

<!-- 1. The BTL model as a measurement model for estimating latent traits -->
Building on the previous section, it is evident that the BTL model commonly functions as the trait's measurement model in CJ experiments [@Andrich_1978; @Bramley_2008]. A measurement model specifies how manifest variables contribute to the estimation of latent variables [@Everitt_et_al_2010]. For example, when evaluating text quality, researchers use the BTL model to process the dichotomous outcomes resulting from the pairwise comparisons (the manifest variables) to estimate scores that reflect the underlying quality level of texts (the latent variable) [@Laming_2004; @Pollitt_2012b; @Whitehouse_2012; @vanDaal_et_al_2016; @Lesterhuis_2018_thesis; @Coertjens_et_al_2017; @Goossens_et_al_2018; @Bouwer_et_al_2023].

<!-- 2. Utilizing BTL Scores for further analysis and hypothesis testing -->
Researchers then typically use the estimated BTL scores, or their transformations, to conduct additional analyses or hypothesis tests. For example, these scores have been used to identify 'misfit' judges and stimuli [@Pollitt_2012b; @vanDaal_et_al_2017; @Goossens_et_al_2018], detect biases in judges' ratings [@Pollitt_et_al_2003; @Pollitt_2012b], calculate correlations with other assessment methods [@Goossens_et_al_2018; @Bouwer_et_al_2023], or test hypotheses related to the underlying trait of interest [@Bramley_et_al_2019; @Boonen_et_al_2020; @Bouwer_et_al_2023; @vanDaal_et_al_2017; @Jones_et_al_2019; @Gijsen_et_al_2021].

<!-- 3. But what are the challenges in using BTL scores for further analysis? -->
However, the statistical literature advises caution when using estimated scores for additional analyses and tests. A key consideration is that BTL scores are parameter estimates that inherently carry uncertainty. Ignoring this uncertainty can bias the analysis and reduce the precision of hypothesis tests. Notably, the direction and magnitude of such biases are often unpredictable. Results may be attenuated, exaggerated, or remain unaffected depending on the degree of uncertainty in the scores and the actual effects being tested [@Kline_et_al_2023, pp. 25; @Hoyle_et_al_2023, pp. 137]. Finally, the reduced precision in hypothesis tests diminishes their statistical power, increasing the likelihood of committing type-I or type-II errors [@McElreath_2020].

<!-- A type-I error occurs when a *true* null hypothesis is incorrectly rejected, while a type-II error occurs when a *false* null hypothesis is incorrectly accepted [@Everitt_et_al_2010]. -->

<!-- 4. So, how do we mitigate these risks? -->
To mitigate these risks, principles from Structural Equation Modeling (SEM) [@Hoyle_et_al_2023, pp. 138] and Item Response Theory (IRT) [@Fox_2010, chap. 6; @vanderLinden_et_al_2017_I, chap. 24] recommend conducting these analyses and tests within a structural model. A structural model specifies how different manifest or latent variables influence the latent variable of interest [@Everitt_et_al_2010]. This approach allows analyses that can account for both the BTL scores and their uncertainties simultaneously, rather than treating them as separate elements. Therefore, an integrated approach that combines CJ’s measurement and structural models can offer significant advantages.

<!-- ## The diverse assessment design features and their role on reliability and validity {#sec-theory-issue3} -->
<!-- Do not include yet!, until you see if you need it when you write about the next section -->
<!-- recording: Sven and Tine 24.11.18; time: 00:47:20 - 00:56:41 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:53:30 - 01:00:00 -->

# An updated theoretical and statistical model for CJ {#sec-theory}

<!-- []{style="color:red;"} -->

<!-- 3. What is the solution? -->

<!-- Notes:  -->
<!-- - There is some gain on considering a more integrated overarching systematic way of looking on what happen in DCJ when people compare two stimuli -->
<!-- - This paper tries to systematically integrate all aspects at play in a single scientific model, we will build the model in a stepwise manner using -->
<!-- - we will build a scientific model (more theoretical model) that integrates different aspects of the method that are at play when people use DCJ. -->

## The theoretical model {#sec-theory-theoretical}

<!-- general development -->
<!-- recording: Sven 24.10.04; time: 00:21:00 - 00:34:10 -->
<!-- recording: Sven 24.10.04; time: 00:41:00 - 00:46:00 -->
<!-- recording: Sven and Tine 24.10.18; time: 00:00:00 - 00:04:05 -->
<!-- recording: Sven and Tine 24.11.25; time: 01:00:00 - 01:08:40 -->

<!-- comparison mechanism as missing mechanism -->
<!-- recording: Sven 24.10.04; time: 00:09:10 - 00:14:30 -->
<!-- recording: Sven 24.10.04; time: 00:34:10 - 00:41:00 -->

<!-- This means that the cancellation of judges' biases may not hold in theory or practice, and that judges' biases may depend on the stimuli, by means of an interaction (but present) or because the judges works as a confounder -->

<!-- no study explicitly proposes that this assumption could also be violated due to the presence of an additional hierarchical (grouping) structure relevant to the texts. One such scenario might arise, for example, when comparing texts produced by university and secondary school students. In this case, university students may consistently (or more precisely) produce higher-quality texts, while secondary school students, who exhibit a broader range of writing abilities, would show greater variability in the quality of their texts. Although this example is somewhat contrived, it effectively illustrates how assuming equal dispersions across texts can overlook meaningful differences in the reliability of text quality across groups or individuals. -->

<!-- The correlation $\rho$, considered above is not the kind of correlation considered by Bock (1958) in studies of preference and choice of objects. The Bock correlation arises from a specific interaction between a subject and an item because the subject has a particular "preference", such as a taste preference, for the object. This preference for the object is then reflected in its comparisons with all other objects. However, it is less likely that subjects would have such particular preferences for some items in a test and, in any case, the assumption is open to empirical checking [@Andrich_1978, pp. 455]. -->

## From theory to statistics {#sec-theory-statistics}

<!-- recording: Sven 24.10.04; time: 00:46:00 - 01:01:00 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:24:50 - 00:30:35 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:49:30 - 00:53:30 -->

<!-- Using assumptions from IRT [@deAyala_2009, p. 20-21], we are assuming a multidimensional model, that violates the local independece assumption -->

<!-- [@deAyala_2009, p. 291-29] think about indeterminancies in MIRT. to solve metric indeterminancy you commonly fix a mu=0, and s=1, but for rotational indeterminancy, you need to define (maybe) orthogonality between units traits and judges bias -->

<!-- @Pritikin_2020 and @Gray_et_al_2024 bayesian modeling attempts -->

<!-- , as well as when eliminating data through ad hoc univariate procedures -->
<!-- Moreover, excluding data using ad hoc univariate procedures can compound these issues by discarding potentially valuable information, further exacerbating the bias [@Zimmerman_1994; @McElreath_2020] -->

# Discussion {#sec-discuss}

## Findings {#sec-discuss-finding}

<!-- recording: Sven and Tine 24.11.25; time: 00:56:10 - 00:56:10 -->

<!-- "If Thurstone’s Law is to be used in support of comparative judgment for assessment, his theory must be extended to cover the uses and applications that are required for assessment purposes." [@Kelly_et_al_2022, pp. 678] -->

<!-- Pollitt (2012a) argues that assessors in pairwise comparisons need less training, given intuitive comparison and decision making. However, there is a distinct lack of research on the role of training.  -->

<!-- Due to efficiency of scoring, most of research only collects only one text. This assumes that there is significant homogeneity between text of the same individual, compared to the between individual variability. However, this cannot be known ad-hoc.  -->

<!-- (Is this a validity contention?) -->
<!-- Furthermore, it is not inconceivable that the selection of a particular group of judges used to perform the pairwise comparison could result in an incomplete depiction of the dimensions and complexity of the trait. Previous research has highlighted that factors such as age, culture, and education [@Kelly_et_al_2022, pp. 683], as well as individual differences among judges [@Gill_et_al_2013; @vanDaal_et_al_2017; @vanDaal_2020], could influence judgment accuracy, thereby affecting the "shared consensus" of the trait measurement. As @Kelly_et_al_2022 noted [pp. 683], “Would the aggregate view of young, British-Asian men always be the same as the aggregate view of older, black women?” -->

<!-- Such differences may not be detected through analyses of bias and misfit. McMahon and Jones (2015) found differences between students and teachers in what was valued in assessments of understanding of a chemistry experiment. Where students prioritized factual recall, teachers prioritized scientific explanation. Both groups produced highly reliable scales (reliability estimates of .893 for students, and .874 for teachers). Although, in this case, a clear argument as to whose consensus should take priority can be made, it demonstrates the existence of different, but equally reliable, sets of consensuses depending on who is asked. [@Kelly_et_al_2022, pp. 683] -->

<!-- "Another challenge to the intrinsic validity argument is that the aggregate approach and valuing of diversity in expert views (Van Daal et al., 2019) is in tension with the idea of removing misfitting judges (Pollitt, 2012a). If someone is considered an expert, and the validity of comparative judgment rests on the collective expertise of a community of practice, the decision to remove them and discard their judgments can only be justifiable on the assumption that those judgments do not reflect their expertise. ... It is thus unclear whether or not this divergence is welcomed." [@Kelly_et_al_2022, pp. 681] -->

<!-- We advocate for the no removal of judges, but rather identify them and treat the data with the appropriate model, so these 'misfits' do not affect the measurement of the trait -->

## Limitations and further research {#sec-discuss-limitations}

<!-- The field of comparative judgment, while growing, is still relatively small, and at present we do not have compelling answers to fundamental questions such as: how many judges do we need? Who should judge? How many judgments should they make? And crucially, how can we be confident that they are basing their judgments on acceptable criteria? If judges are not using construct-relevant criteria, then the argument that comparative judgments are necessarily valid cannot be upheld. We recognise that there is unlikely to be a single answer to these questions for all the cases in which comparative judgment may be used. However, work to derive some general principles for good practice would be valuable.[@Kelly_et_al_2022, pp. 683] -->

<!-- @Bramley_2008 proposes ranking instead of comparisons -->

# Conclusion {#sec-conclusion}

{{< pagebreak >}}

# Declarations {.appendix .unnumbered}

**Funding:** The project was founded through the Research Fund of the University of Antwerp (BOF).

**Financial interests:** The authors have no relevant financial interest to disclose.

**Non-financial interests:** The authors have no relevant non-financial interest to disclose.

**Ethics approval:** The University of Antwerp Research Ethics Committee has confirmed that no ethical approval is required.

**Consent to participate:** Not applicable

**Consent for publication:** All authors have read and agreed to the published version of the manuscript.

**Availability of data and materials:** No data was utilized in this study.

**Code availability:** All the code utilized in this research is available in the digital document located at: [https://jriveraespejo.github.io/paper2_manuscript/](https://jriveraespejo.github.io/paper2_manuscript/).

**AI-assisted technologies in the writing process:** The authors used ChatGPT, an AI language model, during the preparation of this work. They occasionally employed the tool to refine phrasing and optimize wording, ensuring appropriate language use and enhancing the manuscript's clarity and coherence. The authors take full responsibility for the final content of the publication.

**CRediT authorship contribution statement:** *Conceptualization:* S.G., S.DM., T.vD., and J.M.R.E; *Methodology:* S.DM., T.vD., and J.M.R.E; *Software:* J.M.R.E.; *Validation:* J.M.R.E.; *Formal Analysis:* J.M.R.E.; *Investigation:* J.M.R.E; *Resources:* S.G., S.DM., and T.vD.; *Data curation:* J.M.R.E.; *Writing - original draft:* J.M.R.E.; *Writing - review and editing:* S.G., S.DM., and T.vD.; *Visualization:* J.M.R.E.; *Supervision:* S.G. and S.DM.; *Project administration:* S.G. and S.DM.; *Funding acquisition:* S.G. and S.DM.

<!-- **Acknowledgements:** -->

{{< pagebreak >}}

# References {.unnumbered}

:::{#refs}

:::