### 2024-12-06 _tex/index.tex
Built site for gh-pages

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  1p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Psychometrika}

\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Let's talk about Thurstone \& Co.: An information-theoretical model for comparative judgments, and its statistical translation},
  pdfauthor={Jose Manuel Rivera Espejo; Tine van van Daal; Sven De De Maeyer; Steven Gillis},
  pdfkeywords={Probability, Directed Acyclic Graphs, Bayesian
methods, Thurstonian model, Comparative judgement, Structural Causal
Models, Statistical modeling},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Let's talk about Thurstone \& Co.: An information-theoretical
model for comparative judgments, and its statistical translation}
\author[1]{Jose Manuel Rivera Espejo%
\corref{cor1}%
}
 \ead{JoseManuel.RiveraEspejo@uantwerpen.be} 
\author[1]{Tine van Daal%
%
}
 \ead{tine.vandaal@uantwerpen.be} 
\author[1]{Sven De Maeyer%
%
}
 \ead{sven.demaeyer@uantwerpen.be} 
\author[2]{Steven Gillis%
%
}
 \ead{steven.gillis@uantwerpen.be}

\affiliation[1]{organization={University of Antwerp, Training and
education sciences},,postcodesep={}}
\affiliation[2]{organization={University of
Antwerp, Linguistics},,postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
(to do)
\end{abstract}

\begin{keyword}
    Probability \sep Directed Acyclic Graphs \sep Bayesian
methods \sep Thurstonian model \sep Comparative
judgement \sep Structural Causal Models \sep 
    Statistical modeling
\end{keyword}
\end{frontmatter}

\section{Introduction}\label{sec-introduction}

In \emph{comparative judgment} (CJ) studies, judges assess a specific
trait or attribute across various stimuli by performing pairwise
comparisons \citep{Thurstone_1927a, Thurstone_1927b}. Each comparison
produces a dichotomous outcome, indicating which stimulus is perceived
to exhibit a higher trait level. For example, when assessing text
quality, judges compare pairs of written texts (the stimuli) to
determine the relative quality each text exhibit (the trait)
\citep{Laming_2004, Pollitt_2012b, Whitehouse_2012, vanDaal_et_al_2016, Lesterhuis_2018_thesis, Coertjens_et_al_2017, Goossens_et_al_2018, Bouwer_et_al_2023}.

Numerous studies have documented the effectiveness of CJ in assessing
traits and competencies over the past decade. These studies have
emphasized three aspects of the method's effectiveness: its reliability,
validity, and practical applicability. Research on reliability indicates
that CJ requires a relatively small number of pairwise comparisons
\citep{Verhavert_et_al_2019, Crompvoets_et_al_2022} to produce trait
scores that are as precise and consistent as those generated by other
assessment methods
\citep{Coertjens_et_al_2017, Goossens_et_al_2018, Bouwer_et_al_2023}.
Furthermore, evidence suggests that the reliability and time efficiency
of CJ are comparable, if not superior, to those of other assessment
methods when employing adaptive comparison algorithms
\citep{Pollitt_2012b, Verhavert_et_al_2022, Mikhailiuk_et_al_2021}.
Meanwhile, research on validity suggests that scores generated by CJ can
accurately represent the traits under measurement
\citep{Whitehouse_2012, vanDaal_et_al_2016, Lesterhuis_2018_thesis, Bartholomew_et_al_2018, Bouwer_et_al_2023},
while research on practical applicability highlights the method's
versatility across both educational and non-educational contexts
\citep{Kimbell_2012, Jones_et_al_2015, Bartholomew_et_al_2018, Jones_et_al_2019, Marshall_et_al_2020, Bartholomew_et_al_2020, Boonen_et_al_2020}.

Nevertheless, despite the increasing number of CJ studies, unsystematic
and fragmented research approaches have left several critical issues
unaddressed. The present study primarily focuses on two: the
overreliance on the assumptions of Thurstone's Case V in the statistical
analysis of CJ data, and the apparent disconnect between CJ's approach
to trait measurement and hypothesis testing. The following sections
begin with a brief overview of Thurstone's theory and a detailed
examination of these issues. Subsequently, the study introduces a
theoretical model for CJ that builds upon Thurstone's theory, alongside
its statistical translation, designed to address the two concerns
simultaneously.

\section{Thurstone's theory}\label{sec-thurstone_theory}

In its most general form, Thurstone's theory addresses pairwise
comparisons of stimuli made by a single judge
\citep[pp.~267]{Thurstone_1927b}. The theory posits that two key factors
determine the dichotomous outcome of these comparisons: the discriminal
process of each stimulus and their discriminal difference. The
\emph{discriminal process} represents the psychological impact each
stimulus has on judges or, more simply, their perception of the stimulus
trait. According to the theory, the discriminal process for each
stimulus follows a Normal distribution, where its mode (mean), called
the \emph{modal discriminal process}, indicates the stimulus position on
the trait continuum, while its dispersion, known as the
\emph{discriminal dispersion}, reflects variability in the perceived
trait of the stimulus.

Figure~\ref{fig-discriminal_process} illustrates the discriminal
processes along a quality trait continuum for two written texts. The
figure shows that these processes individually follow a Normal
distribution. Furthermore, it shows that the modal discriminal process
for Text B is positioned further along the continuum than Text A
\((S_{B} > S_{A})\). Additionally, the figure highlights the broader
distribution of Text B compared to Text A as a result of its larger
discriminal dispersion \((\sigma_{B} > \sigma_{A})\).

\begin{figure}

\centering{

\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{./images/figures/discriminal_process.png}

}

\caption{\label{fig-discriminal_process}Discriminal processes for two
written texts}

\end{figure}%

However, since the discriminal processes of the stimuli are not directly
observable, the theory introduces the \emph{law of comparative
judgment}. This law posits that in pairwise comparisons, a judge
perceives the stimulus positioned further along the trait continuum as
having a higher level of that trait. This emphasizes that the relative
distance between stimuli, rather than their absolute positions on the
continuum, determines the outcome of the pairwise comparison. Indeed,
the theory assumes that the observed dichotomous outcome arises from the
difference between the underlying discriminal processes of the stimuli,
referred to as the \emph{discriminal difference}. Since the individual
discriminal processes follow a Normal distribution, the discriminal
difference also follows a Normal distribution \citep{Andrich_1978}, with
its mode (mean) representing the relative separation between the
stimuli, while its dispersion reflects the variability in that relative
separation.

Figure~\ref{fig-discriminal_difference} shows the distribution of the
discriminal difference for the texts depicted in
Figure~\ref{fig-discriminal_process}. The figure illustrates that the
judge perceives Text B as having significantly higher quality than Text
A, as indicated by their discriminal difference \((S_{B} - S_{A} > 0)\)
and the shaded gray area under the curve \(P(B > A)\). Therefore, the
dichotomous outcome of this comparison would likely favor Text B over
Text A.

\begin{figure}

\centering{

\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{./images/figures/discriminal_difference.png}

}

\caption{\label{fig-discriminal_difference}Discriminal difference for
two written texts}

\end{figure}%

\section{The two critical issues in CJ
literature}\label{sec-theory-issues}

This section examines the two critical issues in the CJ literature that
serve as the primary focus of this study. The first is the overreliance
on Thurstone's Case V assumptions in the statistical analysis of CJ
data. The second is the apparent disconnect between CJ's approach to
trait measurement and hypothesis testing.

\subsection{The Case V and the statistical analysis of CJ
data}\label{sec-theory-issue1}

The previous section outlines the general form of Thurstone's theory,
which applies to a CJ design where a single judge evaluates multiple
stimuli. For the practical application of the theory, Thurstone
developed four additional cases derived from this general form, where
each successive case incorporates additional simplifying assumptions.
Case I represents the general form of the theory. Case II extends this
by allowing multiple judges to make comparisons rather than restricting
the comparisons to a single judge. Case III introduces the assumption of
zero correlation between stimuli. Case IV builds on this by assuming
that the stimuli have similar dispersions. Finally, Case V replaces this
assumption with the condition that the stimuli have equal discriminal
dispersions. Table~\ref{tbl-thurstone_cases} summarizes these cases and
their assumptions. For a detailed discussion of this progression, refer
to \citet{Thurstone_1927b} and \citet[pp.~248-253]{Bramley_2008}.

\begin{table}

\caption{\label{tbl-thurstone_cases}Thurstones cases and asumptions}

\centering{

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{./images/tables/thurstone_cases.png}

}

\end{table}%

Despite its reliance on the largest number of simplifying assumptions
\citetext{\citealp[pp.~253]{Bramley_2008}; \citealp[pp.~677]{Kelly_et_al_2022}},
Case V remains the most widely used case in the CJ literature. This
popularity stems mainly from its simplified statistical representation
in the Bradley-Terry-Luce (BTL) model
\citep{Bradley_et_al_1952, Luce_1959}. The BTL model mirrors the
assumptions of Case V, with one key difference: while Case V assumes a
Normal distribution for the discriminal processes of the stimuli, the
BTL model uses the more mathematically tractable Logistic distribution
\citep[pp.~254]{Andrich_1978, Bramley_2008} (see
Table~\ref{tbl-thurstone_cases}). This substitution has little impact on
the model's estimation or interpretation, as the Normal and Logistic
distributions share similar statistical properties, differing only by a
scaling factor of approximately \(1.7\)
\citep[pp.~16]{vanderLinden_et_al_2017_I} (see
Figure~\ref{fig-logistic_vs_normal}).

\begin{figure}

\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{./images/figures/density.png}

}

\subcaption{\label{fig-density}Probability density}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{./images/figures/cummulative.png}

}

\subcaption{\label{fig-cumulative}Cummulative probability}

\end{minipage}%

\caption{\label{fig-logistic_vs_normal}Probability density and
cumulative probability of the logistic and Normal distributions.
Extracted from \citet[pp.~254-255]{Bramley_2008}.}

\end{figure}%

Nevertheless, Thurstone originally developed Case V to provide a
``rather coarse scaling'' of traits \citep[pp.~269]{Thurstone_1927b},
prioritizing statistical simplicity over precision in trait measurement
\citep[pp.~677]{Kelly_et_al_2022}. Moreover, he warned against its
untested application, stating that its use ``should not be made without
(an) experimental test'' \citep[pp.~270]{Thurstone_1927b}, acknowledging
that some assumptions might be problematic when complex traits or
heterogeneous stimuli are involved \citep[pp.~376]{Thurstone_1927a}.
Consequently, given that modern CJ applications frequently involve such
traits and stimuli, two main assumptions of Case V and, by extension, of
the BTL model may not consistently hold in theory or practice: the
assumption of equal dispersion and zero correlation between stimuli.

\subsubsection{The assumption of equal dispersions between
stimuli}\label{the-assumption-of-equal-dispersions-between-stimuli}

The discriminal dispersions of stimuli are crucial in determining the
dichotomous outcomes of pairwise comparisons. Discrepancies in these
dispersions shape the distribution of the discriminal difference,
directly influencing the comparison outcome. For instance,
Figure~\ref{fig-dispersion} illustrates how more uncertainty in the
trait perception of a text over another broadens the distribution of
their discriminal difference. This broadening affects the area under the
curve, \(P(B>A)\), determining the comparison outcome. Additionally, the
figure shows that when there are no discrepancies in the texts'
discriminal dispersions, meaning the \emph{texts have equal dispersions}
\((\sigma_{B}-\sigma_{A}=0)\), the likelihood that the discriminal
difference distinctly favors one text over another increases.

\begin{figure}

\centering{

\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{./images/figures/dispersion.png}

}

\caption{\label{fig-dispersion}The effect of different dispersions
between stimuli on the distribution of the discriminal difference}

\end{figure}%

However, Thurstone contended that this assumption may not hold when
researchers assess complex traits or heterogeneous stimuli
\citep[pp.~376]{Thurstone_1927a}, as these traits and stimuli can
introduce judgment discrepancies due to their unique features. Indeed,
evidence of such violation may already exist in the CJ literature as
misfit statistics. These statistics measure the judgment discrepancies
associated with a given stimulus
\citetext{\citealp[pp.~12]{Pollitt_2004}; \citealp[pp.~20]{Goossens_et_al_2018}}.
For instance, labeling texts as ``misfits'' indicates that comparisons
involving these texts result in more judgment discrepancies than
comparisons involving other texts
\citep{Pollitt_2012a, Pollitt_2012b, vanDaal_et_al_2016, Goossens_et_al_2018}.
This finding implies that the discriminal differences associated with
``misfits'' texts usually display a broader dispersion, suggesting that
their discriminal processes also exhibit more variation than other
texts. Notably, this reasoning also applies to ``misfit'' judges, whose
evaluations reflect substantial deviations from the shared consensus due
to the unique characteristics of the stimuli or the judges themselves.

\subsubsection{The assumption of zero correlation between
stimuli}\label{the-assumption-of-zero-correlation-between-stimuli}

Similar to the discriminal dispersions, the correlation between the
discriminal processes is also crucial in determining the dichotomous
outcomes of pairwise comparisons. This correlation shapes the
distribution of the discriminal difference, directly influencing the
comparison outcome. For instance, Figure~\ref{fig-correlation}
illustrates how the stronger dependence of the trait perception in one
text over another narrows the distribution of their discriminal
difference. This narrowing impacts the area under the curve, \(P(B>A)\),
determining the comparison outcome. Additionally, the figure shows that
when the trait perception in one text is independent of another, meaning
the \emph{texts have zero correlation}, the likelihood that the
discriminal difference distinctly favors one text over another is lower
compared to when the texts are highly correlated.

\begin{figure}

\centering{

\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{./images/figures/correlation.png}

}

\caption{\label{fig-correlation}The effect of correlation on the
distribution of the discriminal difference}

\end{figure}%

Notably, Thurstone's Case V and the BTL model assume independent trait
perceptions across stimuli. Thurstone attributed this independence to
the cancellation of potential judges' biases, driven by two opposing and
equally weighted effects occurring during the pairwise comparisons
\citep[pp.~268]{Thurstone_1927b}. \citet{Andrich_1978} mathematically
demonstrated this cancellation using the BTL model under the assumption
of discriminal processes with additive biases. However, it is easy to
imagine at least two scenarios where the zero correlation assumption
almost certainly does not hold: when the pairwise comparison involves
multidimensional, complex traits with heterogeneous stimuli and when an
additional hierarchical structure is relevant to the stimuli.

In the first scenario, the intricate aspects of multidimensional,
complex traits may introduce dependencies between the stimuli due to
certain judges' biases that resist cancellation. Research on text
quality suggests that when judges evaluate these traits, they often rely
on various intricate characteristics of the stimuli to form their
judgments
\citep{vanDaal_et_al_2016, Lesterhuis_2018, Chambers_et_al_2022}. These
additional relevant characteristics, which are unlikely to be equally
weighted or opposing, can unevenly influence judges' perceptions,
creating biases in their judgments and, ultimately, introducing
dependencies between stimuli
\citep[pp.~346]{vanderLinden_et_al_2017_II}. For example, this could
occur when a judge assessing the argumentative quality of a text places
more weight on grammatical accuracy than other judges, ultimately
favoring texts with fewer errors but weaker arguments. While direct
evidence for this specific scenario is lacking, studies such as
\citet{Pollitt_et_al_2003} demonstrate the presence of such biases,
supporting the idea that the factors influencing pairwise comparisons
may not always cancel out.

In the second scenario, the shared context or inherent connections
created by additional hierarchical structures may further introduce
dependencies between stimuli, a statistical phenomenon commonly known as
clustering \citep{Everitt_et_al_2010}. Although the CJ literature
acknowledges the presence of such hierarchical structures, the
statistical handling of this extra source of dependency between stimuli
has been inadequate. For example, when CJ data includes multiple samples
of stimuli from the same individuals, researchers often rely on
(average) estimated BTL scores to conduct subsequent analyses and tests
at the individual hierarchical level
\citep{Bramley_et_al_2019, Boonen_et_al_2020, Bouwer_et_al_2023, vanDaal_et_al_2017, Jones_et_al_2019, Gijsen_et_al_2021}.
This approach, however, has the significant limitation of ignoring the
uncertainty associated with the BTL scores, which generates additional
statistical and measurement issues, as discussed in section
Section~\ref{sec-theory-issue2}.

\subsection{The disconnect between trait measurement and hypothesis
testing}\label{sec-theory-issue2}

Building on the previous section, it is evident that, besides its
inconveniences, the BTL model commonly functions as the trait's
measurement model in CJ assessments. A measurement model specifies how
manifest variables contribute to the estimation of latent variables
\citep{Everitt_et_al_2010}. For example, when evaluating text quality,
researchers use the BTL model to process the dichotomous outcomes
resulting from the pairwise comparisons (the manifest variables) to
estimate scores that reflect the underlying quality level of texts (the
latent variable)
\citep{Laming_2004, Pollitt_2012b, Whitehouse_2012, vanDaal_et_al_2016, Lesterhuis_2018_thesis, Coertjens_et_al_2017, Goossens_et_al_2018, Bouwer_et_al_2023}.

Researchers then typically use these estimated BTL scores, or their
transformations, to conduct additional analyses or hypothesis tests. For
example, these scores have been used to identify `misfit' judges and
stimuli \citep{Pollitt_2012b, vanDaal_et_al_2016, Goossens_et_al_2018},
detect biases in judges' ratings
\citep{Pollitt_et_al_2003, Pollitt_2012b}, calculate correlations with
other assessment methods \citep{Goossens_et_al_2018, Bouwer_et_al_2023},
or test hypotheses related to the underlying trait of interest
\citep{Bramley_et_al_2019, Boonen_et_al_2020, Bouwer_et_al_2023, vanDaal_et_al_2017, Jones_et_al_2019, Gijsen_et_al_2021}.

However, the statistical literature advises caution when using estimated
scores for additional analyses and tests. A key consideration is that
BTL scores are parameter estimates that inherently carry uncertainty.
Ignoring this uncertainty can bias the analysis and reduce the precision
of hypothesis tests. Notably, the direction and magnitude of such biases
are often unpredictable. Results may be attenuated, exaggerated, or
remain unaffected depending on the degree of uncertainty in the scores
and the actual effects being tested
\citetext{\citealp[pp.~25]{Kline_et_al_2023}; \citealp[pp.~137]{Hoyle_et_al_2023}}.
Finally, the reduced precision in hypothesis tests diminishes their
statistical power, increasing the likelihood of committing type-I or
type-II errors \citep{McElreath_2020}.

\section{An updated theoretical and statistical model for
CJ}\label{sec-theory}

\subsection{The theoretical model}\label{sec-theory-theoretical}

\subsection{From theory to statistics}\label{sec-theory-statistics}

\section{Discussion}\label{sec-discuss}

\subsection{Findings}\label{sec-discuss-finding}

\subsection{Limitations and further
research}\label{sec-discuss-limitations}

\section{Conclusion}\label{sec-conclusion}

\newpage{}

\section*{Declarations}\label{declarations}
\addcontentsline{toc}{section}{Declarations}

\textbf{Funding:} The project was founded through the Research Fund of
the University of Antwerp (BOF).

\textbf{Financial interests:} The authors have no relevant financial
interest to disclose.

\textbf{Non-financial interests:} The authors have no relevant
non-financial interest to disclose.

\textbf{Ethics approval:} The University of Antwerp Research Ethics
Committee has confirmed that no ethical approval is required.

\textbf{Consent to participate:} Not applicable

\textbf{Consent for publication:} All authors have read and agreed to
the published version of the manuscript.

\textbf{Availability of data and materials:} No data was utilized in
this study.

\textbf{Code availability:} All the code utilized in this research is
available in the digital document located at:
\url{https://jriveraespejo.github.io/paper2_manuscript/}.

\textbf{AI-assisted technologies in the writing process:} The authors
used ChatGPT, an AI language model, during the preparation of this work.
They occasionally employed the tool to refine phrasing and optimize
wording, ensuring appropriate language use and enhancing the
manuscript's clarity and coherence. The authors take full responsibility
for the final content of the publication.

\textbf{CRediT authorship contribution statement:}
\emph{Conceptualization:} S.G., S.DM., T.vD., and J.M.R.E;
\emph{Methodology:} S.DM., T.vD., and J.M.R.E; \emph{Software:}
J.M.R.E.; \emph{Validation:} J.M.R.E.; \emph{Formal Analysis:} J.M.R.E.;
\emph{Investigation:} J.M.R.E; \emph{Resources:} S.G., S.DM., and T.vD.;
\emph{Data curation:} J.M.R.E.; \emph{Writing - original draft:}
J.M.R.E.; \emph{Writing - review and editing:} S.G., S.DM., and T.vD.;
\emph{Visualization:} J.M.R.E.; \emph{Supervision:} S.G. and S.DM.;
\emph{Project administration:} S.G. and S.DM.; \emph{Funding
acquisition:} S.G. and S.DM.

\newpage{}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\renewcommand{\bibsection}{}
\bibliography{references.bib}

\end{document}