### 2024-12-13 index.qmd
section:issues; process:write

---
title: "Let's talk about Thurstone & Co.: An information-theoretical model for comparative judgments, and its statistical translation"
author:
  - name: 
      given: Jose Manuel
      family: Rivera Espejo
    orcid: 0000-0002-3088-2783
    url: https://www.uantwerpen.be/en/staff/jose-manuel-rivera-espejo_23166/
    email: JoseManuel.RiveraEspejo@uantwerpen.be
    corresponding: true
    affiliation:
      - name: University of Antwerp
        department: Training and education sciences
        group: Edubron
  - name: 
      given: Tine
      family: van Daal
      non-dropping-particle: van
    orcid: https://orcid.org/0000-0001-9398-9775
    url: https://www.uantwerpen.be/en/staff/tine-vandaal/
    email: tine.vandaal@uantwerpen.be
    corresponding: false
    affiliation:
      - name: University of Antwerp
        department: Training and education sciences
        group: Edubron
  - name: 
      given: Sven
      family: De Maeyer
      non-dropping-particle: De
    orcid: 0000-0003-2888-1631
    url: https://www.uantwerpen.be/en/staff/sven-demaeyer/
    email: sven.demaeyer@uantwerpen.be
    corresponding: false
    affiliation:
      - name: University of Antwerp
        department: Training and education sciences
        group: Edubron
  - name: 
      given: Steven
      family: Gillis
    orcid: 
    url: https://www.uantwerpen.be/nl/personeel/steven-gillis/
    email: steven.gillis@uantwerpen.be
    corresponding: false
    affiliation:
      - name: University of Antwerp
        department: Linguistics
        group: Centre for computational linguistics, psycholinguistics, and sociolinguistics (CLiPS)
funding: 
  statement: "The project was founded through the Research Fund of the University of Antwerp (BOF)."
keywords:
  - Probability
  - Directed Acyclic Graphs
  - Bayesian methods
  - Thurstonian model
  - Comparative judgement
  - Structural Causal Models
  - Statistical modeling
abstract: |
  (to do)
key-points:
  - (to do)
date: last-modified
bibliography: references.bib
notebook-links: global
lightbox: true
citation: true
  # type: article-journal
  # container-title: "Psychometrika"
  # doi: ""
  # # url: https://example.com/summarizing-output
execute: 
  cache: true
  # eval: true
  echo: false
  # output: true
  # include: true
  warning: false
  error: false
  message: false
---

<!-- abstract -->
<!-- recording: Sven and Tine 24.11.25; time: 00:58:20 - 00:59:20 -->

# Introduction {#sec-introduction}

<!-- recording: Sven 24.10.04; time: 00:00:00 - 00:09:10 -->
<!-- recording: Sven 24.10.04; time: 00:14:30 - 00:21:00 -->
<!-- recording: Sven and Tine 24.10.18; time: 00:04:05 - 00:08:30 -->
<!-- recording: Sven and Tine 24.10.25; time: 00:00:00 - 00:17:00 -->

<!-- 1. Overview of CJ method in assessing traits -->
In *comparative judgment* (CJ) studies, judges assess a specific trait or attribute across various stimuli by performing pairwise comparisons [@Thurstone_1927a; @Thurstone_1927b]. Each comparison produces a dichotomous outcome, indicating which stimulus is perceived to exhibit a higher trait level. For example, when assessing text quality, judges compare pairs of written texts (the stimuli) to determine the relative quality each text exhibit (the trait) [@Laming_2004; @Pollitt_2012b; @Whitehouse_2012; @vanDaal_et_al_2016; @Lesterhuis_2018_thesis; @Coertjens_et_al_2017; @Goossens_et_al_2018; @Bouwer_et_al_2023].

<!-- and spoken language [@Boonen_et_al_2020]. Additionally, it has been applied to evaluate various competencies, including mathematical problem-solving skills [@Jones_et_al_2015], engineering design skills [@Bartholomew_et_al_2018], build real-time web-based portfolios of performance [@@Kimbell_2012], conceptual understanding in algebra [@Jones_et_al_2019], statistical and English knowledge [@Marshall_et_al_2020], and STEM knowledge and skills [@Bartholomew_et_al_2020]. -->

<!-- 2. What is the effectiveness of CJ -->
Numerous studies have documented the effectiveness of CJ in assessing traits and competencies over the past decade. These studies have emphasized three aspects of the method's effectiveness: its reliability, validity, and practical applicability. Research on reliability indicates that CJ requires a relatively small number of pairwise comparisons [@Verhavert_et_al_2019; @Crompvoets_et_al_2022] to produce trait scores that are as precise and consistent as those generated by other assessment methods [@Coertjens_et_al_2017; @Goossens_et_al_2018; @Bouwer_et_al_2023]. Furthermore, evidence suggests that the reliability and time efficiency of CJ are comparable, if not superior, to those of other assessment methods when employing adaptive comparison algorithms [@Pollitt_2012b; @Verhavert_et_al_2022; @Mikhailiuk_et_al_2021]. Meanwhile, research on validity suggests that scores generated by CJ can accurately represent the traits under measurement [@Whitehouse_2012; @vanDaal_et_al_2016; @Lesterhuis_2018_thesis; @Bartholomew_et_al_2018; @Bouwer_et_al_2023], while research on practical applicability highlights the method's versatility across both educational and non-educational contexts [@Kimbell_2012; @Jones_et_al_2015; @Bartholomew_et_al_2018; @Jones_et_al_2019; @Marshall_et_al_2020; @Bartholomew_et_al_2020; @Boonen_et_al_2020].

<!-- rubrics [@Coertjens_et_al_2017; @Goossens_et_al_2018]  -->
<!-- holistic benchmark ratings [@Bouwer_et_al_2023] -->

<!-- Adaptive comparison algorithms dynamically present stimuli to judges, based on the results of previous comparisons. They aim to enhance the informativeness of comparisons while optimizing judges' time [@Bramley_2015].  -->

<!-- 3. But what critical issues remain in CJ Research?, what is the organization of the study? -->
Nevertheless, despite the increasing number of CJ studies, unsystematic and fragmented research approaches have left several critical issues unaddressed. The present study primarily focuses on two: the overreliance on the assumptions of Thurstone's Case V in the statistical analysis of CJ data, and the apparent disconnect between CJ's approach to trait measurement and hypothesis testing. The following sections begin with a brief overview of Thurstone's theory and a detailed examination of these issues. Subsequently, the study introduces a theoretical model for CJ that builds upon Thurstone's theory, alongside its statistical translation, designed to address the two concerns simultaneously.

# Thurstone's theory {#sec-thurstone_theory}
<!-- recording: Sven and Tine 24.11.18; time: 00:04:30 - 00:14:50 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:19:04 - 00:21:37 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:00:35 - 00:06:55 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:30:35 - 00:35:10 -->
<!-- recording: Sven and Tine 24.12.09; time: 00:00:00 - 00:03:20 -->

<!-- 1. Thurstone's Theory of CJ: discriminal process -->
In its most general form, Thurstone’s theory addresses pairwise comparisons where a single judge evaluates multiple stimuli [@Thurstone_1927b, pp. 267]. The theory posits that two key factors determine the dichotomous outcome of these comparisons: the discriminal process of each stimulus and their discriminal difference. The *discriminal process* captures the psychological impact each stimulus exerts on the judge or, more simply, his perception of the stimulus trait. The theory assumes that the discriminal process for any given stimulus forms a Normal distribution along the trait continuum [@Thurstone_1927b, pp. 266]. The mode (mean) of this distribution, known as the *modal discriminal process*, indicates the stimulus position on this continuum, while its dispersion, referred to as the *discriminal dispersion*, reflects variability in the perceived trait of the stimulus.

<!-- The psychological continuum or scale is so constructed or defined that the frequencies of the respective discriminal processes for any given stimulus form a normal distribution on the psychological scale. This involves no assumption of a normal distribution or of anything else. The psychological scale is at best an artificial construct. If it has any physical reality, we certainly have not the remotest idea of what it may be like. ... We define the scale in terms of the frequencies of the discriminal processes for any stimulus. This artificial construct, the psychological scale, is so spaced off that the frequencies of the discriminal processes for any given stimulus form a normal distribution on the scale.[@Thurstone_1927b, pp. 266] -->

<!-- 2. Example of discriminal processes -->
@fig-discriminal_process illustrates hypothetical discriminal processes along a quality trait continuum for two written texts. The figure indicates that the modal discriminal process for Text B is positioned further along the continuum than that of Text A $(S_{B} > S_{A})$, suggesting that Text B exhibits higher quality. Additionally, the figure highlights that Text B has a broader distribution compared to Text A, which arises from its larger discriminal dispersion $(\sigma_{B} > \sigma_{A})$.

![Discriminal processes for two written texts](/images/png/discriminal_process.png){#fig-discriminal_process width=70%}

<!-- 3. Thurstone's Theory of CJ: discriminal difference -->
However, since the individual discriminal processes of the stimuli are not directly observable, the theory introduces *the law of comparative judgment*. This law posits that in pairwise comparisons, a judge perceives the stimulus with a discriminal process positioned further along the trait continuum as possessing more of the trait [@Bramley_2008, pp. 251]. This suggests that the relative distance between stimuli, rather than their absolute positions on the continuum, likely defines the outcome of pairwise comparisons. Indeed, the theory assumes that the difference between the underlying discriminal processes of the stimuli, referred to as *the discriminal difference*, determines the observed dichotomous outcome. Moreover, the theory assumes that because the individual discriminal processes form a Normal distribution on the continuum, the discriminal difference will also conform to a Normal distribution [@Andrich_1978]. In this distribution, the mode (mean) represents the relative separation between the stimuli, and its dispersion indicates the variability of that separation.

<!-- The mode (mean) of this distribution, representing the (average) relative separation, is given by the difference between the modal discriminal processes of the stimuli $S_{BA}=S_{B}-S_{A}$. Meanwhile, the dispersion of the distribution, reflecting the variability in the relative separation, is calculated as $\sigma_{BA} = \sqrt{\sigma_{B}^{2} + \sigma_{A}^{2} - \rho\sigma_{B}\sigma_{A}}$. Here, $\sigma_{B}$ and $\sigma_{A}$ denote the discriminal dispersions of the stimuli, while $\rho$ represents the correlation between their discriminal processes. This correlation quantifies the dependence of the judge's perception of the trait in one stimulus on his perception of the same trait in another. -->

<!-- 4. Example of discriminal difference -->
@fig-discriminal_difference illustrates the distribution of the discriminal difference for the hypothetical texts depicted in @fig-discriminal_process. The figure indicates that the judge perceives Text B as having significantly higher quality than Text A. This conclusion rests on two key observations: the positive difference between their modal discriminal processes $(S_{B} - S_{A} > 0)$ and the probability area where the discriminal difference distinctly favors Text B over Text A, represented by the shaded gray area denoted as $P(B > A)$. As a result, the dichotomous outcome of this comparison is more likely to favor Text B over Text A.

![Discriminal difference for two written texts](/images/png/discriminal_difference.png){#fig-discriminal_difference width=70%}

# The two critical issues in CJ literature {#sec-theory-issues}

This section examines the two critical issues in the CJ literature that serve as the primary focus of this study. The first is the overreliance on Thurstone’s Case V assumptions in the statistical analysis of CJ data. The second is the apparent disconnect between CJ's approach to trait measurement and hypothesis testing.

## The Case V and the statistical analysis of CJ data {#sec-theory-issue1}

<!-- recording: Sven and Tine 24.10.25; time: 00:31:10 - 00:52:20 -->
<!-- recording: Sven and Tine 24.11.12; time: 00:00:00 - 00:00:00 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:00:25 - 00:04:30 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:14:50 - 00:19:04 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:21:37 - 00:47:20 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:06:55 - 00:24:50 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:35:10 - 00:49:30 -->
<!-- recording: Sven and Tine 24.12.09; time: 00:03:20 - 00:08:10 -->

<!-- 1. What is the problem with the general form -->
Thurstone observed that the general form of the theory, outlined in @sec-thurstone_theory, created a trait scaling problem. The model required estimating more "unknown" parameters than the available pairwise comparisons [@Thurstone_1927b, pp. 267]. To address this issue and facilitate the practical application of the theory, he developed five cases derived from this general form. Each case progressively incorporated additional simplifying assumptions into the model.

<!-- 2. Thurstone's Cases and their simplifying assumptions -->
In Case I, Thurstone assumed that pairs of stimuli maintained a constant correlation across all comparisons. In Case II, he allowed multiple judges to make comparisons instead of restricting evaluations to a single judge. In Case III, he introduced the assumption of zero correlation between stimuli. In Case IV, he assumed stimuli exhibited similar dispersions. Finally, in Case V, he replaced this assumption with the condition that stimuli had equal discriminal dispersions. @tbl-thurstone_cases summarizes the assumptions of the general form and the five cases. For an in-depth discussion of these cases and their progression, refer to @Thurstone_1927b and @Bramley_2008 [pp. 248–253].

<!-- In Case II, Thurstone assumes that "each judge compares each stimulus with every other stimulus only once" [@Thurstone_1927b, pp. 268]. -->

![Thurstones cases and their asumptions](/images/png/thurstone_cases.png){#tbl-thurstone_cases width=100%}

<!-- 3. The Case V and its simpler statistical representation: the BTL model -->
Despite relying on the most extensive set of simplifying assumptions [@Bramley_2008, pp. 253; @Kelly_et_al_2022, pp. 677], Case V remains the most widely used case in the CJ literature. This popularity stems mainly from its simplified statistical representation in the Bradley-Terry-Luce (BTL) model [@Bradley_et_al_1952; @Luce_1959]. The BTL model mirrors the assumptions of Case V, with one key difference: while Case V assumes a Normal distribution for the stimuli's discriminal processes, the BTL model uses the more mathematically tractable Logistic distribution [@Andrich_1978; @Bramley_2008, pp. 254] (see @tbl-thurstone_cases). This substitution has little impact on the model's estimation or interpretation, as the Normal and Logistic distributions share similar statistical properties, differing only by a scaling factor of approximately $1.7$ [@vanderLinden_et_al_2017_I, pp. 16].

<!-- the mathematical function linking dichotomous outcomes to trait scores -->

<!-- Model misspecification happens when the set of probability distributions considered by the statistician does not include the distribution that generated the observed data. -->

<!-- ::: {#fig-logistic_vs_normal layout-nrow=2} -->

<!-- ![Probability density](/images/png/density.png){#fig-density width=70%} -->

<!-- ![Cummulative probability](/images/png/cummulative.png){#fig-cumulative width=70%} -->

<!-- Probability density and cumulative probability of the logistic and Normal distributions. Extracted from @Bramley_2008 [pp. 254-255]. -->
<!-- ::: -->

<!-- 4. But what are limitations of Case V (and the BTL model)? -->
However, Thurstone originally developed Case V to provide a "rather coarse scaling" of traits [@Thurstone_1927b, pp. 269], prioritizing statistical simplicity over precision in trait measurement [@Kelly_et_al_2022, pp. 677]. He explicitly warned against its untested application, stating that its use "should not be made without (an) experimental test" [@Thurstone_1927b, pp. 270], acknowledging that some assumptions could prove problematic when researchers asesss complex traits or heterogeneous stimuli [@Thurstone_1927a, pp. 376]. Consequently, given that modern CJ applications frequently involve such traits and stimuli, two main assumptions of Case V and, by extension, of the BTL model may not consistently hold in theory or practice: the assumption of equal dispersion and zero correlation between stimuli.

<!-- Adding assumptions simplifies the formula and improves its tractability, but one cannot place as much confidence in the outputs of the formula if those assumptions are not met [@Kelly_et_al_2022, pp. 677]. -->

### The assumption of equal dispersions between stimuli {#sec-theory-issue1a}

<!-- recording: Sven and Tine 24.12.09; time: 00:08:10 - 00:37:00 -->

<!-- 1. What does the equal dispersion between stimuli imply? -->
According to the theory, the discriminal dispersions of stimuli play a critical role in determining the outcome of pairwise comparisons. Specifically, discrepancies in these dispersions shape the distribution of the discriminal difference, directly influencing the comparison outcome. @fig-dispersion illustrates this idea, assuming a researcher can observe the discriminal processes for the texts shown in @fig-discriminal_process. The figure also considers that the discriminal dispersion for Text A remains constant and that the texts have no correlation $(\rho=0)$.

@fig-dispersion reveals that more uncertainty in the trait perception of Text B compared to Text A, $(\sigma_{B}-\sigma_{A})$, broadens the distribution of their discriminal difference. This broadening affects the probability area where the discriminal difference distinctly favors Text B over Text A, expressed as $P(B > A)$, ultimately influencing the comparison outcome. Additionally, the figure reveals that when the discriminal dispersions of the texts are equal $(\sigma_{B}-\sigma_{A}=0)$, the discriminal difference is more likely to favor Text B over Text A (shaded gray area), compared to situations where their dispersions differ.

![The discrepancy in the dispersions of stimuli and their effect on the distribution of the discriminal difference](/images/png/dispersion.png){#fig-dispersion width=70%}

<!-- 2. But this is not the scenario we observe -->
In experimental practice, however, this process occurs in reverse. Researchers first observe the comparison outcome and then use the BTL model to infer the discriminal difference between the stimuli and their discriminal processes [@Thurstone_1927a, pp. 373]. For example, when researchers observe a large sample of outcomes favoring Text B over Text A and correctly assume equal dispersions between the texts, the BTL model estimates a discriminal difference distribution that accurately represents the "true" discriminal difference of the texts. This scenario is illustrated with @fig-dispersion when the discriminal difference distribution of the model aligns with the "true" distribution, represented by the thick continuous line corresponding to $\sigma_{B}-\sigma_{A}=0$. The estimation accuracy of this discriminal difference, in turn, ensures reliable discriminal process estimates for the texts (citation needed?). It then intuitively follows that the outcome's ability to represent the "true" differences between stimuli largely depends on the validity of the model's assumptions, particularly the assumption of equal dispersions.

<!-- In experimental practice the procedure is the reverse of this hypothesis because the frequencies are known first experimentally, and from these frequencies we construct the psychological continuum. [@Thurstone_1927a, pp. 373] -->

<!-- 3. But does it hold?  -->
However, Thurstone contended that the assumption of equal dispersions may not hold when researchers assess complex traits or heterogeneous stimuli [@Thurstone_1927a, pp. 376], as these traits and stimuli can introduce judgment discrepancies due to their unique features [@vanDaal_et_al_2016; @Lesterhuis_2018; @Chambers_et_al_2022]. Indeed, evidence of this violation may already exist in the CJ literature as misfit statistics, which measure judgment discrepancies associated with specific stimuli [@Pollitt_2004, pp. 12; @Goossens_et_al_2018, pp. 20]. For example, labeling texts as "misfits" indicates that comparisons involving these texts result in more judgment discrepancies than those involving other texts [@Pollitt_2012a; @Pollitt_2012b; @vanDaal_et_al_2016; @Goossens_et_al_2018]. These discrepancies, in turn, suggest that the discriminal differences for "misfit" texts have broader distributions, indicating that their discriminal processes may also exhibit more variation than that of other texts. A similar reasoning applies to "misfit" judges, whose evaluations deviate substantially from the shared consensus due to the unique characteristics of the stimuli or the judges themselves. Moreover, these "misfit" judges and their deviations can introduce additional statistical and measurement issues, which we discuss in @sec-theory-issue1b.

<!-- It is probably true that this variability of the discriminal dispersion on the psychological continuum is of relatively less serious importance in dealing with strictly homogeneous stimulus series, but it becomes a serious factor in dealing with less conspicuous attributes or with less homogeneous stimulus series, such as handwriting specimens, English compositions, sewing samples, oriental rugs. In measurements of the type known as judgment scales the discriminal dispersion on the psychological continuum becomes one of the unknowns to be determined as well as the scale value of the specimen. Every specimen in such a series presents two unknown values to be determined, namely, the scale value of its modal discriminal process on the psychological continuum and its discriminal dispersion. [@Thurstone_1927a, pp. 376] -->

<!-- Of these, the most questionable is the one that the discriminal dispersions of all the objects are equal. Thurstone clearly did not expect this to hold for any but the most simple stimuli. The scripts used in cross-moderation exercises are obviously far more complex than any used by Thurstone, so it seems naive to expect this assumption to hold here [@Bramley_2008]. -->

<!-- If a portfolio’s WMS exceeds the criterion of mean plus two standard deviations, this means the judges did not judge it consistently, some considering it ‘better’ than others did. [@Pollitt_2012a, pp. 165] -->

<!-- Representations with a large infit are representations which lead to more inconsistent judgments [@Goossens_et_al_2018, pp. 20] -->

<!-- a misfit statistic. Essentially this reports the amount of inconsistency in the various judgements that have been made of that script. It can act as a flag identifying scripts that judges find difficult to judge [@Pollitt_2004, pp. 12] -->

<!-- For instance, in evaluating creativity, a student might complete an assignment in a highly creative but unexpected manner. Judges, unprepared for such originality, may produce more divergent evaluations of the assignment. These increased discrepancies in judgments suggest that the discriminal difference associated with the assignment has a broader dispersion, indicating greater variation in the discriminal process for this assignment compared to others. -->

<!-- "weak" writers consistently produce texts of low quality in all aspects that determine the quality of a text, such as grammatical accuracy, organization, and language use. In such cases, judges typically agree on their inferior quality, leading to a clear disfavoring of these texts compared to others. This agreement ultimately imply narrow dispersions for the discriminal differences related to these texts and, by extension, narrow discriminal processes for them. However, as texts improve along the trait continuum, with certain aspects showing improvement and others reaming unaltered, these variations introduce uncertainty into the judgments. Now, since judges often rely on various intricate characteristics of the texts to form their judgments [@vanDaal_et_al_2016; @Lesterhuis_2018; @Chambers_et_al_2022], these changes lead to greater discrepancies in their evaluations. This divergence ultimately imply broader discriminal differences and, by extension, more varied discriminal processes for the texts. -->

<!-- So it seem that the extreme of the discriminal process distribution of the stimuli have higher reliability than the center? -->

<!-- 4. So, what happens when the assumption does not hold? -->
Then, incorrectly assuming equal dispersions between stimuli can lead the BTL model to introduce various statistical and measurement issues. For example, the model could overestimate the accuracy of the outcome in reflecting the "true" discriminal differences between stimuli. This overestimation may result in spurious inferences about these differences [@McElreath_2020, pp. 370] and, by extension, about the stimuli's discriminal processes. @fig-dispersion also illustrates this scenario when the model's discriminal difference distribution aligns with the thick continuous line for $\sigma_{B}-\sigma_{A}=0$, while the "true" discriminal difference follows any discontinuous line where $\sigma_{B}-\sigma_{A} \neq 0$. Moreover, if researchers recognize that misfit statistics highlight critical differences in dispersions, the common practice in CJ literature of excluding stimuli based on these statistics [@Pollitt_2012a; @Pollitt_2012b; @vanDaal_et_al_2016; @Goossens_et_al_2018] may inadvertently discard valuable information, introducing bias into trait estimates [@Zimmerman_1994; @McElreath_2020, chap. 12]. These biases are often unpredictable, as they depend on the specific stimuli excluded from the analysis.

<!-- But ignoring over-dispersion can also lead to all of the same problems as ignoring any predictor variable. Heterogeneity in counts can be a confound, hiding effects of interest or producing spurious inferences. [@McElreath_2020, pp. 370] -->

### The assumption of zero correlation between stimuli {#sec-theory-issue1b}

<!-- 1. What does the zero correlation between stimuli imply? -->
Denoted by $\rho$, the correlation measures the dependence of a judge's perception of the trait in one stimulus on his perception of the same trait in another. Like the discriminal dispersions, this correlation shapes the distribution of the discriminal difference and directly influences the outcomes of pairwise comparisons. @fig-correlation illustrates this concept, assuming the researcher can observe the discriminal processes for the texts shown in @fig-discriminal_process. The figure also considers that the discriminal dispersions for both texts remain constant.

@fig-correlation reveals that as the correlation between the texts increases, the distribution of their discriminal difference becomes narrower. This narrowing affects the area under the curve where the discriminal difference distinctly favors Text B over Text A, denoted as $P(B > A)$, thus influencing the comparison outcome. Furthermore, the figure shows that when two texts are independent or uncorrelated $(\rho=0)$, their discriminal difference is less likely to favor Text B over Text A (shaded gray area), compared to scenarios when the texts are highly correlated.

<!-- "It is a safe assumption that when the stimulus series is very homogeneous with no distracting attributes, the correlation between discriminal deviations is low and possible even zero" [@Thurstone_1927b, pp. 268]. -->

![The correlation between stimuli and its effect on the distribution of the discriminal difference](/images/png/correlation.png){#fig-correlation width=70%}

<!-- 2. But this is not the scenario we observe -->
However, in experimental practice, researchers typically follow this process in reverse. For example, when they observe a large sample of outcomes favoring Text B over Text A and correctly assume zero correlation between the texts, the BTL model estimates a discriminal difference distribution that accurately represents the "true" discriminal difference of the texts. This scenario is illustrated with @fig-correlation when the discriminal difference distribution of the model aligns with the "true" distribution, represented by the thick continuous line corresponding to $\rho=0$. The accuracy of this discriminal difference estimation, in turn, ensures reliable estimates for the discriminal process of the texts (citation needed?).

<!-- It then follows that the outcome's ability to represent the "true" differences between stimuli largely depends on the validity of the assumption of zero correlation. -->

<!-- 2. But does it hold?  -->
Notably, Thurstone's Case V and the BTL model assume independent discriminal processes across comparisons. Thurstone attributed this independence to the cancellation of potential judges' biases, driven by two opposing and equally weighted effects occurring during the pairwise comparisons [@Thurstone_1927b, pp. 268]. @Andrich_1978 mathematically demonstrated this cancellation using the BTL model under the assumption of discriminal processes with additive biases. However, it is easy to imagine at least two scenarios where the zero correlation assumption almost certainly does not hold: when the pairwise comparison involves multidimensional, complex traits with heterogeneous stimuli and when an additional hierarchical structure is relevant to the stimuli.

<!-- 3. The first scenario where the assumption does not hold -->
In the first scenario, the intricate aspects of multidimensional, complex traits may introduce dependencies between the stimuli due to certain judges' biases that resist cancellation. Research on text quality suggests that when judges evaluate these traits, they often rely on various intricate characteristics of the stimuli to form their judgments [@vanDaal_et_al_2016; @Lesterhuis_2018; @Chambers_et_al_2022]. These additional relevant characteristics, which are unlikely to be equally weighted or opposing, can unevenly influence judges' perceptions, creating biases in their judgments and, ultimately, introducing dependencies between stimuli [@vanderLinden_et_al_2017_II, pp. 346]. For example, this could occur when a judge assessing the argumentative quality of a text places more weight on its grammatical accuracy than other judges, ultimately favoring texts with fewer errors but weaker arguments. While direct evidence for this specific scenario is lacking, studies such as @Pollitt_et_al_2003 demonstrate the presence of such biases, supporting the idea that the factors influencing pairwise comparisons may not always cancel out.

<!-- If the test taker’s position on one latent trait is fixed, the assumption of local stochastic independence requires the association between the responses to the items to vanish. Suppose the assumption is violated and more than one dimension is necessary to describe the test takers’ position in the latent space. The association between the responses to the items given one proficiency parameter does then not vanish. [@vanderLinden_et_al_2017_II, pp. 346]. -->

<!-- research shows that, in the presence of these traits and stimuli, the accuracy of the judges' trait perception is influenced by the rank-order distance of the stimuli [@vanDaal_et_al_2017; @Gijsen_et_al_2021], a coarse measure of the relative positioning of the stimuli within the trait continuum. -->

<!-- 4. The second scenario where the assumption does not hold -->
In the second scenario, the shared context or inherent connections created by additional hierarchical structures may further introduce dependencies between stimuli, a statistical phenomenon commonly known as clustering [@Everitt_et_al_2010]. Although the CJ literature acknowledges the presence of such hierarchical structures, the statistical handling of this extra source of dependency between stimuli has been inadequate. For example, when CJ data includes multiple samples of stimuli from the same individuals, researchers often rely on (average) estimated BTL scores to conduct subsequent analyses and tests at the individual hierarchical level [@Bramley_et_al_2019; @Boonen_et_al_2020; @Bouwer_et_al_2023; @vanDaal_et_al_2017; @Jones_et_al_2019; @Gijsen_et_al_2021]. However, this approach can introduce additional statistical and measurement issues, which we discuss in @sec-theory-issue2.

<!-- 5. But what are the challenges when this assumption is violated? -->
[In any case, similar to @sec-theory-issue1a, incorrectly assuming zero correlation between stimuli can lead the BTL model to introduce various statistical and measurement issues. For instance, the model could over- or underestimate the accuracy of the outcome in reflecting the "true" discriminal differences between stimuli. This over- underestimation may result in spurious inferences about these differences and, by extension, about the stimuli's discriminal processes [@Hoyle_et_al_2023, pp. 341]. @fig-correlation also illustrates this scenario when the model's discriminal difference distribution aligns with the thick continuous line for $\rho=0$, while the "true" discriminal difference follows any discontinuous line where $\rho \neq 0$. This missaligment can be due to the overlook of additional relevant traits, such as judges' biases, which cause dimensional mismatches in the BTL model, artificially inflating the reliability of the trait [@Hoyle_et_al_2023, pp. 341] or, even worse, introduce bias into the trait's estimates [@Ackerman_1989]. Furthermore, researchers who exclude judges based on misfit statistics can risk discarding valuable information, further biasing the trait's estimates [@Zimmerman_1994; @McElreath_2020, chap. 12]. Lastly, researchers who fail to account for hierarchical (grouping) structures can reduce the precision of model parameter estimates, which may amplify the overestimation of the trait's reliability [@Hoyle_et_al_2023, pp. 482].]{style="color:red;"}

<!-- To show the practical consequences, in Figure 18.1, we present the test information for the two models; clearly, the information is positively biased when the multidimensional data are forced into a unidimensional IRT model. Moreover, in this same figure, we show the test response curves (how expected scores change as a function of trait level) for the two models; again, they, clearly, are not overlapping. The estimated implication is that if one applied a unidimensional IRT model, psychometric information is inflated due to the multidimensionality and examinee standard errors, which are the inverse square root of information, are too small. [@Hoyle_et_al_2023, pp. 340-341] -->

<!-- Accounting for measurement error in defining latent variables improves the accuracy of the model’s estimated structural parameters between the latent constructs, which in turn can provide a stronger test of a proposed theoretical model. [@Hoyle_et_al_2023, pp. 482] -->

## The disconnect between trait measurement and hypothesis testing {#sec-theory-issue2}

<!-- recording: Sven 24.10.04; time: 00:14:30 - 00:21:00 -->
<!-- recording: Sven and Tine 24.10.18; time: 00:08:30 - 00:14:55 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:49:30 - 00:53:30 -->

<!-- 1. The BTL model as a measurement model for estimating latent traits -->
Building on the previous section, it is clear that, despite its limitations, the BTL model is commonly used as the measurement model in CJ assessments. A measurement model specifies how manifest variables contribute to the estimation of latent variables [@Everitt_et_al_2010]. For example, when evaluating text quality, researchers use the BTL model to process the dichotomous outcomes resulting from the pairwise comparisons (the manifest variables) to estimate scores that reflect the underlying quality level of the texts (the latent variable) [@Laming_2004; @Pollitt_2012b; @Whitehouse_2012; @vanDaal_et_al_2016; @Lesterhuis_2018_thesis; @Coertjens_et_al_2017; @Goossens_et_al_2018; @Bouwer_et_al_2023].

<!-- 2. Utilizing BTL Scores for further analysis and hypothesis testing -->
Researchers then typically use these estimated BTL scores, or their transformations, to conduct additional analyses or hypothesis tests. For example, these scores have been used to identify 'misfit' judges and stimuli [@Pollitt_2012b; @vanDaal_et_al_2016; @Goossens_et_al_2018], detect biases in judges' ratings [@Pollitt_et_al_2003; @Pollitt_2012b], calculate correlations with other assessment methods [@Goossens_et_al_2018; @Bouwer_et_al_2023], or test hypotheses related to the underlying trait of interest [@Bramley_et_al_2019; @Boonen_et_al_2020; @Bouwer_et_al_2023; @vanDaal_et_al_2017; @Jones_et_al_2019; @Gijsen_et_al_2021].

<!-- 3. But what are the challenges in using BTL scores for further analysis? -->
However, the statistical literature advises caution when using estimated scores for additional analyses and tests. A key consideration is that BTL scores are parameter estimates that inherently carry uncertainty. Ignoring this uncertainty can bias the analysis and reduce the precision of hypothesis tests. Notably, the direction and magnitude of such biases are often unpredictable. Results may be attenuated, exaggerated, or remain unaffected depending on the degree of uncertainty in the scores and the actual effects being tested [@Kline_et_al_2023, pp. 25; @Hoyle_et_al_2023, pp. 137]. Finally, the reduced precision in hypothesis tests diminishes their statistical power, increasing the likelihood of committing type-I or type-II errors [@McElreath_2020].

<!-- A type-I error occurs when a *true* null hypothesis is incorrectly rejected, while a type-II error occurs when a *false* null hypothesis is incorrectly accepted [@Everitt_et_al_2010]. -->

[In aggregate, researchers' inadequate handling of violations to the assumptions of equal dispersion and zero correlation between stimuli, along with the apparent disconnect between CJ's approach to trait measurement and hypothesis testing, can undermine the reliability of the trait estimates and ultimately compromise its validity [@Perron_et_al_2015, pp. 2]. Consequently, adopting a more systematic and integrated approach to examining what happens when judges compare two stimuli could offer several statistical and measurement benefits, including addressing these issues.]{style="color:red;"}

<!-- Reliability is a necessary but not sufficient condition for validity. Reliability can exist without validity but validity cannot exist without reliability [@Perron_et_al_2015, pp. 2]. -->

<!-- People have been employing a model with assumptions that they do not know if those assumptions apply for their cases. -->

<!-- ## The diverse assessment design features and their role on reliability and validity {#sec-theory-issue3} -->
<!-- Do not include yet!, until you see if you need it when you write about the next section -->
<!-- recording: Sven and Tine 24.11.18; time: 00:47:20 - 00:56:41 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:53:30 - 01:00:00 -->

# An updated theoretical and statistical model for CJ {#sec-theory}

This section presents a theoretical model for CJ that extends Thurstone's theory. The model systematically incorporates all factors involved when judges make pairwise comparisons. Additionally, the section develops the statistical translation of the theoretical model based on assumptions informed by the CJ theory.

<!-- in a stepwise manner -->

## The theoretical model {#sec-theory-theoretical}

<!-- []{style="color:red;"} -->

<!-- general development -->
<!-- recording: Sven 24.10.04; time: 00:21:00 - 00:34:10 -->
<!-- recording: Sven 24.10.04; time: 00:41:00 - 00:46:00 -->
<!-- recording: Sven and Tine 24.10.18; time: 00:00:00 - 00:04:05 -->
<!-- recording: Sven and Tine 24.11.25; time: 01:00:00 - 01:08:40 -->

<!-- comparison mechanism as missing mechanism -->
<!-- recording: Sven 24.10.04; time: 00:09:10 - 00:14:30 -->
<!-- recording: Sven 24.10.04; time: 00:34:10 - 00:41:00 -->

The (latent) discriminal difference of the stimuli directly determines the (manifest) outcome of the pairwise comparisons

![Theoretical model A1$](/images/png/CJ_TM_A1.png){#fig-CJ_TM_A1 width=29%}

The (latent) "perceived" discriminal processes for the stimuli directly determines their discriminal difference

![Theoretical model A2$](/images/png/CJ_TM_A2.png){#fig-CJ_TM_A2 width=40%}

The (latent) "true" discriminal processes for the stimuli and the judges' biases directly determines their (latent) "perceived" discriminal processes

![Theoretical model A3$](/images/png/CJ_TM_A3.png){#fig-CJ_TM_A3 width=60%}

without loosing generality, the (latent) "perceived" and "true" discriminal processes for the stimuli can be depicted in a vector for each judge, as in

<!-- This means that the cancellation of judges' biases may not hold in theory or practice, and that judges' biases may depend on the stimuli, by means of an interaction (but present) or because the judges works as a confounder -->

<!-- no study explicitly proposes that this assumption could also be violated due to the presence of an additional hierarchical (grouping) structure relevant to the texts. One such scenario might arise, for example, when comparing texts produced by university and secondary school students. In this case, university students may consistently (or more precisely) produce higher-quality texts, while secondary school students, who exhibit a broader range of writing abilities, would show greater variability in the quality of their texts. Although this example is somewhat contrived, it effectively illustrates how assuming equal dispersions across texts can overlook meaningful differences in the reliability of text quality across groups or individuals. -->

<!-- include subject specific covariates [@Casalicchio_et_al_2015] -->

<!-- Notably, the theory assumes the judge compares each pair of stimuli a "sufficient number of times" to accurately calculate the relative separation probability [@Thurstone_1927b, pp. 267]. -->

<!-- It is assumed that the single observer compares each pair of stimuli sufficient number of times so that a proportion, P(A>B), may be determined for each pair of stimuli [@Thurstone_1927b, pp. 267] -->

## From theory to statistics {#sec-theory-statistics}

<!-- recording: Sven 24.10.04; time: 00:46:00 - 01:01:00 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:24:50 - 00:30:35 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:49:30 - 00:53:30 -->

<!-- @Pritikin_2020 and @Gray_et_al_2024 bayesian modeling attempts -->

<!-- Using assumptions from IRT [@deAyala_2009, p. 20-21], we are assuming a multidimensional model, that violates the local independence assumption -->

<!-- [@deAyala_2009, p. 291-29] think about indeterminancies in MIRT. to solve metric indeterminancy you commonly fix a mu=0, and s=1, but for rotational indeterminancy, you need to define (maybe) orthogonality between units traits and judges bias -->

<!-- , as well as when eliminating data through ad hoc univariate procedures -->
<!-- Moreover, excluding data using ad hoc univariate procedures can compound these issues by discarding potentially valuable information, further exacerbating the bias [@Zimmerman_1994; @McElreath_2020] -->

<!-- 6. So, how do we mitigate these risks? -->
<!-- Fortunately, the statistical literature offers a solution to address the abovementioned issues. This solution involves using models that extend traditional approaches to account for the different variability in the stimuli, referred to as over-dispersed models [@McElreath_2020, chap. 12]. -->

<!-- 11. How to solve the violation of the assumption? -->
<!-- Fortunately, the same literature offers solutions for addressing these issues. @Andrich_1978 and @Wainer_et_al_1978 recommend integrating judges' biases into the BTL model. Moreover, the literature advocates for the incorporation of relevant hierarchical structures into the statistical model to account for these dependencies. Together, these additions can result in a model resembling a Multilevel Structural Equation Model (MSEM) [@Hoyle_et_al_2023, chap. 26] combined with a multidimensional or two-parameter logistic Item Response Theory (IRT) model [@Hoyle_et_al_2023, chap. 15], depending on the theoretical and statistical treatment of judges' biases. -->

<!-- Brennen generalizability theory, generalize conclusions over groups. To generalize over judges, you need to estimate them. Akin to Frisch-Wough-Lovel theorem for latent variables?. -->

<!-- 4. So, how do we mitigate these risks? -->
<!-- To mitigate these risks, principles from Structural Equation Modeling (SEM) [@Hoyle_et_al_2023, pp. 138] and IRT [@Fox_2010, chap. 6; @vanderLinden_et_al_2017_I, chap. 24] recommend conducting these analyses and tests within a structural model. A structural model specifies how different manifest or latent variables influence the latent variable of interest [@Everitt_et_al_2010]. This approach allows analyses that can account for both the BTL scores and their uncertainties simultaneously, rather than treating them as separate elements.  -->

<!-- Remember: the "mitigate" paragraph have the purpose of being the design principles that will guide the development of the theoretical and statistical model. -->

<!-- The correlation $\rho$, considered above is not the kind of correlation considered by Bock (1958) in studies of preference and choice of objects. The Bock correlation arises from a specific interaction between a subject and an item because the subject has a particular "preference", such as a taste preference, for the object. This preference for the object is then reflected in its comparisons with all other objects. However, it is less likely that subjects would have such particular preferences for some items in a test and, in any case, the assumption is open to empirical checking [@Andrich_1978, pp. 455]. -->

# Discussion {#sec-discuss}

## Findings {#sec-discuss-finding}

<!-- recording: Sven and Tine 24.11.25; time: 00:56:10 - 00:56:10 -->

<!-- "If Thurstone’s Law is to be used in support of comparative judgment for assessment, his theory must be extended to cover the uses and applications that are required for assessment purposes." [@Kelly_et_al_2022, pp. 678] -->

<!-- Pollitt (2012a) argues that assessors in pairwise comparisons need less training, given intuitive comparison and decision making. However, there is a distinct lack of research on the role of training.  -->

<!-- Due to efficiency of scoring, most of research only collects only one text. This assumes that there is significant homogeneity between text of the same individual, compared to the between individual variability. However, this cannot be known ad-hoc.  -->

<!-- (Is this a validity contention?) -->
<!-- Furthermore, it is not inconceivable that the selection of a particular group of judges used to perform the pairwise comparison could result in an incomplete depiction of the dimensions and complexity of the trait. Previous research has highlighted that factors such as age, culture, and education [@Kelly_et_al_2022, pp. 683], as well as individual differences among judges [@Gill_et_al_2013; @vanDaal_et_al_2017; @vanDaal_2020], could influence judgment accuracy, thereby affecting the "shared consensus" of the trait measurement. As @Kelly_et_al_2022 noted [pp. 683], “Would the aggregate view of young, British-Asian men always be the same as the aggregate view of older, black women?” -->

<!-- Such differences may not be detected through analyses of bias and misfit. McMahon and Jones (2015) found differences between students and teachers in what was valued in assessments of understanding of a chemistry experiment. Where students prioritized factual recall, teachers prioritized scientific explanation. Both groups produced highly reliable scales (reliability estimates of .893 for students, and .874 for teachers). Although, in this case, a clear argument as to whose consensus should take priority can be made, it demonstrates the existence of different, but equally reliable, sets of consensuses depending on who is asked. [@Kelly_et_al_2022, pp. 683] -->

<!-- "Another challenge to the intrinsic validity argument is that the aggregate approach and valuing of diversity in expert views (Van Daal et al., 2019) is in tension with the idea of removing misfitting judges (Pollitt, 2012a). If someone is considered an expert, and the validity of comparative judgment rests on the collective expertise of a community of practice, the decision to remove them and discard their judgments can only be justifiable on the assumption that those judgments do not reflect their expertise. ... It is thus unclear whether or not this divergence is welcomed." [@Kelly_et_al_2022, pp. 681] -->

<!-- We advocate for the no removal of judges, but rather identify them and treat the data with the appropriate model, so these 'misfits' do not affect the measurement of the trait -->

## Limitations and further research {#sec-discuss-limitations}

<!-- The field of comparative judgment, while growing, is still relatively small, and at present we do not have compelling answers to fundamental questions such as: how many judges do we need? Who should judge? How many judgments should they make? And crucially, how can we be confident that they are basing their judgments on acceptable criteria? If judges are not using construct-relevant criteria, then the argument that comparative judgments are necessarily valid cannot be upheld. We recognise that there is unlikely to be a single answer to these questions for all the cases in which comparative judgment may be used. However, work to derive some general principles for good practice would be valuable.[@Kelly_et_al_2022, pp. 683] -->

<!-- @Bramley_2008 proposes ranking instead of comparisons -->

# Conclusion {#sec-conclusion}

{{< pagebreak >}}

# Declarations {.appendix .unnumbered}

**Funding:** The project was founded through the Research Fund of the University of Antwerp (BOF).

**Financial interests:** The authors have no relevant financial interest to disclose.

**Non-financial interests:** The authors have no relevant non-financial interest to disclose.

**Ethics approval:** The University of Antwerp Research Ethics Committee has confirmed that no ethical approval is required.

**Consent to participate:** Not applicable

**Consent for publication:** All authors have read and agreed to the published version of the manuscript.

**Availability of data and materials:** No data was utilized in this study.

**Code availability:** All the code utilized in this research is available in the digital document located at: [https://jriveraespejo.github.io/paper2_manuscript/](https://jriveraespejo.github.io/paper2_manuscript/).

**AI-assisted technologies in the writing process:** The authors used ChatGPT, an AI language model, during the preparation of this work. They occasionally employed the tool to refine phrasing and optimize wording, ensuring appropriate language use and enhancing the manuscript's clarity and coherence. The authors take full responsibility for the final content of the publication.

**CRediT authorship contribution statement:** *Conceptualization:* S.G., S.DM., T.vD., and J.M.R.E; *Methodology:* S.DM., T.vD., and J.M.R.E; *Software:* J.M.R.E.; *Validation:* J.M.R.E.; *Formal Analysis:* J.M.R.E.; *Investigation:* J.M.R.E; *Resources:* S.G., S.DM., and T.vD.; *Data curation:* J.M.R.E.; *Writing - original draft:* J.M.R.E.; *Writing - review and editing:* S.G., S.DM., and T.vD.; *Visualization:* J.M.R.E.; *Supervision:* S.G. and S.DM.; *Project administration:* S.G. and S.DM.; *Funding acquisition:* S.G. and S.DM.

<!-- **Acknowledgements:** -->

{{< pagebreak >}}

# References {.unnumbered}

:::{#refs}

:::