### 2025-06-05 paper2.qmd
section:6; process:write

---
title: "Let's talk about Thurstone & Co.: An information-theoretical model for comparative judgments, and its statistical translation"
author:
  - name: 
      given: Jose Manuel
      family: Rivera Espejo
    orcid: 0000-0002-3088-2783
    url: https://www.uantwerpen.be/en/staff/jose-manuel-rivera-espejo_23166/
    email: JoseManuel.RiveraEspejo@uantwerpen.be
    corresponding: true
    affiliation:
      - name: University of Antwerp
        department: Training and education sciences
        group: Edubron
  - name: 
      given: Tine
      family: van Daal
      non-dropping-particle: van
    orcid: 0000-0001-9398-9775
    url: https://www.uantwerpen.be/en/staff/tine-vandaal/
    email: tine.vandaal@uantwerpen.be
    corresponding: false
    affiliation:
      - name: University of Antwerp
        department: Training and education sciences
        group: Edubron
  - name: 
      given: Sven
      family: De Maeyer
      non-dropping-particle: De
    orcid: 0000-0003-2888-1631
    url: https://www.uantwerpen.be/en/staff/sven-demaeyer/
    email: sven.demaeyer@uantwerpen.be
    corresponding: false
    affiliation:
      - name: University of Antwerp
        department: Training and education sciences
        group: Edubron
  - name: 
      given: Steven
      family: Gillis
    orcid: 
    url: https://www.uantwerpen.be/nl/personeel/steven-gillis/
    email: steven.gillis@uantwerpen.be
    corresponding: false
    affiliation:
      - name: University of Antwerp
        department: Linguistics
        group: Centre for computational linguistics, psycholinguistics, and sociolinguistics (CLiPS)
funding: 
  statement: "The project was founded through the Research Fund of the University of Antwerp (BOF)."
keywords:
  - causal inference
  - directed acyclic graphs
  - structural causal models
  - bayesian statistical methods
  - thurstonian model
  - comparative judgement
  - probability
  - statistical modeling
abstract: |
  This study revisits Thurstone's law of comparative judgment (CJ), focusing on two prominent issues of traditional approaches. First, it critiques the heavy reliance on Thurstone's Case V assumptions and, by extension, the Bradley-Terry-Luce (BTL) model when analyzing CJ data. Specifically, the study raises concerns about the assumptions of equal discriminal dispersions and zero correlation between the stimuli. While these assumptions simplify the trait measurement model, they may fail to capture the complexity of CJ data, potentially leading to unreliable and inaccurate trait estimates. Second, the study highlights the apparent disconnect between CJ's trait measurement and hypothesis testing processes. Although separating these processes simplifies the analysis of CJ data, it may also undermine the reliability of various statistical results derived from these processes.

  To address these issues, the study extends Thurstone's general form using a systematic and integrated approach based on Causal and Bayesian inference methods. This extension integrates core theoretical principles alongside key CJ assessment design features, such as the selection of judges, stimuli, and comparisons. It then translates these elements into a probabilistic statistical model for analyzing dichotomous CJ data, overcoming the rigid assumptions of Case V and the BTL model.

  Finally, the study emphasizes the relevance of this extension for contemporary empirical CJ research, particularly stressing the need for bespoke CJ models tailored to the experiments and data assumptions. It also lays the foundation for broader applications, encouraging researchers across the social sciences to adopt more robust and interpretable methodologies.
key-points:
  - (to do)
date: last-modified
bibliography: references.bib
---

<!-- commands for d-separation -->
\newcommand{\dsep}{\:\bot\:}
\newcommand{\ndsep}{\:\not\bot\:}

<!-- abstract -->
<!-- recording: Sven and Tine 24.11.25; time: 00:58:20 - 00:59:20 -->
<!-- recording: Sven and Tine 24.12.17; time: 00:11:50 - 00:26:10, (Here we did it!) -->

# Introduction {#sec-introduction}

<!-- recording: Sven 24.10.04; time: 00:00:00 - 00:09:10 -->
<!-- recording: Sven 24.10.04; time: 00:14:30 - 00:21:00 -->
<!-- recording: Sven and Tine 24.10.18; time: 00:04:05 - 00:08:30 -->
<!-- recording: Sven and Tine 24.10.25; time: 00:00:00 - 00:17:00 -->
<!-- recording: Sven and Tine 24.10.25; time: 00:00:00 - 00:17:00 -->
<!-- recording: Sven and Tine 25.01.22; time: 00:06:30 - 00:08:25 -->

<!-- 1. Overview of CJ method in assessing traits -->
In *comparative judgment* (CJ) studies, judges assess a specific trait or attribute across different stimuli by performing pairwise comparisons [@Thurstone_1927a; @Thurstone_1927b]. Each comparison produces a dichotomous outcome, indicating which stimulus is perceived to have a higher trait level. For example, when assessing writing quality, judges compare pairs of written texts (the stimuli) to determine the relative writing quality each text exhibit (the trait) [@Laming_2004; @Pollitt_2012b; @Whitehouse_2012; @vanDaal_et_al_2016; @Lesterhuis_2018_thesis; @Coertjens_et_al_2017; @Goossens_et_al_2018; @Bouwer_et_al_2023].

<!-- and spoken language [@Boonen_et_al_2020]. Additionally, it has been applied to evaluate various competencies, including mathematical problem-solving skills [@Jones_et_al_2015], engineering design skills [@Bartholomew_et_al_2018], build real-time web-based portfolios of performance [@@Kimbell_2012], conceptual understanding in algebra [@Jones_et_al_2019], statistical and English knowledge [@Marshall_et_al_2020], and STEM knowledge and skills [@Bartholomew_et_al_2020b]. -->

<!-- 2. What is the effectiveness of CJ -->
Numerous studies have documented the effectiveness of CJ in assessing traits and competencies over the past decade. These studies have highlighted three aspects of the method's effectiveness: its reliability, validity, and practical applicability. Research on reliability suggests that CJ requires a relatively modest number of pairwise comparisons [@Verhavert_et_al_2019; @Crompvoets_et_al_2022] to generate trait scores that are as precise and consistent as those generated by other assessment methods [@Coertjens_et_al_2017; @Goossens_et_al_2018; @Bouwer_et_al_2023]. In addition, the evidence suggests that the reliability and time efficiency of CJ are comparable, if not superior, to those of other assessment methods when employing adaptive comparison algorithms [@Pollitt_2012b; @Verhavert_et_al_2022; @Mikhailiuk_et_al_2021]. Meanwhile, research on validity indicates the capacity of CJ scores to represent accurately the traits under measurement [@Whitehouse_2012; @vanDaal_et_al_2016; @Lesterhuis_2018_thesis; @Bartholomew_et_al_2018; @Bouwer_et_al_2023]. Lastly, research on its practical applicability highlights CJ's versatility across both educational and non-educational contexts [@Kimbell_2012; @Jones_et_al_2015; @Bartholomew_et_al_2018; @Jones_et_al_2019; @Marshall_et_al_2020; @Bartholomew_et_al_2020b; @Boonen_et_al_2020].

<!-- rubrics [@Coertjens_et_al_2017; @Goossens_et_al_2018]  -->

<!-- holistic benchmark ratings [@Bouwer_et_al_2023] -->

<!-- The A for ‘adaptivity’ in ACJ means that the choice of which objects are presented to the judges depends on the outcomes of judgments made so far – the idea being to use judge time efficiently by not presenting objects to compare where the result of the comparison is almost certain. [@Bramley_2015, pp. 4].  -->

<!-- 3. But what critical issues remain in CJ Research? -->
Nevertheless, despite the increasing number of CJ studies, research in this domain remains unsystematic and fragmented, leaving several critical issues unresolved. This study identifies and discusses two prominent issues of traditional approaches that can undermine the reliability and validity of CJ's trait estimates. First, it critiques the heavy reliance on Thurstone's Case V assumptions [@Thurstone_1927b] and, by extension, the Bradley-Terry-Luce (BTL) model [@Bradley_et_al_1952; @Luce_1959] when analyzing CJ data. Specifically, the study raises concerns about the assumptions of equal discriminal dispersions and zero correlation between the stimuli. While these assumptions simplify the trait measurement model, they may fail to capture the complexity of some traits or account for heterogeneous stimuli, potentially leading to unreliable and inaccurate trait estimates. Second, the study highlights the disconnect between CJ's trait measurement and hypothesis testing processes. Although separating these processes simplifies the analysis of CJ data, it may also undermine the reliability of various statistical inferences derived from these processes.

<!-- Reliability is a necessary but not sufficient condition for validity. Reliability can exist without validity but validity cannot exist without reliability. [@Perron_et_al_2015, pp. 2] -->

<!-- 5. So, what are we doing? -->
To address these issues, this study extends Thurstone's general form through a systematic and integrated approach that combines causal and Bayesian inference methods. In addition to potentially enhancing measurement reliability and validity, and improving statistical accuracy in hypothesis testing, this approach offers two key advantages. First, it clarifies the interactions among all actors and processes involved in CJ assessments. Second, it shifts the current comparative data analysis paradigm from passively accepting the BTL model assumptions to actively testing whether those assumptions fit the data under analysis.

<!-- Reliability is a necessary but not sufficient condition for validity. Reliability can exist without validity but validity cannot exist without reliability. [@Perron_et_al_2015, pp. 2] -->

<!-- Many cross-cultural scientists already pay close attention to concerns of causal inference and comparison without use of a formal framework (e.g., Norenzayan & Heine, 2005; Pollet et al., 2014). For these researchers, a formalized framework can provide a vocabulary to articulate their concerns and work toward solutions in a more systematic manner. [@Deffner_et_al_2022, pp. 2] -->

<!-- Specifically, a framework is needed which lets us reason about the causes of non-invariance. [@Sterner_et_al_2024, pp. 751] -->

<!-- 4. what is the organization of the study? -->
As a result, the study divides its content into six main sections. @sec-thurstone_theory provides an overview of Thurstone's theory. @sec-theory-issues discusses the identified issues in detail. @sec-theoretical extends Thurstone's general form to address these challenges. The extension integrates core theoretical principles alongside key CJ assessment design features, such as the selection of judges, stimuli, and comparisons. @sec-statistical translates these theoretical and practical elements into a probabilistic statistical model to analyze dichotomous pairwise comparison data. Finally, @sec-discussion discusses the findings, explores avenues for future research, and detail the challenges for future researchers.

# Thurstone's theory {#sec-thurstone_theory}
<!-- recording: Sven and Tine 24.11.18; time: 00:04:30 - 00:14:50 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:19:04 - 00:21:37 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:00:35 - 00:06:55 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:30:35 - 00:35:10 -->
<!-- recording: Sven and Tine 24.12.09; time: 00:00:00 - 00:03:20 -->

<!-- 1. Thurstone's Theory of CJ: discriminal process -->
In its most general form, Thurstone’s theory addresses pairwise comparisons wherein a single judge evaluates multiple stimuli [@Thurstone_1927b]. The theory posits that two key factors determine the dichotomous outcome of these comparisons: the discriminal process of each stimulus and their discriminal difference. The *discriminal process* captures the psychological impact each stimulus exerts on the judge or, more simply, his perception of the stimulus trait. The theory assumes that the discriminal process for any given stimulus forms a Normal distribution along the trait continuum [@Thurstone_1927b]. The mode (mean) of this distribution, known as the *modal discriminal process*, indicates the stimulus position on this continuum, while its dispersion, referred to as the *discriminal dispersion*, reflects variability in the perceived trait of the stimulus.

<!-- The object of this paper is to describe a new psychophysical law which may be called the law of comparative judgment and to show some of its special applications in the measurement of psychological values. [@Thurstone_1927b, pp. 266] -->

<!-- It applies fundamentally to the judgments of a single observer who compares a series of stimuli by the method of paired comparison when no 'equal' judgments are allowed [@Thurstone_1927b, pp. 267] -->

<!-- The psychological continuum or scale is so constructed or defined that the frequencies of the respective discriminal processes for any given stimulus form a normal distribution on the psychological scale. This involves no assumption of a normal distribution or of anything else. The psychological scale is at best an artificial construct. If it has any physical reality, we certainly have not the remotest idea of what it may be like. ... We define the scale in terms of the frequencies of the discriminal processes for any stimulus. This artificial construct, the psychological scale, is so spaced off that the frequencies of the discriminal processes for any given stimulus form a normal distribution on the scale.[@Thurstone_1927b, pp. 266] -->

<!-- 2. Example of discriminal processes -->
@fig-discriminal_process illustrates the hypothetical discriminal processes along a quality trait continuum for two written texts. The figure indicates that the modal discriminal process for Text B is positioned further along the continuum than that of Text A $(T_{B} > T_{A})$, suggesting that Text B exhibits higher quality. Additionally, the figure highlights that Text B has a broader distribution compared to Text A, which arises from its larger discriminal dispersion $(\sigma_{B} > \sigma_{A})$.

::: {#fig-thurstone_theory layout-ncol=2}

![Discriminal processes](/images/png/discriminal_process.png){#fig-discriminal_process width=100%}

![Discriminal difference](/images/png/discriminal_difference.png){#fig-discriminal_difference width=100%}

Hypothetical discriminal processes and discriminant difference along a quality trait continuum for two written texts.
:::

<!-- 3. Thurstone's Theory of CJ: discriminal difference -->
However, since the individual discriminal processes of the stimuli are not directly observable, the theory introduces the *law of comparative judgment*. This law posits that in pairwise comparisons, a judge perceives the stimulus with a discriminal process positioned further along the trait continuum as possessing more of the trait [@Bramley_2008]. This suggests that pairwise comparison outcomes depend on the relative distance between stimuli, not their absolute positions on the continuum. Indeed, the theory assumes that the difference between the underlying discriminal processes of the stimuli, referred to as the *discriminal difference*, determines the observed dichotomous outcome. Furthermore, the theory assumes that because the individual discriminal processes form a Normal distribution on the continuum, the discriminal difference will also conform to a Normal distribution [@Andrich_1978]. In this distribution, the mode (mean) represents the average relative separation between the stimuli, and its dispersion indicates the variability of that separation.

<!-- The mode (mean) of this distribution, representing the (average) relative separation, is given by the difference between the modal discriminal processes of the stimuli $S_{BA}=S_{B}-S_{A}$. Meanwhile, the dispersion of the distribution, reflecting the variability in the relative separation, is calculated as $\sigma_{BA} = \sqrt{\sigma_{B}^{2} + \sigma_{A}^{2} - \rho\sigma_{B}\sigma_{A}}$. Here, $\sigma_{B}$ and $\sigma_{A}$ denote the discriminal dispersions of the stimuli, while $\rho$ represents the correlation between their discriminal processes. This correlation quantifies the dependence of the judge's perception of the trait in one stimulus on his perception of the same trait in another. -->

<!-- The discriminal mode and dispersion corresponding to a single stimulus are inaccessible to observation. They can only be estimated when two objects are compared. Thurstone assumed that when two objects are compared with respect to a specified attribute, the object evoking the discriminal process further along the psychological continuum would be judged as possessing more of the attribute. ... The outcome of the paired comparison judgement is therefore related to the distribution of the difference between the two distributions of discriminal processes for essay A and essay B. [@Bramley_2008, pp. 251] -->

<!-- When subject v compares two statements, j and i, he/she reports that item j reflects more of the property of D_{vj}, - D_{vi}, > 0. Since D_{vj} and D_{vi}, are assumed normally distributed, the difference process ... is normally distributed with mean a_{j} - a_{i}, and variance s^{2}_{ji} [@Andrich_1978, pp. 452] -->

<!-- 4. Example of discriminal difference -->
@fig-discriminal_difference illustrates the distribution of the discriminal difference for the two hypothetical texts. The figure indicates that the judge perceives Text B as having significantly higher quality than Text A. Two key observations support this conclusion: the positive difference between their modal discriminal processes $(T_{B} - T_{A} > 0)$ and the probability area where the discriminal difference distinctly favors Text B over Text A, represented by the shaded gray area denoted as $P(B > A)$. As a result, the dichotomous outcome of this comparison is more likely to favor Text B over A.

# Two Prominent Issues in Traditional CJ Practice {#sec-theory-issues}

<!-- 1. What is the problem with the general form -->
Thurstone noted from the outset that his general formulation, described in @sec-thurstone_theory, led to a *trait scaling problem*. Specifically, the model required estimating more "unknown" parameters than the number of available pairwise comparisons [@Thurstone_1927b]. For instance, in a CJ assessment with five texts, the general form would require estimating $20$ parameters: five modal discriminal processes, five discriminal dispersions, and $10$ correlations--one per comparison (see @tbl-thurstone_cases). However, a single judge could only provide ${5 \choose 2} = 10$ unique comparisons, an insufficient data set to estimate the required parameters.

<!-- According to this equation every pair of stimuli presents the possibility of a different correlation between the discriminal deviations. If this degree of freedom is allowed, the problem of psychological scaling would be insoluble because every observation equation would introduce a new unknown and the number of unknowns would then always be greater that the number of observation equations. In order to make the problem soluble it is necessary to make at least one assumption, namely that the correlation between discriminal deviations is practically constant throughout the stimulus series and for the single observer. [@Thurstone_1927b, pp. 267] -->

<!-- 2. Thurstone's Cases and their simplifying assumptions -->
To address this issue and facilitate the practical implementation of the theory, Thurstone developed five cases derived from this general form, each progressively incorporating additional simplifying assumptions [@Thurstone_1927b]. In Case I, Thurstone postulated that pairs of stimuli would maintain a constant correlation across all comparisons. In Case II, he allowed multiple judges to undertake comparisons instead of confining evaluations to a single judge. In Case III, he posited that there was no correlation between stimuli. In Case IV, he assumed that the stimuli exhibited similar dispersions. Finally, in Case V, he replaced this assumption with the condition that stimuli had equal discriminal dispersions. @tbl-thurstone_cases summarizes the assumptions of the general form and the five cases. For a detailed discussion of these cases and their progression, refer to @Thurstone_1927b and @Bramley_2008 [pp. 248–253].

<!-- For the practical application of the law of comparative judgment we shall consider five cases which differ in assumptions, approximations, and degree of simplification. The more assumptions we care to make, the simpler will be the observation equations. [@Thurstone_1927b, pp. 267] -->

<!-- In Case II, Thurstone assumes that "each judge compares each stimulus with every other stimulus only once" [@Thurstone_1927b, pp. 268]. -->

![Thurstones cases and their asumptions](/images/png/thurstone_cases.png){#tbl-thurstone_cases width=100%}

<!-- 3. But are our specific problems? -->
However, Thurstone developed Case V prioritizing statistical simplicity over precise trait measurement and offering no guidance on how to use its trait estimates for statistical inference or hypothesis testing. Specifically, Thurstone cautioned that its use "should not be made without (an) experimental test" [@Thurstone_1927b, pp. 270], as it imposes the most extensive set of simplifying assumptions [@Bramley_2008; @Kelly_et_al_2022] (see @tbl-thurstone_cases). Moreover, because Thurstone's primary goal was to produce a "rather coarse scaling" of traits and "allocate the compared stimuli on this continuum" [@Thurstone_1927b, pp. 269], his theory did not support formal statistical inference. Despite these limitations, it is surprising that CJ research has predominantly relied on Case V to measure different traits, which raises significant concerns about the reliability and validity of such measurements in contexts where the case's assumptions may not hold [@Kelly_et_al_2022; @Andrich_1978]. Furthermore, although the CJ tradition has attempted to address the gap of hypothesis testing by using the point estimates of the traits--or their transformations--a critical question remains: Does this approach provide a valid foundation for statistical inference?

<!-- Case V - This is the simplest formulation of the law and it involves, in addition to the previous assumptions, the assumption that all the discriminal dispersions are equal. This assumption should not be made without experimental test. [@Thurstone_1927b, pp. 270] -->

<!-- Case 5 makes the greatest simplification by assuming that the discriminal dispersions are all equal [@Bramley_2008, pp. 253] -->

<!-- Adding assumptions simplifies the formula and improves its tractability, but one cannot place as much confidence in the outputs of the formula if those assumptions are not met [@Kelly_et_al_2022, pp. 677]. -->

<!-- This is a simple observation equation which may be used for rather coarse scaling . .... This law defines the psychological scale or continuum. It allocates the compared stimuli on the continuum [@Thurstone_1927b, pp. 269] -->

<!-- Consequently, different samples, whether drawn randomly or otherwise from the same empirical population, should give the same relative scale value estimates. In this sense the models are also "sample-free." 
However, this property of the model is true in empirical data only if the model holds in the data. Therefore, "sample-freeness" is always only a hypothesis until checked, as is the assumption that the model holds. [@Andrich_1978, pp.459] -->

<!-- 4. Section division -->
Thus, this section discusses these two prominent issues. Specifically, @sec-theory-issue1 examines the heavy reliance on Thurstone's Case V assumptions in the statistical analysis of CJ data. Conversely, @sec-theory-issue2 focuses on the apparent disconnect between the approaches to trait measurement and hypothesis testing in CJ.

## The Case V and the statistical analysis of CJ data {#sec-theory-issue1}

<!-- recording: Sven and Tine 24.10.25; time: 00:31:10 - 00:52:20 -->
<!-- recording: Sven and Tine 24.11.12; time: 00:00:00 - 00:00:00 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:00:25 - 00:04:30 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:14:50 - 00:19:04 -->
<!-- recording: Sven and Tine 24.11.18; time: 00:21:37 - 00:47:20 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:06:55 - 00:24:50 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:35:10 - 00:49:30 -->
<!-- recording: Sven and Tine 24.12.09; time: 00:03:20 - 00:08:10 -->
<!-- recording: Sven and Tine 24.12.17; time: 00:00:00 - 00:06:10 -->

<!-- 1. The Case V and its simpler statistical representation: the BTL model -->
As previously discussed, Case V remains the most widely used model in CJ literature. This preference largely stems from the widespread adoption of the BTL model, which provides a simplified statistical representation of the case. The BTL model mirrors most of Case V's assumptions, with one notable distinction. While Case V assumes a Normal distribution for the stimuli' discriminal processes, the BTL model uses the more mathematically tractable Logistic distribution [@Andrich_1978; @Bramley_2008] (see @tbl-thurstone_cases). However, this substitution has minimal impact on trait estimation or model interpretation because the scale of the discriminal process (i.e., the latent trait) is arbitrary up to a non-monotonic transformation [@vanderLinden_et_al_2017_I; @McElreath_2021]. That is, as long as the substitution (transformation) preserves the data rank order, the choice of distribution for the discriminal processes is inconsequential. This condition is satisfied in this case, as the Normal and Logistic distributions exhibit analogous statistical properties, differing only by a scaling factor of approximately $1.7$ [@vanderLinden_et_al_2017_I].

<!-- Model misspecification happens when the set of probability distributions considered by the statistician does not include the distribution that generated the observed data. -->

<!-- An alternative to the normal density function is the logistic. The substitution of the logistic for the normal is commonly made in circumstances such as the above (cf. Birnbaum, 1968; Bock, 1975) because it is more tractable. [@Andrich_1978, pp. 455] -->

<!-- It will perhaps not come as a great surprise to the reader to discover that it is the Case 5 version of Thurstone’s law that has usually been used in cross-moderation exercises. The only difference is in the mathematical function linking the observed proportions to the difference in scale values – the more tractable logistic function is used rather than the cumulative Normal. [@Bramley_2008, pp. 254] -->

<!-- Finally, it is still not uncommon to find a version of the model with a scale constant D = 1.7 added to it, giving it 1.7a_{i}(θ_{p} − b_{i}) as its core structure. The original purpose of this practice was to bring the shape of the response functions close to that of the normal-ogive model, generally believed to be the “true model” that would give us a scale with “equal units of measurement” in the early days of test theory (Volume One, Chapter 1). However, as discussed below, the scale for the θp parameters is arbitrary up to a non monotonic transformation, and the practice of bringing it close to the scale of a normal ogive is a relict from a distant past, bound to create communication problems between users of item response theory (IRT) rather than add to their understanding of it. [@vanderLinden_et_al_2017_I, pp. 16] -->

<!-- the scale of the latent variable is arbitrary, so there is no loss of generality [@McElreath_2021] 02:40:30 - 02:40:45  -->

<!-- ::: {#fig-logistic_vs_normal layout-nrow=2} -->

<!-- ![Probability density](/images/png/density.png){#fig-density width=70%} -->

<!-- ![Cumulative probability](/images/png/cumulative.png){#fig-cumulative width=70%} -->

<!-- Probability density and cumulative probability of the logistic and Normal distributions. -->
<!-- ::: -->

<!-- 2. But what are limitations of Case V (and the BTL model)? -->
However, Thurstone acknowledged that some assumptions of Case V could be problematic when researchers assess complex traits or heterogeneous stimuli [@Thurstone_1927a]. Thus, given that modern CJ applications often involve such traits and stimuli, two key assumptions of Case V, and by extension, the BTL model, may not always hold in theory or practice. These assumptions are the equal dispersion and zero correlation between stimuli.

### The assumption of equal dispersions between stimuli {#sec-theory-issue1a}

<!-- recording: Sven and Tine 24.12.09; time: 00:08:10 - 00:37:00 -->
<!-- recording: Sven and Tine 24.12.17; time: 00:06:10 - 00:08:50 -->
<!-- recording: Sven and Tine 25.03.31; time: 00:01:20 - 00:34:00 -->

<!-- 1. What does the equal dispersion between stimuli imply? -->
According to the theory, discrepancies in the discriminal dispersions of stimuli shape the distribution of the discriminal difference, directly influencing the outcome of pairwise comparisons. A thought experiment can help illustrate this idea. In it, researchers observe the discriminal processes for two texts, A and B, assuming that the dispersion for Text A remains constant and that the two texts are uncorrelated $(\rho=0)$. @fig-dispersion demonstrates that an increase in the uncertainty associated with the perception of Text B relative to Text A $(\sigma_{B} - \sigma_{A})$, broadens the distribution of their discriminal difference. This broadening affects the probability area where the discriminal difference distinctly favors Text B over Text A, expressed as $P(B>A)$, ultimately influencing the comparison outcome. Additionally, the figure reveals that when the discriminal dispersions of the texts are equal, as in the BTL model $(\sigma_{B} - \sigma_{A}=0)$, the discriminal difference distribution is more narrow than when the dispersions differ. As a result, the discriminal difference is more likely to favor Text B over Text A, as it is represented by the shaded gray area.

::: {#fig-casev_issues layout-ncol=2}

![Discriminal Difference distribution under varying discrepancies in stimuli dispersions](/images/png/dispersion.png){#fig-dispersion width=100%}

![Discriminal Difference distribution under varying levels of correlation between stimuli](/images/png/correlation.png){#fig-correlation width=100%}

The effect of dispersion discrepancies and stimuli correlation on the distribution of the discriminal difference.
:::

<!-- 2. But this is not the scenario we observe -->
In experimental practice, however, the thought experiment occurs in reverse. Researchers first observe the comparison outcome and then use the BTL model to infer the discriminal difference between stimuli and their respective discriminal processes [@Thurstone_1927a]. Consequently, the outcome's ability to reflect *true* differences between stimuli largely depends on the validity of the model's assumptions [@Kohler_et_al_2019], in this case, the assumption of equal dispersions. For instance, when the assumption accurately captures the complexity of the data, the BTL model estimates a discriminal difference distribution that accurately represents the *true* discriminal difference between the texts. This scenario is illustrated in @fig-dispersion, when the model's discriminal difference distribution aligns with the *true* discriminal difference distribution, represented by the thick continuous line corresponding to $\sigma_{B}-\sigma_{A}=0$. The accuracy of this discriminal difference then ensures reliable estimates for the texts’ discriminal processes.

<!-- In experimental practice the procedure is the reverse of this hypothesis because the frequencies are known first experimentally, and from these frequencies we construct the psychological continuum. [@Thurstone_1927a, pp. 373] -->

<!-- Statements about the real world from observed data require assumptions. This is true no matter what kind of data we have, and it is explicitly also true for statements based on data stemming from SRSg. The validity of our statements therefore critically hinges on the validity of those assumptions. If our statistical assumptions are wrong for the analysis at hand, the inferential statements will be incorrect no matter how sophisticated the statistical methods are. [@Kohler_et_al_2019, pp. 150] -->

<!-- 3. But does it hold in practice? -->
Notably, while assuming equal dispersions simplifies the trait measurement model, evidence from the CJ literature suggests that this assumption may fail accommodate for heterogeneous stimuli, such as handwritten texts or English compositions [@Thurstone_1927a; @Andrich_1978; @Bramley_2008; @Kelly_et_al_2022]. Indeed, the presence of the so-called *misfit texts*--texts that elicit more judgment discrepancies than others [@Pollitt_2004; @Pollitt_2012b; @Pollitt_2012a; @Goossens_et_al_2018]--may signal this limitation, as such discrepancies can arise from larger discriminal dispersions due to stimulus heterogeneity or because the texts are genuine outliers--that is, texts with distinctive characteristics that deviate markedly from the rest of the sample [@Grubbs_1969]. In either case, the BTL model's assumptions prevent it from adequately accounting for or addressing these anomalies, leaving exclusion of the "problematic" texts as the primary remedy.

<!-- It is probably true that this variability of the discriminal dispersion on the psychological continuum is of relatively less serious importance in dealing with strictly homogeneous stimulus series, but it becomes a serious factor in dealing with less conspicuous attributes or with less homogeneous stimulus series, such as handwriting specimens, English compositions, sewing samples, oriental rugs. In measurements of the type known as judgment scales the discriminal dispersion on the psychological continuum becomes one of the unknowns to be determined as well as the scale value of the specimen. Every specimen in such a series presents two unknown values to be determined, namely, the scale value of its modal discriminal process on the psychological continuum and its discriminal dispersion. [@Thurstone_1927a, pp. 376] -->

<!-- Of these, the most questionable is the one that the discriminal dispersions of all the objects are equal. Thurstone clearly did not expect this to hold for any but the most simple stimuli. The scripts used in cross-moderation exercises are obviously far more complex than any used by Thurstone, so it seems naive to expect this assumption to hold here [@Bramley_2008, pp. 273]. -->

<!-- It also seems unclear whether Thurstone’s work really can be applied to the stimuli typically used in assessment. While Thurstone used his Law to create measurement scales for purely psychological or subjective entities, he did so only with stimuli whose values could be evaluated relatively instantaneously, such as short statements and handwriting samples. He was also skeptical about its use for very heterogeneous stimuli (Thurstone, 1927b) [@Kelly_et_al_2022, pp. 678]. -->

<!-- As well as an estimate of its value parameter and standard error the analysis will report, for each script, a misfit statistic. Essentially this reports the amount of inconsistency in the various judgements that have been made of that script. [@Pollitt_2004, pp. 11]  -->

<!-- A similar statistic is calculated for every script. A misfitting script is one that judges find difficult for some reason, one that some judges value more than others do. [@Pollitt_2012b, pp. 290] -->

<!-- If a portfolio’s WMS exceeds the criterion of mean plus two standard deviations, this means the judges did not judge it consistently, some considering it ‘better’ than others did. [@Pollitt_2012a, pp. 165] -->

<!-- Representations with a large infit are representations which lead to more inconsistent judgments [@Goossens_et_al_2018, pp. 20] -->

<!-- An outlying observation, or "outlier," is one that appears to deviate markedly from other members of the sample in which it occurs. [@Grubbs_1969, pp. 1] -->

<!-- For instance, in evaluating creativity, a student might complete an assignment in a highly creative but unexpected manner. Judges, unprepared for such originality, may produce more divergent evaluations of the assignment. These increased discrepancies in judgments suggest that the discriminal difference associated with the assignment has a broader dispersion, indicating greater variation in the discriminal process for this assignment compared to others. -->

<!-- "weak" writers consistently produce texts of low quality in all aspects that determine the quality of a text, such as grammatical accuracy, organization, and language use. In such cases, judges typically agree on their inferior quality, leading to a clear disfavoring of these texts compared to others. This agreement ultimately imply narrow dispersions for the discriminal differences related to these texts and, by extension, narrow discriminal processes for them. However, as texts improve along the trait continuum, with certain aspects showing improvement and others reaming unaltered, these variations introduce uncertainty into the judgments. Now, since judges often rely on various intricate characteristics of the texts to form their judgments [@vanDaal_et_al_2016; @Lesterhuis_et_al_2018; @Chambers_et_al_2022], these changes lead to greater discrepancies in their evaluations. This divergence ultimately imply broader discriminal differences and, by extension, more varied discriminal processes for the texts. -->

<!-- So it seem that the extreme of the discriminal process distribution of the stimuli have higher reliability than the center? -->

<!-- 5. So, what happens when the assumption does not hold? -->
Significant statistical and measurement issues can arise when the assumption of equal dispersions between stimuli does not hold. Specifically, the BTL model may overestimate the trait's reliability, that is, the degree to which the outcome accurately reflects the *true* discriminal differences between stimuli. This overestimation, in turn, results in spurious conclusions about these differences [@McElreath_2020; @Wu_et_al_2022] and, by extension, about the underlying discriminal processes of stimuli. @fig-dispersion also illustrates this scenario when the model's discriminal difference distribution aligns with the thick continuous line for $\sigma_{B}-\sigma_{A}=0$, while the *true* discriminal difference follows any discontinuous line where $\sigma_{B}-\sigma_{A} \neq 0$. Furthermore, if researchers acknowledge that misfit statistics may represent texts with different dispersions or outlying observations, the common CJ practice of excluding stimuli based on these statistics, as seen in @Pollitt_2012a, @Pollitt_2012b, @vanDaal_et_al_2016, and @Goossens_et_al_2018, may unintentionally discard valuable information, introducing bias into the trait estimates [@Zimmerman_1994; @McElreath_2020]. The direction and magnitude of these biases remain unpredictable, as they depend on which stimuli researchers exclude from the analysis.

<!-- The results that I have described in the present article reveal that outliers in mixed-normal distributions influence both parametric and nonparametric tests. The influence on the parametric test is greater, but the direction of the changes in probability of Type 1 and Type I1 errors is the same for both. The typical power function is depressed. irrespective of whether the analysis is based on parametric or nonparametric methods [@Zimmerman_1994, pp. 397-398] -->

<!-- What can be done about this? There is a tradition of dropping outliers. People sometimes drop outliers even before a model is fit, based only on standard deviations from the mean outcome value. You should never do that—a point can only be unexpected and highly influential in light of a model. After you fit a model, the picture changes. If there are only a few outliers, and you are sure to report results both with and without them, dropping outliers might be okay. But if there are several outliers and we really need to model them, what then? ... One way to both use these extreme observations and reduce their influence is to employ some kind of robust regression. [@McElreath_2020, pp. 232-233] -->

<!-- But ignoring over-dispersion can also lead to all of the same problems as ignoring any predictor variable. Heterogeneity in counts can be a confound, hiding effects of interest or producing spurious inferences. [@McElreath_2020, pp. 370] -->

### The assumption of zero correlation between stimuli {#sec-theory-issue1b}

<!-- recording: Sven and Tine 24.12.17; time: 00:08:50 - 00:10:30 -->

<!-- 1. What does the zero correlation between stimuli imply? -->
The correlation $\rho$, illustrated in @fig-discriminal_difference, measures how much the judges' perception of a specific trait in one stimulus depends on their perception of the same trait in another. Similar to the discriminal dispersions, this correlation shapes the distribution of the discriminal difference, directly impacting the outcomes of pairwise comparisons. A thought experiment, akin to the one presented in @sec-theory-issue1a, can illustrate this idea. Assuming that the discriminal dispersions for both texts remain constant, @fig-correlation shows that as the correlation between the two texts increases, the distribution of their discriminal difference becomes narrower. This narrowing, in turn, affects the probability that the discriminal difference distinctly favors Text B over Text A--denoted as $P(B > A)$--and thus directly influences the comparison outcome. Furthermore, the figure shows that when two texts are independent or uncorrelated, as assumed in the BTL model $(\rho=0)$, the distribution of their discriminal difference is less narrow than in scenarios where the texts are positively correlated. As a result, it becomes less likely for the comparison to favor Text B over Text A, as indicated by the larger shaded area.

<!-- "It is a safe assumption that when the stimulus series is very homogeneous with no distracting attributes, the correlation between discriminal deviations is low and possible even zero" [@Thurstone_1927b, pp. 268]. -->

<!-- 2. But this is not the scenario we observe -->
<!-- As outlined in the previous section, researchers approach this process in reverse. They begin by observing the outcomes and using the BTL model to estimate the stimuli' discriminal differences and discriminal processes. Given that the BTL model assumes independent discriminal processes across comparisons, if this assumption holds in the data, the model estimates a discriminal difference distribution that accurately reflects the *true* discriminal difference between the texts. Once more, the accuracy of the discriminal difference ensures reliable estimates for the discriminal processes of the texts. -->

<!-- It then follows that the outcome's ability to represent the "true" differences between stimuli largely depends on the validity of the assumption of zero correlation. -->

<!-- 3. But does it hold?  -->
Despite these notable differences in the distribution of the discriminal difference under various correlational assumptions, in practice, assessment designs often adopt the assumption of no correlation between stimuli based on an old theoretical justification. Specifically, @Thurstone_1927b argued that stimuli could be treated as uncorrelated because judges' biases--arising from two opposing and equally weighted effects occurring during the pairwise comparisons--would cancel each other out. This idea was later formalized by @Andrich_1978, who provided a mathematical demonstration of this cancellation using the BTL model under the assumption of discriminal processes with additive biases. However, evidence from the CJ literature indicates that the assumption of zero correlation does not hold in practice in at least two scenarios: when intricate aspects of multidimensional, complex traits or heterogeneous stimuli influence judges' perceptions or when additional hierarchical structures are relevant to the stimuli.

<!-- If we accept that correlation as zero, we are really assuming that the degree of excellence which an observer perceives in one of the specimens has no influence on the degree of excellence that he perceives in the comparison specimen. There are two effects that may be operative here and which are agnostic to each other. ... mood and the opposite effect of simultaneous contrast .... [@Thurstone_1927b, pp. 268] -->

<!-- The correlation $\rho$, considered above is not the kind of correlation considered by Bock (1958) in studies of preference and choice of objects. The Bock correlation arises from a specific interaction between a subject and an item because the subject has a particular "preference", such as a taste preference, for the object. This preference for the object is then reflected in its comparisons with all other objects. However, it is less likely that subjects would have such particular preferences for some items in a test and, in any case, the assumption is open to empirical checking [@Andrich_1978, pp. 455]. -->

<!-- 2. The first scenario where the assumption does not hold -->
In the first scenario, research on text quality suggests that when judges evaluate complex, multidimensional traits or heterogeneous stimuli, they often rely on a variety of intricate stimulus characteristics to inform their judgments [@vanDaal_et_al_2016; @Lesterhuis_et_al_2018; @Chambers_et_al_2022]. Regardless of their relevance, these characteristics may not receive equal weight or consistently oppose one another across comparisons. As a result, they may exert a disproportionate influence on judges' perceptions, generating biases that persist rather than cancel out. For example, this could occur when a judge assessing the argumentative quality of a text may place disproportionate emphasis on handwriting clarity, thereby favoring neatly written texts despite their weaker arguments. Moreover, because the discriminal process of stimuli becomes an observable outcome only through the judges' perceptions, these biases could introduce dependencies between the stimuli [@vanderLinden_et_al_2017_II]. While direct evidence for this exact scenario is limited, existing studies document the presence of judge bias in CJ [@Pollitt_et_al_2003, @vanDaal_et_al_2016, @Bartholomew_et_al_2020a], reinforcing the argument that the factors influencing pairwise comparisons do not always cancel each other out.

<!-- research shows that, in the presence of these traits and stimuli, the accuracy of the judges' trait perception is influenced by the rank-order distance of the stimuli [@vanDaal_et_al_2017; @Gijsen_et_al_2021], a coarse measure of the relative positioning of the stimuli within the trait continuum. -->

<!-- If the test taker’s position on one latent trait is fixed, the assumption of local stochastic independence requires the association between the responses to the items to vanish. Suppose the assumption is violated and more than one dimension is necessary to describe the test takers’ position in the latent space. The association between the responses to the items given one proficiency parameter does then not vanish. [@vanderLinden_et_al_2017_II, pp. 346]. -->

<!-- The elimination of the subject parameters in the Rasch model has given rise to such expressions "sample-free" and "population-free" In this sense, then, so is the Thurstone Case V model. Although these phrases suggest desirable properties, they are not always fully appreciated; therefore, these terms have been avoided so far. A clarification seems in order and possible in this context. 
Because subject parameters are eliminated explicitly, their values (individually or collectively) need not be known. As a result, no assumptions need be made about the distribution of these parameters in the population, for example, that they are normally distributed. This does not mean, of course, that the "empirical" subject population need not be specified or described or that the scale values will be the same in different empirical populations. In fact, it is generally vital that the empirical populations be described. But because it does not matter what the actual values of the subject parameters are in the population, the models are, in the statistical distribution sense of population,
"population-free" This is a property of the model and independent of empirical data.
It is important to note that this holds only with respect to the estimation of the relative scale values of the items. It does not matter if the subject parameters are large, small, or otherwise; the subjects' impressions of the relative positions of the items should be the same. That is, items which appear more intense to one group should appear similarly more intense to the other groups. Consequently, different samples, whether drawn randomly or otherwise from the same empirical population, should give the same relative scale value estimates. In this sense the models are also "sample-free."
However, this property of the model is true in empirical data only if the model holds in the data. Therefore, "sample-freeness" is always only a hypothesis until checked, as is the assumption that the model holds. Conversely, if the relative scale values are not sample-free, the model does not hold across these samples. This feature is often used in checks on the Rasch model, and the usual sample formation is according to total scores on the items. Therefore, the check is with respect to an interaction between relative scale values and the level of the subject parameters. This is one of the best checks since it is then known that the subject parameter values are different. However, other samples may also be formed (e.g., on the basis of sex); then the check is with respect to a sex by relative scale value interaction. If there is no interaction, then it can be inferred that the items are working in the same relative way in both groups. As a result, generalizability of the scale across the groups is demonstrated. [@Andrich_1978, pp.459] -->

<!-- 3. The second scenario where the assumption does not hold -->
In the second scenario, the shared context or inherent connections introduced by additional hierarchical structures may create dependencies between stimuli--a statistical phenomenon known as clustering [@Everitt_et_al_2010]. For instance, when the same individual produces multiple texts, those texts often exhibit common features, such as writing style or overall quality, that judges can easily recognize. In this regard, although the CJ literature acknowledges the existence of such hierarchical structures, the statistical approaches to account for this additional source of dependence have been insufficient. For instance, when CJ data incorporates multiple samples of stimuli from the same individuals, researchers frequently rely on (averaged) point estimates of the BTL scores to conduct subsequent analyses and tests at the individual level [@Bramley_et_al_2019; @Boonen_et_al_2020; @Bouwer_et_al_2023; @vanDaal_et_al_2017; @Jones_et_al_2019; @Gijsen_et_al_2021]. However, this approach can introduce additional statistical and measurement issues, similar to the ones discussed in @sec-theory-issue2.

<!-- for example, when comparing texts produced by university and secondary school students. In this case, university students may consistently (or more precisely) produce higher-quality texts, while secondary school students, who exhibit a broader range of writing abilities, would show greater variability in the quality of their texts. Although this example is somewhat contrived, it effectively illustrates how assuming equal dispersions across texts can overlook meaningful differences in the reliability of text quality across groups or individuals. -->

<!-- Clustered data: A term applied to both data in which the sampling units are grouped into clusters sharing some common feature, for example, animal litters, families or geographical regions, and longitudinal data in which a cluster is deﬁned by a set of repeated measures on the same unit. [@Everitt_et_al_2010, pp. 86] -->

<!-- 3. But what are the challenges when this assumption is violated? -->
Thus, erroneously assuming zero correlation between stimuli can also lead to significant statistical and measurement issues. Specifically, neglecting judges' biases or relevant hierarchical structures can create dimensional mismatches in the model, leading to the over- or underestimation of trait reliability [@Ackerman_1989; @Hoyle_et_al_2023] and even biases [@Wu_et_al_2022]. This issue is illustrated in @fig-correlation when the discriminal difference distribution of the BTL scores follows the thick continuous line $(\rho = 0)$, while the *true* discriminal difference follows any discontinuous line where $\rho \neq 0$. Moreover, these inaccuracies can result in spurious conclusions about the discriminal differences [@McElreath_2020] and, by extension, the underlying discriminal processes of the stimuli--potentially causing the model to incorrectly classify stimuli or judges as *misfits*.

<!-- The whole paper [@Ackerman_1989] -->

<!-- To show the practical consequences, in Figure 18.1, we present the test information for the two models; clearly, the information is positively biased when the multidimensional data are forced into a unidimensional IRT model. Moreover, in this same figure, we show the test response curves (how expected scores change as a function of trait level) for the two models; again, they, clearly, are not overlapping. The estimated implication is that if one applied a unidimensional IRT model, psychometric information is inflated due to the multidimensionality and examinee standard errors, which are the inverse square root of information, are too small. [@Hoyle_et_al_2023, pp. 340-341] -->

<!-- But ignoring over-dispersion can also lead to all of the same problems as ignoring any predictor variable. Heterogeneity in counts can be a confound, hiding effects of interest or producing spurious inferences. [@McElreath_2020, pp. 370] -->

Finally, similar to what we discussed in the previous section, removing *misfit* judges--that is, judges whose assessments deviate markedly from the shared consensus [@Pollitt_2012a; @Pollitt_2012b; @vanDaal_et_al_2016; @Goossens_et_al_2018; @Wu_et_al_2022], and may appear as outliers under the BTL model [@Wu_et_al_2022]--risk discarding valuable information and introducing bias into trait estimates. The direction and magnitude of these biases remain unpredictable, as they depend on which judges researchers exclude from the analysis [@Zimmerman_1994; @OHagan_2018; @McElreath_2020].

<!-- When a judge compares two portfolios, Eq. 2 can be used to measure the degree of ‘surprise’ in their decision. ... [@Pollitt_2012a, pp. 164-165]  -->

<!-- Conventionally, any figure greater than the mean plus two standard deviations – 1.14 here – is taken to indicate significant misfit. In this case, only Judge ‘46 N-M 2’ exceeded the criterion. To a significant degree, this judge interpreted the trait ‘good writing’ differently from the others. [@Pollitt_2012b, pp. 289-290] -->

<!-- Furthermore, in previous studies of CJ, judges are identified as taking decisions that deviate from the shared consensus (Pollitt, 2012b). These judges are labelled as ‘misfit’ and statistically identified by examining the extent to which judges make unexpected decisions (Pollitt, 2012b. See Pollitt, 2012a, 2012b for an extended explanation of the identification of misfitting judges). A misfitting judge can either be judging too inconsistently compared to the average judge by taking too many unexpected decisions or judging too consistently compared to the average judge. However, within CJ, misfit statistics have to be interpreted relatively. Misfit statistics quantify the extent to which judges are misfit compared to the whole of judges (Pollitt, 2012b). It is suggested that judges with the highest misfit statistic may have a diverging conceptualisation of the competence to be assessed (Bramley, 2007; Whitehouse & Pollitt, 2012). [@vanDaal_et_al_2016, pp. 4] -->

<!-- Pollitt [9] states that a large infit for raters suggests that they consistently judge away from the group consensus. Those raters are called misfits. As can be seen in Fig. 3 no rater misfits in this assessment. Representations with a large infit are representations which lead to more inconsistent judgments [33]. [@Goossens_et_al_2018, pp. 20] -->

<!-- Under the Bradley–Terry model, the object residual ri asymptotically follows a standard normal distribution, and is independent of \beta_i. Note that the asymptotic distribution of Rij holds when Vij grows large, so it is important to exclude residuals show Rij for which Vij = 0.
If the absolute value of ri is very large, object i is regarded as an outlier. By definition, this happens if object i unexpectedly wins over a popular object j (i.e., \beta_j is positive) and Rij is large and positive, or if object i unexpectedly loses to an unpopular object j (\beta_j negative) and Rij is large and negative.
Too many outlying objects may imply that the model suffers from a lack of fit. As Liner and Amin (2004) pointed out, this is more likely to happen if some of the objects are involved in only a few comparisons with most of the others. [@Wu_et_al_2022, pp. 466] -->

<!-- @Zimmerman_1994 the whole paper -->

<!-- Furthermore, these points tend to be outliers with unlikely values, according to the model. As a result, they are highly influential and make it difficult to estimate out-of-sample predictive accuracy. Why? Because any new sample is unlikely to contain these same outliers, and since these outliers were highly influential, they could make out-of-sample predictions worse than expected. [@McElreath_2020, pp. 231] -->

<!-- What can be done about this? There is a tradition of dropping outliers. People sometimes drop outliers even before a model is fit, based only on standard deviations from the mean outcome value. You should never do that—a point can only be unexpected and highly influential in light of a model. After you fit a model, the picture changes. If there are only a few outliers, and you are sure to report results both with and without them, dropping outliers might be okay. But if there are several outliers and we really need to model them, what then? [@McElreath_2020, pp. 232] -->

<!-- The problem isn’t necessarily that “outliers” are bad data. Rather processes are often variable mixtures and this results in thicker tails. [@McElreath_2020, pp. 369] -->

<!-- Consider, for example, the simplest location and scale problem-inference for a normal distribution with unknown mean \theta and variance \sigma^{2}. If we look at the marginal posterior distribution for \theta we know that if \sigma^{2} is known this posterior exhibits outlier-resistance, and we might expect similar behavior for unknown \sigma^{2}, but a simple example refutes this. Let the prior distribution give probability p (0<p<1) to (\theta = 1, \sigma^{2} = 1) and 1-p to (\theta = 0, \sigma^{2} = 4). Then the posterior mean of \theta is a decreasing function of any sufficiently large observation. [@OHagan_2018, pp. 366] -->

## The disconnect between trait measurement and hypothesis testing {#sec-theory-issue2}

<!-- recording: Sven 24.10.04; time: 00:14:30 - 00:21:00 -->
<!-- recording: Sven and Tine 24.10.18; time: 00:08:30 - 00:14:55 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:49:30 - 00:53:30 -->

<!-- 1. Utilizing BTL Scores for further analysis and hypothesis testing -->
Building on the previous section, it is clear that, researchers in CJ studies typically use the BTL model to measure traits and position the compared stimuli along a latent continuum [@Thurstone_1927b]. The CJ literature also show that they frequently rely on point estimates of these traits--typically the BTL scores or its transformations--to conduct statistical inference or hypothesis testing. For example, researchers have used these scores to identify 'misfit' judges and stimuli [@Pollitt_2012b; @vanDaal_et_al_2016; @Goossens_et_al_2018], detect biases in judges' ratings [@Pollitt_et_al_2003; @Pollitt_2012b], calculate correlations with other assessment methods [@Goossens_et_al_2018; @Bouwer_et_al_2023], or test hypotheses related to the underlying trait of interest [@Casalicchio_et_al_2015; @Bramley_et_al_2019; @Boonen_et_al_2020; @Bouwer_et_al_2023; @vanDaal_et_al_2017; @Jones_et_al_2019; @Gijsen_et_al_2021].

<!-- This is a simple observation equation which may be used for rather coarse scaling . .... This law defines the psychological scale or continuum. It allocates the compared stimuli on the continuum [@Thurstone_1927b, pp. 269] -->

<!-- 3. But what are the challenges in using BTL scores for further analysis? -->
Nevertheless, while separating the trait measurement and hypothesis testing processes simplifies the analysis of CJ data, the statistical literature cautions against relying solely on the point estimates of BTL scores to conduct statistical inference or hypothesis tests, as this practice can undermine the resulting statistical conclusions. A key consideration is that BTL scores are parameter estimates that inherently carry uncertainty (measurement error). Ignoring this uncertainty can bias the analysis and reduce the precision of hypothesis tests. The direction and magnitude of such biases are often unpredictable. Results may be attenuated, exaggerated, or remain unaffected depending on the degree of uncertainty in the scores and the actual effects being tested [@McElreath_2020; @Kline_et_al_2023; @Hoyle_et_al_2023]. Furthermore, the reduced precision in hypothesis tests diminishes their statistical power, increasing the likelihood of committing type-I or type-II errors [@McElreath_2020].

<!-- @fig-measurement_error illustrates these issues, showing that neglecting measurement error $(\sigma_{T})$--by relying solely on outcome averages--reduces the precision of a predictor's estimated effect $\beta_{X}$, as indicated by the solid lines. -->

<!-- ![The effect of outcome uncertainty $\sigma_{T}$ on the estimation of an effect $\beta_{X}$ linked to the predictor. The example assumes a sample size $n=100$ and an uncertainty that increases from left to right.](/images/png/measurement_error.png){#fig-measurement_error width=100%} -->

<!-- A type-I error occurs when a *true* null hypothesis is incorrectly rejected, while a type-II error occurs when a *false* null hypothesis is incorrectly accepted. [@Everitt_et_al_2010] -->

<!-- Measurement error in the criterion only does not bias unstandardized regression coefficients, if that error is unrelated to the predictors. But there is downward bias in the standardized coefficients, which also makes the value of R2 decrease as error in the criterion increases under the same condition (Williams et al., 2013). Things are more complicated when there is error in the predictors: Measurement error in a single predictor that is not shared with any other variable will attenuate the absolute value of the regression coefficient for that predictor. But regression coefficients for the other, error-free predictors can also be affected. Whether those other coefficients are attenuated, exaggerated, or unbiased depends on the values and signs of (1) correlations between the underlying latent variable imperfectly measured by a predictor and all other explanatory variables, and (2) the effect of the conceptual variable just mentioned on the criterion (Bollen, 1989; Kenny, 1979). Values of these correlations and their potential distorting effects are rarely known in practice, so it difficult to anticipate the magnitudes and directions of this propagation of measurement error in a single predictor. [@Kline_et_al_2023, pp. 25]  -->

<!-- In manifest variable path models, it is assumed that scores on exogenous variables, such as X1 and X2 in Figures 7.1(a) and 7.1(b), are perfectly reliable. Unlike endogenous variables, which have error terms that can “absorb” random measurement error—see Y1 and Y2 in the figures—there is no place in the model for measurement error in exogenous variables to “go” except downstream. This means that unmodeled measurement error can bias estimates of path coefficients or disturbance variances (Cole & Preacher, 2014). Specifically, values of absolute path coefficients for multiple exogenous causes of the same outcome are negatively biased (i.e., the estimates are too small) when measurement error is uncorrelated (independent) over
those causes. But when measurement error is correlated over multiple predictors, that bias can be either negative or positive (i.e., the estimates are too large). If scores on exogenous variables are perfectly precise, then measurement error in endogenous variables results in truncated values of R2, or proportions of explained variation, and also truncated absolute values of standardized—but not unstandardized—regression coefficients, assuming this measurement error is unrelated to the exogenous variables (Williams, Grajales, & Kurkiewicz, 2013). [@Hoyle_et_al_2023, pp. 137] -->

<!-- Probably the majority of regressions are really using proxies like Dobs , because most variables are measurements with some error. But even though it doesn’t necessarily open a non-causal path, using a proxy can introduce systematic bias, distorting the estimates [@McElreath_2020, pp. 492]. -->

<!-- The effect that measurement error has depends upon the context. Sometimes it exaggerates effects, as in this example. Other times it hides them. But you can’t safely assume that measurement error makes estimates conservative. [@McElreath_2020, pp. 494] -->

<!-- The big take home point for this section is that when you have a distribution of values, don't reduce it down to a single value to use in a regression. Instead, use the entire distribution. Anytime we use an average value, discarding the uncertainty around that average, we risk overconfidence and spurious inference. This doesn't only apply to measurement error, but also to cases in which data are averaged before analysis. [@McElreath_2020, pp. 497] -->

<!-- 4. segway for next section -->
In aggregate, the heavy reliance on Thurstone's Case V assumptions in the statistical analysis of comparative data can compromise the reliability of trait estimates. This overreliance may also undermine their validity [@Perron_et_al_2015], particularly when coupled with the disconnect between the trait measurement and hypothesis testing processes. However, the structural approach to causal inference can address these issues by offering a systematic and integrated framework that strengthens measurement reliability and validity while enhancing the statistical accuracy of hypothesis tests.

<!-- Reliability is a necessary but not sufficient condition for validity. Reliability can exist without validity but validity cannot exist without reliability [@Perron_et_al_2015, pp. 2]. -->

<!-- ## The diverse assessment design features and their role on reliability and validity {#sec-theory-issue3} -->
<!-- Do not include yet!, until you see if you need it when you write about the next section -->
<!-- recording: Sven and Tine 24.11.18; time: 00:47:20 - 00:56:41 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:53:30 - 01:00:00 -->

# Extending Thurstone's general form {#sec-theoretical}

<!-- recording: Sven and Tine 25.01.22; time: 00:05:40 - 00:06:30 -->
<!-- recording: Sven and Tine 25.01.22; time: 00:08:25 - 00:31:56 -->
<!-- recording: Sven and Tine 25.01.30 p1; time: 00:00:00 - 00:23:25 -->
<!-- recording: Sven and Tine 25.02.07; time: 00:02:20 - 00:06:20 -->

<!-- 1. The structural approach to causal inference -->
The *structural approach* to causal inference provides a formal framework for identifying causes and estimating their effects using data. The approach uses structural causal models (SCMs) and directed acyclic graphs (DAGs) [@Pearl_2009; @Pearl_et_al_2016; @Gross_et_al_2018; @Neal_2020] to formally and graphically represent the assumed causal structure of a system, such as the one found in CJ assessments. Essentially, SCMs and DAGs function as *conceptual models* on which identification analysis rests. *Identification analysis* determines whether an estimator can accurately compute an estimand based solely on its (causal) assumptions, regardless of random variability [@Schuessler_et_al_2023]. Here, *estimands* represent the specific quantities researchers aim to determine (i.e., a parameter) [@Everitt_et_al_2010]. *Estimators* denote the methods or functions that transform data into an estimate (e.g., a statistical model), while *estimates* are the numerical values approximating the estimand [@Neal_2020; @Everitt_et_al_2010].

<!-- In general terms, causal graphs encode a researcher’s assumptions about the causal process that generated the data in a population of interest. ... The causal assumptions encoded in a DAG constitute the theoretical model on which identification analysis rests. Identification analysis is concerned with determining whether or not the true value of a parameter can be estimated using the assumptions in the model (graph), regardless of random variability in the data due to small samples [@Schuessler_et_al_2023, pp. 4]. -->

<!-- Estimand: The quantity (usually a parameter) to be estimated using an estimator. -->
<!-- Estimate: See estimation. -->
<!-- Estimation: The process of providing a numerical value for a population parameter on the basis of information collected from a sample. If a single ﬁgure is calculated for the unknown parameter the process is called point estimation. If an interval is calculated which is likely to contain the parameter, then the procedure is called interval estimation. -->
<!-- Estimator: A statistic used to provide an estimate for a parameter. The sample mean, for example, is an unbiased estimator of the population mean. -->
<!-- [@Everitt_et_al_2010, pp. 154] -->

<!-- 3. Example of a CJ estimand and estimator, and its limitations -->
A motivating example that will appear throughout this document clarifies these concepts. In this example, researchers aim to answer the question: "To what extent do different teaching methods influence students' ability to produce high-quality written texts?" To investigate this, a researcher designs a CJ assessment by randomly assigning students (individuals) to two groups, each receiving a different teaching method. Judges then compare pairs of students' written texts (stimuli) to produce a dichotomous outcome reflecting the relative quality of each text (trait). Based on this setup, researchers can reformulate the research question as the estimand: "*On average*, is there a difference in the ability to produce high-quality written texts between the two groups of students?".

Following standard CJ practices, researchers typically use estimates from the BTL model--or its transformations--to approximate this estimand. However, as discussed in @sec-theory-issues, Thurstone's Case V and the BTL model exhibit several statistical and measurement limitations. These limitations hinder the model's ability to identify various estimands relevant to CJ inquiries, including the one described in the example.

<!-- 5. So, how do SCMs and DAGs support identification? -->
Fortunately, SCMs and DAGs support identification analysis through two key advantages[^1]. First, regardless of complexity, they can represent various causal structures using only five fundamental building blocks [@Neal_2020; @McElreath_2024]. This feature allows researchers to decompose complex structures into manageable components, facilitating their analysis. Second, they depict causal relationships in a non-parametric way. This flexibility enables feasible identification strategies without requiring specification of the types of variables, the functional forms relating them, or the parameters of those functional forms [@Pearl_et_al_2016].

<!-- We can understand this flow in general DAGs by understanding the flow in the minimal building blocks of graphs. These minimal building blocks consist of chains (Figure 3.9a), forks (Figure 3.9b), immoralities (Figure 3.9c), two unconnected nodes (Figure 3.10), and two connected nodes (Figure 3.11). [@Neal_2020, pp. 23]  -->

<!-- Four elemental confounds [@McElreath_2024] -->

<!-- The net result is that a researcher who has scientific knowledge in the form of a structural equation model is able to predict patterns of independencies in the data, based solely on the structure of the model's graph, without relying on any quantitative information carried by the equations or by the distributions of the errors. Conversely, it means that observing patterns of independencies in the data enables us to say something about whether a hypothesized model is correct. [@Pearl_et_al_2016, pp. 35] -->

[^1]: These topics are beyond the scope of this study, thus, readers seeking a more profound understanding can refer to introductory papers such as @Pearl_2010, @Rohrer_2018, @Pearl_2019, and @Cinelli_et_al_2020, and introductory books like @Pearl_et_al_2018, @Neal_2020, and @McElreath_2020 are useful. For more advanced study, seminal papers such as @Neyman_et_al_1923, @Rubin_1974, @Spirtes_et_al_1991, and @Sekhon_2009, along with books such as @Pearl_2009, @Morgan_et_al_2014, and @Hernan_et_al_2025, are recommended.

<!-- 6. What the section will do? -->
Thus, this section addresses the issues identified in @sec-theory-issues by extending Thurstone's general form using the structural approach to causal inference. Specifically, it combines the core theoretical principles outlined in @sec-thurstone_theory with key CJ assessment design features, such as the selection of judges, stimuli, and comparisons. In addition to improving statistical accuracy and strengthening measurement reliability and validity, the approach offers two key advantages. First, it clarifies the interactions among all actors and processes involved in CJ assessments. Second, it shifts the current comparative data analysis paradigm from passively accepting the model assumptions to actively testing whether those assumptions fit the data under analysis.

<!-- 7. How the section is divided? -->
Accordingly, @sec-theory-theoretical_P incorporates the theoretical principles into what we refer to as the *conceptual-population model*. This model assumes an idealized scenario where researchers have access to a *conceptual population* of comparative data, that is, data representing all repeated judgments made by every available judge for each pair of stimuli produced by each pair of individuals in the population. Conversely, @sec-theory-theoretical_SC integrates the assessment design features into what we refer to as the *sample-comparison model*. This model assumes a more realistic scenario where researchers only have access to a sample of judges, individuals, stimuli, and comparisons from the conceptual population.

<!-- Generalizability is the degree to which you can apply the results of your study to a broader context.  -->
<!-- https://www.scribbr.com/research-bias/generalizability/ -->

<!-- A theoretical model is a framework that researchers create to structure a study process and plan how to approach a specific research inquiry. It can allow you to define the purpose of your research and develop an informed perspective. Creating a theoretical model typically involves analyzing a subject's literature, including previous research studies and journal articles. Researchers often include a written description of a theoretical model in a beginning section of a dissertation or similar document so a reader can gain more context. -->
<!-- url: https://www.indeed.com/career-advice/career-development/theoretical-model -->

## The conceptual-population model {#sec-theory-theoretical_P}

<!-- recording: Sven 24.10.04; time: 00:21:00 - 00:34:10 -->
<!-- recording: Sven 24.10.04; time: 00:41:00 - 00:46:00 -->
<!-- recording: Sven and Tine 24.10.18; time: 00:00:00 - 00:04:05 -->
<!-- recording: Sven and Tine 24.11.25; time: 01:00:00 - 01:08:40 -->
<!-- recording: Sven and Tine 24.12.17; time: 00:26:10 - 00:54:00 -->
<!-- recording: Sven and Tine 25.01.30 p1; time: 00:23:25 - 00:48:00 -->
<!-- recording: Sven and Tine 25.02.07; time: 00:00:00 - 00:02:20 -->
<!-- recording: Sven and Tine 25.03.31; time: 00:47:05 - 00:48:40 -->

<!-- 1. What is the assumption on the conceptual-population model -->
In the conceptual-population model, the idealized scenario of a *conceptual population* of comparative data enables the integration of Thurstone's theoretical principles and provides a foundation for proposing innovations aimed at addressing some of the issues discussed in @sec-theory-issues.

### Integrating the first theoretical principles {#sec-theory-theoretical_P1}

<!-- 1. The SCMs -->
Before incorporating the first theoretical principles of Thurstone's theory, it is essential to further define SCMs. SCMs are formal mathematical models characterized by a set of *endogenous* variables $V$, a set of *exogenous* variables $E$, and a set of functions $F$ [@Pearl_2009; @Pearl_et_al_2016; @Cinelli_et_al_2020]. Endogenous variables are those whose causal mechanisms a researcher chooses to model [@Neal_2020]. In contrast, exogenous variables represent *errors* or *disturbances* arising from omitted factors that the investigator chooses not to model explicitly [@Pearl_2009]. Lastly, the functions, referred to as *structural equations*, express the endogenous variables as non-parametric functions of other endogenous and exogenous variables. These functions use the symbol '$:=$' to denote the asymmetrical causal dependence between variables and the symbol '$\dsep$' to represent *d-separation*, a concept akin to statistical (conditional) independence.

<!-- Consider the SCM M shown in Figure 13(a). The variables V = {Z, X, Y } are called the endogenous variables, and stand for those variables that the investigator chose to model their cause-effect relationships; the variables U = {Uz , Ux , Uy } are called the exogenous variables and represent everything else that the investigator chose not to explicitly model (these are also usually called disturbances). The functions F = {fz , fx , fy } are called structural equations, and each function represents a causal process that assigns to its respective endogenous variable a value based on the values of the other variables. [@Cinelli_et_al_2020, pp. 15] -->

<!-- A graph consists of a set V of vertices (or nodes) and a set E of edges (or links) that connect some pairs of vertices. The vertices in our graphs will correspond to variables (whence the common symbol V), and the edges will denote a certain relationship that holds in pairs of variables, the interpretation of which will vary with the application. [@Pearl_2009, pp. 42] -->

<!-- where pa_{i} (connoting parents) stands for the set of variables that directly determine the value of X_{i} and where the U_{i} represent errors (or "disturbances") due to omitted factors. [@Pearl_2009, pp. 65] -->

<!-- These disturbances reflect "hidden" or unmeasurable conditions that Nature governs by some undisclosed probability function. [@Pearl_2009, pp. 89] -->

<!-- Formally, a structural causal model consists of two sets of variables U and V, and a set of functions f that assigns each variable in V a value based on the values of the other variables in the model. ... The variables in U are called exogenous variables, meaning, roughly, that they are external to the model; we choose, for whatever reason, not to explain how they are caused. The variables in V are endogenous. Every endogenous variable in a model is a descendant of at least one exogenous variable. Exogenous variables cannot be descendants of any other variables, and in particular, cannot be a descendant of an endogenous variable; they have no ancestors and are represented as root nodes in graphs. If we know the value of every exogenous variable, then using the functions in f , we can determine with perfect certainty the value of very endogenous variable. [@Pearl_et_al_2016, pp. 26-27] -->

<!-- 2. the preliminary SCM for CJ -->
[SCM -@fig-cj03_scm] presents the first theoretical principles embedded in the conceptual-population model, which evaluates the impact of different teaching methods on students' writing ability. This SCM outlines the relationship between the conceptual-population outcome $(O^{cp}_{iahbjk})$ and several related variables. The subscripts $i$ and $h$ identify the students who authored the texts (i.e., the individuals). The indices $a$ and $b$ represent the texts under comparison (i.e., the stimuli). The index $j$ indicates the judge conducting the comparison, while the index $k$ accounts for assessment conditions where a judge compares the same pair of stimuli multiple times, i.e., a *repeated measures designs* [@Lawson_2015, pp. 366-376]. Thus, the indexing system supports comparisons between different texts written by the same student $(i = h;$ $a \neq b)$ and between texts written by distinct students $(i \neq h;$ where $a = b$ is permitted$)$, each compared once or repeatedly by all judges $(j = 1,\dots,n_{J};$ $k = 1,\dots,n_K;$ where $n_{J}>1$ and $n_{K}\geq1)$. However, it excludes cases where a judge compares a student's text to itself, whether once or multiple times $(i = h;$ $a = b;$ $j = 1,\dots,n_{J};$ $k = 1,\dots,n_{K};$ where $n_{J}>1$ and $n_{K}\geq1)$, as such comparison lacks practical relevance within the CJ framework. Here, $n_{J}$ indicates the total number of judges, and $n_{K}$ denotes the number of repeated judgments each judge performs.

::: {#fig-cj03 layout-nrow=2}

::: {#fig-cj03_scm fig-align="center" fig-pos="H"}
$$
\begin{aligned}
  O^{cp}_{iahbjk} & := f_{O}(D_{iahbjk}) \\
  D_{iahbjk} & := f_{D}(T_{ia}, T_{hb}, B_{jk})
\end{aligned}
$$

SCM
:::

![DAG](/images/png/CJ_TM_03.png){#fig-cj03_dag fig-align="center" fig-pos="H" width=35%}

Conceptual-population model, scalar form.
:::

In line with Thurstone's theory, [SCM -@fig-cj03_scm] depicts the texts' discriminal processes $(T_{ia}, T_{hb})$ and their discriminal difference $(D_{iahbjk})$ (see @sec-thurstone_theory). Additionally, the SCM incorporates a key CJ design feature: the judges' biases $(B_{kj})$. This extension builds on the arguments presented in @sec-theory-issue1b, contending that the discriminal difference becomes an observable outcome only through judges' perceptions. Given that such perceptions may be imperfect--and that each judge may carry some degree of bias [see @Pollitt_et_al_2003; @vanDaal_et_al_2016]--it is reasonable that judges' perceptions (bias) should be treated as an integral component of the CJ system from the outset, as this leads to a more accurate representation of the data-generating process underlying the pairwise comparisons. This model defines the preliminary set of endogenous variables, $V = \{ O_{iahbjk}, D_{iahbjk}, T_{ia}, T_{hb}, B_{kj} \}$, and the preliminary set of structural equations, $F = { f_{O}, f_{D} }$, which capture the non-parametric dependencies among these variables.

<!-- Moreover, the literature advocates for the incorporation of relevant hierarchical structures into the statistical model to account for these dependencies. Together, these additions can result in a model resembling a Multilevel Structural Equation Model (MSEM) [@Hoyle_et_al_2023, chap. 26] combined with a multidimensional or two-parameter logistic Item Response Theory (IRT) model [@Hoyle_et_al_2023, chap. 15], depending on the theoretical and statistical treatment of judges' biases.  -->

<!-- Brennen generalizability theory, generalize conclusions over groups. To generalize over judges, you need to estimate them. Akin to Frisch-Wough-Lovel theorem for latent variables?. -->

<!-- 3. The DAGs -->
Notably, every SCM has an associated DAG [@Pearl_et_al_2016; @Cinelli_et_al_2020]. A DAG is a *graph* consisting of nodes connected by edges, where nodes represent random variables. The term *directed* indicates that edges or arrows extend from one node to another, indicating the direction of causal influence. The absence of an edge implies no direct relationship between the nodes. The term *acyclic* means that the causal influences do not form loops, ensuring the influences do not cycle back on themselves [@McElreath_2020]. DAGs conventionally depict observed variables as solid black circles and unobserved (latent) variables as open circles [@Morgan_et_al_2014]. Although DAGs conventionally omit exogenous variables for simplicity, the DAGs presented in this section includes exogenous variables to improve clarity and reveal potential issues related to conditioning and confounding [@Cinelli_et_al_2020].

<!-- the *standard representation* omit exogenous variables for simplicity -->
<!-- the *magnified representation* includes exogenous variables -->

<!-- Every SCM is associated with a graphical causal model, referred to informally as a “graphical model” or simply “graph.” Graphical models consist of a set of nodes representing the variables in U and V, and a set of edges between the nodes representing the functions in f. [@Pearl_et_al_2016, pp. 27] -->

<!-- Every SCM has an associated graph G, usually called its causal diagram. In the types of models we consider here, which do not exhibit cycles, the causal diagram will be a directed acyclic graph (DAG). [@Cinelli_et_al_2020, pp. 15] -->

<!-- DAG, short for directed acyclic graph. Graph means it is nodes and connections. Directed means the connections have arrows that indicate directions of causal influence. And acyclic means that causes do not eventually flow back on themselves. [@McElreath_2020, pp. 128]  -->

<!-- Nodes that are represented by a solid circle are observed random variables, whereas nodes that are represented by a hollow circle are unobserved random variables. Causal effects are represented by directed edges (i.e., single-headed arrows), [@Morgan_et_al_2014, pp. 79] -->

<!-- Here we explicitly show the exogenous variables, but, conventionally, these are omitted from the graph for brevity. [@Cinelli_et_al_2020] -->

<!-- 4. the preliminary DAG for CJ -->
@fig-cj03_dag displays the DAG corresponding to [SCM -@fig-cj03_scm], illustrating the expected causal relationships outlined in Thurstone's theory. The graph shows that the discriminal processes of the texts $(T_{ia}, T_{hb})$ influence their discriminal difference $(D_{iahbjk})$, which in turn determines the outcome $(O^{cp}_{iahbjk})$. It also highlights the influence of judges' biases $(B_{kj})$ on the discriminal difference. Additionally, the DAG differentiates between observed endogenous variables, such as the outcome (solid black circle), and latent endogenous variables, including the texts' discriminal processes, their discriminal difference, and the judges' biases (open circles).

### The *conceptual-population* data structure {#sec-theory-theoretical_P2}

<!-- recording: Sven and Tine 25.03.28; time: 00:10:00 - 00:11:00 -->

<!-- 1. Why is useful to declare the data structure? -->
Although specifying a data structure is not mandatory when using SCMs and DAGs, defining one in this case can improve clarity and facilitate the description of the system. Thus, to re-express the scalar form of the CJ system shown in @fig-cj03 into an equivalent vectorized form, we first define the vectors $I$ and $J$, along with the matrices $IA$ and $JK$, as in Equation ([-@eq-mat11]). Here, each element of $I$ represents a unique individual $i$ or $h$, where $n_{I}$ denotes the total number of individuals. Similarly, each element of $J$ corresponds to a unique judge $j$, with $n_{J}$ indicating the total number of judges. Moreover, each row of $IA$ represents a unique pairing of individuals $i, h$ with stimuli $a, b$. As a result, the matrix $IA$ contains $n_{I} \cdot n_{A}$ rows and $2$ columns, where $n_{A}$ specifies the number of stimuli available per individual. Likewise, each row of $JK$ associates a judge $j$ with a (repeated) judgment index $k$. Consequently, the matrix $JK$ has $n_{J} \cdot n_{K}$ rows and $2$ columns, where $n_{K}$ indicates the number of repeated judgments each judge makes.
$$
I = \begin{bmatrix}
1 \\
\vdots \\
i \\
\vdots \\
h \\
\vdots \\
n_{I}
\end{bmatrix} ; \;
J = \begin{bmatrix}
1 \\
\vdots \\
j \\
\vdots \\
n_{J}
\end{bmatrix} ; \;
IA = \begin{bmatrix}
1 & 1 \\
\vdots & \vdots \\
1 & n_{A} \\
\vdots & \vdots \\
i & a \\
\vdots & \vdots \\
h & b \\
\vdots & \vdots \\
n_{I} & 1 \\
\vdots & \vdots \\
n_{I} & n_{A}
\end{bmatrix} ; \;
JK = \begin{bmatrix}
1 & 1 \\
\vdots & \vdots \\
1 & n_{K} \\
\vdots & \vdots \\
j & k \\
\vdots & \vdots \\
n_{J} & 1 \\
\vdots & \vdots \\
n_{J} & n_{K}
\end{bmatrix}
$$ {#eq-mat11}

Additionally, we construct the matrix $R$ to map each row of the $IA$ matrix with a corresponding row from the $JK$ matrix. This matrix has $n$ rows and $6$ columns, where $n = {n_{I} \cdot n_{A} \choose 2} \cdot n_{J} \cdot n_{K}$. Here, the term ${n_{I} \cdot n_{A} \choose 2}$ represents the binomial coefficient, which quantifies the total number of unique comparisons possible between every pair of stimuli generated by each pair of individuals in the population. Thus, we define the matrix as in Equation ([-@eq-mat12]). 
$$
R = \begin{bmatrix}
1 & 1 & 1 & 2 & 1 & 1 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
1 & 1 & 1 & 2 & 1 & n_{K} \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
i & a & h & b & j & k \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
n_{I} & n_{A}-1 & n_{I} & n_{A} & n_{J} & 1 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
n_{I} & n_{A}-1 & n_{I} & n_{A} & n_{J} & n_{K}
\end{bmatrix}
$$ {#eq-mat12}

It is easier to visualize the structure of the previously defined vectors and matrices by considering an example. Assuming $n_{I} = 5$, $n_{A} = 2$, $n_{J} = 3$, and $n_{K} = 3$, the vectors and matrices described in Equations ([-@eq-mat11]) and ([-@eq-mat12]) take the form as in Equation ([-@eq-mat13]).

$$
I = \begin{bmatrix}
1 \\
2 \\
3 \\
4 \\
5 
\end{bmatrix} ; \;
J = \begin{bmatrix}
1 \\
2 \\
3 
\end{bmatrix} ; \;
IA = \begin{bmatrix}
1 & 1 \\
1 & 2 \\
2 & 1 \\
2 & 2 \\
3 & 1 \\
3 & 2 \\
4 & 1 \\
4 & 2 \\
5 & 1 \\
5 & 2 
\end{bmatrix} ; \;
JK = \begin{bmatrix}
1 & 1 \\
1 & 2 \\
1 & 3 \\
2 & 1 \\
2 & 2 \\
2 & 3 \\
3 & 1 \\
3 & 2 \\
3 & 3 
\end{bmatrix} ; \;
R = \begin{bmatrix}
1 & 1 & 1 & 2 & 1 & 1 \\
1 & 1 & 1 & 2 & 1 & 2 \\
1 & 1 & 1 & 2 & 1 & 3 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
1 & 1 & 5 & 2 & 1 & 1 \\
1 & 1 & 5 & 2 & 1 & 2 \\
1 & 1 & 5 & 2 & 1 & 3 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
4 & 2 & 5 & 2 & 3 & 1 \\
4 & 2 & 5 & 2 & 3 & 2 \\
4 & 2 & 5 & 2 & 3 & 3 \\
5 & 1 & 5 & 2 & 3 & 1 \\
5 & 1 & 5 & 2 & 3 & 2 \\
5 & 1 & 5 & 2 & 3 & 3 
\end{bmatrix}
$$ {#eq-mat13}

::: {#fig-cj04 layout-nrow=2}

::: {#fig-cj04_scm fig-align="center" fig-pos="H"}
$$
\begin{aligned}
  O^{cp}_{R} & := f_{O}(D_{R}) \\
  D_{R} & := f_{D}(T_{IA}, B_{JK})
\end{aligned}
$$

SCM
:::

![DAG](/images/png/CJ_TM_04.png){#fig-cj04_dag fig-align="center" fig-pos="H" width=44%}

Conceptual-population model, initial vectorized form.
:::

<!-- 3. how de we use that data structure? -->
Now, using Equations ([-@eq-mat11]) and ([-@eq-mat12]), we can re-express [SCM -@fig-cj03_scm] and [DAG -@fig-cj03_dag] in an equivalent vectorized form, as shown in @fig-cj04. In this depiction, the outcome $O^{cp}_{R}$, the texts' discriminal difference $D_{R}$, their discriminal processes $T_{IA}$, and the judges' biases $B_{JK}$ are represented as vectors rather than scalar values. These vectors capture all the observations from the conceptual population. Specifically, $O^{cp}_{R}$ and $D_{R}$ are observed and latent vectors of length $n$, respectively. Moreover, $T_{IA}$ and $B_{JK}$ are latent vectors of lengths $n_{I} \cdot n_{A}$ and $n_{J} \cdot n_{K}$, respectively.

### Integrating hierarchical structural components {#sec-theory-theoretical_P3}

<!-- recording: Sven and Tine 25.03.28; time: 00:12:45 - 00:33:00 -->

<!-- 1. What are we going to do? -->
Building on the principles of Structural Equation Modeling (SEM) [@Hoyle_et_al_2023] and Item Response Theory (IRT) [@Fox_2010; @vanderLinden_et_al_2017_I], the conceptual-population model integrates two *hierarchical structural components* to examine how different *relevant* [^2] variables--whether observed or latent--affect the primary latent variable of interest [@Everitt_et_al_2010]. This hierarchical design enables researchers to formulate and test hypotheses that account for both the nested structure of stimuli and the uncertainties inherent in trait estimation (see @sec-theory-issue1b and @sec-theory-issue2 for a discussion of these considerations).

[^2]: *Relevant variables* are those that satisfy the *backdoor criterion* [@Neal_2020, pp 37], that is, they belong to a *sufficient adjustment set* [@Pearl_2009; @Pearl_et_al_2016; @Morgan_et_al_2014]. A *sufficient* set (potentially empty) blocks all non-causal paths between a predictor and an outcome without opening new ones [@Pearl_2009]. Refer also to footnote 1.

<!-- The assumption of no measurement error in exogenous variables is generally unrealistic, such as for self-report measures of psychological traits such as anxiety, extraversion, or leadership. An alternative is to respecify a manifest variable path model as a latent variable model with single indicators such that values of score reliability coefficients for all variables are explicitly represented in the model. Doing so relaxes the requirement for perfect measurement of exogenous variables while also controlling for measurement error in endogenous variables. [@Hoyle_et_al_2023, pp. 138] -->

<!-- Accounting for measurement error in defining latent variables improves the accuracy of the model’s estimated structural parameters between the latent constructs, which in turn can provide a stronger test of a proposed theoretical model. [@Hoyle_et_al_2023, pp. 482] -->

<!-- This is a whole section [@Fox_2010, chap. 6] -->

<!-- This is a whole section  [@vanderLinden_et_al_2017_I, chap. 24] -->

<!-- 2 Explaining what we did -->
The top branch of [DAG -@fig-cj09_dag] illustrates the first component, where *relevant* [^3] student-related variables $X_{I}$, such as teaching method, and students' idiosyncratic errors $e_{I}$ causally influence the latent variable representing students' writing-quality trait $T_{I}$. The error term $e_{I}$ captures variations in students' traits unexplained by $X_{I}$. Here, $X_{I}$ is an observed matrix with $n_{I}$ rows and $q_{I}$ independent columns (variables), and both $e_{I}$ and $T_{I}$ are latent vectors of length $n_{I}$. Additionally, this branch shows how $T_{I}$, along with *relevant* [^4] text-related variables $X_{IA}$ (e.g., text length), and texts' idiosyncratic errors $e_{IA}$ causally influence the texts' written-quality trait $T_{IA}$, the first primary latent variable of interest. The error term $e_{IA}$ captures variations in the texts' traits that remain unexplained by $T_{I}$ or $X_{IA}$. Here, $X_{IA}$ is an observed matrix with dimensions $n_{I} \cdot n_{A}$ rows and $q_{IA}$ independent columns (variables), while $e_{IA}$ and $T_{IA}$ are latent matrices with $n_{I}$ rows and $n_{A}$ columns.

[^3]: refer to footnote 2.
[^4]: refer to footnote 2.

::: {#fig-cj09 layout-nrow=2}

::: {#fig-cj09_scm fig-align="center" fig-pos="H"}
$$
\begin{aligned}
  O^{cp}_{R} & := f_{O}(D_{R}) \\
  D_{R} & := f_{D}(T_{IA}, B_{JK}) \\
  T_{IA} & := f_{T}(T_{I}, X_{IA}, e_{IA}) \\
  T_{I} & := f_{T}(X_{I}, e_{I}) \\
  B_{JK} & := f_{B}(B_{J}, Z_{JK}, e_{JK}) \\
  B_{J} & := f_{B}(Z_{J}, e_{J}) \\
  e_{I} & \dsep \{ e_{J}, e_{IA}, e_{JK} \} \\
  e_{J} & \dsep \{ e_{IA}, e_{JK} \} \\
  e_{IA} & \dsep e_{JK} 
\end{aligned}
$$

SCM
:::

![DAG](/images/png/CJ_TM_09.png){#fig-cj09_dag fig-align="center" fig-pos="H" width=70%}

Conceptual-population model, final vectorized form.
:::

Similarly, the bottom branch of [DAG -@fig-cj09_dag] depicts the second component, where *relevant* [^5] judge-related variables $Z_{J}$, such as judgment expertise, and judges' idiosyncratic errors $e_{J}$ causally influence the latent variable representing judges' bias $B_{J}$. The error $e_{J}$ captures variations in judges' bias unexplained by $Z_{J}$. Here, $Z_{J}$ is an observed matrix with $n_{J}$ rows and $q_{J}$ independent columns (variables), and both $e_{J}$ and $B_{J}$ are latent vectors of length $n_{J}$. Furthermore, the branch shows how $B_{J}$, along with *relevant* [^6] judgment-related variables $Z_{JK}$ (e.g., the number of judgments a judge makes), and judgments' idiosyncratic errors $e_{JK}$ causally influence the judges' biases associated with each text $B_{JK}$, the second primary latent variable of interest. The error $e_{JK}$ captures variations in judgments unexplained by $B_{J}$ or $Z_{JK}$. Here, $Z_{JK}$ is an observed matrix with dimension $n_{J} \cdot n_{K}$ rows and $q_{JK}$ independent columns (variables), while $e_{JK}$ and $B_{JK}$ are latent latent matrices with $n_{J}$ rows and $n_{K}$ columns

[^5]: refer to footnote 2.
[^6]: refer to footnote 2.

Notably, all variables and functions shown in [SCM -@fig-cj09_scm] and [DAG -@fig-cj09_dag] are part of the set of endogenous variables $V$, structural equations $F$, and exogenous variables $E$ for the conceptual-population model. Additionally, the figures demonstrate that all exogenous variables are independent of one another, as indicated by the relationships $e_{IA} \dsep \{ e_{I}, e_{JK}, e_{J} \}$, $e_{I} \dsep \{ e_{JK}, e_{J} \}$ and $e_{JK} \dsep e_{J}$ and the absence of connecting arrows.

<!-- 3. segway for next section -->
Overall, the conceptual-population model extends Thurstone's general form by introducing key innovations to address the limitations discussed in @sec-theory-issue1b and @sec-theory-issue2. These enhancements include accounting for judges' biases and integrating hierarchical structural components. Nevertheless, despite its promise of enhancing measurement accuracy and precision, the model still depends on the unrealistic assumption that researchers have access to data from the *conceptual population*. Since researchers rarely meet this assumption in practice, they must consider a more realistic scenario.

## The sample-comparison model {#sec-theory-theoretical_SC}

<!-- recording: Sven 24.10.04; time: 00:09:10 - 00:14:30 -->
<!-- recording: Sven 24.10.04; time: 00:34:10 - 00:41:00 -->
<!-- recording: Sven 25.03.31; time: 00:34:00 - 00:38:50 -->

<!-- 1. What is the more realistic scenario: the sample mechanism -->
The sample-comparison model presents a more realistic scenario than the conceptual-population model. First, in @sec-theory-theoretical_SC1, it explicitly assumes researchers work with a data sample consisting of a limited number of repeated judgments $(n^{s}_{K})$ from a sample of judges $(n^{s}_{J})$ and a specific number of texts $(n^{s}_{A})$ from a sample of students $(n^{s}_{I})$, all drawn from the conceptual population. Second, in @sec-theory-theoretical_SC2, the model assumes that judges do not perform *all repeated judgments* within the data sample. Instead, they conduct a sufficient number of stimuli comparisons, $n_{C}$, to ensure an accurate estimation of the proportion $P(B>A)$, as proposed by @Thurstone_1927b.

<!-- It is assumed that the single observer compares each pair of stimuli a "sufficient number of times" so that a proportion, P(A>B), may be determined for each pair of stimuli [@Thurstone_1927b, pp. 267] -->

<!-- Do not depict the variances, nor the sample sizes. Or if you do explain that it is with illustrative purposes and then remove them to make the reading of the DAGs easier. -->

### The sample mechanism {#sec-theory-theoretical_SC1}

<!-- recording: Sven and Tine 25.03.28; time: 00:09:00 - 00:10:00 -->
<!-- recording: Sven and Tine 25.03.28; time: 00:11:30 - 00:12:25 -->

<!-- 1. start by defining the sampling vector and matrices  -->
To incorporate the sampling mechanism and facilitate the interpretation of the sample-comparison model, we first define the *data sampling process* using the binary vector variables $S_{I}$, $S_{J}$, $S_{IA}$, and $S_{JK}$ as follows:
$$
S_{I} = \begin{bmatrix}
i_{(1)} \\
\vdots \\
i_{(i)} \\
\vdots \\
i_{(h)} \\
\vdots \\
i_{(nI)}
\end{bmatrix} ; \;
S_{J} = \begin{bmatrix}
j_{(1)} \\
\vdots \\
j_{(j)} \\
\vdots \\
j_{(nJ)}
\end{bmatrix} ; \;
S_{IA} = \begin{bmatrix}
ia_{(1,1)} \\
\vdots \\
ia_{(1,n_{A})} \\
\vdots \\
ia_{(i,a)} \\
\vdots \\
ia_{(h,b)} \\
\vdots \\
ia_{(nI,1)} \\
\vdots \\
ia_{(nI,nA)}
\end{bmatrix} ; \;
S_{JK} = \begin{bmatrix}
jk_{(1,1)} \\
\vdots \\
jk_{(1,n_{K})} \\
\vdots \\
jk_{(j,k)} \\
\vdots \\
jk_{(nJ,1)} \\
\vdots \\
jk_{(nJ,nK)}
\end{bmatrix}
$$ {#eq-mat21}

Where each element of $S_{I}$ is a binary value indicating the presence or absence of corresponding elements in the vector $I$, as in Equation ([-@eq-mat22]). We apply the same logic to $S_{J}$ using vector $J$ (not shown). Thus, the vectors $S_{I}$ and $S_{J}$ contains $n_{I}$ and $n_{J}$ elements, respectively.
$$
i_{(i)} = \begin{cases} 
1 & \text{if data element } i \text{ from } I \text{ is sampled} \\
0 & \text{if data element } i \text{ from } I \text{ is missing}
\end{cases}
$$ {#eq-mat22}

Similarly, each element of $S_{IA}$ is a binary value indicating the presence or absence of data rows in the matrices $IA$, as defined in Equation ([-@eq-mat23]). We apply the same logic to $S_{JK}$ using the matrix $JK$ (not shown). Thus, the vectors $S_{IA}$ and $S_{JK}$ contains $n_{I} \cdot n_{A}$ and $n_{J} \cdot n_{K}$ elements, respectively. 
$$
ia_{(i,a)} = \begin{cases} 
1 & \text{if data elements } i,a \text{ from } IA \text{ are sampled} \\
0 & \text{if data elements } i,a \text{ from } IA \text{ are missing}
\end{cases}
$$ {#eq-mat23}

We can illustrate the structure of these vectors more clearly with an example. Suppose researchers exclude the second student, the second text from each student, and the third judge from the setup shown in Equation ([-@eq-mat13]). Given $n_{I} = 5$, $n_{A} = 2$, $n_{J} = 3$, and $n_{K} = 3$, the resulting vectors would have the following structure:

$$
S_{I} = \begin{bmatrix}
1 \\
0 \\
1 \\
1 \\
1
\end{bmatrix} ; \;
S_{J} = \begin{bmatrix}
1 \\
1 \\
0
\end{bmatrix} ; \;
S_{IA} = \begin{bmatrix}
1 \\
0 \\
0 \\
0 \\
1 \\
0 \\
1 \\
0 \\
1 \\
0 
\end{bmatrix} ; \;
S_{JK} = \begin{bmatrix}
1 \\
1 \\
1 \\
1 \\
1 \\
1 \\
0 \\
0 \\
0 
\end{bmatrix}
$$ {#eq-mat24}

Notably, Equation ([-@eq-mat24]) shows that missing observations in the vectors $S_{I}$ and $S_{J}$--which represent unsampled students and judges--directly determine which observations are missing in $S_{IA}$ and $S_{JK}$. In other words, researchers can only observe texts and judgments from students and judges initially included in the sample. The equation also shows that the sum of observed elements in $S_{I}$ equals the number of sampled students $(n^{s}_{I})$ and that a similar sum in vector $S_{J}$ equals the sampled judges $(n^{s}_{J})$. Conversely, the sum of observed elements in $S_{IA}$ represents the total sampled texts across all sampled students $(n^{s}_{I} \cdot n^{s}_{A})$, while a similar sum in vector $S_{JK}$ represents the total sampled repeated judgments across all sampled judges $(n^{s}_{J} \cdot n^{s}_{K})$. Notice that in this example, because the design systematically excludes every third repeated judgment, researchers can also express $S_{JK}$ using $n_{K} = n^{s}_{K} = 2$.

Finally, we define the *sample mechanism* $S$ in Equation ([-@eq-mat25]), which maps each element of $S_{IA}$ to every element of $S_{JK}$. Each element $s_{(i,a,h,b,j,k)}$ is a binary value indicating the presence or absence of data rows in the matrix $R$ resulting from the sample mechanism, as in Equation ([-@eq-mat26]). Thus, the vector contains $n$ elements, matching the number of rows in $R$, and the sum of its elements represents the total data sample: $n^{s} = \binom{n^{s}_{I} \cdot n^{s}_{A}}{2} \cdot n^{s}_{J} \cdot n^{s}_{K}$. Here, the term ${n^{s}_{I} \cdot n^{s}_{A} \choose 2}$ represents the binomial coefficient, which quantifies the total number of unique comparisons possible between every pair of sampled stimuli generated by each pair of sampled individuals.

$$
s_{(i,a,h,b,j,k)} = \begin{cases} 
1 & \text{if data elements } i,a,h,b,j,k \text{ from } R \text{ are sampled} \\
0 & \text{if data elements } h,i,a,b,j,k \text{ from } R \text{ are missing}
\end{cases}
$$ {#eq-mat26}

$$
S = \begin{bmatrix}
s_{(1,1,1,2,1,1)} \\
\vdots \\
s_{(1,1,1,2,1,n_{K})} \\
\vdots \\
s_{(i,a,h,b,j,k)} \\
\vdots \\
s_{(n_{I},n_{A}-1,n_{I},n_{A},n_{J},1)} \\
\vdots \\
s_{(n_{I},n_{A}-1,n_{I},n_{A},n_{J},1)}
\end{bmatrix}
$$ {#eq-mat25}

<!-- 2. Integrate the sample mechanism -->
With the definition of $S$, we incorporate the sample mechanism into the conceptual-population model. Following the convention of @McElreath_2020 and @Deffner_et_al_2022, [DAG -@fig-cj14_dag] represents the conceptual-population outcome $O^{cp}_{R}$ as unobserved, emphasizing that researchers cannot directly access this outcome due to the sampling mechanism. The DAG also depicts the *sample design* vector $S$ as a causal factor influencing the sample-comparison outcome $O^{sc}_{R}$. A square encloses $S$, indicating that it is a conditioned variable. In this context, *conditioning* means that researchers restrict their focus to the elements of $O^{cp}_{R}$ that satisfy $s_{(i,a,h,b,j,k)}=1$ [@Neal_2020; @McElreath_2020]. In essence, $S$ is a vector that selects *all repeated judgments made by a subset of judges for a subset of stimuli produced by the sampled individuals*.

<!-- Follow conventions of @McElreath_2020 NOT @Schuessler_et_al_2023 regarding selection variables and missing data. -->

<!-- This is a whole section [@McElreath_2020, pp. 499-516] -->

<!-- This is a whole section [@Deffner_et_al_2022, pp. 6-7] -->

<!-- What does “conditioning” mean? Informally, conditioning on a variable Z means learning its value and then asking if X adds any additional information about Y. [@McElreath_2020, pp. 130] -->

<!-- In the regular notation for probability, we have conditioning, but that isn’t the same as intervening. Conditioning on T=t just means that we are restricting our focus to the subset of the population to those who received treatment t. In contrast, an intervention would be to take the whole population and give everyone treatment t. [@Neal_2020, pp. 32] -->

Notably, the DAG shows that $S$ is independent of all other variables in the model. This implies that [DAG -@fig-cj14_dag] applies exclusively to Simple Random Sampling (SRSg) designs. In these designs, each repeated judgment, judge, stimulus, and individual has the same probability of being included in the sample as any other observation within their respective groups [@Lawson_2015].

<!-- This is a whole section [@Lawson_2015, pp. 16-112] -->

<!-- 3. segway for next section -->
However, due to concerns about the practical feasibility of the comparison task [@Boonen_et_al_2020; ], CJ assessments rarely implement an exhaustive pairings of sampled judges, stimuli, and individuals. Thus, a realistic scenario must account for the fact that judges typically compare only a subset of stimuli authored by a sample of individuals.

<!-- An exhaustive pairing of all stimuli would lead to a total of 7875 possible pairs for each listener group. Obviously, such a large number of pairs would undermine the practical feasibility of the experiment. However, in order to arrive at a reliable ranking, not all possible combinations of stimuli had to be assessed. More specifically, each stimulus had to be judged by the listeners of each particular listener group – together, not individually – at least 20 times. [@Boonen_et_al_2020, p. 5] -->

::: {#fig-cj14 layout-nrow=2}

::: {#fig-cj14_scm fig-align="center" fig-pos="H"}
$$
\begin{aligned}
  O_{R} & := f_{C}(O^{sc}_{R}, C) \\
  O^{sc}_{R} & := f_{S}(O^{cp}_{R}, S) \\
  O^{cp}_{R} & := f_{O}(D_{R}) \\
  D_{R} & := f_{D}(T_{IA}, B_{JK}) \\
  T_{IA} & := f_{T}(T_{I}, X_{IA}, e_{IA}) \\
  T_{I} & := f_{T}(X_{I}, e_{I}) \\
  B_{JK} & := f_{B}(B_{J}, Z_{JK}, e_{JK}) \\
  B_{J} & := f_{B}(Z_{J}, e_{J}) \\
  e_{I} & \dsep \{ e_{J}, e_{IA}, e_{JK} \} \\
  e_{J} & \dsep \{ e_{IA}, e_{JK} \} \\
  e_{IA} & \dsep e_{JK} 
\end{aligned}
$$

SCM
:::

![DAG](/images/png/CJ_TM_14.png){#fig-cj14_dag fig-align="center" fig-pos="H" width=100%}

Sample-comparison model, final vectorized form
:::

### The comparison mechanism {#sec-theory-theoretical_SC2}

<!-- recording: Sven 24.10.04; time: 00:09:10 - 00:14:30 -->
<!-- recording: Sven and Tine 25.04.10; time: 00:16:30 - 00:19:00 -->

<!-- 1. start by defining the comparison vector  -->
As in the previous section, we begin defining the *comparison mechanism* using the binary vector variable $C$ to facilitate the interpretation of the sample-comparison model. Equation ([-@eq-mat27]) shows that $C$ contains $n$ elements corresponding to the number of rows in the $R$ matrix, with each element $c_{(i,a,h,b,j,k)}$ being a binary value indicating the presence or absence of data rows in $R$, a definition similar to that of $s_{(i,a,h,b,j,k)}$ in Equation ([-@eq-mat26]).
$$
C = \begin{bmatrix}
c_{(1,1,1,2,1,1)} \\
\vdots \\
c_{(1,1,1,2,1,n_{K})} \\
\vdots \\
c_{(i,a,h,b,j,k)} \\
\vdots \\
c_{(n_{I},n_{A}-1,n_{I},n_{A},n_{J},1)} \\
\vdots \\
c_{(n_{I},n_{A}-1,n_{I},n_{A},n_{J},1)}
\end{bmatrix}
$$ {#eq-mat27}

<!-- 2. What is the effect of the comparison mechanism -->
The [DAG -@fig-cj14_dag] also incorporates the *comparison mechanism* $C$ into the conceptual-population model. It shows the sample-comparison outcome $O^{sc}_{R}$ as unobserved, emphasizing that researchers cannot directly access this variable because of the comparison mechanism. The DAG further shows $C$ as a conditioned variable (enclosed in a square) that causally influences the observed outcome $O_{R}$. This structure implies that $C$ determines *which repeated judgments judges make for the stimuli produced by the individuals*. In essence, $C$ reflects the assumption that judges *do not* perform all possible repeated judgments but instead complete a sufficient number, $n_{C}$, to enable the accurate estimation of the proportion $P(B>A)$ for each stimulus pair [@Thurstone_1927b, pp. 267].

<!-- It is assumed that the single observer compares each pair of stimuli a "sufficient number of times" so that a proportion, P(A>B), may be determined for each pair of stimuli [@Thurstone_1927b, pp. 267] -->

Notably, [DAG -@fig-cj14_dag] also shows that $C$ is independent of all other variables in the model. This independence implies that the conceptual model represented by the DAG applies exclusively to Random Allocation Comparative Designs [@Bramley_2015], or Incomplete Block Designs [@Lawson_2015], where every repeated judgment has an equal probability of being included in the sample.

<!-- The different analyses were based on the same two underlying datasets. First of all, observations from these datasets were sampled in order to mimic either random allocation of pairs of scripts to judges, or adaptive allocation (where the probability of being sampled depends on the outcome of earlier comparisons). In all cases the selection of observations was carefully planned to ensure that i) each script was compared the same number of times; and ii) the complete set of judgments formed a fully linked design whereby each script was compared directly or indirectly with every other script5. For the ‘random only’ analyses, this could be achieved by systematically sampling the required observations. [@Bramley_2015, pp. 8] -->

<!-- This is a whole section [@Lawson_2015, pp. 259-271] -->

<!-- Actually algorithm can be thought as the observation mechanism that "transforms" the outcome in the population of comparisons to the outcome in the sample of comparisons, see fig. 2, [@Deffner_et_al_2022] -->

<!-- 3. preparing to transform the conceptual model into an statistical one -->
Finally, since it is standard to assume that the distribution of the conceptual-population outcome $O^{cp}_{R}$ also holds for $O^{sc}_{R}$ and $O_{R}$, we can reformulate the sample-comparison model in @fig-cj14 into the equivalent form shown in @fig-cj15. This reformulation produces a model that applies directly to a sample of comparative data. In this version, the unobserved outcomes $O^{cp}_{R}$ and $O^{sc}_{R}$ are omitted, and $O_{R}$ inherits the structural equation $f_{O}$ that originally defined $O^{cp}_{R}$. Moreover, the definition of $O_{R}$ now reflects its direct dependence on the discriminal difference $D_{R}$ and the sample and comparison mechanisms, $S$ and $C$.

<!-- move the assumption for the conceptual population outcome to the data outcome based on the fact that it is logical to think that the data outcome would have the same distribution as the population outcome, if the data comes from that population. For example, if I have the distribution of the population outcome, and it is a Bernoulli, and even if I do not have it because of S and C, the distributional assumption still holds -->

::: {#fig-cj15 layout-nrow=2}

::: {#fig-cj15_scm fig-align="center" fig-pos="H"}
$$
\begin{aligned}
  O_{R} & := f_{O}(D_{R}, S, C) \\
  D_{R} & := f_{D}(T_{IA}, B_{JK}) \\
  T_{IA} & := f_{T}(T_{I}, X_{IA}, e_{IA}) \\
  T_{I} & := f_{T}(X_{I}, e_{I}) \\
  B_{JK} & := f_{B}(B_{J}, Z_{JK}, e_{JK}) \\
  B_{J} & := f_{B}(Z_{J}, e_{J}) \\
  e_{I} & \dsep \{ e_{J}, e_{IA}, e_{JK} \} \\
  e_{J} & \dsep \{ e_{IA}, e_{JK} \} \\
  e_{IA} & \dsep e_{JK} 
\end{aligned}
$$

SCM
:::

![DAG](/images/png/CJ_TM_15.png){#fig-cj15_dag fig-align="center" fig-pos="H" width=72%}

Comparative judgment model
:::

<!-- 4. segway for next section -->
In summary, the [SCM -@fig-cj15_scm] and [DAG -@fig-cj15_dag] extend Thurstone's general form to address several limitations of the BTL model. These extensions account for judge biases (see @sec-theory-theoretical_P1), reflect the hierarchical structure of stimuli and incorporate measurement error in trait estimation and hypothesis testing (see @sec-theory-theoretical_P3), and even clarify the role of the sample and comparison mechanisms in CJ assessments (see @sec-theory-theoretical_SC). However, they do not resolve concerns about the assumption of equal dispersions among stimuli discussed in @sec-theory-issue1a. Since this concern relates to the statistical assumption underlying the distribution of the discriminal process, we develop a formal statistical model to address it in the next section.

# From SCM to statistical model  {#sec-statistical}

<!-- recording: Sven 24.10.04; time: 00:46:00 - 01:01:00 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:24:50 - 00:30:35 -->
<!-- recording: Sven and Tine 24.11.25; time: 00:49:30 - 00:53:30 -->
<!-- recording: Sven and Tine 25.03.28; time: 00:33:55 - 00:37:50 -->
<!-- recording: Sven and Tine 25.03.31; time: 00:38:50 - 00:47:00 -->
<!-- recording: Sven and Tine 25.04.10; time: 00:00:00 - 00:01:10 -->
<!-- recording: Sven and Tine 25.04.10; time: 00:01:10 - 00:01:35 -->
<!-- recording: Sven and Tine 25.04.10; time: 00:01:35 - 00:16:30 -->
<!-- recording: Sven and Tine 25.04.10; time: 00:19:00 - 00:26:40 -->
<!-- recording: Sven and Tine 25.04.10; time: 00:52:00 - 00:53:40 -->
<!-- recording: Sven and Tine 25.05.21; time: 00:42:20 - 00:46:10 -->

<!-- 1. How to derive an stat model from an SCM? -->
Using the structural causal model (SCM) [-@fig-cj15_scm], we can derive a statistical model that addresses violations of the equal dispersion assumption (see @sec-theory-issue1a). This derivation is possible because a fully specified SCM encodes functional and probabilistic information, which we can replace with suitable functions and probabilistic assumptions [@Pearl_et_al_2016]. Specifically, [SCM -@fig-cj15_scm] allows us to express the joint distribution of our complex CJ system as a product of simpler conditional probability distributions (CPDs)[^7], as shown in Equation ([-@eq-mat51]). For clarity, we treat expressions such as $Y := f_{Y}(X)$, $P(Y \mid X)$, and $Y \sim f(Y \mid X)$ as equivalent, where $P(Y \mid X)$ and $f(Y \mid X)$ represent the CPD of $Y$ given $X$.
$$
\begin{aligned}
  P(O_{R}, & S, C, D_{R}, T_{IA}, X_{IA}, e_{IA}, T_{I}, X_{I}, e_{I}, B_{JK}, Z_{JK}, e_{JK}, B_{J}, Z_{J}, e_{J} ) & \\
  &= P(O_{R} \mid D_{R}, S, C) \cdot P(S) \cdot P(C) \cdot P(D_{R} \mid T_{IA}, B_{JK}) \\
  & \quad \cdot P(T_{IA} \mid T_{I}, X_{IA}, e_{IA}) \cdot P(T_{I} \mid X_{I}, e_{I}) \\
  & \quad \cdot P(B_{JK} \mid B_{J}, Z_{JK}, e_{JK}) \cdot P(B_{J} \mid Z_{J}, e_{J}) \\
  & \quad \cdot P(X_{IA}) \cdot P(X_{I}) \cdot P(Z_{JK}) \cdot P(Z_{J}) \\
  & \quad \cdot P(e_{IA}) \cdot P(e_{I}) \cdot P(e_{JK}) \cdot P(e_{J})  
\end{aligned}
$$ {#eq-mat51}

<!-- Second, in a fully specified SCM, we can represent the joint distributions of n variables with greater efficiency: We need only to specify the n functions that govern the relationships between the variables, and then from the probabilities of the error terms, we can discover all the probabilities that govern the joint distribution. For any model whose graph is acyclic, the joint distribution of the variables in the model is given by the product of the conditional distributions P(child|parents) over all the "families" in the graph. [@Pearl_et_al_2016, pp. 29] -->

<!-- The functions F = {fz, fx, fy} are called structural equations, and each function represents a causal process that assigns to its respective endogenous variable a value based on the values of the other variables. [@Cinelli_et_al_2020, pp. 15] -->

[^7]: This re-expression is possible because the *chain rule* of probability and the *Bayesian Network Factorization (BNF)* property. For further details, see @Pearl_et_al_2016 and @Neal_2020.

<!-- We will denote the parents of a node X with pa(X). We’ll use an even simpler shorthand when the nodes are ordered so that we can denote the i^{th} node by X_{i}; in that case, we will also denote the parents of X_{i} by pa_{i} [@Neal_2020, pp. 19] Definition 3.1 (Bayesian Network Factorization) Given a probability distribution P and a DAG G, P factorizes according to G if: P(x_{1}, \dots, x_{n}) = \prod_{i}P(x_{i} | pa_{i}). [@Neal_2020, pp. 21] -->

<!-- For any model whose graph is acyclic, the joint distribution of the variables in the model is given by the product of the conditional distributions P(child | parents) over all the "families" in the graph. Formally, we write this rule as P(x_{1}, \dots, x_{n}) = \prod_{i}P(x_{i} | pa_{i}) where pa_{i} stands for the values of the parents of variable X_{i}, and the product i runs over all i, from 1 to n. [@Pearl_et_al_2016, pp. 29] -->

<!-- 2. assumptions about the outcome -->
Each CPD in Equation ([-@eq-mat51]) rests on specific assumptions, which we outline in the statistical model [-@fig-cj16_stat]. We begin by assuming that $O_{R}$ follows a Bernoulli distribution[^8], reflecting the binary nature of CJ outcomes. Furthermore, following the conventions of Generalized Linear Models (GLMs) [@Nelder_et_al_1983; @Nelder_et_al_1996; @Agresti_2015], the distribution links $O_{R}$ to the latent discriminal difference vector $D_{R}$ using an inverse-logit function: $\text{inv\_logit}(x) = 1/(1 + \exp(-x))$.

[^8]: The binomial distribution--including its special case, the Bernoulli distribution--represent a maximum entropy distribution for binary events [@McElreath_2020, pp. 34]. This means that the Bernoulli distribution is the most consistent alternative when only two un-ordered outcomes are possible and their expected frequencies are assumed to be constant [@McElreath_2020, pp. 310]. For a detailed discussion of the binomial as a maximum entropy distribution, see @McElreath_2020 [sec. 10.1.2].

<!-- The binomial distribution is rather special, because it represents the maximum entropy way to count binary events. [@McElreath_2020, pp. 34] -->

<!-- Choosing the distribution with the largest entropy means spreading probability as evenly as possible, while still remaining consistent with anything we think we know about a process. [@McElreath_2020, pp. 299] -->

<!-- The principle of maximum entropy helps us choose likelihood functions, by providing a way to use stated assumptions about constraints on the outcome variable to choose the likelihood function that is the most conservative distribution compatible with the known constraints. [@McElreath_2020, pp. 300] -->

<!-- There are a few conceptual lessons to take away from this example. First, hopefully it reinforces the maximum entropy nature of the binomial distribution. When only two un-ordered outcomes are possible—such as blue and white marbles—and the expected numbers of each type of event are assumed to be constant, then the distribution that is most consistent with these constraints is the binomial distribution.  -->
<!-- Second, of course usually we do not know the expected value, but wish to estimate it. If only two un-ordered outcomes are possible and you think the process generating them is invariant in time—so that the expected value remains constant at each combination of predictor values—then the distribution that is most conservative is the binomial. -->
<!-- Third, back in Chapter 2, we derived the binomial distribution just by counting how many paths through the garden of forking data were consistent with our assumptions. ... The likelihoods derived in that way turn out to be exactly the same as the likelihoods we get by maximizing entropy. This is not a coincidence. Entropy counts up the number of different ways a process can produce a particular result, according to our assumptions. [@McElreath_2020, pp. 310-311, chap. 10.1.2] -->

::: {#fig-cj16 layout-ncol=3 layout-valign="top"}

::: {#fig-cj16_scm}
$$
\begin{aligned}
  O_{R} & := f_{O}(D_{R}, S, C) \\ 
  D_{R} & := f_{D}(T_{IA}, B_{JK}) \\
  T_{IA} & := f_{T}(T_{I}, X_{IA}, e_{IA}) \\
  T_{I} & := f_{T}(X_{I}, e_{I}) \\
  B_{JK} & := f_{B}(B_{J}, Z_{JK}, e_{JK}) \\
  B_{J} & := f_{B}(Z_{J}, e_{J}) \\ \\
  e_{I} & \dsep \{ e_{J}, e_{IA}, e_{JK} \} \\
  e_{J} & \dsep \{ e_{IA}, e_{JK} \} \\
  e_{IA} & \dsep e_{JK}
\end{aligned}
$$

SCM
:::

::: {#fig-cj16_prob}
$$
\begin{aligned}
  & P( O_{R} \mid D_{R}, S, C ) \\
  & P( D_{R} \mid T_{IA}, B_{JK} ) \\
  & P( T_{IA} \mid T_{I}, X_{IA}, e_{IA} ) \\
  & P( T_{I} \mid X_{I}, e_{I} ) \\
  & P( B_{JK} \mid B_{J}, Z_{JK}, e_{JK} ) \\
  & P( B_{J} \mid Z_{J}, e_{J} ) \\ \\
  & P( e_{I} ) P( e_{IA} ) P( e_{J} ) P( e_{JK} ) \\ \\ \\
\end{aligned}
$$

Probabilistic model
:::

::: {#fig-cj16_stat}
$$
\begin{aligned}
  O_{R} & \overset{iid}{\sim} \text{Bernoulli} \left[ \text{inv\_logit}( D_{R} ) \right] \\
  D_{R} & = \left( T_{IA}[i,a] - T_{IA}[h,b] \right) + B_{JK}[j,k] \\
  T_{IA} & = T_{I} + \beta_{XA} X_{IA} + e_{IA} \\
  T_{I} & = \beta_{XI} X_{I} + e_{I} \\
  B_{JK} & = B_{J} + \beta_{ZK} Z_{JK} + e_{JK} \\
  B_{J} & = \beta_{ZJ} Z_{J} + e_{J} \\ \\
  \boldsymbol{e} & \sim \text{Multi-Normal}( \boldsymbol{\mu}, \boldsymbol{\Sigma} )
  \\
  \boldsymbol{\Sigma} &= \boldsymbol{V} \boldsymbol{Q} \boldsymbol{V} \\ \\
\end{aligned}
$$

Statistical model
:::

Comparative judgment model, SCM, probabilistic and statistical model assuming different discriminal dispersions for the student's traits
:::

<!-- 3. Note about the sample and comparison mechanism, as well as the predictor's distribution -->
Note that while the joint distribution in Equation ([-@eq-mat51]) includes the probability distributions of the sampling and comparison mechanisms, $P(S)$ and $P(C)$, as well as those of the predictor variables--$P(X_{IA})$, $P(X_{I})$, $P(Z_{JK})$, and $P(Z_{J})$--we omit all of these probabilities from the statistical model [-@fig-cj16_stat]. This omission is justified because, while these distributions contribute to the overall joint distribution of the data, the variables $S$, $C$, $X_{IA}$, $X_{I}$, $Z_{JK}$, and $Z_{J}$ are observed and independent of any other variable in the model. As observed variables, they do not require distributional assumptions in the same way the idiosyncratic errors do. Furthermore, their independence follows from the underlying random selection procedures that govern the variables[^9].

[^9]: Randomization ensures that data--and, by extension, an estimator--satisfies several key identification properties, such as common support, no interference, and consistency. The most critical property, however, is the elimination of confounding. *Confounding* occurs when an external variable, such as $X_{I}$, simultaneously influences both the outcome (e.g., $O_{R}$) and a variable of interest (e.g., $S$), resulting in spurious associations between the latter two [@Everitt_et_al_2010]. Randomization ensure the absence of confounding by effectively decoupling the association between the variable of interest and any other variable, except for the outcome itself. For a more detailed discussion on the benefits of randomization, see @Pearl_2009, @Morgan_et_al_2014, @Neal_2020, and @Hernan_et_al_2025.

<!-- 4. functional form for D_{R} -->
Next we define $D_{R}$ as the difference between the discriminal processes $T_{IA}[i, a]$ and $T_{IA}[h, b]$, representing the underlying written-quality trait of the compared texts, plus the corresponding repeated judge bias $B_{JK}[j, k]$. Note that if we assume that $B_{JK}[j,k]$ reflects the difference in stimulus-specific biases, i.e., $B_{JK}[j,k] = B_{JK}[i,a,j,k] - B_{JK}[h,b,j,k]$, we can re-write the discriminal difference as: 
$$
\begin{aligned}
D_{R} &= \left( T_{IA}[i,a] - T_{IA}[h,b] \right) + B_{JK}[j,k] \\
&= \left( T_{IA}[i, a] + B_{JK}[i, a, j, k] \right) - \left( T_{IA}[h, b] + B_{JK}[h, b, j, k] \right) \\
& = T^{*}_{IA}[i,a] - T^{*}_{IA}[h,b]
\end{aligned}
$$ {#eq-mat52}

This formulation reveals that the discriminal difference captures a *pure interaction effect*, in which neither the texts' discriminal processes nor the judges' biases alone determine the outcome, but their interaction does [@Attia_et_al_2022]. Put simply, this mathematical description captures the idea that the stimuli' discriminal processes become an observable outcome only through the lens of judges' perceptions (i.e., their biases). For clarity, the square brackets in $D_{R}$ indicate the relevant indices for each trait vector; they do not imply any subsetting of the data.

<!-- Weinberg[4] proposed a further diagram to capture ‘pure’ interaction in which neither G nor E has a direct effect on D and they only have an effect when present together (see Figure 4a). We suggest that this ‘pure’ interaction could be captured by the DAG shown in Figure 4b. An example of this is phenylketonuria in which phenylalanine in the diet (E) and the absence of the enzyme to break it down (phenylalanine hydroxylase, G) are both needed in order for a disease to manifest (intellectual disability, D). [@Attia_et_al_2022, pp. 1052] -->

<!-- This means that the cancellation of judges' biases may not hold in theory or practice, and that judges' biases may depend on the stimuli, by means of an interaction (but present) or because the judges works as a confounder? -->

<!-- Judges discrimination ability can be integrated in the models if we assume that the bias is not additive (as in Bramley explanation) but multiplicative. In the latter case the bias is not eliminated by taking just the difference. True, but it is false to think that because the biases are additive they get just eliminated. For instance, if we assume that the bias is additive but different for each stimuli, this means that the DAG should be: {T_{IA}, B_{JK}} -> O_{R}, but mathematically you can write it as it is above  -->

<!-- 5. Functional form and assumptions -->
We now specify the functional forms for $T_{IA}$, $T_{I}$, $B_{JK}$, and $B_{J}$. We model $T_{IA}$ as a linear combination of the students' underlying writing-quality traits $T_{I}$, the effects of relevant text-related variables on quality assessment $\beta_{XA}X_{IA}$ (such as the influence of text length), and the text-specific idiosyncratic errors $e_{IA}$. Similarly, we express $T_{I}$ as a linear combination of relevant student-related variables affecting the quality assessment $\beta_{XI} X_{I}$, and student-specific idiosyncratic errors $e_{I}$. For the judge-specific terms, we model $B_{JK}$ as a linear combination of the judge's individual bias $B_{J}$, the influence of relevant judgment-related variables on quality assessment $\beta_{ZK}Z_{JK}$ (e.g., how the number of judgments affect the evaluation), and judgment-specific idiosyncratic errors $e_{JK}$. Finally, we define $B_{J}$ as a linear combination of relevant judge-level variables influencing the quality assessment $\beta_{ZJ}Z_{J}$ (such as judgment expertise) and judge-specific idiosyncratic errors $e_{J}$.

<!-- 6. error assumptions -->
Next, we specify the probabilistic assumptions for the idiosyncratic errors $e_{I}$, $e_{IA}$, $e_{J}$, and $e_{JK}$. Unlike other variables in the model, these error terms exhibit indeterminacies in their *location*, *orientation*, and *scale* due to the lack of an inherent scale in the associated latent variables $T_{I}$, $T_{IA}$, $B_{J}$, and $B_{JK}$. Thus, to identify the latent variable model, we must resolve these indeterminacies [@Depaoli_2021; @deAyala_2009]. Drawing on principles from SEM [@Hoyle_et_al_2023], we assume that the vector of idiosyncratic errors $\boldsymbol{e} = [e_{I}, e_{IA}, e_{J}, e_{JK}]^{T}$, follows a Multivariate Normal distribution with mean vector $\boldsymbol{\mu}$ and a covariance matrix $\boldsymbol{\Sigma} = \boldsymbol{V} \boldsymbol{Q} \boldsymbol{V}$, with $\boldsymbol{V}$ denoting a diagonal matrix of standard deviations and $\boldsymbol{Q}$ a correlation matrix. To address the *location* indeterminacy, we set the errors' mean vector to zero:
$$
\boldsymbol{\mu} = [0, 0, 0, 0]^{T}
$$ {#eq-mat53}

Following [SCM -@fig-cj16_scm], we resolve the *orientation* indeterminacy by assuming that the errors are uncorrelated. This assumption leads us to define the error correlation matrix, $\boldsymbol{Q}$, as the identity matrix:
$$
\boldsymbol{Q} = \begin{bmatrix}
    1 & 0 & 0 & 0 \\
    0 & 1 & 0 & 0 \\
    0 & 0 & 1 & 0 \\
    0 & 0 & 0 & 1 
  \end{bmatrix}
$$ {#eq-mat54}

To resolve the *scale* indeterminacy, we define the diagonal matrix $\boldsymbol{V}$ as:
$$
\boldsymbol{V} = \begin{bmatrix}
    s_{XI} & 0 & 0 & 0 \\
    0 & p_{IA} & 0 & 0 \\
    0 & 0 & s_{ZJ} & 0 \\
    0 & 0 & 0 & p_{JK} 
  \end{bmatrix}
$$ {#eq-mat55}

Here, $s_{XI}$ represents the standard deviation for the individuals, $p_{IA}$ for the stimuli, $s_{ZJ}$ for the judges, and $p_{JK}$ for the judgments. We assume $s_{XI}$ varies depending on the teaching method group to which each student belongs. Using the example from @sec-theoretical, where the teaching method $X_{I} = \{1,2\}$, the model sets the constraint according to Equation ([-@eq-mat56]). This constraint anchors the scale of the individuals' latent trait while relaxing the assumption of equal dispersion for the stimuli, thereby addressing the concerns raised in @sec-theory-issue1a. 
$$
\sum_{g=1}^{2} s_{XI}[g]/2 = 1
$$ {#eq-mat56}

Because the error vector $\boldsymbol{e}$ follows an uncorrelated Multivariate Normal distribution, the marginal distribution of $e_{IA}$ is a univariate Normal distribution with mean zero and standard deviation $p_{IA}$. We set $P_{IA}$ as a proportion of $1$ to establish the scale of the stimuli' latent trait relative to the scale of the individuals' trait. Note that as a result, $T_{IA}$ is also normally distributed. This configuration effectively reinstates Thurstone’s original assumption of Normal discriminal processes for the stimuli (see @tbl-thurstone_cases).

Similarly, we assume $s_{ZJ}$ varies depending on the groups to which each judge belongs. For instance, if $Z_{J} = \{1,2,3\}$ represents three groups of judges with varying expertise, the model sets the constraint according to Equation ([-@eq-mat57]). This constraint anchors the scale of the judges' latent trait and relaxes the assumption of equal dispersion for the judgments. 
$$
\sum_{g=1}^{2} s_{ZJ}[g]/3 = 1
$$ {#eq-mat57}

Conversely, $p_{JK}$ is defined as a proportion of $1$ to establish the scale of the judgments' latent trait relative to the scale of the judges' trait.

<!-- Using assumptions from IRT [@deAyala_2009, p. 20-21], we are assuming a multidimensional model, that violates the local independence assumption -->

<!-- when you talk about the variability, refer that the equal dispersion is just assuming 1 for identification. -->

<!-- Within CFA, it is also necessary to resolve a series of indeterminacies, which are largely linked to the fact that the latent factors do not have a scale of their own. Indeterminacies exist in terms of the scale, location, and orientation of the factors. In order to handle these indeterminacies, we must set the scale of each of the latent factors. This can be handled in one of two ways. The first method that can be used is to standardize the latent factors such that the variances are set to 1 and the latent factor means are set to 0. The second method, and the one employed in the examples below, is to fix an otherwise free factor loading to a value of 1. [@Depaoli_2021, pp. 93] -->

<!-- [@deAyala_2009, p. 291-29] think about indeterminancies in MIRT. to solve metric indeterminancy you commonly fix a mu=0, and s=1, but for rotational indeterminancy, you need to define (maybe) orthogonality between units traits and judges bias -->

<!-- The difficulty and discrimination parameters for the questions should then be given a diffuse, or ideally a hierarchical prior, which will identify these parameters by scaling and locating relative to the student ability parameters. (https://mc-stan.org/docs/stan-users-guide/regression.html#priors-for-identification.section) -->

<!-- 9. Computational implementation -->
Finally, we use *Bayesian inference methods* to convert the statistical model [-@fig-cj16_stat] into a practical statistical tool for analyzing paired comparison data. Bayesian inference offers three main advantages in this context. First, it handles complex and overparameterized models, where the number of parameters exceeds the number of observations [@Baker_1998; @Kim_et_al_1999]. This feature is essential for our implementation, as the proposed model is indeed overparameterized. Second, it incorporates prior information to constrain parameter estimates within plausible bounds, thereby mitigating estimation issues like non-convergence or improper solutions that often affect frequentist methods [@Martin_et_al_1975; @Seaman_et_al_2011]. We primarily use this capability to define the error distribution and set the scale of latent variables [@Depaoli_2014]. Third, Bayesian inference supports robust inferences from small samples, where the asymptotic properties underlying frequentist methods are less reliable [@Baldwin_et_al_2013; @Lambert_et_al_2006; @Depaoli_2014]. This feature is particularly relevant in CJ assessments, as researchers often collect large volumes of paired comparisons but work with relatively small samples of judges, stimuli, and individuals to test hypotheses.

The **Declarations** section of this document provides a link to the model code, along with an alternative specification that assumes equal discriminal dispersions. We tested both versions of the model with success using `Stan` [@Stan_2020, version 2.26.1].

<!-- Proposal: Make the model to be bX %*% X 
Decision: Although this makes it easier for the software, the way it is declared is more explicit. -->

<!-- Proposal: Add new way of declaring non-centered parametrization and optional variables (See section 'Affinely transformed real' and 'Declaring optional variables' sections in https://mc-stan.org/docs/reference-manual/types.html) 
Decision: interesting, but I think I like the model as it is, explicit! -->

# Discussion {#sec-discussion}

<!-- recording: Sven and Tine 24.11.25; time: 00:56:10 - 00:58:20 -->
<!-- recording: Sven and Tine 25.04.10; time: 00:12:00 - 00:14:30 -->
<!-- recording: Sven and Tine 25.04.10; time: 00:41:55 - 00:47:00 -->
<!-- recording: Sven and Tine 25.04.22; time: 00:00:00 - 00:13:15 -->
<!-- recording: Sven and Tine 25.05.21; time: 00:00:00 - 00:03:30 -->
<!-- recording: Sven and Tine 25.05.21; time: 00:32:55 - 00:33:50 -->
<!-- recording: Sven and Tine 25.06.04; time: 00:00:00 - 00:05:40 -->

<!-- 1. Summary of Thurstone's general theory -->
Thurstone introduced the Law of Comparative Judgment to measure psychological traits of stimuli through pairwise comparisons [@Thurstone_1927a; @Thurstone_1927b]. In its general form, the theory models single-judge comparisons across multiple, potentially correlated stimuli. Each comparison produces a dichotomous outcome indicating which stimulus the judge perceives as having a higher trait level. However, Thurstone identified one key challenge in this general formulation: the measurement model required estimating more "unknown" parameters than the number of available pairwise comparisons [@Thurstone_1927b]. To address this issue and facilitate the theory's practical applicability, he formulated five cases, each progressively incorporating several simplifying assumptions.

<!-- 2. Summary of the why the literature uses Case V, and its problems -->
Among these, Case V remains the most widely used model in empirical CJ research, mainly due to the widespread adoption of the BTL model. The BTL model mirrors the core assumptions of Case V--namely, equal discriminal dispersions and zero correlation among stimuli' discriminal processes--but replaces the processes' normal distribution with the more mathematically tractable logistic distribution [@Andrich_1978; @Bramley_2008]. Although this substitution has minimal impact on trait estimation or model interpretation [@vanderLinden_et_al_2017_I; @McElreath_2021], the simplifying assumptions of the BTL model--and by extension, of Case V--may fail to capture the complexity of some traits or account for heterogeneous stimuli [@Thurstone_1927a; @Andrich_1978; @Bramley_2008; @Kelly_et_al_2022], potentially leading to unreliable and inaccurate trait estimates [@Ackerman_1989; @Zimmerman_1994; @McElreath_2020; @Hoyle_et_al_2023].

Moreover, because Thurstone's original goal was to produce a "coarse scaling" of traits and allocate stimuli along this continuum [@Thurstone_1927a, pp. 269], his theory offered no guidance on how to use trait estimates for statistical inference. The CJ tradition has attempted to address this gap by separating trait estimation from hypothesis testing, relying on point estimates--such as BTL scores or their transformations--for inference. Notably, while this approach simplifies analysis, it can also introduce bias and compromise the reliability of the resulting inferences [@McElreath_2020; @Kline_et_al_2023; @Hoyle_et_al_2023].

<!-- 4. General summary of what we have done. -->
Thus, to address the limitations of Thurstone's Case V and the BTL model, this study extends Thurstone's general form using a systematic, integrated approach that combines causal and Bayesian inference methods. The approach begins by formulating a conceptual model represented by a Structural Causal Model (SCM) and a Directed Acyclic Graph (DAG). This model integrates Thurstone's core theoretical principles, such as the discriminal processes of stimuli, alongside key CJ assessment design features, including judges' bias, sampling procedures, and comparison mechanisms, thereby disentangling the causal processes in the CJ system.

The approach then translates the SCM into a bespoke statistical model that allows researchers to analyze CJ data when violations to the assumptions of equal dispersion and zero correlation occur (see @sec-theory-issue1a and @sec-theory-issue1b) and when statistical inference is necessary (refer to @sec-theory-issue2). In particular, this model accounts for judge biases (see @sec-theory-theoretical_P1), captures the hierarchical structure of stimuli, incorporates measurement error into the hypothesis testing process (refer to @sec-theory-theoretical_P3), and accommodates heterogeneity in discriminal dispersions (see @sec-statistical). By addressing all these issues, these methodological innovations have the potential to enhance the reliability and validity of trait measurement in CJ [@Perron_et_al_2015], while also improving the accuracy of statistical inferences.

<!-- Reliability is a necessary but not sufficient condition for validity. Reliability can exist without validity but validity cannot exist without reliability. [@Perron_et_al_2015, pp. 2] -->

<!-- 5. Are there any additional advantages? -->
Beyond these potential benefits, the approach offers two additional advantages. First, it clarifies the roles and interactions of all actors and processes involved in CJ assessments. Second, it shifts the analytic paradigm from passively accepting the assumptions of the BTL model to actively testing their fit with observed data. Together, these advantages establish a principled framework for evaluating best practices in designing CJ assessments, one that better aligns with the demands of contemporary CJ contexts [@Kelly_et_al_2022], offering new insights into existing research and opening promising avenues for future inquiry.

<!-- If Thurstone's Law is to be used in support of comparative judgment for assessment, his theory must be extended to cover the uses and applications that are required for assessment purposes. [@Kelly_et_al_2022, pp. 678] -->

<!-- We recognise that there is unlikely to be a single answer to these questions for all the cases in which comparative judgment may be used. However, work to derive some general principles for good practice would be valuable. [@Kelly_et_al_2022, pp. 683] -->

<!-- These procedures also bear on a larger topic of discussion in social science. The proper response, when faced with low predictive capacity of presumed independent variables, is not only to search for new independent variables and/or alternative theoretical models; the proper recourse may also be to use more sophisticated measurement models. [@Wainer_et_al_1978, pp. 332] -->

## Future research directions based on this study's approach {#sec-discussion_RA}
<!-- recording: Sven and Tine 25.06.04; time: 00:05:40 - 00:07:25 -->

Among the many potential directions for future research--both current and new--three avenues deserve particular attention due to their direct impact on the reliability and validity of trait estimates, as well as on the accuracy of statistical inferences. The following sections outline these avenues and explain how our approach provides a rigorous research framework to investigate them.

### The impact of sampling and comparison mechanisms {#sec-discussion_RA1}
<!-- recording: Sven and Tine 25.05.21; time: 00:03:30 - 00:24:00 -->
<!-- recording: Sven and Tine 25.06.04; time: 00:07:25 - 00:11:55 -->

<!-- 1. The overlooked role of the sampling and comparison mechanisms in CJ -->
Although sampling and comparison mechanisms are central to modern CJ assessments, it is striking that most of the CJ literature has examined them only within a limited scope. Researchers have primarily investigated the effects of adaptive comparative judgment (ACJ) designs on trait reliability [@Pollitt_2012a; @Pollitt_2012b; @Bramley_2015; @Verhavert_et_al_2022; @Mikhailiuk_et_al_2021; @Gray_et_al_2024] or proposed practical guidelines for the number of comparisons judges should make [@Verhavert_et_al_2019; @Crompvoets_et_al_2022]. While these studies offer valuable insights, they often overlook the broader role these mechanisms play within the CJ system. Moreover, as this oversight likely stems from a broader lack of conceptual clarity about how these mechanisms function within the system, this study formally incorporates these mechanisms into CJ's conceptual model.

<!-- 2. The missing data perspective on sampling and comparison mechanisms -->
The explicit integration of the sampling and comparison mechanisms offers a new perspective on how these mechanisms shape the CJ process. In particular, it clarifies their role in the data-generating process by framing them as sources of missing data--that is, as mechanisms that determine which observations are missing from the final data sample. This new perspective encourages the application of Little and Rubin's principled missing data framework [-@Little_et_al_2020], enabling a more rigorous evaluation of existing claims about these missing data mechanisms, their influence on CJ outcomes, and their implications for designing and assessing more complex assessments setups.

<!-- 3. Leveraging MCAR assumptions to simplify CJ analysis in this study -->
Notably, this study circumvents the need to apply this missing data framework by deliberately structuring the sampling and comparison mechanisms to be independent of any observed or unobserved variables, including the outcome. In other words, we designed these mechanisms so they produce data that are *missing completely at random* (MCAR) [@Little_et_al_2020]. This design offers one key advantage: it generates simple random samples that satisfy the condition of *ignorability*, allowing researchers to legitimately *ignore* missing data during analysis without introducing bias [@Everitt_et_al_2010; @Kohler_et_al_2019; @Neal_2020].

<!-- Well, what assumption(s) would make it so that the ATE is simply the associational difference? This is equivalent to saying “what makes it valid to calculate the ATE by taking the average of the Y(0) column, ignoring the question marks, and subtracting that from the average of the Y(1) column, ignoring the question marks?” This ignoring of the question marks (missing data) is known as ignorability. Assuming ignorability is like ignoring how people ended up selecting the treatment they selected and just assuming they were randomly assigned their treatment; we depict this graphically in Figure 2.2 by the lack of a causal arrow from X to T. [@Neal_2020, pp. 9] -->

<!-- The condition of case selection to be independent from Y is always met in experiments because the Y is not realized before the experiment starts. [@Kohler_et_al_2019, pp. 160] -->

<!-- 4. More complex CJ designs and the implications of MAR or MNAR -->
However, many modern CJ applications rely on more complex assessment designs in which the sampling and comparison mechanisms introduce more intricate forms of missingness--such as *missing at random* (MAR) or *missing not at random* (MNAR) [@Little_et_al_2020]. A prominent example is the previously discussed ACJ designs, where prior judgment outcomes inform the selection of stimulus pairs for subsequent comparisons [@Pollitt_2012a; @Pollitt_2012b; @Bramley_2015]. This outcome-informed pair selection suggests that the comparison mechanism is outcome-dependent, potentially classifying ACJ as a generator of MNAR data. This, in turn, facilitates a more nuanced understanding of the mixed evidence surrounding ACJ's capabilities. While some studies report that the method improves trait reliability [@Pollitt_et_al_2003; @Pollitt_2012a; @Pollitt_2012b], others argue that it artificially inflates these gains [@Bramley_2015; @Bramley_et_al_2019; @Crompvoets_et_al_2020; @Crompvoets_et_al_2022], potentially as a result of outcome-dependent missingness.

<!-- The A for ‘adaptivity’ in ACJ means that the choice of which objects are presented to the judges depends on the outcomes of judgments made so far – the idea being to use judge time efficiently by not presenting objects to compare where the result of the comparison is almost certain. The analogy is with computerised adaptive testing (CAT) where the choice of the next item to present is based on the examinee’s success or failure on items presented so far. [@Bramley_2015, pp. 4]. -->

<!-- In contrast, case-control studies [@Everitt_et_al_2010] use sampling mechanisms tied to non-assignable variables that likely influence--or are influenced by--the trait of interest. For example, @Boonen_et_al_2020 formed comparison groups based on children’s hearing status, a non-assignable variable likely affecting speech quality [@Boonen_et_al_2023; @Rivera_et_al_2024]. Notably, although the authors carefully constructed comparable groups, they noted that some subgroups--such as children with hearing aids--were not representative of their broader population. This is a common challenge in such designs, care must be taken to ensure that the excluded individuals are not systematically different in trait performance. Otherwise, the resulting trait estimates or the statistical inferences may be biased. -->

<!-- (Done) It is more likely that we can decide the characteristics of the objects we want to include in the study (independent variables), ensuring groups that are more balanced. However, in the case of subjects sometimes the independent variables cannot be assigned (e.g., hearing status) -->

<!-- [another example: For instance, consider a CJ experiment assessing speech quality using a case-control design [@Everitt_et_al_2010] that samples individuals with normal hearing (NH) and those with hearing impairment (HI). Here, the sampling mechanism depends on hearing status--a variable likely to influence the trait of interest, as HI individuals may exhibit lower speech quality [@Boonen_et_al_2023; @Rivera_et_al_2024]. Suppose a subgroup within the HI population, such as individuals with genetically induced hearing loss, is difficult to reach. If this group is systematically excluded and happens to over- or under-perform in speech quality, the resulting trait estimates may be biased. It is therefore essential to determine whether such biases occur--and if so, whether they undermine the validity of trait estimates and the robustness of statistical inferences--and our framework provides a rigorous setting to test this claim.]{style="color:blue;"} -->

<!-- [another example: For instance, consider a high-stakes non-mandatory summative CJ assessment, where the results the individuals have in this assessment have some implications for them, however, their participation is not mandatory. In this case, it not hard to imagine that some of these individuals will not participate in the CJ assessment, thus affecting the sample, and ultimately the trait estimates: now a probable under performing group will not available in the data sample, thus biasing the resulting trait estimates. It is therefore essential to determine whether such biases occur--and if so, whether they undermine the validity of trait estimates and the robustness of statistical inferences--and our framework provides a rigorous setting to test this claim.]{style="color:blue;"} -->

<!-- Comment for myself: when the sampling or comparison mechanism depend on the precision of the measured trait, the DAG can be depicted as: 
{S, C} -> {Y}
e_{A} -> {S, C, Y} 
this means that e_{A} is a confounder?,  
then should you: "control for" the variable?, or the equivalent, break the relationship e_{A} -> {C, S} by taking a random sample of stimuli?. Does this mean that random algorithms per se will always be unbiased?. -->

<!-- 5. The importance of clarifying assumptions behind CJ assessment designs -->
Nevertheless, regardless of the underlying missingness mechanisms, any CJ assessment design would benefit from explicitly defining its assumptions--a practice supported by our proposed approach. This clarity enables researchers to evaluate how the sampling and comparison mechanisms affect trait estimation and statistical inference within each design. Such assessments are particularly relevant given the common misconception in the CJ literature that Thurstone's model can naturally handle even non-random missing data without compromising the reliability or validity of trait estimates [@Bramley_2008].

<!-- Thus Thurstone’s model (in both its Case 5 and Rasch formulation) achieves the goal of sample-free calibration (Wright, 1977) in the sense that the estimate of the distance between A and B does not depend on which other objects they are compared with. The practical importance of this is that it is not necessary to have a completely crossed or balanced design of paired comparisons because the estimation process naturally handles missing data. [@Bramley_2008, pp. 256] -->

<!-- if the data fit the Case 5 Thurstone model, the estimate of the separation between any two scripts does not depend on which other scripts they are compared with. This means that data can be missing in a non-random way without affecting the results. However, the precision (standard error) of each script’s estimate depends largely on how many comparisons it has been involved in, so some effort needs to be made to ensure that each script is compared a similar number of times in total. [@Bramley_2008, pp. 260] -->

<!-- Since the analysis can cope with non-random missing data (see earlier), the collection of data can be very flexible – tailored to the needs of the particular situation. Because the comparative judgements are all independent, data can be accumulated as needed. The investigator is thus not (in principle) bound by a prior design. At any point more data could be collected and added to the existing data to recover from accidents, to replace an unacceptable judge, to add extra scripts, to resolve conflicts, to investigate possible bias or to deepen the study of any aspect that proves especially interesting. [@Bramley_2008, pp. 261] -->

### The effects of judges' bias on the reliability of traits {#sec-discussion_RA2}
<!-- recording: Sven and Tine 25.05.21; time: 00:24:00 - 00:31:45 -->
<!-- recording: Sven and Tine 25.06.04; time: 00:11:55 - 00:16:50 -->

<!-- 1. Toward a more realistic CJ model: recognizing judges' bias -->
Despite the growing notion that various stimulus-related factors can influence judges' perceptions [@vanDaal_et_al_2016; @Lesterhuis_et_al_2018; @Chambers_et_al_2022] and that these influences may not always cancel each other out, few studies in the CJ literature provide empirical evidence for judges' biases [@Pollitt_et_al_2003; @vanDaal_et_al_2016; @Bartholomew_et_al_2020a]. This gap likely persists not due to a lack of interest or research but because researchers often rely on ad-hoc detection methods--such as 'misfit' statistics--that are poorly suited for the task [@Kelly_et_al_2022]. To address this limitation, this study treats judges' biases as an integral component of the CJ system from the outset. This approach offers one key advantage: it provides a more accurate representation of the data-generating process behind pairwise comparisons, one that acknowledges that the discriminal processes of stimuli become an observable outcome only through judges' perceptions, which may exhibit bias.

<!-- Such differences may not be detected through analyses of bias and misfit. McMahon and Jones (2015) found differences between students and teachers in what was valued in assessments of understanding of a chemistry experiment. Where students prioritized factual recall, teachers prioritized scientific explanation. Both groups produced highly reliable scales (reliability estimates of .893 for students, and .874 for teachers). Although, in this case, a clear argument as to whose consensus should take priority can be made, it demonstrates the existence of different, but equally reliable, sets of consensuses depending on who is asked. [@Kelly_et_al_2022, pp. 683] -->

<!-- 2. Relevant research question paved by recognizing judges' bias (part I) -->
The explicit integration of judges' bias into CJ's conceptual model then paves the way for investigating several relevant research questions. One key question is whether researchers can validly analyze CJ data under the assumption of "sample-free" trait calibration, specifically under the hypothesis that judges exhibit no systematic bias. This question is particularly relevant because many researchers still regard "sample-freeness" as an inherent property of the BTL model [@Bramley_2008; @Andrich_1978] despite growing evidence of persistent biases. Another critical question, discussed in @sec-theory-issue1b, is whether training or expertise can help judges avoid focusing on irrelevant stimulus features--such as handwriting--over more central criteria like argumentative quality in writing assessments [@Kelly_et_al_2022]. Notably, exploring these questions may also provide insights into what it truly means to be an "expert" within the CJ context [@Kelly_et_al_2022].

<!-- The elimination of the subject parameters in the Rasch model has given rise to such expressions as "sample-free"" and "population-free." In this sense, then, so is the Thurstone Case V model. Although these phrases suggest desirable properties, they are not always fully appreciated; therefore, these terms have been avoided so far. A clarification seems in order and possible in this context.
Because subject parameters are eliminated explicitly, their values (individually or collectively) need not be known. As a result, no assumptions need be made about the distribution of these parameters in the population, for example, that they are normally distributed. This does not mean, of course, that the "empirical" subject population need not be specified or described or that the scale values will be the same in different empirical populations. In fact, it is generally vital that the empirical populations be described. But because it does not matter what the actual values of the subject parameters are in the population, the models are, in the statistical distribution sense of population, "population-free." This is a property of the model and independent of empirical data.
It is important to note that this holds only with respect to the estimation of the relative scale values of the items. It does not matter if the subject parameters are large, small, or otherwise; the subjects’ impressions of the relative positions of the items should be the same. That is, items which appear more intense to one group should appear similarly more intense to the other groups. Consequently, different samples, whether drawn randomly or otherwise from the same empirical population, should give the same relative scale value estimates. In this sense the models are also "sample-free." 
However, this property of the model is true in empirical data only if the model holds in the data. Therefore, "sample-freeness" is always only a hypothesis until checked, as is the assumption that the model holds. ...
Note that it would be missing a crucial point to choose two random samples from which to check "sample-freeness" and that it would not be putting the models to any special test. Two random samples should give the same relative scale values because it is an inherent assumption of all statistical models that parameter estimates are "random sample-free." For the greatest benefits, the samples should be systematically different. [@Andrich_1978, pp.459] -->

<!-- In order to make it possible to estimate the parameters satisfactorily with the relatively small amount of data available it would probably be necessary to reduce the number of parameters estimated in some other way, for example, by representing all the scripts from one board with a single parameter, as in the logistic regression model used in Greatorex et al. (2002). Such an approach shifts the philosophy of the exercise from that of Thurstone and Rasch (constructing psychological scales capable of yielding sample-free measurement) to the philosophy of statistical modelling (finding a model that optimises parsimony and variance explained). [@Bramley_2008, pp. 274] -->
<!-- Comment: the sample-free measurement philosophy is an illusory one, because in this contexts, the measurements depend on the perception of judges. -->

<!-- However, the definition of expert in this assessment context is further complicated by studies suggesting that training is unnecessary for comparative judgment exercises to be conducted well. [@Kelly_et_al_2022, pp. 681] -->

<!-- The primary writing study showed that both trained markers and teachers with no marker training can use the method effectively to assess the quality of children’s writing, and the e-scape project has demonstrated the same for the assessment of e-portfolios in Design & Technology. We do not yet know how widely it can be, or should be, applied. [@Pollitt_2012b, pp. 297]-->

<!-- 2. Relevant research question paved by recognizing judges' bias (part II) -->
Moreover, since judges draw on a variety of intricate stimulus characteristics when evaluating complex, multidimensional traits [@vanDaal_et_al_2016; @Lesterhuis_et_al_2018; @Chambers_et_al_2022], and these characteristics account for variation in judgment accuracy [@Gill_et_al_2013; @vanDaal_et_al_2017; @vanDaal_2020; @Gijsen_et_al_2021], it is reasonable to expect that judges' assessments also vary with individual attributes such as gender, age, culture, income, education, training, or expertise [@Kelly_et_al_2022; @Bartholomew_et_al_2020a]. Thus, building on the discussion in @sec-discussion_RA1, researchers could investigate how the selection of judges influences the formation of a "shared consensus" and whether these individual attributes introduce systematic biases or distortions in the observed trait distribution [@Deffner_et_al_2022]. Furthermore, if such attributes indeed undermine the assumption of "sample-freeness," then it becomes essential to explore strategies to control their influence and to determine how many judges--and how many judgments per judge--are needed to produce reliable trait estimates in these contexts. Additionally, it is worth considering whether *repeated measures designs*, in which judges evaluate the same stimulus pairs multiple times [@Lawson_2015], can improve judgment consistency and accuracy.

<!-- The logic is that in aggregating the varied judgments, one comes to an overall view of the community of practice. However, this raises important questions about diversity of the judging pool. If it is acknowledged that judges may differ in what they value, it is not inconceivable that judges may differ as a function of their gender, age, ethnicity, culture, class, income, education and so on. Would the aggregate view of young, British-Asian men always be the same as the aggregate view of older, black women? For example, Bartholomew et al. (2020) found differences between what judges from the UK, Sweden, and the USA valued when assessing design portfolios, providing evidence to support divergence along demographic lines. [@Kelly_et_al_2022, pp. 683] -->

<!-- Such differences may not be detected through analyses of bias and misfit. McMahon and Jones (2015) found differences between students and teachers in what was valued in assessments of understanding of a chemistry experiment. Where students prioritized factual recall, teachers prioritized scientific explanation. Both groups produced highly reliable scales (reliability estimates of .893 for students, and .874 for teachers). Although, in this case, a clear argument as to whose consensus should take priority can be made, it demonstrates the existence of different, but equally reliable, sets of consensuses depending on who is asked. [@Kelly_et_al_2022, pp. 683] -->

<!-- Samples from different sites often differ in terms of their demographic profiles (here, their age and gender distribution), and these demographic variables might in turn affect the distribution of the trait of interest. How should researchers deal with these differences? The answer depends on the processes that generated the observed disparities. Demographic disparities among samples may result from (a) differences in the actual populations from which the samples are taken or (b) sampling procedures that differ among sites. [@Deffner_et_al_2022, pp. 6] -->

<!-- These examples demonstrate that the generalizability of experimental effects does not depend on the presence of population differences per se but on the exact mechanisms by which populations differ. While some differences—especially those concerning the independent variable—are inconsequential for intended generalizations, differences concerning effect modifiers or mediators require statistical adjustment. [@Deffner_et_al_2022, pp. 11] -->

<!-- To date, there is no evidence to support a minimum number of judges and judgments required to ensure an acceptable level of validity [@Kelly_et_al_2022, pp. 682]. -->

<!-- The field of comparative judgment, while growing, is still relatively small, and at present we do not have compelling answers to fundamental questions such as: how many judges do we need? Who should judge? How many judgments should they make? And crucially, how can we be confident that they are basing their judgments on acceptable criteria? If judges are not using construct-relevant criteria, then the argument that comparative judgments are necessarily valid cannot be upheld. We recognise that there is unlikely to be a single answer to these questions for all the cases in which comparative judgment may be used. However, work to derive some general principles for good practice would be valuable. [@Kelly_et_al_2022, pp. 683] -->

<!-- It is assumed that the single observer compares each pair of stimuli a "sufficient number of times" so that a proportion, P(A>B), may be determined for each pair of stimuli [@Thurstone_1927b, pp. 267] -->

### The identification of 'misfitting' stimuli and judges {#sec-discussion_RA3}
<!-- recording: Sven and Tine 25.05.21; time: 00:31:45 - 00:32:20 -->
<!-- recording: Sven and Tine 25.06.04; time: 00:33:55 - 00:53:15 -->

[LAST THING TO DO]{style="color:blue;"}

<!-- Although the CJ literature clearly defines *misfit judges* as those whose assessments deviate markedly from the "shared consensus" [@Pollitt_2012a; @Pollitt_2012b; @vanDaal_et_al_2016; @Goossens_et_al_2018; @Wu_et_al_2022], there has been little discussion about how these misfit judges relate to Thurstonian theory. In particular, it remains unclear which aspects of Thurstone's theory might explain the occurrence of misfit judges. To address this, we propose three statistical perspectives: (1) misfit judges are simply judges who exhibit bias; (2) misfit judges are those whose judgments display greater variability than their peers, and (2) misfit judges are indeed outliers. -->

<!-- Regarding the first perspective, @sec-theory-issue1b explains in detail how, due to the assumption of zero correlation, judges' biases can be miss classified as *misfits*. Meanwhile, @sec-discussion_RA2 discusses how our approach explicitly account for such biases, aiming to reduce the reliance on ad-hoc detection methods--such as misfit statistics--which are poorly suited to the task [@Kelly_et_al_2022]. -->

<!-- Regarding the second perspective -->

<!-- Conversely, regarding the second perspective, the very definition of *misfit judges* leads to their identification as outliers--that is, observations that deviate significantly from the rest of the sample in which they occur [@Grubbs_1969].  -->

<!-- Such differences may not be detected through analyses of bias and misfit. McMahon and Jones (2015) found differences between students and teachers in what was valued in assessments of understanding of a chemistry experiment. Where students prioritized factual recall, teachers prioritized scientific explanation. Both groups produced highly reliable scales (reliability estimates of .893 for students, and .874 for teachers). Although, in this case, a clear argument as to whose consensus should take priority can be made, it demonstrates the existence of different, but equally reliable, sets of consensuses depending on who is asked. [@Kelly_et_al_2022, pp. 683] -->

<!-- An outlying observation, or "outlier," is one that appears to deviate markedly from other members of the sample in which it occurs. [@Grubbs_1969, pp. 1] -->

<!-- In the previous subsection we argue that ad-hoc detection methods, such as 'misfit' statistics, could be poorly suited for the task of detecting judges bias [@Kelly_et_al_2022]. This hinges in the assumption that the CJ data does have judges biases, but the method does not manages to identify them, leading to false negatives? or lack of sensitivity in the tests. However, there is a related problem which is: these ad-hoc method could be detecting judges' biases that are not there. as in @Wu_et_al_2022  -->

<!-- These procedures also bear on a larger topic of discussion in social science. The proper response, when faced with low predictive capacity of presumed independent variables, is not only to search for new independent variables and/or alternative theoretical models; the proper recourse may also be to use more sophisticated measurement models. [@Wainer_et_al_1978, pp. 332] -->

<!-- "Another challenge to the intrinsic validity argument is that the aggregate approach and valuing of diversity in expert views (Van Daal et al., 2019) is in tension with the idea of removing misfitting judges (Pollitt, 2012a). If someone is considered an expert, and the validity of comparative judgment rests on the collective expertise of a community of practice, the decision to remove them and discard their judgments can only be justifiable on the assumption that those judgments do not reflect their expertise. ... It is thus unclear whether or not this divergence is welcomed." [@Kelly_et_al_2022, pp. 681] -->

<!-- A similar point can be made for *misfit stimuli*--that is, those that elicit more judgment discrepancies than others [@Pollitt_2004; @Pollitt_2012b; @Pollitt_2012a; @Goossens_et_al_2018]. It remains unclear which aspects of Thurstone's theory might explain the occurrence of misfit stimuli. To address this, we propose two statistical perspectives: (1) misfit stimuli are those whose discriminal processes exhibit greater variability compared to other stimuli, and (2) are statistical outliers. -->

<!-- Regarding the first perspective, @sec-theory-issue1a  -->

<!-- , possibly due to larger discriminal dispersions or because they are genuine outliers--meaning, texts with distinctive characteristics that deviate markedly from the rest of the sample [@Grubbs_1969]. In either case, the BTL model neither accounts for these anomalies nor provides tools for addressing them, apart from excluding the "problematic" texts from analysis. -->

<!-- The same could be said for the stimuli. These misfit texts likely arise from two underlying mechanisms. First, a text may exhibit a larger discriminal dispersion, increasing the overlap between its discriminal process and those of multiple other texts. This greater overlap leads to more judgment discrepancies, increasing the likelihood of classifying the text as a misfit. @fig-discriminal_process illustrates this scenario. In the figure, Text B has a wider discriminal dispersion than Text A, enlarging the region where their discriminal processes intersect. This larger overlap represents an area where judges face greater difficulty in determining which text exhibits a higher level of the trait. Second, a text may represent an outlier --one with distinct characteristics that deviate markedly from the rest of the sample [@Grubbs_1969]. Such texts are harder to compare and place on the trait scale, resulting in more judgment discrepancies and a misfit classification. In either case, the BTL model is inherently agnostic to these mechanisms and does not provide any means to address them. -->

<!-- Indeed, the presence of the so-called *misfit* texts may already signal these limitations, as these are texts that elicit more judgment discrepancies than others [@Pollitt_2004; @Pollitt_2012b; @Pollitt_2012a; @Goossens_et_al_2018], possibly due to larger discriminal dispersions or because they are genuine outliers--meaning, texts with distinctive characteristics that deviate markedly from the rest of the sample [@Grubbs_1969]. In either case, the BTL model neither accounts for these anomalies nor provides tools for addressing them, apart from excluding the "problematic" texts from analysis. -->

<!-- The outcomes of the analysis can be considered in two parts: the ‘model’ and the ‘misfit’. [@Bramley_2008, pp. 260] -->

<!-- It has a good reporting structure, and the pattern of both judge and script fit and misfit enables further investigation of the biases which may be taking place. [@Pollitt_et_al_2003, pp. 7]. -->

<!-- We advocate for the no removal of judges, but rather identify them and treat the data with the appropriate model, so these 'misfits' do not affect the measurement of the trait -->

<!-- Fortunately, the statistical literature offers a solution to address violations of the equal dispersion assumption in CJ experiments. This solution involves using models that extend traditional approaches to account for the different variability in the stimuli, referred to as over-dispersed models [@McElreath_2020, chap. 12]. -->

<!-- Furthermore, these points tend to be outliers with unlikely values, according to the model. As a result, they are highly influential and make it difficult to estimate out-of-sample predictive accuracy. Why? Because any new sample is unlikely to contain these same outliers, and since these outliers were highly influential, they could make out-of-sample predictions worse than expected. [@McElreath_2020, pp. 231] -->

<!-- What can be done about this? There is a tradition of dropping outliers. People sometimes drop outliers even before a model is fit, based only on standard deviations from the mean outcome value. You should never do that—a point can only be unexpected and highly influential in light of a model. After you fit a model, the picture changes. If there are only a few outliers, and you are sure to report results both with and without them, dropping outliers might be okay. But if there are several outliers and we really need to model them, what then? [@McElreath_2020, pp. 232] -->

<!-- The problem isn’t necessarily that “outliers” are bad data. Rather processes are often variable mixtures and this results in thicker tails. [@McElreath_2020, pp. 369] -->

<!-- Consider, for example, the simplest location and scale problem-inference for a normal distribution with unknown mean \theta and variance \sigma^{2}. If we look at the marginal posterior distribution for \theta we know that if \sigma^{2} is known this posterior exhibits outlier-resistance, and we might expect similar behavior for unknown \sigma^{2}, but a simple example refutes this. Let the prior distribution give probability p (0<p<1) to (\theta = 1, \sigma^{2} = 1) and 1-p to (\theta = 0, \sigma^{2} = 4). Then the posterior mean of \theta is a decreasing function of any sufficiently large observation. [@OHagan_2018, pp. 366] -->

<!-- The problem isn’t necessarily that “outliers” are bad data. Rather processes are often variable mixtures and this results in thicker tails. [@McElreath_2020, pp. 369] -->

<!-- Second, and complementary to the previous point, the statistical literature consistently highlights the risks of using ad hoc procedures to identify "outliers" when the model does not accurately represent the data-generating process. For instance, McElreath (2020, pp. 232) expressed that: "There is a tradition of dropping outliers. People sometimes drop outliers even before a model is fit, based only on standard deviations from the mean outcome value. You should never do that--a point can only be unexpected and highly influential in light of a model. After you fit a model, the picture changes. If there are only a few outliers, and you are sure to report results both with and without them, dropping outliers might be okay. But if there are several outliers, we really need to model them". Similarly, O'Hagan (2018, pp. 366) makes a small example explaining how the average of a parameter--much like a trait score you could get from the BTL model--depends on outliers. Miller (2023), in psychology, explores a similar topic. -->

<!-- You might notice I use the term "outlier" here to describe biased judges. This is because, like statistical outliers, these judges deviate significantly from the rest of the sample in which they occur (Grubbs, 1969). However, the key reason for using this terminology lies in that any ad hoc method for flagging biased judges based on the BTL model--such as misfit statistics--effectively acts as an outlier detection tool. This is due to the fact that the BTL model relies on several strong assumptions, which limit its capacity to capture other features of the data unless supplemented by additional procedures--procedures that, in essence, resemble outlier detection.  -->

<!-- Moreover, from McElreath's text, you might recognize that the practice of removing outliers mirrors the act of discarding biased judges in CJ using misfit statistics, where these statistics are essentially transformations of residual probabilities linked to trait estimates, or as McElreath puts it "standard deviations from the mean outcome values." But what are the potential consequences of wrongfully identifying and discarding misfit judges, you might ask?. Well, this practice risks removing valuable information and introducing bias into trait estimates [Zimmerman, 1994; O'Hagan, 2018; McElreath, 2020]. The direction and magnitude of these biases are unpredictable, depending on which judges are excluded from the analysis. -->

## Study limitations and practical challenges for applied CJ researchers {#sec-discussion_challenges}
<!-- recording: Sven and Tine 25.05.21; time: 00:33:50 - 00:42:20 -->
<!-- recording: Sven and Tine 25.06.04; time: 00:18:45 - 00:33:55 -->

The process of deriving conclusions from observed data always requires assumptions, whether the data are observational or experimental [@Kohler_et_al_2019; @Deffner_et_al_2022]. Our approach is not an exception to this fundamental principle. As with all approaches grounded on causal inference, ours has the limitation of relying on expert knowledge and assumptions about the variables' causal structure that are often untestable at the outset [@Hernan_et_al_2025]. As such, the approach does not seek to yield automatic answers when applied to a given CJ dataset. Instead, it aims to encourage researchers to formulate precise questions and to make their assumptions explicit, fostering a generalizable understanding of the CJ system under study [@Rohrer_et_al_2022; @Deffner_et_al_2022; @Sterner_et_al_2024]. This clarity is critical because a model's ability to produce accurate estimates and valid inferences depends heavily on how well the data and inferential goals align with its underlying assumptions [@Kohler_et_al_2019]. Although this alignment remains empirically untested for the proposed models, the theory-driven nature of our approach provides a solid foundation for future empirical evaluation of its causal assumptions [@Deffner_et_al_2022], which are well supported by both theory and existing evidence (see @sec-theory-issues and @sec-theoretical).

<!-- Statements about the real world from observed data require assumptions. This is true no matter what kind of data we have, and it is explicitly also true for statements based on data stemming from SRSg. The validity of our statements therefore critically hinges on the validity of those assumptions. If our statistical assumptions are wrong for the analysis at hand, the inferential statements will be incorrect no matter how sophisticated the statistical methods are. [@Kohler_et_al_2019, pp. 150] -->

<!-- This approach makes strong causal assumptions about the nature of confounding and our ability to measure shared history. However, strong assumptions are always necessary in observational settings. What is important is that the assumptions are transparent and logically connected to data analysis. [@Deffner_et_al_2022, pp. 13] -->

<!-- Causal inference generally requires expert knowledge and untestable assumptions about the causal network linking treatment, outcome, and other variables. ...
Because the scenarios we considered were so simple, there was really no need to make the causal network explicit. As we start to turn our attention towards more complex situations, however, it will become crucial to be explicit about what we know and what we assume about the variables relevant to our particular causal inference problem. [@Hernan_et_al_2025, pp. 73] -->

<!-- Many cross-cultural scientists already pay close attention to concerns of causal inference and comparison without use of a formal framework (e.g., Norenzayan & Heine, 2005; Pollet et al., 2014). For these researchers, a formalized framework can provide a vocabulary to articulate their concerns and work toward solutions in a more systematic manner. [@Deffner_et_al_2022, pp. 2] -->

<!-- A generalizable understanding of a given phenomenon, therefore, cannot be based only on the accumulation of data but requires the theory-driven testing of causal assumptions. [@Deffner_et_al_2022, pp. 13] -->

<!-- A structural causal framework encourages researchers to explicitly spell out their assumptions, removing verbal ambiguity and facilitating communication, and it calls for a cumulative approach to science as one study’s findings become the scaffolding assumptions of the next. [@Deffner_et_al_2022, pp. 14] -->

<!-- An explicit causal-inference framework makes it (at times painfully) transparent how strong the assumptions are that we need to arrive at substantive conclusions and how little we collectively know about many real-world phenomena. But this is no reason to embrace the status quo that often avoids causal language (Grosz et al., 2020)—assumptions do not disappear just because we ignore them. [@Deffner_et_al_2022, pp. 14] -->

<!-- Causal inference frameworks do not magically guarantee value-free answers, but they force us to be precise about the questions we ask, and to be transparent about the assumptions that we are willing to make [@Rohrer_et_al_2022, pp. 4-5] -->

<!-- Specifically, a framework is needed which lets us reason about the causes of non-invariance. [@Sterner_et_al_2024, pp. 751] -->

<!-- 1. Challenges for the approach: background knowledge and research context -->
Conversely, this theoretical commitment to causal inference also introduces several practical challenges that applied CJ researchers must navigate. These fall into two main categories: first, acquiring the foundational knowledge necessary to apply the approach effectively; and second, dedicating greater attention to both conceptual and statistical modeling.

### Required foundational knowledge {#sec-discussion_challenges1}
<!-- recording: Sven and Tine 25.06.04; time: 00:24:30 - 00:25:15 -->
<!-- recording: Sven and Tine 25.06.04; time: 00:26:05 - 00:26:30 -->

To apply our approach effectively, CJ researchers must tackle two main challenges related to the required foundational knowledge. First, they need a solid understanding of causal inference principles. Second, they must learn to translate the functional and probabilistic content of conceptual models into bespoke statistical models. One example that illustrates the importance of a sound knowledge of causal inference is the recurrent assumption that predictor variables are "relevant" to the research context--interpreting this relevance as their inclusion in a *sufficient adjustment set* [@Pearl_2009; @Pearl_et_al_2016; @Morgan_et_al_2014]. However, we do not fully explore what this assumption entails or its implications for model specification, estimation, and inference. As a result, CJ researchers must deeply engage with these and other complex ideas, including how SCMs encode functional and probabilistic information. To assist this effort, this study includes key references throughout the study to guide motivated aaplied CJ researchers toward a deeper understanding of these concepts [^10].

[^10]: refer to footnote 1, 2, and the detailed online document referenced in the Declarations section

Developing the skills to translate conceptual models into bespoke statistical models also presents challenges. Although Bayesian inference methods offer a more accessible path for many applied CJ researchers to develop these skills--by reducing the need for specialized knowledge in areas like optimization theory, which frequentist methods often require--they still demand a working knowledge of other technical concepts such as probabilistic programming languages (PPLs), probability distributions, and convergence. These requirements can be non-trivial to master. To support CJ researchers in developing these skills, this study provides a link to the statistical model code and alternative model specifications in the **Declarations** section of this document [^11].

[^11]: Seminal texts on Bayesian inference methods, such as @Gelman_et_al_2014 and @McElreath_2020, offer valuable support for developing a deeper understanding of these models.

### Attention to conceptual and statistical modeling {#sec-discussion_challenges2}
<!-- recording: Sven and Tine 25.06.04; time: 00:26:30 - 00:33:55 -->

Even after acquiring the necessary foundational knowledge, CJ researchers face two additional challenges when applying our approach to their specific research context. First, they must verify whether the conceptual model provides a *faithful* [^12] representation of the CJ system under study. Second, they need to assess whether the statistical translation of the model can accurately estimate the intended estimands (i.e., parameters) from empirical data. To ensure a faithful representation of the CJ system, we encourage researchers to treat our conceptual and statistical models as starting points rather than universal solutions for all CJ designs or datasets. While these models may prove adequate in some contexts, researchers cannot assume this apriori. Instead, they need to adapt these models to their specific CJ contexts, paying close attention to assessment design features and causal assumptions. As discussed in @sec-discussion_challenges, our approach facilitates this process by offering a transparent framework for articulating new assumptions and guiding the design of CJ assessments.

[^12]: Avoid confusing this term with the *faithfulness assumption* [@Neal_2020; @Hernan_et_al_2025] described in the causal inference literature. For further details, see footnote 1.

<!-- Our DAG is, of course, incomplete and possibly wrong, in particular when it comes to the nodes that have not been experimentally manipulated. But an incomplete model is still an improvement over no model at all. [@Deffner_et_al_2022, pp. 5] -->

<!-- A causal framework is not only useful for analysis but also aids research design [@Deffner_et_al_2022, pp. 11] -->

<!-- To use graphs to reason about confounding control and other forms of bias requires that the representation on the graph be such that there is: -->
<!-- 1. No ungraphed confounding: Any variable that affects two or more variables in the graph via distinct direct effects must be shown on the graph as an ancestor of the affected variables, even if it is unmeasured. -->
<!-- 2. No ungraphed collider stratification: Any variable that affects selection or is used for stratification and is affected by two or more variables in the graph (including its random disturbance) must be shown on the graph as a descendant of the variables affecting it, even if it is unmeasured. -->
<!-- As may become clear from the descriptions below, these requirements are needed to ensure that any association between two variables in the graph can be atributed solely to open paths between the variables within the graph, and thus can be traced to either variation in graphed causal variables, or conditioning on effects of such variables, or both. Conditions 1 and 2 enforce the connectivity requirement of DAGs by demanding the graph we draw show all variables which can causally explain associations (either as uncontrolled confounders or as controlled colliders), even if those variables are unmeasured or hypothetical. [@Lash_et_al_2021, pp. 113] -->

<!-- If we are willing to assume that the depicted DAG is a causal DAG—which means that it includes all common causes of any pair of variables (Elwert, 2013)—we can algorithmically derive which variables need to be “conditioned” (see Box 1) on to identify the causal effect of interest. [@Deffner_et_al_2022, pp. 4] -->

<!-- DAGs are heuristic—they are not detailed statistical models. But they allow us to deduce which statistical models can provide valid causal inferences, assuming the DAG is true. [@McElreath_2020, pp. 17] -->

<!-- DAGs are not about adding assumptions—they are about revealing the assumptions that are otherwise made implicitly [@Sterner_et_al_2024, pp. 755; @Deffner_et_al_2022; @Pearl_et_al_2014] -->

<!-- It has been argued that causal DAGs can be considered graphical representations of nonparametric SEMs [@Pearl_2012; @Kunicki_et_al_2023, pp. 4; @Sterner_et_al_2024, pp. 749]. -->

<!-- DAG can be thought of as a “qualitative schematic” for a particular class of structural equation modeling (i.e., linear structural equation modeling), or a nonparametric SEM. Conversely, structural equation modeling may be thought of as an algebraic system for a causal DAG [@VanderWeele_et_al_2021; Kunicki_et_al_2023, pp. 12]. -->

Conversely, to evaluate the estimation capabilities of the statistical model, researchers must tackle the challenge of performing identification analysis. As outlined in @sec-theoretical, *identification analysis* determines whether a statistical model can accurately compute a given estimand (e.g., a parameter) based solely on its (causal) assumptions, independent of random variability [@Schuessler_et_al_2023]. Identification is crucial because it is a necessary condition for consistency. *Consistency* is the property of an estimator (e.g., a statistical model) whose estimates converge to the "true" value of an estimand as data size approaches infinity [@Everitt_et_al_2010]. Without identification, consistency is impossible--even with infinite, error-free data--and thus, meaningful inference from finite samples cannot be achieved [@Schuessler_et_al_2023].

Although performing identification analysis through formal derivations may seem like a natural next step, we argue that the complexity of the CJ system renders such an approach impractical at the outset. Instead, simulation-based methods, such as power analysis, provide a more practical and flexible alternative, enabling researchers to examine the consistency of estimates without relying on cumbersome mathematical proofs. Nonetheless, the development of formal identification proofs remains a significant goal for future CJ research. Importantly, our approach supports both strategies by providing the probabilistic foundation for formal derivations and the necessary statistical structure for simulation-based methods.

<!-- Consistency: An estimator is said to be consistent if it converges to its estimand as the sample size tends to infinity. [@Everitt_et_al_2010, pp. 100] -->

<!-- The main reason to focus on identification is that if these parameters cannot be identified using "infinite" and error free data, we certainly cannot learn anything about it using finite data. So, for the time being, we assume large samples and error-free measurements [@Schuessler_et_al_2023, pp. 5]. -->

<!-- we not only underline the value of causal graphs as an intuitive and powerful tool for survey researchers to communicate their assumptions, be it for inferences on descriptive or causal parameters, but also highlight their value in uncovering nontrivial insights for improving more complicated statistical practices. [@Schuessler_et_al_2023, pp. 3] -->

<!-- (do not do this yet!) Develop how you multiply the predictions by the inverse of the sample probability to get the estimates. Should I use the probability itself?: Remember that in a random sample the sample probability is n/N, so the the inverse of the sample probability is N/n. So, multiplying the estimate by the n * N/n gives you the prediction for the population?, yes. -->
<!-- (See https://timss2023.org/methods/, and pdf in mail, and https://discourse.mc-stan.org/t/survey-weighted-regression/1654 ) -->

<!-- (do not do this yet!) Then to get the differences between groups you need to aggregate the latent variable of interest using the variable of interest without considering any other variable (marginalization) (see https://www.r-bloggers.com/2017/06/sampling-weights-and-multilevel-modeling-in-r/ ) -->

<!-- Add to future research how would be to use the model when you do not know the weights of the sample, but you can use external data to approximate them. -->

<!-- Develop how you get the vectors S and C and sampling weights when you know the extension of the population. (See https://mc-stan.org/docs/stan-users-guide/poststratification.html ) -->

{{< pagebreak >}}

# Declarations {.unnumbered .appendix appendix-style=plain}

**Funding:** The Research Fund (BOF) of the University of Antwerp funded this project.

**Financial interests:** The authors declare no relevant financial interests.

**Non-financial interests:** The authors declare no relevant non-financial interests.

**Ethics approval:** The University of Antwerp Research Ethics Committee confirmed that this study does not require ethical approval.

**Consent to participate:** Not applicable

**Consent for publication:** All authors have read and approved the final version of the manuscript for publication.

**Data availability:** This study did not use any data.

**Materials and code availability:** The `CODE LINK` section at the top of the digital document located at: [https://jriveraespejo.github.io/paper2_manuscript/](https://jriveraespejo.github.io/paper2_manuscript/) provides access to all materials and code.

**AI-assisted technologies in the writing process:** The authors used various AI-based language tools to refine phrasing, optimize wording, and enhance clarity and coherence throughout the manuscript. They take full responsibility for the final content of the publication.

**CRediT authorship contribution statement:** *Conceptualization:* J.M.R.E, T.vD., S.DM., and S.G.; *Methodology:* J.M.R.E, T.vD., and S.DM.; *Software:* J.M.R.E.; *Validation:* J.M.R.E.; *Formal Analysis:* J.M.R.E.; *Investigation:* J.M.R.E; *Resources:* T.vD. and S.DM.; *Data curation:* J.M.R.E.; *Writing - original draft:* J.M.R.E.; *Writing - review and editing:* T.vD., S.DM., and S.G.; *Visualization:* J.M.R.E.; *Supervision:* S.G. and S.DM.; *Project administration:* S.G. and S.DM.; *Funding acquisition:* S.G. and S.DM.

<!-- **Acknowledgements:** -->

{{< pagebreak >}}

# Appendix {#sec-appendix .appendix appendix-style=plain}

## Statistical and Causal inference {#sec-appendixB .appendix appendix-style=plain}

<!-- recording: Sven and Tine 25.01.22; time: 00:00:00 - 00:05:40 -->

This section introduces fundamental statistical and causal inference concepts necessary for understanding the core theoretical principles described in this document. It does not, however, offer a comprehensive overview of causal inference methods. Readers seeking more in-depth understanding may wish to explore introductory papers such as @Pearl_2010, @Rohrer_2018, @Pearl_2019, and @Cinelli_et_al_2020. They may also find it helpful to consult introductory books like @Pearl_et_al_2018, @Neal_2020, and @McElreath_2020. For more advanced study, readers may refer to seminal intermediate papers such as @Neyman_et_al_1923, @Rubin_1974, @Spirtes_et_al_1991, and @Sekhon_2009, as well as books such as @Pearl_2009, @Morgan_et_al_2014, and @Hernan_et_al_2025.

### Empirical research and randomized experiments {#sec-appendixB1 .appendix appendix-style=plain}

<!-- 1. What does empirical research do first? -->
Empirical research uses evidence from observation and experimentation to address real-world challenges. In this context, researchers typically formulate their research questions as *estimands* or *targets of inference*, i.e., the specific quantities they seek to determine [@Everitt_et_al_2010]. For instance, researchers might be interested in answering the following question: "To what extent do different teaching methods $(T)$ influence students' ability to produce high-quality written texts $(Y)$?" To investigate this, researchers could randomly assign students to two groups, each exposed to a different teaching method $(T_{i} = \{1,2\})$. Then, they would perform pairwise comparisons, generating a dichotomous outcome $(Y_{i} = \{0,1\})$ showing which student exhibits more of the ability. In this scenario, the research question can be rephrased as the estimand, "*On average*, is there a difference in the ability to produce high-quality written texts between the two groups of students?" and this estimand can be mathematically represented by the random associational quantity in @eq-group_diff, where $E[\cdot]$ denotes the expected value.

$$
E[Y_{i} \mid T_{i}=1] - E[Y_{i} \mid T_{i}=2]
$$ {#eq-group_diff}

<!-- 2. What do they do next? -->
Researchers then proceed to identify the estimands. *Identification* determines whether an estimator can accurately compute the estimand based solely on its assumptions, regardless of random variability [@Schuessler_et_al_2023, pp. 4]. An *estimator* refers to a method or function that transforms data into an estimate [@Neal_2020]. *Estimates* are numerical values that approximate the estimand derived through the process of *estimation*, which integrates data with an estimator [@Everitt_et_al_2010]. The Identification-Estimation flowchart [@McElreath_2020; @Neal_2020], shown in @fig-IEflow, visually represents the transition from estimands to estimates.

<!-- The causal assumptions encoded in a DAG constitute the theoretical model on which identification analysis rests. Identification analysis is concerned with determining whether or not the true value of a parameter can be estimated using the assumptions in the model (graph), regardless of random variability in the data due to small samples [@Schuessler_et_al_2023, pp. 4]. -->

![ Identification-Estimation flowchart. Extracted and slightly modified from @Neal_2020 [pp. 32] ](images/png/IEflow.png){#fig-IEflow fig-align="center" width=35%}

<!-- 3. the importance of identification -->
Identification is a necessary condition to ensure *consistent* estimators. An estimator achieves *consistency* when it converges to the "true" value of an estimand as the data size approaches infinity [@Everitt_et_al_2010]. Without identification, researchers cannot achieve consistency, even with "infinite" and error-free data. As a result, deriving meaningful insights about an estimand from finite data becomes impossible [@Schuessler_et_al_2023, pp. 5]. Therefore, to ensure accurate and reliable estimates, researchers prioritize estimators with desirable identification properties. For instance, the Z-test is a widely used estimator for comparing group proportions, yielding accurate estimates when its underlying assumptions are satisfied [@Kanji_2006]. Furthermore, researchers can interpret estimates from the Z-test as causal, provided the data is collected through a randomized experiment.

<!-- The main reason to focus on identification is that if these parameters cannot be identified using "infinite" and error free data, we certainly cannot learn anything about it using finite data. So, for the time being, we assume large samples and error-free measurements [@Schuessler_et_al_2023, pp. 5]. -->

<!-- If both assumptions are met, the Z-test is expressed as a signal-to-noise statistic $Z = (\hat{p}_{1} - \hat{p}_{2})/ \hat{s}_{p}$. The signal is defined as the difference between the groups' sample proportions, $\hat{p}_{1} = \sum_{i=1}^{n_{1}}{Y_{i}/n_{1}}$ and $\hat{p}_{2} = \sum_{i=1}^{n_{2}}{Y_{i}/n_{2}}$, analogous to $E[Y_{i} \mid T_{i}=1]$ and $E[Y_{i} \mid T_{i}=2]$, respectively. The noise, represented by $\hat{s}_{p}$, is defined as the unpooled sample variability observed between the two groups.  -->

<!-- the BTL model [@Bradley_et_al_1952; @Luce_1959] is an estimator known for its effectiveness in modeling pairwise comparisons, yielding accurate estimates for the comparison of two groups, as long as the assumptions of the model are met. Usually these estimates are expressed in terms of the proportions of groups  -->

<!-- However, researchers often seek to uncover the mechanisms underlying specific data and establish causal relationships rather than simply estimate associations. -->

<!-- 4. Why does causality stem from randomize experiments? -->
Randomized experiments are widely recognized as the gold standard in evidence-based science [@Hariton_et_al_2018; @Hansson_2014]. This recognition stems from their ability to enable researchers interpret associational estimates as causal. They achieve this by ensuring data, and by extension an estimator, satisfies several key identification properties, such as common support, no interference, and consistency [@Morgan_et_al_2014; @Neal_2020]. The most critical property, however, is the elimination of confounding. *Confounding* occurs when an external variable $X$ simultaneously influences the outcome $Y$ and the variable of interest $T$, resulting in spurious associations [@Everitt_et_al_2010]. Randomization addresses this issue by decoupling the association between the intervention allocation $T$ from any other variable $X$ [@Morgan_et_al_2014; @Neal_2020].

<!-- 5. Experiments are not available to everyone, then causal inference comes to the rescue -->
Nevertheless, researchers often face constraints that limit their ability to conduct randomized experiments. These constraints include ethical concerns, such as the assignment of individuals to potentially harmful interventions, and practical limitations, such as the infeasibility of, for example, assigning individuals to genetic modifications or physical impairments [@Neal_2020]. In these cases, causal inference offers a valuable alternative for generating causal estimates and understanding the mechanisms underlying specific data. In addition, the framework can provide significant theoretical insights that can enhance the design of experimental and observational studies [@McElreath_2020].

### Identification under causal inference {#sec-appendixB2 .appendix appendix-style=plain}

<!-- 1. What is causal inference? -->
Unlike classical statistical modeling, which focuses primarily on summarizing data and inferring associations, the *causal inference* framework is designed to identify causes and estimate their effects using data [@Shaughnessy_et_al_2010; @Neal_2020]. The framework uses rigorous mathematical techniques to address the *fundamental problem of causality* [@Pearl_2009; @Pearl_et_al_2016; @Morgan_et_al_2014]. This problem revolves around the question, "What would have happened 'in the world' under different circumstances?" This question introduces the concept of counterfactuals, which are instrumental in defining and identifying causal effects.

<!-- 2. What are counterfactuals? -->
*Counterfactuals* are hypothetical scenarios that are *contrary to fact*, where alternative outcomes resulting from a given cause are neither observed nor observable [@Neal_2020; @Counterfactual_2024]. The structural approach to causal inference [@Pearl_2009; @Pearl_et_al_2016] provides a formal framework for defining counterfactuals. For instance, in the scenario described in @sec-appendixB1, the approach begins by defining the *individual causal effect* (ICE) as the difference between each student's potential outcomes, as in @eq-ICE.

$$
\tau_{i} = Y_{i} \mid do(T_{i}=1) - Y_{i} \mid do(T_{i}=2)
$$ {#eq-ICE}

where $do(T_{i}=t)$ represents the intervention operator, $Y_{i} \mid do(T_{i}=1)$ represents the potential outcome under intervention $T_{i}=1$, and $Y_{i} \mid do(T_{i}=1)$ represents the potential outcome under intervention $T_{i}=2$. Here, an *intervention* involves assigning a constant value to the treatment variable for each student's potential outcomes. Note that if a student is assigned to intervention $T_{i}=1$, the potential outcome under $T_{i}=2$ becomes a counterfactual, as it is no longer observed nor observable. To address this challenge, the structural approach extends the ICE to the *average causal effect* (ACE, @eq-ACE), representing the average difference between the students' observed potential outcomes and their counterfactual counterparts.

$$
\begin{aligned}
\tau & = E[\tau_{i}] \\
  & = E[Y_{i} \mid do(T_{i}=1)]- E[Y_{i} \mid do(T_{i}=2)]
\end{aligned}
$$ {#eq-ACE}

<!-- [^1]: The potential outcomes approach [@Neyman_et_al_1923; @Rubin_1974] defines the ICE as $\tau_{i} = Y^{1}_{i} - Y^{2}_{i}$, while it defines the ACE as $\tau = E[\tau_{i}] = E[Y^{1}_{i}]- E[Y^{2}_{i}]$. -->

<!-- 3. But how does it work? -->
Even though counterfactuals are unobservable, researchers can still identify the ACE from associational estimates by leveraging the structural approach. The approach identifies the ACE by statistically conditioning data on a *sufficient adjustment set* of variables $X$ [@Pearl_2009; @Pearl_et_al_2016; @Morgan_et_al_2014]. This *sufficient* set (potentially empty) must block all non-causal paths between $T$ to $Y$ without opening new ones. When such a set exists, then $T$ and $Y$ are *d-separated* by $X$ ($T \dsep Y \mid X$) [@Pearl_2009], and $X$ satisfies the *backdoor criterion* [@Neal_2020, pp 37]. Here, *conditioning* describes the process of restricting the focus to the subset of the population defined by the conditioning variable [@Neal_2020, pp. 32] (see @eq-CACE).

<!-- Conditioning on T=t just means that we are restricting our focus to the subset of the population to those who received treatment t. In contrast, an intervention would be to take the whole population and give everyone treatment t [@Neal_2020, pp. 32]. -->

Conditioning on a sufficient adjustment set enables researchers to estimate the ACE, even when the data comes from an observational study. This process is feasible because such conditioning ensures that the ACE estimator satisfies several critical properties, including confounding elimination [@Morgan_et_al_2014]. Naturally, the validity of claims about the causal effects of $T$ on $Y$ now hinges on the assumption that $X$ serves as a sufficient adjustment set. However, as @Kohler_et_al_2019 [pp. 150] noted, drawing conclusions about the real world from observed data inevitably requires assumptions. This requirement holds true for both observational and experimental data.

<!-- Statements about the real world from observed data require assumptions. This is true no matter what kind of data we have, and it is explicitly also true for statements based on data stemming from SRSg. The validity of our statements therefore critically hinges on the validity of those assumptions. If our statistical assumptions are wrong for the analysis at hand, the inferential statements will be incorrect no matter how sophisticated the statistical methods are. [@Kohler_et_al_2019, pp. 150] -->

For instance, if researchers cannot conduct the randomized experiments described in @sec-appendixB1 and must instead rely on observational data, they can still identify the ACE as long as an observed variable $X$, such as the socio-economic status of the school, satisfies the backdoor criterion. Under these circumstances, researchers first identify the *conditional average causal effect* (CACE, @eq-CACE)

$$
CACE_{t} = E[Y_{i} \mid T_{i}=t, X]
$$ {#eq-CACE}

From the CACE, researchers can identify the ACE from associational quantities as in @eq-mACE1. This identification process is commonly known as the *backdoor adjustment*. Here, $E_{X}[\cdot]$ represents the marginal expected value over $X$ [@Morgan_et_al_2014].

$$
\begin{aligned}
  \tau & = E[Y_{i} \mid do(T_{i}=1)]- E[Y_{i} \mid do(T_{i}=2)] \\
  & = E_{X}[CACE_{1} - CACE_{2}] \\
  & = E_{X}\left[ E[Y_{i} \mid T_{i}=1, X] - E[Y_{i} \mid T_{i}=2, X] \right]
\end{aligned}
$$ {#eq-mACE1}

Notably, the approach extends the ACE identification for a continuous variable $T$ as in @eq-mACE_cont, ensuring broad applicability across different causal scenarios [@Neal_2020, pp. 45]

$$
\begin{aligned}
  \tau &= E[Y_{i} \mid do(T_{i}=t)] \\
  & = d E_{X}\left[ E[Y_{i} \mid T_{i}=t, X]\right]/ dt
  \end{aligned}
$$ {#eq-mACE_cont}

### Diving into the specifics {#sec-appendixB3 .appendix appendix-style=plain}

<!-- 1. The benefits of SCMs and DAGs -->
The structural approach to causal inference uses SCMs and DAGs to formally and graphically represent the presumed causal structure underlying the ACE [@Pearl_2009; @Pearl_et_al_2016; @Gross_et_al_2018; @Neal_2020]. Essentially, these tools serve as *conceptual (theoretical) models* on which identification analysis rests [@Schuessler_et_al_2023, pp. 4]. Thus, using these tools, researchers can determine which statistical models can identify (ACE, CACE, or other), assuming the depicted causal structure is correct [@McElreath_2020], thus enabling valid causal inference. @fig-IEflow shows the role of theoretical models in the inference process.

<!-- The causal assumptions encoded in a DAG constitute the theoretical model on which identification analysis rests. Identification analysis is concerned with determining whether or not the true value of a parameter can be estimated using the assumptions in the model (graph), regardless of random variability in the data due to small samples [@Schuessler_et_al_2023, pp. 4]. -->

<!-- Causal diagrams are a simple way to encode our subject-matter knowledge, and our assumptions, about the qualitative causal structure of a problem. But, as described in the next sections, causal diagrams also encode information about potential associations between the variables in the causal network. It is precisely this simultaneous representation of association and causation that makes causal diagrams such an attractive tool. [@Hernan_et_al_2025, pp. 75] -->

SCMs and DAGs support identification analysis through two key advantages. First, regardless of complexity, they can represent various causal structures using only five fundamental building blocks [@Neal_2020; @McElreath_2020]. This feature allows researchers to decompose complex structures into manageable components, facilitating their analysis [@McElreath_2020]. Second, they depict causal relationships in a non-parametric and fully interactive way. This flexibility enables feasible ACE identification strategies without defining the variables' data types, the functional form between them, or their parameters [@Pearl_et_al_2016, pp. 35].

Thus, @sec-appendixB31 and @sec-appendixB32 elaborate on the first advantage, while @sec-appendixB32 and @sec-appendixB33 do so for the second. Finally, @sec-appendixB34 explains how researchers use SCMs and DAGs alongside Bayesian inference methods in the estimation process.

<!-- A heuristic is a strategy that simplifies information to make decisions more quickly, efficiently, and sometimes more accurately than complex methods [@Chow_2015]. -->

#### The five fundamental block for SCMs and DAGs {#sec-appendixB31 .appendix appendix-style=plain}

<!-- 2. the SCMs -->
Figures [-@fig-dags_scms1], [-@fig-dags_scms2], [-@fig-dags_scms3], [-@fig-dags_scms4], and [-@fig-dags_scms5] display the five fundamental building blocks for SCMs and DAGs. The left panels of the figures show the formal mathematical models, represented by the SCMs, defined in terms of a set of *endogenous* variables $V=\{X_{1},X_{2},X_{3}\}$, a set of *exogenous* variables $E=\{e_{X1},e_{X2},e_{X3}\}$, and a set of functions $F=\{f_{X1},f_{X2},f_{X3}\}$ [@Pearl_2009; @Cinelli_et_al_2020]. Endogenous variables are those whose causal mechanisms a researcher chooses to model [@Neal_2020]. In contrast, exogenous variables represent *errors* or *disturbances* arising from omitted factors that the investigator chooses not to model explicitly [@Pearl_2009, pp. 27,68]. Lastly, the functions, referred to as *structural equations*, express the endogenous variables as non-parametric functions of other variables. These functions use the symbol '$:=$' to denote the asymmetrical causal dependence of the variables and the symbol '$\dsep$' to represent *d-separation*, a concept akin to (conditional) independence.

<!-- 3. The DAGs -->
Notably, every SCM has an associated DAG [@Pearl_et_al_2016; @Cinelli_et_al_2020]. The right panels of the figures display these DAGs. A DAG is a graph consisting of nodes connected by edges, where the nodes represent random variables. The term *directed* means that the edges extend from one node to another, with arrows indicating the direction of causal influence. The term *acyclic* implies that the causal influences do not form loops, ensuring the influences do not cycle back on themselves [@McElreath_2020]. DAGs represent observed variables as solid black circles, while they use open circles for unobserved (latent) variables [@Morgan_et_al_2014]. Although the *standard representation* of DAGs typically omits exogenous variables for simplicity, the *magnified representation* depicted in the figures offers one key advantage: including exogenous variables can help researchers highlight potential issues related to conditioning and confounding [@Cinelli_et_al_2020].

::: {#fig-dags_scms1 layout-ncol=2}

::: {#fig-scm_bb1 fig-align="center" fig-pos="H"}
$$
\begin{aligned}
  X_{1} & := f_{X1}(e_{X1}) \\
  X_{3} & := f_{X3}(e_{X3}) \\
  e_{X1} & \dsep e_{X3}
\end{aligned}
$$

SCM
:::

![DAG](images/png/mdag_bb1.png){#fig-mdag_bb1 fig-align="center" fig-pos="H" width=54%}

Two unconnected nodes
:::

::: {#fig-dags_scms2 layout-ncol=2}

::: {#fig-scm_bb2 fig-align="center" fig-pos="H"}
$$
\begin{aligned}
  X_{1} & := f_{X1}(e_{X1}) \\
  X_{3} & := f_{X3}(X_{1},e_{X3}) \\
  e_{X1} & \dsep e_{X3}
\end{aligned}
$$

SCM
:::

![DAG](images/png/mdag_bb2.png){#fig-mdag_bb2 fig-align="center" fig-pos="H" width=54%}

Two connected nodes or descendant
:::

::: {#fig-dags_scms3 layout-ncol=2}

::: {#fig-scm_bb3 fig-align="center" fig-pos="H"}
$$
\begin{aligned}
  X_{1} & := f_{X1}(e_{X1}) \\
  X_{2} & := f_{X2}(X_{1},e_{X2}) \\
  X_{3} & := f_{X3}(X_{2},e_{X3}) \\
  e_{X1} & \dsep e_{X2} \\
  e_{X1} & \dsep e_{X3} \\
  e_{X2} & \dsep e_{X3}
\end{aligned}
$$

SCM
:::

![DAG](images/png/mdag_bb3.png){#fig-mdag_bb3 fig-align="center" fig-pos="H" width=65%}

Chain or mediator
:::

::: {#fig-dags_scms4 layout-ncol=2}

::: {#fig-scm_bb4 fig-align="center" fig-pos="H"}
$$
\begin{aligned}
  X_{1} & := f_{X1}(X_{2},e_{X1}) \\
  X_{2} & := f_{X2}(e_{X2}) \\
  X_{3} & := f_{X3}(X_{2},e_{X3}) \\
  e_{X1} & \dsep e_{X2} \\
  e_{X1} & \dsep e_{X3} \\
  e_{X2} & \dsep e_{X3}
\end{aligned}
$$

SCM
:::

![DAG](images/png/mdag_bb4.png){#fig-mdag_bb4 fig-align="center" fig-pos="H" width=65%}

Fork or confounder
:::

::: {#fig-dags_scms5 layout-ncol=2}

::: {#fig-scm_bb5 fig-align="center" fig-pos="H"}
$$
\begin{aligned}
  X_{1} & := f_{X1}(e_{X1}) \\
  X_{2} & := f_{X2}(X_{1},X_{3},e_{X2}) \\
  X_{3} & := f_{X3}(e_{X3}) \\
  e_{X1} & \dsep e_{X2} \\
  e_{X1} & \dsep e_{X3} \\
  e_{X2} & \dsep e_{X3}
\end{aligned}
$$

SCM
:::

![DAG](images/png/mdag_bb5.png){#fig-mdag_bb5 fig-align="center" fig-pos="H" width=65%}

Collider or inmorality
:::

<!-- 4. Theoretical implications of the basic building block of the SCMS and DAGs -->
A careful examination of these building blocks highlights the theoretical assumptions underlying their observed variables. SCM [-@fig-scm_bb1] and DAG [-@fig-mdag_bb1] depict two unconnected nodes, representing a scenario where variables $X_{1}$ and $X_{3}$ are independent or not causally related. SCM [-@fig-scm_bb2] and DAG [-@fig-mdag_bb2] illustrate two connected nodes, representing a scenario where a *parent* node $X_{1}$ exerts a causal influence on a *child* node $X_{3}$. In this setup, $X_{3}$ is considered a *descendant* of $X_{1}$. Additionally, $X_{1}$ and $X_{3}$ are described as *adjacent* because there is a *direct path* connecting them. SCM [-@fig-scm_bb3] and DAG [-@fig-mdag_bb3] depict a *chain*, where $X_{1}$ influences $X_{2}$, and $X_{2}$ influences $X_{3}$. In this configuration, $X_{1}$ is a parent node of $X_{2}$, which is a parent node of $X_{3}$. This structure creates a *directed path* between $X_{1}$ and $X_{3}$. Consequently, $X_{1}$ is an *ancestor* of $X_{3}$, and $X_{2}$ fully *mediates* the relationship between the two. SCM [-@fig-scm_bb4] and DAG [-@fig-mdag_bb4] illustrate a *fork*, where variables $X_{1}$ and $X_{3}$ are both influenced by $X_{2}$. Here, $X_{2}$ is a parent node that *confounds* the relationship between $X_{1}$ and $X_{3}$. Finally, SCM [-@fig-scm_bb5] and DAG [-@fig-mdag_bb5] show a *collider*, where variables $X_{1}$ and $X_{3}$ are concurrent causes of $X_{2}$. In this configuration, $X_{1}$ and $X_{3}$ are not causally related to each other but both influence $X_{2}$ (an *inmorality*). Notably, all building blocks assume the errors are independent of each other and from all other variables in the graph, as evidenced by the pairwise relations $e_{X1} \dsep e_{X2}$, $e_{X1} \dsep e_{X3}$, and $e_{X2} \dsep e_{X3}$.

<!-- 5. Continuing with the motivating example -->
Researchers can then use these building blocks to represent the scenario outlined in @sec-appendixB2. SCM [-@fig-scm_example1] and DAG [-@fig-mdag_example1] depict the plausible causal structure for this example. In this context, the variable $X$ (socio-economic status of the school) is thought to be a confounder in the relationship between the teaching method $T$ and the outcome $Y$. The figures display multiple descendant relationships such as $X \rightarrow T$, $X \rightarrow Y$, and $T \rightarrow Y$. They also highlight unconnected node pairs, evident from the relationships $e_{T} \dsep e_{X}$, $e_{T} \dsep e_{Y}$, and $e_{X} \dsep e_{Y}$. Additional, the figures show one fork, $X \rightarrow \{T, Y\}$, and two colliders: $\{X, e_{T}\} \rightarrow T$ and $\{X, T, e_{Y}\} \rightarrow Y$.

::: {#fig-example1 layout-ncol=2}

::: {#fig-scm_example1 fig-align="center"}
$$
\begin{aligned}
  X & := f_{X}(e_{X}) \\
  T & := f_{T}(X,e_{T}) \\
  Y & := f_{Y}(T,X,e_{Y}) \\
  e_{T} & \dsep e_{X} \\
  e_{T} & \dsep e_{Y} \\
  e_{X} & \dsep e_{Y}
\end{aligned}
$$

SCM
:::

![DAG](images/png/mdag_example1.png){#fig-mdag_example1 fig-align="center" width=65%}

Plausible causal structure the scenario outlined in @sec-appendixB2.
:::

#### The probabilistic implications of these blocks {#sec-appendixB32 .appendix appendix-style=plain}

<!-- 1. How the SCMS and DAGs encode probabilistic information? -->
Beyond their graphical capabilities, researchers can translate SCMs and DAGs into statistical models because these conceptual models encode functional and probabilistic information, which they can then replace with suitable functions and probabilistic assumptions [@Pearl_et_al_2016]. Notably, this encoding relies on two fundamental assumptions: the causal Markov and the faithfulness assumption. The *causal Markov assumption* links the graph structure to the joint probability distribution of variables by asserting that any variable in a graph is independent of all its non-descendants conditional on its parents [@Neal_2020, pp. 20; @Hernan_et_al_2025, pp. 80]. The *faithfulness assumption*, in turn, states that the observed statistical (in)dependencies in the data reflect the underlying causal structure among variables--except in rare degenerate cases [@Neal_2020, pp. 100; @Hernan_et_al_2025, pp. 81].

<!-- Causal DAGs are of no practical use unless we make an assumption linking the causal structure represented by the DAG to the data obtained in a study. This assumption, referred to as the causal Markov assumption, states that, conditional on its direct causes, a variable Vj is independent of any variable for which it is not a cause. That is, conditional on its parents, Vj is independent of its non-descendants; hence, a causal DAG is Markov with respect to the DAG G. [@Hernan_et_al_2025, pp. 74] -->

<!-- causal diagrams also encode information about potential associations between the variables in the causal network. It is precisely this simultaneous representation of association and causation that makes causal diagrams such an attractive tool. [@Hernan_et_al_2025, pp. 75] -->

<!-- This and the previous section review three structural reasons why two variables may be associated: one causes the other, they share common causes, or they share a common effect and the analysis is restricted to certain level of that common effect (or of its descendants). Along the way we introduced a number of graphical rules that can be applied to any causal diagram to determine whether two variables are (conditionally) independent. The arguments we used to support these graphical rules were heuristic and relied on our causal intuitions. These arguments, however, have been formalized and mathematically proven. See Fine Point 6.1 for a systematic summary of the graphical rules, and Fine Point 6.2 for an introduction to the concept of faithfulness. [@Hernan_et_al_2025, pp. 79] -->

<!-- (from probability to graph): The relationship between statistical independence and the purely graphical concept of d-separation relies on the causal Markov assumption (Technical Point 6.1): In a causal DAG, any variable is independent of its non-descendants conditional on its parents. Pearl (1988) proved the following fundamental theorem: The causal Markov assumption implies that, given any three disjoint sets A, B, C of variables, if A is d-separated from B conditional on C, then A is statistically independent of B given C. The assumption that the converse holds, i.e., that A is d-separated from B conditional on C if A is statistically independent of B given C, is a separate assumption—the faithfulness assumption described in Fine Point 6.2. [@Hernan_et_al_2025, pp. 80] -->

<!-- (from graph to probability): Formally, faithfulness is the assumption that, for three disjoint sets A, B, C on a causal DAG (where C may be the empty set), A independent of B given C implies A is d-separated from B given C. When, as in our example, the causal diagram makes us expect a non-null association that does not actually exist in the data, we say that the joint distribution of the data is not faithful to the causal DAG. In our example the unfaithfulness was the result of effect modification (by sex) with opposite effects of exactly equal magnitude in each half of the population. Such perfect cancellation of effects is rare, and thus we will assume faithfulness throughout this book. Because unfaithful distributions are rare, in practice lack of d-separation (See Fine Point 6.1) can be almost always equated to non-zero association. There are, however, instances in which faithfulness is violated by design. For example, consider the prospective study in Section 4.5. The average causal effect of A on Y was computed after matching on L. In the matched population, L and A are not associated because the distribution of L is the same in the treated and the untreated. That is, individuals are selected into the matched population because they have a particular combination of values of L and A. The causal diagram in Figure 6.9 represents the setting of a matched study in which selection S (1: yes, 0: no) is determined by both A and L. The box around S indicates that the analysis is restricted to those selected into the matched cohort (S = 1). According to d-separation rules, there are two open paths between A and L when conditioning on S: L -> A and L -> S <- A. Thus one would expect L and A to be associated conditionally on S. However, matching ensures that L and A are not associated (see Chapter 4). Why the discrepancy? Matching creates an association via the path L -> S <- A that is of equal magnitude, but opposite direction, as the association via the path L -> A. The net result is a perfect cancellation of the associations. Matching leads to unfaithfulness. [@Hernan_et_al_2025, pp. 81] -->

<!-- ![ The flow of association and causation in graphs. Extracted and slightly modified from @Neal_2020 [pp. 31] ](images/png/ACflow.png){#fig-ACflow fig-align="center" width=80%} -->

<!-- 2. What are the general implications of this? -->
A notable implication of the assumptions underlying the probabilistic encoding is that any conceptual model described by an SCM and DAG can represent the joint distribution of variables more efficiently [@Pearl_et_al_2016, pp. 29]. This expression takes the form of a product of conditional probability distributions (CPDs) of the type $P(child \mid parents)$. This property is formally known as the *Bayesian Network factorization* (BNF, @eq-net_factor) [@Pearl_et_al_2016, pp. 29; @Neal_2020, pp. 21]. In this expression, $pa(X_{i})$ denotes the set of variables that are the parents of $X_{i}$.
$$
\begin{aligned}
P(X_{1}, X_{2}, \dots, X_{P}) & = X_{1} \cdot \prod^{P}_{p=2} P(X_{i} \mid X_{i-1}, \dots, X_{1}) & (\small{\text{by chain rule})}\\
& = X_{1} \cdot \prod^{P}_{p=2} P(X_{i} \mid pa(X_{i}) ) & (\small{\text{by BNF}})
\end{aligned}
$$ {#eq-net_factor}

<!-- 3. What is the net result of this? -->
This encoding enables researchers with conceptual (theoretical) knowledge in the form of an SCM and DAG to predict patterns of (in)dependencies in the data. As highlighted by @Pearl_et_al_2016 [pp. 35], these predictions depend solely on the structure of these conceptual models without requiring the quantitative details of the equations or the distributions of the errors. Moreover, once researchers observe empirical data, the patterns of (in)dependencies in the data can provide significant insights into the validity of the proposed conceptual model.

<!-- The net result is that a researcher who has scientific knowledge in the form of a structural equation model is able to predict patterns of independencies in the data, based solely on the structure of the model’s graph, without relying on any quantitative information carried by the equations or by the distributions of the errors [@Pearl_et_al_2016, pp. 35]. -->

<!-- 4. can you illustrate this concept? -->
The five fundamental building blocks described in @sec-appendixB31 clearly illustrate which (in)dependencies can SMCs and DAGs predict. For instance, applying the BNF to the causal structure shown in the SCM [-@fig-scm_bb1] and DAG [-@fig-mdag_bb1] enables researchers to express the joint probability distribution of the observed variables as $P(X_{1}, X_{3}) = P(X_{1}) P(X_{3})$, supporting the theoretical assumption that the observed variables $X_{1}$ and $X_{3}$ are unconditionally independent ($X_{1} \dsep X_{3}$) [@Neal_2020, pp. 24]. Conversely, when $X_{3}$ is unconditionally dependent on $X_{1}$ ($X_{1} \ndsep X_{3}$), as depicted in the SCM [-@fig-scm_bb2] and DAG [-@fig-mdag_bb2], the BNF express their joint probability distribution as $P(X_{1}, X_{3}) = P(X_{3} \mid X_{1}) P(X_{1})$. Notably, these descriptions demonstrate the clear correspondence between the structural equations illustrated in @sec-appendixB31 and the CPDs.

Beyond the insights gained from two-node structures, researchers can uncover more nuanced patterns of(in)dependencies from chains, forks, and colliders. These (in)dependencies apply to any data set generated by a causal model with those structures, regardless of the specific functions attached to the SCM [@Pearl_et_al_2016, pp. 36]. For instance, applying the BNF to the chain structure depicted in the SCM [-@fig-scm_bb3] and DAG [-@fig-mdag_bb3] allow researchers to represent the joint distribution for the observed variables as $P(X_{1},X_{2},X_{3}) =$ $P(X_{1}) P(X_{2} \mid X_{1}) P(X_{3} \mid X_{2})$. This expression implies that $X_{1}$ and $X_{3}$ are unconditionally dependent $(X_{1} \ndsep X_{3})$, but conditionally independent when controlling for $X_{2}$ $(X_{1} \dsep X_{3} \mid X_{2})$. Moreover, in the fork structure shown in the SCM [-@fig-scm_bb4] and DAG [-@fig-mdag_bb4], researchers can express the joint distribution of the observed variables as $P(X_{1},X_{2},X_{3}) =$ $P(X_{1} \mid X_{2}) P(X_{2}) P(X_{3} \mid X_{2})$. Similar to the chain structure, this expression allows researchers to further infer that $X_{1}$ and $X_{3}$ are unconditionally dependent $(X_{1} \ndsep X_{3})$, but conditionally independent when controlling for $X_{2}$ $(X_{1} \dsep X_{3} \mid X_{2})$. Finally, researchers analyzing the collider structure illustrated in the SCM [-@fig-scm_bb5] and DAG [-@fig-mdag_bb5] can express the joint distribution of the observed variables as $P(X_{1},X_{2},X_{3}) =$ $P(X_{1}) P(X_{2} \mid X_{1}, X_{3}) P(X_{3})$. This representation allows researchers to infer that $X_{1}$ and $X_{3}$ are unconditionally independent $(X_{1} \dsep X_{3})$, but conditionally dependent when controlling for $X_{2}$ $(X_{1} \ndsep X_{3} \mid X_{2})$. The authors @Pearl_et_al_2016 [pp. 37, 40, 41] and @Neal_2020 [pp. 25–26] provide the mathematical proofs for these conclusions.

<!-- These (in)dependencies will be true of every data set generated by a causal model with that graphical structure, regardless of the specific functions attached to the SCM [@Pearl_et_al_2016, pp. 36]. -->

<!-- 5. Continuing with the motivating example -->
Using these additional probabilistic insights, researchers can revisit the scenario in @sec-appendixB2. In this context, applying the BNF to the SCM [-@fig-scm_example2] structure, enables the representation of the joint probability distribution of the observed variables as $P(Y, T, X) =$ $P(Y \mid T, X) P(T \mid X) P(X)$. From this expression, researchers can infer that the outcome $Y$ is unconditionally dependent on the teaching method $T$ $(Y \ndsep T)$. This dependency arises from two key structures: a direct causal path from the teaching method $T$ to the outcome $Y$, represented by the two-connected-nodes structure $T \rightarrow Y$ (black path in DAG [-@fig-mdag_example2]), and a confounding non-causal path from the teaching method $T$ to the outcome $Y$ through the socio-economic status of the school $X$, represented by the fork structure $T \leftarrow X \rightarrow Y$ (gray path in DAG [-@fig-mdag_example2]).

::: {#fig-example2 layout-ncol=2}

::: {#fig-scm_example2 fig-align="center"}
$$
\begin{aligned}
  X & := x \\
  T & := f_{T}(x,e_{T}) \\
  Y & := f_{Y}(T,x,e_{Y}) \\
  e_{T} & \dsep e_{X} \\
  e_{T} & \dsep e_{Y} \\
  e_{X} & \dsep e_{Y}
\end{aligned}
$$

SCM
:::

![Conditioned DAG](images/png/mdag_example1a.png){#fig-mdag_example2 fig-align="center" width=65%}

Plausible causal structure the scenario outlined in @sec-appendixB2.
:::

#### From probability to causality {#sec-appendixB33 .appendix appendix-style=plain}

<!-- 1. Transforming probabilistic association into causation?  -->
The structural approach to causal inference translates probabilistic insights into actionable strategies seeking to identify the ACE from associational quantities. The approach achieves this by relying on the *modularity assumption*, which posits that intervening on a node alters only the causal mechanism of that node, leaving others unchanged [@Neal_2020, pp. 34].

<!-- 2. additional necessary concepts  -->
The modularity assumption underpins the concepts of manipulated graphs and Truncated Factorization, which are essential for representing interventions $P(Y_{i} \mid do(T_{i}=t))$ within SCMs and DAGs. *Manipulated graphs* simulate physical interventions by removing specific edges from a DAG, while preserving the remaining structure unchanged [@Neal_2020, pp. 34]. In parallel, *Truncated Factorization* (TF) achieves a similar simulation by removing specific functions from the conceptual model and replacing them with constants, while keeping the rest of the structure unchanged [@Pearl_2010]. The probabilistic implications of this factorization are formalized in @eq-trunc_factor, where $S$ represents the subset of variables $X_{p}$ directly influenced by the intervention, while an example illustrating these concepts follows below.

$$
P(X_{1}, X_{2}, \dots, X_{P} \mid do(S)) =
\begin{cases}
  \prod P(X_{p} \mid pa(X_{p}) ) & \text{if} \: p \not\in S \\
  1 \quad & \text{otherwise}
\end{cases}
$$ {#eq-trunc_factor}

<!-- 3. How to use these concepts to finally define the ACE -->
Using the TF, researchers can define the *backdoor adjustment* to identify the ACE. This adjustment states that if a variable $X_{p} \in S$ serves as a *sufficient adjustment set* for the effect of $X_{a}$ on $X_{b}$, then the ACE can be identified using @eq-backdoor [@Pearl_2009; @Pearl_et_al_2016; @Morgan_et_al_2014]. The sufficient adjustment set (potentially empty) must block all non-causal paths between $X_{a}$ and $X_{b}$ without introducing new paths. If such a set exists, then $X_{a}$ and $X_{b}$ are *d-separated* by $X_{p}$ ($X_{a} \dsep X_{b} \mid X_{p}$) [@Pearl_2009], and $X_{p}$ satisfies the *backdoor criterion* [@Neal_2020, pp. 37].

$$
P(X_{a} \mid do(X_{b}=x)) = \sum_{Xp} P(X_{a} \mid X_{b}=x, X_{p}) P(X_{p})
$$ {#eq-backdoor}

Ultimately, the backdoor adjustment enables researchers to express the ACE as:

$$
\begin{aligned}
\tau & = E[X_{a} \mid do(X_{b}=1)]- E[X_{a} \mid do(X_{b}=2)] \\
  & = E_{Xp}\left[ E[X_{a} \mid do(X_{b}=1), X_{p}]- E[X_{a} \mid do(X_{b}=2), X_{p}] \right] \\
  & = \sum_{Xp} X_{a} \cdot P(X_{a} \mid X_{b}=1, X_{p}) \cdot P(X_{p}) - \sum_{Xp} X_{a} \cdot P(X_{a} \mid X_{b}=2, X_{p}) \cdot P(X_{p})
\end{aligned}
$$ {#eq-backdoor_adjustment}

<!-- 4. revisiting the motivating example -->
With these new insights, researchers revisiting the scenario in @sec-appendixB32 can infer that the socio-economic status of the school, $X$, satisfies the backdoor criterion, assuming the causal structure depicted by the SCM [-@fig-scm_example2] and DAG [-@fig-mdag_example2] is correct. This means that $X$ serves as a sufficient adjustment set, as it effectively blocks all confounding non-causal paths introduced by the fork structure. Nevertheless, since $Y$ remains dependent on $T$ even after conditioning $(Y \ndsep T \mid X)$, this dependency can only be attributed to the direct causal effect $T \rightarrow Y$. Notably, for the purpose of identification, the conditioned DAG [-@fig-mdag_example2] is equivalent to the manipulated DAG [-@fig-mdag_example3], because $X$ satisfies the backdoor criterion.

::: {#fig-example3 layout-ncol=2}

::: {#fig-scm_example3 fig-align="center"}
$$
\begin{aligned}
  X & := f_{X}(e_{X}) \\
  T & := t \\
  Y & := f_{Y}(t,X,e_{Y}) \\
  e_{T} & \dsep e_{X} \\
  e_{T} & \dsep e_{Y} \\
  e_{X} & \dsep e_{Y}
\end{aligned}
$$

SCM
:::

![Manipulated DAG](images/png/mdag_example1b.png){#fig-mdag_example3 fig-align="center" width=65%}

Plausible causal structure the scenario outlined in @sec-appendixB32.
:::

Researchers can then apply the *backdoor adjustment* to identify the ACE of $T$ on $Y$. They achieve this by first identifying the CACE of $T$ on $Y$ by conditioning on $X$, and then marginalizing this effect over $X$ to obtain the ACE. This process is expressed in @eq-mACE2 (see @sec-appendixB2).

$$
\begin{aligned}
  \tau & = E[Y_{i} \mid do(T_{i}=1)]- E[Y_{i} \mid do(T_{i}=2)] \\
  & = E_{X}\left[ E[Y_{i} \mid T_{i}=1, X] - E[Y_{i} \mid T_{i}=2, X] \right] \\
  & = \sum_{X} Y_{i} \cdot P( Y_{i} \mid T_{i}=1, X) \cdot P(X) - \sum_{X} Y_{i} \cdot P( Y_{i} \mid T_{i}=2, X) \cdot P(X)
\end{aligned}
$$ {#eq-mACE2}

#### The estimation process {#sec-appendixB34 .appendix appendix-style=plain}

<!-- 1. How do we use Bayesian inference? -->
Ultimately, researchers can use Bayesian inference methods to estimate the ACE. The approach begins by defining two probability distributions: the likelihood of the data, $P(X_{1}, X_{2}, \dots, X_{P} \mid \theta)$, and the prior distribution, $P(\theta)$ [@Everitt_et_al_2010], where $X_{P}$ represents a random variable, and $\theta$ represents a one-dimensional parameter space for simplicity. After observing empirical data, researchers can update the priors to posterior distributions using Bayes' rule in @eq-bayes_rule [@Jeffreys_1998]:

$$
P(\theta \mid X_{1}, X_{2}, \dots, X_{P}) = \frac{P(X_{1}, X_{2}, \dots, X_{P} \mid \theta) \cdot P(\theta)}{P(X_{1}, X_{2}, \dots, X_{P})}
$$ {#eq-bayes_rule}

Given that the denominator on the right-hand side of @eq-bayes_rule serves as a normalizing constant independent of the parameter $\theta$, researchers can simplify the posterior updating process into three steps. First, they integrate new empirical data through the likelihood. Second, they update the parameters' priors to a posterior distribution according to @eq-prop_rule. Ultimately, they normalize these results to obtain a valid probability distribution.

$$
P(\theta \mid X_{1}, X_{2}, \dots, X_{P}) \propto P(X_{1}, X_{2}, \dots, X_{P}\mid \theta) \cdot P(\theta)
$$ {#eq-prop_rule}

<!-- check: -->
<!-- https://en.wikipedia.org/wiki/Maximum_a_posteriori_estimation -->
<!-- https://en.wikipedia.org/wiki/Bayesian_inference -->

<!-- 2. But what is the problem? -->

Temporarily setting aside the definition of prior distributions $P(\theta)$, note that the posterior updating process depends heavily on the assumptions underlying the likelihood of the data. However, as the number of random variables, $P$, increases, this joint distribution quickly becomes intractable [@Neal_2020]. This intractability is evident from @eq-like_chain, where the likelihood distribution is expressed by multiple chained CPDs.

$$
P(X_{1}, X_{2}, \dots, X_{P} \mid \theta) = P(X_{1} \mid \theta) \prod^{P}_{p=2} P(X_{i} \mid X_{i-1}, \dots, X_{1}, \theta )
$$ {#eq-like_chain}

<!-- 3. How do DAGs can help? -->
Nevertheless, researchers can manage the complexity of the likelihood by assuming specific local (in)dependencies among variables. SCMs and DAGs provide a formal framework to represent these assumptions, as detailed in @sec-appendixB32. These assumptions improve model tractability and simplify the estimation process by enabling the derivation of the BNF of the likelihood (@eq-like_BNF). With this simplified structure, any probabilistic programming language can model the system and compute the parameter's posterior distribution using @eq-bayes_rule.

$$
P(X_{1}, X_{2}, \dots, X_{P} \mid \theta) = P(X_{1} \mid \theta) \prod^{P}_{p=2} P(X_{i} \mid pa(X_{i}), \theta )
$$ {#eq-like_BNF}

{{< pagebreak >}}

# References {.unnumbered .appendix appendix-style=plain}

<!-- {.unnumbered} -->

:::{#refs}

:::