### 2025-07-03 _tex/paper2.tex
Built site for gh-pages

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  review,
  1p]{elsarticle}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}

\usepackage{float}
\floatplacement{table}{H}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Psychometrika}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Let's talk about Thurstone \& Co.: An information-theoretical model for comparative judgments, and its statistical translation},
  pdfauthor={Jose Manuel Rivera Espejo; Tine van van Daal; Sven De De Maeyer; Steven Gillis},
  pdfkeywords={causal inference, directed acyclic graphs, structural
causal models, bayesian statistical methods, thurstonian
model, comparative judgement, probability, statistical modeling},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Let's talk about Thurstone \& Co.: An information-theoretical
model for comparative judgments, and its statistical translation}
\author[1]{Jose Manuel Rivera Espejo%
\corref{cor1}%
}
 \ead{JoseManuel.RiveraEspejo@uantwerpen.be} 
\author[1]{Tine van Daal%
%
}
 \ead{tine.vandaal@uantwerpen.be} 
\author[1]{Sven De Maeyer%
%
}
 \ead{sven.demaeyer@uantwerpen.be} 
\author[2]{Steven Gillis%
%
}
 \ead{steven.gillis@uantwerpen.be}

\affiliation[1]{organization={University of Antwerp, Training and
education sciences},,postcodesep={}}
\affiliation[2]{organization={University of
Antwerp, Linguistics},,postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
This study revisits Thurstone's law of comparative judgment (CJ) by
addressing two central issues in the CJ literature. First, it critiques
the widespread reliance on Thurstone's Case V assumptions and, by
extension, the Bradley-Terry-Luce (BTL) model in the analysis of CJ
data. Specifically, this study argues that while the assumptions of
equal discriminal dispersions and zero correlation between stimuli
simplify the trait measurement model, they may fail to capture the
complexity of some traits or account for heterogeneous stimuli,
potentially leading to unreliable and inaccurate trait estimates.
Second, the study highlights the apparent disconnect between trait
measurement and hypothesis testing procedures in CJ applications. It
contends that while separating these procedures simplifies the analysis
of CJ data, this practice can also undermine the resulting statistical
inferences.

To address these issues, this study aims to extend Thurstone's general
form using a systematic and integrated approach based on causal and
Bayesian inference methods. This extension integrates Thurstone's core
theoretical principles alongside key CJ assessment design features. It
then translates these elements into a probabilistic statistical model
for analyzing dichotomous CJ data, overcoming the rigid assumptions of
Case V and the BTL model.

Finally, the study emphasizes the relevance of this extension for
contemporary empirical CJ research, particularly stressing the need for
bespoke CJ models tailored to the experiments and data assumptions. It
also lays the foundation for broader applications, encouraging
researchers across the social sciences to adopt more robust and
interpretable methodologies.
\end{abstract}

\begin{keyword}
    causal inference \sep directed acyclic graphs \sep structural causal
models \sep bayesian statistical methods \sep thurstonian
model \sep comparative judgement \sep probability \sep 
    statistical modeling
\end{keyword}
\end{frontmatter}

\newcommand{\dsep}{\:\bot\:}
\newcommand{\ndsep}{\:\not\bot\:}

\section{Introduction}\label{sec-introduction}

In \emph{comparative judgment} (CJ) studies, judges assess a specific
trait or attribute across different stimuli by performing pairwise
comparisons \citep{Thurstone_1927a, Thurstone_1927b}. Each comparison
produces a dichotomous outcome, indicating which stimulus is perceived
to have a higher trait level. For example, judges can compare pairs of
written texts (the stimuli) to determine the relative writing quality
each text exhibit (the trait)
\citep{Pollitt_2012b, vanDaal_et_al_2016, Lesterhuis_2018_thesis, Coertjens_et_al_2017, Goossens_et_al_2018, Bouwer_et_al_2023}.

Numerous studies have documented the effectiveness of CJ in assessing
traits and competencies over the past decade. These studies have
highlighted three key strengths of the method: its reliability, its
validity, and practical applicability. Research on reliability suggests
that CJ requires a relatively modest number of pairwise comparisons
\citep{Verhavert_et_al_2019, Crompvoets_et_al_2022} to generate trait
scores that are as precise and consistent as those generated by other
assessment methods
\citep{Coertjens_et_al_2017, Goossens_et_al_2018, Bouwer_et_al_2023}. In
addition, the evidence suggests that the reliability and time efficiency
of CJ are comparable, if not superior, to those of other assessment
methods when employing adaptive comparison algorithms
\citep{Pollitt_2012b, Verhavert_et_al_2022, Mikhailiuk_et_al_2021}.
Meanwhile, research on validity indicates the capacity of CJ scores to
represent accurately the traits under measurement
\citep{Whitehouse_2012, vanDaal_et_al_2016, Lesterhuis_2018_thesis, Bartholomew_et_al_2018, Bouwer_et_al_2023}.
Lastly, research on its practical applicability highlights CJ's
versatility across educational and non-educational contexts
\citep{Kimbell_2012, Jones_et_al_2015, Bartholomew_et_al_2018, Jones_et_al_2019, Marshall_et_al_2020, Bartholomew_et_al_2020b, Boonen_et_al_2020}.

Nevertheless, despite the increasing number of CJ studies, research in
this domain remains unsystematic and fragmented, leaving several
critical issues unresolved. This study identifies and discusses two
prominent issues in the CJ literature that can undermine the reliability
and validity of CJ's trait estimates. First, it critiques the widespread
reliance on Thurstone's Case V assumptions \citep{Thurstone_1927b} and,
by extension, the Bradley-Terry-Luce (BTL) model
\citep{Bradley_et_al_1952, Luce_1959} in the analysis of CJ data.
Specifically, this study argues that while the assumptions of equal
discriminal dispersions and zero correlation between stimuli simplify
the trait measurement model, they may fail to capture the complexity of
some traits or account for heterogeneous stimuli, potentially leading to
unreliable and inaccurate trait estimates. Second, the study highlights
the disconnect between trait measurement and hypothesis testing
procedures in CJ applications. In particular, it contends that while
separating these procedures simplifies the analysis of CJ data, this
practice can also undermine the resulting statistical inferences
\citep{McElreath_2020, Kline_et_al_2023, Hoyle_et_al_2023}.

To address these issues, this study aims to extend Thurstone's general
form through a systematic and integrated approach that combines causal
and Bayesian inference methods. This extension integrates Thurstone's
core theoretical principles alongside key CJ assessment design features,
such as the selection of judges, stimuli, and comparisons. In addition
to potentially enhancing measurement reliability and validity, and
improving statistical accuracy in hypothesis testing, this approach
offers two key advantages. First, it clarifies the interactions among
all actors and processes involved in CJ assessments. Second, it shifts
the current comparative data analysis paradigm from passively accepting
Case V and the BTL model assumptions to actively testing whether those
assumptions fit the data under analysis.

The remainder of this study is organized into seven sections.
Section~\ref{sec-thurstone_theory} provides an overview of Thurstone's
theory. Section~\ref{sec-theory-issues} examines the central issues
identified in the CJ literature. Section~\ref{sec-theoretical} extends
Thurstone's general form to address these issues. The extension
integrates core theoretical principles alongside key CJ assessment
design features, such as the selection of judges, stimuli, and
comparisons. Section~\ref{sec-statistical} translates these theoretical
and design elements into a probabilistic statistical model to analyze
dichotomous pairwise comparison data. Section~\ref{sec-discussion}
reviews the findings, outline directions for future research, discusses
the study's limitations and details the challenges that applied
researchers may encounter. Finally, Section~\ref{sec-conclusion}
provides the concluding remarks.

\section{Thurstone's theory}\label{sec-thurstone_theory}

In its \emph{general form}, Thurstone's theory addresses pairwise
comparisons of a single judge who evaluates multiple stimuli
\citep{Thurstone_1927a, Thurstone_1927b}. The theory posits that two key
factors determine the dichotomous outcome of these comparisons: the
discriminal process of each stimulus and their discriminal difference.
The \emph{discriminal process} captures the psychological impact each
stimulus exerts on the judge or, more simply, his perception of the
stimulus trait. The theory assumes that the discriminal process for any
given stimulus forms a Normal distribution along the trait continuum
\citep{Thurstone_1927b}. The mode (mean) of this distribution, known as
the \emph{modal discriminal process}, indicates the stimulus position on
this continuum, while its dispersion, referred to as the
\emph{discriminal dispersion}, reflects variability in the perceived
trait of the stimulus.

These ideas become clearer with an example.
Figure~\ref{fig-discriminal_process} illustrates the hypothetical
discriminal processes for two written texts along a \emph{quality} trait
continuum. The figure shows that the modal discriminal process for Text
B lies further along the continuum than that of Text A
\((T_{B} > T_{A})\), suggesting that Text B exhibits higher quality. The
figure also shows that Text B has a broader distribution than Text A,
indicating a greater variability in perception due to its larger
discriminal dispersion \((\sigma_{B} > \sigma_{A})\).

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./images/png/discriminal_process.png}

}

\subcaption{\label{fig-discriminal_process}Hypothetical discriminal
processes for two written texts}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./images/png/discriminal_difference.png}

}

\subcaption{\label{fig-discriminal_difference}Hypothetical discriminal
difference of two written texts}

\end{minipage}%

\caption{\label{fig-thurstone_theory}Hypothetical discriminal processes
and discriminant difference along a quality trait continuum for two
written texts.}

\end{figure}%

However, since the individual discriminal processes of the stimuli are
not directly observable \citep{Thurstone_1927a}, the theory introduces
the \emph{law of comparative judgment}. This law posits that in pairwise
comparisons, a judge perceives the stimulus with a discriminal process
positioned further along the trait continuum as possessing more of the
trait \citep{Bramley_2008}. This suggests that pairwise comparison
outcomes depend on the relative distance between stimuli, not their
absolute positions on the continuum. Indeed, the theory assumes that the
difference between the underlying discriminal processes of the stimuli,
referred to as the \emph{discriminal difference}, determines the
observed dichotomous outcome. Furthermore, the theory assumes that
because the individual discriminal processes form a Normal distribution
on the continuum, the discriminal difference will also conform to a
Normal distribution \citep{Andrich_1978}. In this distribution, the mode
(mean) represents the average relative separation between the stimuli
\((T_{B} - T_{A})\), and its dispersion \(\sigma_{BA}\) indicates the
variability of that separation.

Figure~\ref{fig-discriminal_difference} illustrates the distribution of
the discriminal difference for the two hypothetical texts. The figure
indicates that the judge perceives Text B as having significantly higher
quality than Text A. Two key observations support this conclusion: the
positive difference between their modal discriminal processes
\((T_{B} - T_{A} > 0)\) and the probability area where the discriminal
difference distinctly favors Text B over Text A, represented by the
shaded gray area denoted as \(P(B > A)\). As a result, the dichotomous
outcome of this comparison is more likely to favor Text B over A.

\section{Two Prominent Issues in Traditional CJ
Practice}\label{sec-theory-issues}

Thurstone noted from the outset that his general formulation, described
in Section~\ref{sec-thurstone_theory}, led to a \emph{trait scaling
problem}. Specifically, the model required estimating more ``unknown''
parameters than the number of available pairwise comparisons
\citep{Thurstone_1927b}. For instance, in a CJ assessment with five
texts, the general form would require estimating \(20\) parameters: five
modal discriminal processes, five discriminal dispersions, and \(10\)
correlations--one per comparison. However, a single judge could only
provide \({5 \choose 2} = 10\) unique comparisons, an ``insufficient''
data set to estimate the required parameters.

To address this issue and facilitate the practical implementation of the
theory, Thurstone developed five cases derived from this general form,
each progressively incorporating additional simplifying assumptions
\citep{Thurstone_1927b}. In Case I, Thurstone postulated that pairs of
stimuli would maintain a constant correlation across all comparisons. In
Case II, he allowed multiple judges to undertake comparisons instead of
confining evaluations to a single judge. In Case III, he posited that
there was no correlation between stimuli. In Case IV, he assumed that
the stimuli exhibited similar dispersions. Finally, in Case V, he
replaced this assumption with the condition that stimuli had equal
discriminal dispersions. Table~\ref{tbl-thurstone_cases} summarizes the
assumptions of the general form and the five cases. For a detailed
discussion of these cases and their growing simplification, refer to
\citet{Thurstone_1927b} and \citet{Bramley_2008}.

\begin{table}

\caption{\label{tbl-thurstone_cases}Thurstones cases and their
asumptions}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./images/png/thurstone_cases.png}

}

\end{table}%

As the table suggest, Thurstone developed Case V with an emphasis on
statistical simplicity, but this simplicity comes at the expense of
accurate and precise trait measurement and practical guidance for
inference. Specifically, Thurstone cautioned that Case V use ``should
not be made without (an) experimental test''
\citep[p.~270]{Thurstone_1927b}, as it imposes the most extensive set of
simplifying assumptions \citep{Bramley_2008, Kelly_et_al_2022} (see
Table~\ref{tbl-thurstone_cases}). Moreover, because Thurstone's primary
goal was to produce a ``rather coarse scaling'' of traits and ``allocate
the compared stimuli on this continuum''
\citep[p.~269]{Thurstone_1927b}, his theory did not support formal
statistical inference. Nevertheless, despite Case V's limitations, CJ
research has predominantly relied on it to measure various traits,
raising significant concerns about the reliability and validity of such
measurements in contexts where its assumptions may not hold
\citep{Kelly_et_al_2022, Andrich_1978}. Furthermore, although the CJ
tradition has attempted to address the gap in statistical inference by
relying on the point estimates of traits--or their transformations--the
statistical literature cautions against using these estimates as the
sole basis for inference, as such practices introduce bias and reduce
the precision of hypothesis tests
\citep{McElreath_2020, Kline_et_al_2023, Hoyle_et_al_2023}. The next
sections discuss both issues in greater depth.

\subsection{The Case V and the statistical analysis of CJ
data}\label{sec-theory-issue1}

Case V is the most widely used model in the CJ literature. This
preference largely stems from the widespread adoption of the BTL model,
which provides a simplified statistical representation of the case. The
BTL model incorporates most of Case V's assumptions, with one notable
exception. While Case V assumes a Normal distribution for the stimuli'
discriminal processes, the BTL model uses the more mathematically
tractable Logistic distribution \citep{Andrich_1978, Bramley_2008} (see
Table~\ref{tbl-thurstone_cases}). However, this substitution has minimal
impact on trait estimation or model interpretation because the scale of
the discriminal process (i.e., the latent trait) is arbitrary up to a
non-monotonic transformation
\citep{vanderLinden_et_al_2017_I, McElreath_2021}. In other words, as
long as the substitution (transformation) preserves the rank order of
the data, the choice of distribution for the discriminal processes is
inconsequential. The BTL model satisfies this condition because the
Normal and Logistic distributions exhibit analogous statistical
properties, differing only by a scaling factor of approximately \(1.7\)
\citep{vanderLinden_et_al_2017_I}.

However, Thurstone acknowledged that some assumptions of Case V are
problematic when assessing complex traits or heterogeneous stimuli
\citep{Thurstone_1927a}. Thus, given that contemporary CJ applications
often involve such traits and stimuli, two key assumptions of Case V,
and by extension, the BTL model, may not always hold in theory or
practice. These assumptions are the equal dispersion and zero
correlation between stimuli.

\subsubsection{The assumption of equal dispersions between
stimuli}\label{sec-theory-issue1a}

According to the theory, discrepancies in the discriminal dispersions of
stimuli shape the distribution of the discriminal difference, directly
influencing the outcome of pairwise comparisons. A thought experiment
can help illustrate this idea. Suppose researchers observe the
discriminal processes for two texts, A and B, assuming that the
dispersion for Text A remains constant and that the two texts are
uncorrelated \((\rho=0)\). Figure~\ref{fig-dispersion} demonstrates that
an increase in the uncertainty associated with the perception of Text B
relative to Text A \((\sigma_{B} - \sigma_{A})\), broadens the
distribution of their discriminal difference. This broadening affects
the probability area where the discriminal difference distinctly favors
Text B over Text A, expressed as \(P(B>A)\), ultimately influencing the
comparison outcome. Additionally, the figure reveals that when the
discriminal dispersions of the texts are equal, as in the BTL model
\((\sigma_{B} - \sigma_{A}=0)\), the discriminal difference distribution
is more narrow than when the dispersions differ. As a result, the
discriminal difference is more likely to favor Text B over Text A, as it
is represented by the shaded gray area.

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./images/png/dispersion.png}

}

\subcaption{\label{fig-dispersion}Discriminal Difference distribution
under varying discrepancies in stimuli dispersions}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./images/png/correlation.png}

}

\subcaption{\label{fig-correlation}Discriminal Difference distribution
under varying levels of correlation between stimuli}

\end{minipage}%

\caption{\label{fig-casev_issues}The effect of dispersion discrepancies
and stimuli correlation on the distribution of the discriminal
difference.}

\end{figure}%

In experimental practice, however, the thought experiment occurs in
reverse. Researchers first observe the comparison outcome and then use
the BTL model to infer the discriminal difference between stimuli and
their respective discriminal processes \citep{Thurstone_1927a}.
Consequently, the effectiveness of the outcome to reflect \emph{true}
differences between stimuli largely depends on the validity of the
model's assumptions \citep{Kohler_et_al_2019}, in this case, the
assumption of equal dispersions. When the assumption accurately captures
the complexity of the data, the BTL model estimates a discriminal
difference distribution that accurately represents the \emph{true}
discriminal difference between the texts. This scenario is illustrated
in Figure~\ref{fig-dispersion}, when the model's discriminal difference
distribution aligns with the \emph{true} discriminal difference
distribution, represented by the thick continuous line corresponding to
\(\sigma_{B}-\sigma_{A}=0\). The accuracy of this discriminal difference
then ensures reliable estimates for the texts' discriminal processes.

Notably, while assuming equal dispersions simplifies the trait
measurement model, evidence from the CJ literature suggests that this
assumption may not hold for heterogeneous stimuli, such as handwritten
texts or English compositions
\citep{Thurstone_1927a, Andrich_1978, Bramley_2008, Kelly_et_al_2022}.
The presence of the so-called misfit texts signals this limitation.
\emph{Misfit texts} are those that elicit more judgment discrepancies
than others
\citep{Pollitt_2004, Pollitt_2012b, Pollitt_2012a, Goossens_et_al_2018},
and these discrepancies may arise from larger discriminal dispersions
caused by the stimulus' heterogeneity or because the texts are genuine
outliers--i.e., texts with distinctive characteristics that deviate
markedly from the rest of the sample in which they occur
\citep{Grubbs_1969}. In either case, the BTL model's assumptions prevent
it from adequately accounting for or addressing these anomalies, leaving
exclusion of such ``problematic'' texts as the primary remedy
\citep{Pollitt_2012a, Pollitt_2012b}.

Significant statistical and measurement issues can arise when the
assumption of equal dispersions between stimuli does not hold.
Specifically, the BTL model may overestimate the trait's reliability,
that is, the degree to which the outcome accurately reflects the
\emph{true} discriminal differences between stimuli. This
overestimation, in turn, results in spurious conclusions about these
differences \citep{McElreath_2020, Wu_et_al_2022} and, by extension,
about the underlying discriminal processes of stimuli.
Figure~\ref{fig-dispersion} also illustrates this scenario when the
model's discriminal difference distribution aligns with the thick
continuous line for \(\sigma_{B}-\sigma_{A}=0\), while the \emph{true}
discriminal difference follows any discontinuous line where
\(\sigma_{B}-\sigma_{A} \neq 0\). Furthermore, if it is acknowledged
that \emph{misfit statistics} may represent texts with different
dispersions or outlying observations, the common CJ practice of
excluding stimuli based on these statistics may unintentionally discard
valuable information \citep{Miller_2023}, and introduce bias into the
trait estimates \citep{Zimmerman_1994, McElreath_2020}. The direction
and magnitude of these biases remain unpredictable, as they depend on
which stimuli are excluded from the analysis.

\subsubsection{The assumption of zero correlation between
stimuli}\label{sec-theory-issue1b}

The correlation between two stimuli \(\rho\) measures how much the
judges' perception of a specific trait in one stimulus depends on their
perception of the same trait in other stimulus. Similar to the
discriminal dispersions, this correlation shapes the distribution of the
discriminal difference, directly impacting the outcomes of pairwise
comparisons. Assuming that the discriminal dispersions for a couple of
texts remain constant, Figure~\ref{fig-correlation} shows that as the
correlation between the two texts increases, the distribution of their
discriminal difference becomes narrower. This narrowing, in turn,
affects the probability that the discriminal difference distinctly
favors Text B over Text A--denoted as \(P(B > A)\)--and thus directly
influences the comparison outcome. Furthermore, the figure shows that
when two texts are independent or uncorrelated, as assumed in the BTL
model \((\rho=0)\), the distribution of their discriminal difference is
less narrow than in scenarios where the texts are positively correlated.
As a result, it becomes less likely for the comparison to favor Text B
over Text A, as indicated by the larger shaded area.

Despite these notable differences in the distribution of the discriminal
difference under various correlational assumptions, in practice,
assessment designs often adopt the assumption of no correlation between
stimuli based on Thurstone's early theoretical justification
\citep{Thurstone_1927b}. He argued that stimuli could be treated as
uncorrelated because judges' biases--arising from two opposing and
equally weighted effects occurring during the pairwise
comparisons--would cancel each other out. This idea was later formalized
by \citet{Andrich_1978}, who provided a mathematical demonstration of
this cancellation using the BTL model under the assumption of
discriminal processes with additive biases. However, evidence from the
CJ literature indicates that the assumption of zero correlation does not
hold in practice in at least two cases: (1) when intricate aspects of
multidimensional, complex traits or heterogeneous stimuli influence
judges' perceptions, (2) when additional hierarchical structures are
relevant to the stimuli.

Regarding the first case, research on text quality assessments suggests
that when judges evaluate complex, multidimensional traits or
heterogeneous stimuli, they often rely on a variety of intricate
stimulus characteristics to inform their judgments
\citep{vanDaal_et_al_2016, Lesterhuis_et_al_2018, Chambers_et_al_2022}.
Regardless of their relevance, these characteristics may not be equally
weighted or consistently opposed across comparisons. As a result, they
may exert a disproportionate influence on judges' perceptions,
generating biases that persist rather than cancel out. For example, this
might occur when a judge assessing the argumentative quality of a text
is disproportionately influenced by the clarity of the handwriting,
thereby favoring neatly written texts even if their arguments are
weaker. Moreover, because the discriminal process of stimuli becomes an
observable outcome only through the judges' perceptions, these biases
could introduce dependencies between the stimuli
\citep{vanderLinden_et_al_2017_II}. While direct evidence for this exact
scenario is limited, existing studies document the presence of judge's
biases in CJ contexts
\citep{Pollitt_et_al_2003, vanDaal_et_al_2016, Bartholomew_et_al_2020a},
reinforcing the argument that the factors influencing pairwise
comparisons do not always cancel each other out.

In the second case, the shared context or inherent connections
introduced by additional hierarchical structures may create dependencies
between stimuli--a statistical phenomenon known as clustering
\citep{Everitt_et_al_2010}. For instance, when the same individual
produces multiple texts, those texts often share several features such
as writing style or overall quality. Although some CJ studies
acknowledge the presence of such hierarchical structures and account for
them \citep[e.g.,][]{Boonen_et_al_2020}, the treatment of this
additional source of dependence in other research has often been
insufficient. For instance, when CJ data include multiple samples of
stimuli from the same individuals, researchers frequently rely on
(averaged) point estimates of the BTL scores to conduct subsequent
analyses and tests at the individual level
\citep{Bramley_et_al_2019, Bouwer_et_al_2023, vanDaal_et_al_2017, Jones_et_al_2019, Gijsen_et_al_2021}.

Thus, erroneously assuming zero correlation between stimuli can also
lead to significant statistical and measurement issues. In particular,
neglecting judges' biases or relevant hierarchical structures can create
dimensional mismatches in the model, leading to the over- or
underestimation of trait reliability
\citep{Ackerman_1989, Hoyle_et_al_2023} and even introduce statistical
biases \citep{Wu_et_al_2022}. These inaccuracies can result in spurious
conclusions about the discriminal differences \citep{McElreath_2020}
and, by extension, the underlying discriminal processes of the stimuli.
One such spurious conclusion could be the incorrect classification of
stimuli (or judges) as \emph{misfits}. Figure~\ref{fig-correlation}
illustrates how assuming zero correlation can undermine trait
reliability: the discriminal difference distribution of the BTL scores
follows the thick continuous line \((\rho = 0)\), while the \emph{true}
discriminal difference may correspond to any discontinuous line where
\(\rho \neq 0\).

Finally, similar to misfit stimuli, removing \emph{misfit judges} risks
discarding valuable information and introducing bias into trait
estimates \citep{Miller_2023}. The direction and magnitude of these
biases remain unpredictable, as they depend on which judges are excluded
from the analysis \citep{Zimmerman_1994, OHagan_2018, McElreath_2020}.
\emph{Misfit} judges are those whose assessments deviate markedly from
the shared consensus
\citep{Pollitt_2012a, Pollitt_2012b, vanDaal_et_al_2016, Goossens_et_al_2018, Wu_et_al_2022},
often appearing as outliers under the BTL model \citep{Wu_et_al_2022}.

\subsection{The disconnect between trait measurement and hypothesis
testing}\label{sec-theory-issue2}

In CJ studies, the BTL model is typically used to measure traits and
position the compared stimuli along a latent continuum
\citep{Thurstone_1927b}. The CJ literature shows that studies frequently
relies on point estimates of these traits--typically the BTL scores or
its transformations--to conduct statistical inference or hypothesis
testing. For example, researchers have used these scores to identify
`misfit' judges and stimuli
\citep{Pollitt_2012b, vanDaal_et_al_2016, Goossens_et_al_2018}, detect
biases in judges' ratings \citep{Pollitt_et_al_2003, Pollitt_2012b},
calculate correlations with other assessment methods
\citep{Goossens_et_al_2018, Bouwer_et_al_2023}, or test hypotheses
related to the underlying trait of interest
\citep{Casalicchio_et_al_2015, Bramley_et_al_2019, Boonen_et_al_2020, Bouwer_et_al_2023, vanDaal_et_al_2017, Jones_et_al_2019, Gijsen_et_al_2021}.

Nevertheless, while separating the trait measurement and hypothesis
testing procedures simplifies the analysis of CJ data, the statistical
literature cautions against relying solely on the point estimates of BTL
scores to conduct statistical inference or hypothesis tests, as this
practice can undermine the resulting statistical inferences. A key
consideration is that BTL scores are parameter estimates that inherently
carry uncertainty (measurement error). Ignoring this uncertainty can
bias the analysis and reduce the precision of hypothesis tests. The
direction and magnitude of such biases are often unpredictable. Results
may be attenuated, exaggerated, or remain unaffected depending on the
degree of uncertainty in the scores and the actual effects being tested
\citep{McElreath_2020, Kline_et_al_2023, Hoyle_et_al_2023}. Furthermore,
the reduced precision in hypothesis tests diminishes their statistical
power, increasing the likelihood of committing type-I or type-II errors
\citep{McElreath_2020}.

In aggregate, the heavy reliance on Thurstone's Case V assumptions in
the statistical analysis of comparative data can compromise the
reliability of trait estimates. This overreliance may also undermine
their validity \citep{Perron_et_al_2015}, particularly when coupled with
the disconnect between the trait measurement and hypothesis testing
procedures. The structural approach to causal inference can address
these issues by offering a systematic and integrated framework to extend
Thurstone's general form. This approach can also strengthen measurement
reliability and validity while enhancing the statistical accuracy of
hypothesis tests.

\section{Extending Thurstone's general form}\label{sec-theoretical}

The \emph{structural approach} to causal inference
\citep{Pearl_2009, Pearl_et_al_2016} offers a formal framework for
identifying causes and estimating their effects using data. The approach
relies on structural causal models (SCMs) and directed acyclic graphs
(DAGs) to formally and graphically represent the assumed causal
structure of a system
\citep{Morgan_et_al_2014, Gross_et_al_2018, Neal_2020}, such as the one
found in CJ assessments. In essence, SCMs and DAGs function as
\emph{conceptual models} on which identification analysis rests
\citep{Schuessler_et_al_2023}. \emph{Identification analysis} determines
whether an estimator can accurately compute an estimand based solely on
its (causal) assumptions, regardless of random variability. Here,
\emph{estimands} represent the specific quantities researchers aim to
determine (i.e., a parameter) \citep{Everitt_et_al_2010}.
\emph{Estimators} denote the methods or functions that transform data
into an estimate (e.g., a statistical model), while \emph{estimates} are
the numerical values approximating the estimand
\citep{Neal_2020, Everitt_et_al_2010}.

As an illustration, consider a researcher seeking to answer the
question: ``To what extent do different teaching methods influence
students' ability to produce high-quality written texts?'' To
investigate this, the researcher designs a CJ assessment by randomly
assigning students (individuals) to two groups, each exposed to a
different teaching method. Judges then compare pairs of students'
written texts (stimuli) to produce a dichotomous outcome reflecting the
relative quality of each text (trait). Based on this setup, researchers
can reformulate the research question as the estimand: \emph{On average,
is there a difference in the ability to produce high-quality written
texts between the two groups of students?} Following standard CJ
practices, the researcher would then use point estimates from the BTL
model, or its transformations, to approximate this estimand. However, as
discussed in Section~\ref{sec-theory-issue1}, Thurstone's Case V and the
BTL model exhibit several statistical and measurement limitations. These
limitations hinder the model's ability to identify and accurately
estimate a range of estimands relevant to CJ inquiries, including the
one described in this illustration.

Fortunately, SCMs and DAGs support identification analysis through two
key advantages\footnote{In depth explanation of these topics is beyond
  the scope of this study, thus, readers seeking a more profound
  understanding can refer to introductory papers such as
  \citet{Pearl_2010}, \citet{Rohrer_2018}, \citet{Pearl_2019}, and
  \citet{Cinelli_et_al_2020}, and introductory books like
  \citet{Pearl_et_al_2018}, \citet{Neal_2020}, and
  \citet{McElreath_2020} are useful. For more advanced study, seminal
  papers such as \citet{Neyman_et_al_1923}, \citet{Rubin_1974},
  \citet{Spirtes_et_al_1991}, and \citet{Sekhon_2009}, along with books
  such as \citet{Pearl_2009}, \citet{Morgan_et_al_2014}, and
  \citet{Hernan_et_al_2025}, are recommended.}. First, regardless of
complexity, they can represent various causal structures using only five
fundamental building blocks \citep{Neal_2020, McElreath_2024}. This
feature allows the decomposition of complex structures into manageable
components, facilitating their analysis. Second, they depict causal
relationships in a non-parametric way. This flexibility enables feasible
identification strategies without requiring specification of the types
of variables, the functional forms relating them, or the parameters of
those functional forms \citep{Pearl_et_al_2016}.

Thus, using SCMs and DAGs, this study extends Thurstone's general form
to address the issues identified in Section~\ref{sec-theory-issues}.
This extension combines Thurstone's core theoretical principles (see
Section~\ref{sec-thurstone_theory}) with key CJ assessment design
features, such as the selection of judges, stimuli, and comparisons.
Specifically, Section~\ref{sec-theory-theoretical_P} introduces the
\emph{conceptual-population model} (henceforth CPM), which incorporates
these theoretical principles and assumes an idealized setting where
researchers observe a \emph{conceptual population} of comparative
judgment data--that is, data representing all repeated judgments made by
every available judge for each pair of stimuli produced by each pair of
individuals in the population. Conversely,
Section~\ref{sec-theory-theoretical_SC} presents the
\emph{sample-comparison model} (hereafter SCM), which integrates the
assessment design features and reflects a more realistic setting where
researchers access only a sample of judges, individuals, stimuli, and
comparisons from the conceptual population.

\subsection{The conceptual-population model
(CPM)}\label{sec-theory-theoretical_P}

In the CPM, the idealized scenario of a \emph{conceptual population} of
comparative data enables the integration of Thurstone's theoretical
principles and provides a foundation for proposing innovations aimed at
addressing some of the issues discussed in
Section~\ref{sec-theory-issues}.

\subsubsection{Integrating the first theoretical
principles}\label{sec-theory-theoretical_P1}

Before incorporating the first theoretical principles of Thurstone's
theory, it is essential to further define SCMs. SCMs are formal
mathematical models characterized by a set of \emph{endogenous}
variables \(V\), a set of \emph{exogenous} variables \(E\), and a set of
functions \(F\)
\citep{Pearl_2009, Pearl_et_al_2016, Cinelli_et_al_2020}. Endogenous
variables are those whose causal mechanisms a researcher chooses to
model \citep{Neal_2020}. In contrast, exogenous variables represent
\emph{errors} or \emph{disturbances} arising from omitted factors that
the investigator chooses not to model explicitly \citep{Pearl_2009}.
Lastly, the functions, referred to as \emph{structural equations},
express the endogenous variables as non-parametric functions of other
endogenous and exogenous variables. These functions use the symbol
`\(:=\)' to denote the asymmetrical causal dependence between variables
and the symbol `\(\:\bot\:\)' to represent \emph{d-separation}, a
concept akin to statistical (conditional) independence.

SCM~\ref{fig-cj03_scm} presents the first theoretical principles
embedded in the CPM evaluating the impact of different teaching methods
on students' writing ability. This SCM outlines the relationship between
the conceptual-population outcome \((O^{cp}_{iahbjk})\) and several
related variables. The subscripts \(i\) and \(h\) identify the students
who authored the texts (i.e., the individuals). The indices \(a\) and
\(b\) represent the texts under comparison (i.e., the stimuli). The
index \(j\) indicates the judge conducting the comparison, while the
index \(k\) accounts for assessment conditions where a judge compares
the same pair of stimuli multiple times, i.e., a \emph{repeated measures
designs} \citep{Lawson_2015}. Thus, the indexing system supports
comparisons between different texts written by the same student
\((i = h;\) \(a \neq b)\) and between texts written by distinct students
\((i \neq h;\) where \(a = b\) is permitted\()\), each compared once or
repeatedly by all judges \((j = 1,\dots,n_{J};\) \(k = 1,\dots,n_K;\)
where \(n_{J}>1\) and \(n_{K}\geq1)\). However, it excludes cases where
a judge compares a student's text to itself, whether once or multiple
times \((i = h;\) \(a = b;\) \(j = 1,\dots,n_{J};\)
\(k = 1,\dots,n_{K};\) where \(n_{J}>1\) and \(n_{K}\geq1)\), as such
comparison lacks practical relevance within the CJ framework. Here,
\(n_{J}\) indicates the total number of judges, and \(n_{K}\) denotes
the number of repeated judgments each judge performs.

\begin{figure}[H]

\begin{minipage}{0.50\linewidth}

\centering{

\[
\begin{aligned}
  O^{cp}_{iahbjk} & := f_{O}(D_{iahbjk}) \\
  D_{iahbjk} & := f_{D}(T_{ia}, T_{hb}, B_{jk})
\end{aligned}
\]

}

\subcaption{\label{fig-cj03_scm}SCM}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=0.77\linewidth,height=\textheight,keepaspectratio]{./images/png/CJ_TM_03.png}

}

\subcaption{\label{fig-cj03_dag}DAG}

\end{minipage}%

\caption{\label{fig-cj03}Conceptual-population model (CPM), scalar
form.}

\end{figure}%

In line with Thurstone's theory, SCM~\ref{fig-cj03_scm} depicts the
texts' discriminal processes \((T_{ia}, T_{hb})\) and their discriminal
difference \((D_{iahbjk})\) (see Section~\ref{sec-thurstone_theory}).
Additionally, the SCM incorporates a key CJ design feature: the judges'
biases \((B_{kj})\). This extension builds on the arguments presented in
Section~\ref{sec-theory-issue1b}, contending that the discriminal
difference becomes an observable outcome only through judges'
perceptions. Given that such perceptions may be imperfect--and that each
judge may carry some degree of bias
\citep[see][]{Pollitt_et_al_2003, vanDaal_et_al_2016}--it is reasonable
that judges' perceptions (bias) should be treated as an integral
component of the CJ system from the outset, as this leads to a more
accurate representation of the data-generating process underlying the
pairwise comparisons. This model defines the preliminary set of
endogenous variables,
\(V = \{ O_{iahbjk}, D_{iahbjk}, T_{ia}, T_{hb}, B_{kj} \}\), and the
preliminary set of structural equations, \(F = { f_{O}, f_{D} }\), which
capture the non-parametric dependencies among these variables.

Notably, every SCM has an associated DAG
\citep{Pearl_et_al_2016, Cinelli_et_al_2020}. A DAG is a \emph{graph}
consisting of nodes connected by edges, where nodes represent random
variables. The term \emph{directed} indicates that edges or arrows
extend from one node to another, indicating the direction of causal
influence. The absence of an edge implies no direct relationship between
the nodes. The term \emph{acyclic} means that the causal influences do
not form loops, ensuring the influences do not cycle back on themselves
\citep{McElreath_2020}. DAGs conventionally depict observed variables as
solid black circles and unobserved (latent) variables as open circles
\citep{Morgan_et_al_2014}. Although DAGs conventionally omit exogenous
variables for simplicity, the DAGs presented in this section includes
exogenous variables to improve clarity and reveal potential issues
related to conditioning and confounding \citep{Cinelli_et_al_2020}.

Figure~\ref{fig-cj03_dag} displays the DAG corresponding to
SCM~\ref{fig-cj03_scm}, illustrating the expected causal relationships
outlined in Thurstone's theory. The graph shows that the discriminal
processes of the texts \((T_{ia}, T_{hb})\) influence their discriminal
difference \((D_{iahbjk})\), which in turn determines the outcome
\((O^{cp}_{iahbjk})\). It also highlights the influence of judges'
biases \((B_{kj})\) on the discriminal difference. Additionally, the DAG
differentiates between observed endogenous variables, such as the
outcome (solid black circle), and latent endogenous variables, including
the texts' discriminal processes, their discriminal difference, and the
judges' biases (open circles).

\subsubsection{\texorpdfstring{The \emph{conceptual-population} data
structure}{The conceptual-population data structure}}\label{sec-theory-theoretical_P2}

Although specifying a data structure is not mandatory when using SCMs
and DAGs, defining one improves clarity and facilitates the description
of the system. Thus, to re-express the scalar form of the CJ system
shown in Figure~\ref{fig-cj03} into an equivalent vectorized form, we
first define the vectors \(I\) and \(J\), along with the matrices \(IA\)
and \(JK\), as in Equation (\ref{eq-mat11}). Here, each element of \(I\)
represents a unique individual \(i\) or \(h\), where \(n_{I}\) denotes
the total number of individuals. Similarly, each element of \(J\)
corresponds to a unique judge \(j\), with \(n_{J}\) indicating the total
number of judges. Moreover, each row of \(IA\) represents a unique
pairing of individuals \(i, h\) with stimuli \(a, b\). As a result, the
matrix \(IA\) contains \(n_{I} \cdot n_{A}\) rows and \(2\) columns,
where \(n_{A}\) specifies the number of stimuli available per
individual. Likewise, each row of \(JK\) associates a judge \(j\) with a
(repeated) judgment index \(k\). Consequently, the matrix \(JK\) has
\(n_{J} \cdot n_{K}\) rows and \(2\) columns, where \(n_{K}\) indicates
the number of repeated judgments each judge makes.

Additionally, we construct the matrix \(R\) to map each row of the
\(IA\) matrix with a corresponding row from the \(JK\) matrix. This
matrix has \(n\) rows and \(6\) columns, where
\(n = {n_{I} \cdot n_{A} \choose 2} \cdot n_{J} \cdot n_{K}\). Here, the
term \({n_{I} \cdot n_{A} \choose 2}\) represents the binomial
coefficient, which quantifies the total number of unique comparisons
possible between every pair of stimuli generated by each pair of
individuals in the population. Thus, we define the matrix as in Equation
(\ref{eq-mat11}).

\begin{equation}\phantomsection\label{eq-mat11}{
I = \begin{bmatrix}
1 \\
\vdots \\
i \\
\vdots \\
h \\
\vdots \\
n_{I}
\end{bmatrix};
J = \begin{bmatrix}
1 \\
\vdots \\
j \\
\vdots \\
n_{J}
\end{bmatrix};
IA = \begin{bmatrix}
1 & 1 \\
\vdots & \vdots \\
1 & n_{A} \\
\vdots & \vdots \\
i & a \\
\vdots & \vdots \\
h & b \\
\vdots & \vdots \\
n_{I} & 1 \\
\vdots & \vdots \\
n_{I} & n_{A}
\end{bmatrix};
JK = \begin{bmatrix}
1 & 1 \\
\vdots & \vdots \\
1 & n_{K} \\
\vdots & \vdots \\
j & k \\
\vdots & \vdots \\
n_{J} & 1 \\
\vdots & \vdots \\
n_{J} & n_{K}
\end{bmatrix};
R = \begin{bmatrix}
1 & 1 & 1 & 2 & 1 & 1 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
1 & 1 & 1 & 2 & 1 & n_{K} \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
i & a & h & b & j & k \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
n_{I} & n_{A}-1 & n_{I} & n_{A} & n_{J} & 1 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
n_{I} & n_{A}-1 & n_{I} & n_{A} & n_{J} & n_{K}
\end{bmatrix}
}\end{equation}

It is easier to visualize the structure of the previously defined
vectors and matrices by considering an example. Assuming \(n_{I} = 5\),
\(n_{A} = 2\), \(n_{J} = 3\), and \(n_{K} = 3\), the vectors and
matrices described in Equation (\ref{eq-mat11}) take the form as in
Equation (\ref{eq-mat13}).
\begin{equation}\phantomsection\label{eq-mat13}{
I = \begin{bmatrix}
1 \\
2 \\
3 \\
4 \\
5 
\end{bmatrix} ; \;
J = \begin{bmatrix}
1 \\
2 \\
3 
\end{bmatrix} ; \;
IA = \begin{bmatrix}
1 & 1 \\
1 & 2 \\
2 & 1 \\
2 & 2 \\
3 & 1 \\
3 & 2 \\
4 & 1 \\
4 & 2 \\
5 & 1 \\
5 & 2 
\end{bmatrix} ; \;
JK = \begin{bmatrix}
1 & 1 \\
1 & 2 \\
1 & 3 \\
2 & 1 \\
2 & 2 \\
2 & 3 \\
3 & 1 \\
3 & 2 \\
3 & 3 
\end{bmatrix} ; \;
R = \begin{bmatrix}
1 & 1 & 1 & 2 & 1 & 1 \\
1 & 1 & 1 & 2 & 1 & 2 \\
1 & 1 & 1 & 2 & 1 & 3 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
1 & 1 & 5 & 2 & 1 & 1 \\
1 & 1 & 5 & 2 & 1 & 2 \\
1 & 1 & 5 & 2 & 1 & 3 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
4 & 2 & 5 & 2 & 3 & 1 \\
4 & 2 & 5 & 2 & 3 & 2 \\
4 & 2 & 5 & 2 & 3 & 3 \\
5 & 1 & 5 & 2 & 3 & 1 \\
5 & 1 & 5 & 2 & 3 & 2 \\
5 & 1 & 5 & 2 & 3 & 3 
\end{bmatrix}
}\end{equation}

Now, using Equation (\ref{eq-mat11}), we can re-express
SCM~\ref{fig-cj03_scm} and DAG~\ref{fig-cj03_dag} in an equivalent
vectorized form, as shown in Figure~\ref{fig-cj04}. In this depiction,
the outcome \(O^{cp}_{R}\), the texts' discriminal difference \(D_{R}\),
their discriminal processes \(T_{IA}\), and the judges' biases
\(B_{JK}\) are represented as vectors rather than scalar values. These
vectors capture all the observations from the conceptual population.
Specifically, \(O^{cp}_{R}\) and \(D_{R}\) are observed and latent
vectors of length \(n\), respectively. Moreover, \(T_{IA}\) and
\(B_{JK}\) are latent vectors of lengths \(n_{I} \cdot n_{A}\) and
\(n_{J} \cdot n_{K}\), respectively.

\begin{figure}[H]

\begin{minipage}{0.50\linewidth}

\centering{

\[
\begin{aligned}
  O^{cp}_{R} & := f_{O}(D_{R}) \\
  D_{R} & := f_{D}(T_{IA}, B_{JK})
\end{aligned}
\]

}

\subcaption{\label{fig-cj04_scm}SCM}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=0.87\linewidth,height=\textheight,keepaspectratio]{./images/png/CJ_TM_04.png}

}

\subcaption{\label{fig-cj04_dag}DAG}

\end{minipage}%

\caption{\label{fig-cj04}Conceptual-population model (CPM), initial
vectorized form.}

\end{figure}%

\subsubsection{Integrating hierarchical structural
components}\label{sec-theory-theoretical_P3}

Building on the principles of Structural Equation Modeling (SEM)
\citep{Hoyle_et_al_2023} and Item Response Theory (IRT)
\citep{Fox_2010, vanderLinden_et_al_2017_I}, the CPM integrates two
\emph{hierarchical structural components} to examine how different
\emph{relevant} \footnote{\emph{Relevant variables} are those that
  satisfy the \emph{backdoor criterion} \citep[pp 37]{Neal_2020}, that
  is, they belong to a \emph{sufficient adjustment set}
  \citep{Pearl_2009, Pearl_et_al_2016, Morgan_et_al_2014}. A
  \emph{sufficient} set (potentially empty) blocks all non-causal paths
  between a predictor and an outcome without opening new ones
  \citep{Pearl_2009}. Refer also to footnote 1.} variables--whether
observed or latent--affect the primary latent variable of interest
\citep{Everitt_et_al_2010}. This hierarchical design supports the
formulation and testing of hypotheses that account for both the nested
structure of stimuli and the uncertainties inherent in trait estimation
(see Section~\ref{sec-theory-issue1b} and
Section~\ref{sec-theory-issue2} for a discussion of these
considerations).

The top branch of DAG~\ref{fig-cj09_dag} illustrates the first
component, where \emph{relevant} \footnote{refer to footnote 2.}
student-related variables \(X_{I}\), such as teaching method, and
students' idiosyncratic errors \(e_{I}\) causally influence the latent
variable representing students' writing-quality trait \(T_{I}\). The
error term \(e_{I}\) captures variations in students' traits unexplained
by \(X_{I}\). Here, \(X_{I}\) is an observed matrix with \(n_{I}\) rows
and \(q_{I}\) independent columns (variables), and both \(e_{I}\) and
\(T_{I}\) are latent vectors of length \(n_{I}\). Additionally, this
branch shows how \(T_{I}\), along with \emph{relevant} \footnote{refer
  to footnote 2.} text-related variables \(X_{IA}\) (e.g., text length),
and texts' idiosyncratic errors \(e_{IA}\) causally influence the texts'
written-quality trait \(T_{IA}\), the first primary latent variable of
interest. The error term \(e_{IA}\) captures variations in the texts'
traits that remain unexplained by \(T_{I}\) or \(X_{IA}\). Here,
\(X_{IA}\) is an observed matrix with dimensions \(n_{I} \cdot n_{A}\)
rows and \(q_{IA}\) independent columns (variables), while \(e_{IA}\)
and \(T_{IA}\) are latent matrices with \(n_{I}\) rows and \(n_{A}\)
columns.

Similarly, the bottom branch of DAG~\ref{fig-cj09_dag} depicts the
second component, where \emph{relevant} \footnote{refer to footnote 2.}
judge-related variables \(Z_{J}\), such as judgment expertise, and
judges' idiosyncratic errors \(e_{J}\) causally influence the latent
variable representing judges' bias \(B_{J}\). The error \(e_{J}\)
captures variations in judges' bias unexplained by \(Z_{J}\). Here,
\(Z_{J}\) is an observed matrix with \(n_{J}\) rows and \(q_{J}\)
independent columns (variables), and both \(e_{J}\) and \(B_{J}\) are
latent vectors of length \(n_{J}\). Furthermore, the branch shows how
\(B_{J}\), along with \emph{relevant} \footnote{refer to footnote 2.}
judgment-related variables \(Z_{JK}\) (e.g., the number of judgments a
judge makes), and judgments' idiosyncratic errors \(e_{JK}\) causally
influence the judges' biases associated with each text \(B_{JK}\), the
second primary latent variable of interest. The error \(e_{JK}\)
captures variations in judgments unexplained by \(B_{J}\) or \(Z_{JK}\).
Here, \(Z_{JK}\) is an observed matrix with dimension
\(n_{J} \cdot n_{K}\) rows and \(q_{JK}\) independent columns
(variables), while \(e_{JK}\) and \(B_{JK}\) are latent latent matrices
with \(n_{J}\) rows and \(n_{K}\) columns

Notably, all variables and functions shown in SCM~\ref{fig-cj09_scm} and
DAG~\ref{fig-cj09_dag} are part of the set of endogenous variables
\(V\), structural equations \(F\), and exogenous variables \(E\) for the
CPM. Additionally, the figures demonstrate that all exogenous variables
are independent of one another, as indicated by the relationships
\(e_{IA} \:\bot\:\{ e_{I}, e_{JK}, e_{J} \}\),
\(e_{I} \:\bot\:\{ e_{JK}, e_{J} \}\) and \(e_{JK} \:\bot\:e_{J}\) and
the absence of connecting arrows.

Overall, the CPM extends Thurstone's general form by introducing key
innovations to address the limitations discussed in
Section~\ref{sec-theory-issue1b} and Section~\ref{sec-theory-issue2}.
These enhancements include accounting for judges' biases and integrating
hierarchical structural components. Nevertheless, despite its promise of
enhancing measurement accuracy and precision, the model still depends on
the unrealistic assumption that researchers have access to data from the
\emph{conceptual population}. Since this assumption is rarely met in
practice, the study must consider a more realistic scenario.

\begin{figure}[H]

\begin{minipage}{\linewidth}

\centering{

\[
\begin{aligned}
  O^{cp}_{R} & := f_{O}(D_{R}) \\
  D_{R} & := f_{D}(T_{IA}, B_{JK}) \\
  T_{IA} & := f_{T}(T_{I}, X_{IA}, e_{IA}) \\
  T_{I} & := f_{T}(X_{I}, e_{I}) \\
  B_{JK} & := f_{B}(B_{J}, Z_{JK}, e_{JK}) \\
  B_{J} & := f_{B}(Z_{J}, e_{J}) \\
  e_{I} & \:\bot\:\{ e_{J}, e_{IA}, e_{JK} \} \\
  e_{J} & \:\bot\:\{ e_{IA}, e_{JK} \} \\
  e_{IA} & \:\bot\:e_{JK} 
\end{aligned}
\]

}

\subcaption{\label{fig-cj09_scm}SCM}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{./images/png/CJ_TM_09.png}

}

\subcaption{\label{fig-cj09_dag}DAG}

\end{minipage}%

\caption{\label{fig-cj09}Conceptual-population model (CPM), final
vectorized form.}

\end{figure}%

\subsection{The sample-comparison model
(SCM)}\label{sec-theory-theoretical_SC}

The SCM presents a more realistic scenario than the CPM. First, it
explicitly assumes a data sample consisting of a limited number of
repeated judgments \((n^{s}_{K})\) from a sample of judges
\((n^{s}_{J})\) and a specific number of texts \((n^{s}_{A})\) from a
sample of students \((n^{s}_{I})\), all drawn from the conceptual
population (Section~\ref{sec-theory-theoretical_SC1}). Second, the model
assumes that judges do not perform \emph{all repeated judgments} within
the data sample (Section~\ref{sec-theory-theoretical_SC2}). Instead,
they conduct a sufficient number of stimuli comparisons, \(n_{C}\), to
ensure an accurate estimation of the proportion \(P(B>A)\), as proposed
by \citet{Thurstone_1927b}.

\subsubsection{The sample mechanism}\label{sec-theory-theoretical_SC1}

To incorporate the sampling mechanism and facilitate the interpretation
of the sample-comparison model, we first define the \emph{data sampling
process} using the binary vector variables \(S_{I}\), \(S_{J}\),
\(S_{IA}\), and \(S_{JK}\) as follows:
\begin{equation}\phantomsection\label{eq-mat21}{
S_{I} = \begin{bmatrix}
i_{(1)} \\
\vdots \\
i_{(i)} \\
\vdots \\
i_{(h)} \\
\vdots \\
i_{(nI)}
\end{bmatrix} ; \;
S_{J} = \begin{bmatrix}
j_{(1)} \\
\vdots \\
j_{(j)} \\
\vdots \\
j_{(nJ)}
\end{bmatrix} ; \;
S_{IA} = \begin{bmatrix}
ia_{(1,1)} \\
\vdots \\
ia_{(1,n_{A})} \\
\vdots \\
ia_{(i,a)} \\
\vdots \\
ia_{(h,b)} \\
\vdots \\
ia_{(nI,1)} \\
\vdots \\
ia_{(nI,nA)}
\end{bmatrix} ; \;
S_{JK} = \begin{bmatrix}
jk_{(1,1)} \\
\vdots \\
jk_{(1,n_{K})} \\
\vdots \\
jk_{(j,k)} \\
\vdots \\
jk_{(nJ,1)} \\
\vdots \\
jk_{(nJ,nK)}
\end{bmatrix}
}\end{equation}

Where each element of \(S_{I}\) is a binary value indicating the
presence or absence of corresponding elements in the vector \(I\), as in
Equation (\ref{eq-mat22}). We apply the same logic to \(S_{J}\) using
vector \(J\) (not shown). Thus, the vectors \(S_{I}\) and \(S_{J}\)
contains \(n_{I}\) and \(n_{J}\) elements, respectively.
\begin{equation}\phantomsection\label{eq-mat22}{
i_{(i)} = \begin{cases} 
1 & \text{if data element } i \text{ from } I \text{ is sampled} \\
0 & \text{if data element } i \text{ from } I \text{ is missing}
\end{cases}
}\end{equation}

Similarly, each element of \(S_{IA}\) is a binary value indicating the
presence or absence of data rows in the matrices \(IA\), as defined in
Equation (\ref{eq-mat23}). We apply the same logic to \(S_{JK}\) using
the matrix \(JK\) (not shown). Thus, the vectors \(S_{IA}\) and
\(S_{JK}\) contains \(n_{I} \cdot n_{A}\) and \(n_{J} \cdot n_{K}\)
elements, respectively. \begin{equation}\phantomsection\label{eq-mat23}{
ia_{(i,a)} = \begin{cases} 
1 & \text{if data elements } i,a \text{ from } IA \text{ are sampled} \\
0 & \text{if data elements } i,a \text{ from } IA \text{ are missing}
\end{cases}
}\end{equation}

We can illustrate the structure of these vectors more clearly with an
example. Suppose researchers exclude the second student, the second text
from each student, and the third judge from the setup shown in Equation
(\ref{eq-mat13}). Given \(n_{I} = 5\), \(n_{A} = 2\), \(n_{J} = 3\), and
\(n_{K} = 3\), the resulting vectors would have the following structure:
\begin{equation}\phantomsection\label{eq-mat24}{
S_{I} = \begin{bmatrix}
1 \\
0 \\
1 \\
1 \\
1
\end{bmatrix} ; \;
S_{J} = \begin{bmatrix}
1 \\
1 \\
0
\end{bmatrix} ; \;
S_{IA} = \begin{bmatrix}
1 \\
0 \\
0 \\
0 \\
1 \\
0 \\
1 \\
0 \\
1 \\
0 
\end{bmatrix} ; \;
S_{JK} = \begin{bmatrix}
1 \\
1 \\
1 \\
1 \\
1 \\
1 \\
0 \\
0 \\
0 
\end{bmatrix}
}\end{equation}

Notably, Equation (\ref{eq-mat24}) shows that missing observations in
the vectors \(S_{I}\) and \(S_{J}\)--which represent unsampled students
and judges--directly determine which observations are missing in
\(S_{IA}\) and \(S_{JK}\). In other words, researchers can only observe
texts and judgments from students and judges initially included in the
sample. The equation also shows that the sum of observed elements in
\(S_{I}\) equals the number of sampled students \((n^{s}_{I})\) and that
a similar sum in vector \(S_{J}\) equals the sampled judges
\((n^{s}_{J})\). Conversely, the sum of observed elements in \(S_{IA}\)
represents the total sampled texts across all sampled students
\((n^{s}_{I} \cdot n^{s}_{A})\), while a similar sum in vector
\(S_{JK}\) represents the total sampled repeated judgments across all
sampled judges \((n^{s}_{J} \cdot n^{s}_{K})\). Notice that in this
example, because the design systematically excludes every third repeated
judgment, \(S_{JK}\) can also be expressed using
\(n_{K} = n^{s}_{K} = 2\).

Finally, we define the \emph{sample mechanism} \(S\) in Equation
(\ref{eq-mat25}), which maps each element of \(S_{IA}\) to every element
of \(S_{JK}\). Each element \(s_{(i,a,h,b,j,k)}\) is a binary value
indicating the presence or absence of data rows in the matrix \(R\)
resulting from the sample mechanism, as in Equation (\ref{eq-mat26}).
Thus, the vector contains \(n\) elements, matching the number of rows in
\(R\), and the sum of its elements represents the total data sample:
\(n^{s} = \binom{n^{s}_{I} \cdot n^{s}_{A}}{2} \cdot n^{s}_{J} \cdot n^{s}_{K}\).
Here, the term \({n^{s}_{I} \cdot n^{s}_{A} \choose 2}\) represents the
binomial coefficient, which quantifies the total number of unique
comparisons possible between every pair of sampled stimuli generated by
each pair of sampled individuals.

\begin{equation}\phantomsection\label{eq-mat26}{
s_{(i,a,h,b,j,k)} = \begin{cases} 
1 & \text{if data elements } i,a,h,b,j,k \text{ from } R \text{ are sampled} \\
0 & \text{if data elements } h,i,a,b,j,k \text{ from } R \text{ are missing}
\end{cases}
}\end{equation}

\begin{equation}\phantomsection\label{eq-mat25}{
S = \begin{bmatrix}
s_{(1,1,1,2,1,1)} \\
\vdots \\
s_{(1,1,1,2,1,n_{K})} \\
\vdots \\
s_{(i,a,h,b,j,k)} \\
\vdots \\
s_{(n_{I},n_{A}-1,n_{I},n_{A},n_{J},1)} \\
\vdots \\
s_{(n_{I},n_{A}-1,n_{I},n_{A},n_{J},1)}
\end{bmatrix}
}\end{equation}

With the definition of \(S\), we incorporate the sample mechanism into
the CPM. Following the convention of \citet{McElreath_2020} and
\citet{Deffner_et_al_2022}, DAG~\ref{fig-cj14_dag} represents the
conceptual-population outcome \(O^{cp}_{R}\) as unobserved, emphasizing
that this outcome cannot be directly accessed due to the sampling
mechanism. The DAG also depicts the \emph{sample design} vector \(S\) as
a causal factor influencing the sample-comparison outcome
\(O^{sc}_{R}\). A square encloses \(S\), indicating that it is a
conditioned variable. In this context, \emph{conditioning} means that
the analysis is restricted to the elements of \(O^{cp}_{R}\) that
satisfy \(s_{(i,a,h,b,j,k)}=1\) \citep{Neal_2020, McElreath_2020}. In
essence, \(S\) is a vector that selects \emph{all repeated judgments
made by a subset of judges for a subset of stimuli produced by the
sampled individuals}.

Notably, the DAG shows that \(S\) is independent of all other variables
in the model. This implies that DAG~\ref{fig-cj14_dag} applies
exclusively to Simple Random Sampling (SRSg) designs. In these designs,
each repeated judgment, judge, stimulus, and individual has the same
probability of being included in the sample as any other observation
within their respective groups \citep{Lawson_2015}.

However, due to concerns about the practical feasibility of the
comparison task \citep{Boonen_et_al_2020}, CJ assessments rarely
implement an exhaustive pairings of sampled judges, stimuli, and
individuals. Thus, a realistic scenario must account for the fact that
judges typically compare only a subset of stimuli authored by a sample
of individuals.

\begin{figure}[H]

\begin{minipage}{\linewidth}

\centering{

\[
\begin{aligned}
  O_{R} & := f_{C}(O^{sc}_{R}, C) \\
  O^{sc}_{R} & := f_{S}(O^{cp}_{R}, S) \\
  O^{cp}_{R} & := f_{O}(D_{R}) \\
  D_{R} & := f_{D}(T_{IA}, B_{JK}) \\
  T_{IA} & := f_{T}(T_{I}, X_{IA}, e_{IA}) \\
  T_{I} & := f_{T}(X_{I}, e_{I}) \\
  B_{JK} & := f_{B}(B_{J}, Z_{JK}, e_{JK}) \\
  B_{J} & := f_{B}(Z_{J}, e_{J}) \\
  e_{I} & \:\bot\:\{ e_{J}, e_{IA}, e_{JK} \} \\
  e_{J} & \:\bot\:\{ e_{IA}, e_{JK} \} \\
  e_{IA} & \:\bot\:e_{JK} 
\end{aligned}
\]

}

\subcaption{\label{fig-cj14_scm}SCM}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./images/png/CJ_TM_14.png}

}

\subcaption{\label{fig-cj14_dag}DAG}

\end{minipage}%

\caption{\label{fig-cj14}Sample-comparison model, final vectorized form}

\end{figure}%

\subsubsection{The comparison
mechanism}\label{sec-theory-theoretical_SC2}

As in the previous section, we begin defining the \emph{comparison
mechanism} using the binary vector variable \(C\) to facilitate the
interpretation of the sample-comparison model. Equation (\ref{eq-mat27})
shows that \(C\) contains \(n\) elements corresponding to the number of
rows in the \(R\) matrix, with each element \(c_{(i,a,h,b,j,k)}\) being
a binary value indicating the presence or absence of data rows in \(R\),
a definition similar to that of \(s_{(i,a,h,b,j,k)}\) in Equation
(\ref{eq-mat26}). \begin{equation}\phantomsection\label{eq-mat27}{
C = \begin{bmatrix}
c_{(1,1,1,2,1,1)} \\
\vdots \\
c_{(1,1,1,2,1,n_{K})} \\
\vdots \\
c_{(i,a,h,b,j,k)} \\
\vdots \\
c_{(n_{I},n_{A}-1,n_{I},n_{A},n_{J},1)} \\
\vdots \\
c_{(n_{I},n_{A}-1,n_{I},n_{A},n_{J},1)}
\end{bmatrix}
}\end{equation}

The DAG~\ref{fig-cj14_dag} also incorporates the \emph{comparison
mechanism} \(C\) into the CPM. It shows the sample-comparison outcome
\(O^{sc}_{R}\) as unobserved, emphasizing that this variable cannot be
directly accessed because of the comparison mechanism. The DAG further
shows \(C\) as a conditioned variable (enclosed in a square) that
causally influences the observed outcome \(O_{R}\). This structure
implies that \(C\) determines \emph{which repeated judgments judges make
for the stimuli produced by the individuals}. In essence, \(C\) reflects
the assumption that judges \emph{do not} perform all possible repeated
judgments but instead complete a sufficient number, \(n_{C}\), to enable
the accurate estimation of the proportion \(P(B>A)\) for each stimulus
pair \citep[p.~267]{Thurstone_1927b}.

Notably, DAG~\ref{fig-cj14_dag} also shows that \(C\) is independent of
all other variables in the model. This independence implies that the
conceptual model represented by the DAG applies exclusively to Random
Allocation Comparative Designs \citep{Bramley_2015}, or Incomplete Block
Designs \citep{Lawson_2015}, where every repeated judgment has an equal
probability of being included in the sample.

Finally, since it is standard to assume that the distribution of the
conceptual-population outcome \(O^{cp}_{R}\) also holds for
\(O^{sc}_{R}\) and \(O_{R}\), we can reformulate the sample-comparison
model in Figure~\ref{fig-cj14} into the equivalent form shown in
Figure~\ref{fig-cj15}. This reformulation produces a model that applies
directly to a sample of comparative data. In this version, the
unobserved outcomes \(O^{cp}_{R}\) and \(O^{sc}_{R}\) are omitted, and
\(O_{R}\) inherits the structural equation \(f_{O}\) that originally
defined \(O^{cp}_{R}\). Moreover, the definition of \(O_{R}\) now
reflects its direct dependence on the discriminal difference \(D_{R}\)
and the sample and comparison mechanisms, \(S\) and \(C\).

\begin{figure}[H]

\begin{minipage}{\linewidth}

\centering{

\[
\begin{aligned}
  O_{R} & := f_{O}(D_{R}, S, C) \\
  D_{R} & := f_{D}(T_{IA}, B_{JK}) \\
  T_{IA} & := f_{T}(T_{I}, X_{IA}, e_{IA}) \\
  T_{I} & := f_{T}(X_{I}, e_{I}) \\
  B_{JK} & := f_{B}(B_{J}, Z_{JK}, e_{JK}) \\
  B_{J} & := f_{B}(Z_{J}, e_{J}) \\
  e_{I} & \:\bot\:\{ e_{J}, e_{IA}, e_{JK} \} \\
  e_{J} & \:\bot\:\{ e_{IA}, e_{JK} \} \\
  e_{IA} & \:\bot\:e_{JK} 
\end{aligned}
\]

}

\subcaption{\label{fig-cj15_scm}SCM}

\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\centering{

\includegraphics[width=0.72\linewidth,height=\textheight,keepaspectratio]{./images/png/CJ_TM_15.png}

}

\subcaption{\label{fig-cj15_dag}DAG}

\end{minipage}%

\caption{\label{fig-cj15}Comparative judgment model}

\end{figure}%

In summary, the SCM~\ref{fig-cj15_scm} and DAG~\ref{fig-cj15_dag} extend
Thurstone's general form to address several limitations of the BTL
model. These extensions account for judges' biases (see
Section~\ref{sec-theory-theoretical_P1}), reflect the hierarchical
structure of stimuli and incorporate measurement error in trait
estimation and hypothesis testing (see
Section~\ref{sec-theory-theoretical_P3}), and even clarify the role of
the sample and comparison mechanisms in CJ assessments (see
Section~\ref{sec-theory-theoretical_SC}). However, they do not resolve
concerns about the assumption of equal dispersions among stimuli
discussed in Section~\ref{sec-theory-issue1a}. Since this concern
relates to the statistical assumption underlying the distribution of the
discriminal process, we develop a formal statistical model to address it
in the next section.

\section{From SCM to statistical model}\label{sec-statistical}

Using the SCM~\ref{fig-cj15_scm}, we can derive a statistical model that
addresses violations of the equal dispersion assumption (see
Section~\ref{sec-theory-issue1a}). This derivation is possible because a
fully specified SCM encodes functional and probabilistic information,
which we can replace with suitable functions and probabilistic
assumptions \citep{Pearl_et_al_2016}. Specifically,
SCM~\ref{fig-cj15_scm} allows us to express the joint distribution of
our complex CJ system as a product of simpler conditional probability
distributions (CPDs)\footnote{This re-expression is possible because the
  \emph{chain rule} of probability and the \emph{Bayesian Network
  Factorization (BNF)} property. For further details, see
  \citet{Pearl_et_al_2016} and \citet{Neal_2020}.}, as shown in Equation
(\ref{eq-mat51}). For clarity, we treat expressions such as
\(Y := f_{Y}(X)\), \(P(Y \mid X)\), and \(Y \sim f(Y \mid X)\) as
equivalent, where \(P(Y \mid X)\) and \(f(Y \mid X)\) represent the CPD
of \(Y\) given \(X\). \begin{equation}\phantomsection\label{eq-mat51}{
\begin{aligned}
  P(O_{R}, & S, C, D_{R}, T_{IA}, X_{IA}, e_{IA}, T_{I}, X_{I}, e_{I}, B_{JK}, Z_{JK}, e_{JK}, B_{J}, Z_{J}, e_{J} ) & \\
  &= P(O_{R} \mid D_{R}, S, C) \cdot P(S) \cdot P(C) \cdot P(D_{R} \mid T_{IA}, B_{JK}) \\
  & \quad \cdot P(T_{IA} \mid T_{I}, X_{IA}, e_{IA}) \cdot P(T_{I} \mid X_{I}, e_{I}) \\
  & \quad \cdot P(B_{JK} \mid B_{J}, Z_{JK}, e_{JK}) \cdot P(B_{J} \mid Z_{J}, e_{J}) \\
  & \quad \cdot P(X_{IA}) \cdot P(X_{I}) \cdot P(Z_{JK}) \cdot P(Z_{J}) \\
  & \quad \cdot P(e_{IA}) \cdot P(e_{I}) \cdot P(e_{JK}) \cdot P(e_{J})  
\end{aligned}
}\end{equation}

Each CPD in Equation (\ref{eq-mat51}) rests on specific assumptions,
which we outline in the statistical model presented in
Figure~\ref{fig-cj16_stat}. The model starts by assuming that \(O_{R}\)
follows a Bernoulli distribution\footnote{The binomial
  distribution--including its special case, the Bernoulli
  distribution--represent a maximum entropy distribution for binary
  events \citep[p.~34]{McElreath_2020}. This means that the Bernoulli
  distribution is the most consistent alternative when only two
  un-ordered outcomes are possible and their expected frequencies are
  assumed to be constant \citep[p.~310]{McElreath_2020}. For a detailed
  discussion of the binomial as a maximum entropy distribution, see
  \citet[sec.~10.1.2]{McElreath_2020}.}, reflecting the binary nature of
CJ outcomes. Furthermore, following the conventions of Generalized
Linear Models (GLMs)
\citep{Nelder_et_al_1983, Nelder_et_al_1996, Agresti_2015}, the
distribution links \(O_{R}\) to the latent discriminal difference vector
\(D_{R}\) using an inverse-logit function:
\(\text{inv\_logit}(x) = 1/(1 + \exp(-x))\).

\begin{figure}

\begin{minipage}{0.33\linewidth}

\centering{

\[
\begin{aligned}
  O_{R} & := f_{O}(D_{R}, S, C) \\ 
  D_{R} & := f_{D}(T_{IA}, B_{JK}) \\
  T_{IA} & := f_{T}(T_{I}, X_{IA}, e_{IA}) \\
  T_{I} & := f_{T}(X_{I}, e_{I}) \\
  B_{JK} & := f_{B}(B_{J}, Z_{JK}, e_{JK}) \\
  B_{J} & := f_{B}(Z_{J}, e_{J}) \\ \\
  e_{I} & \:\bot\:\{ e_{J}, e_{IA}, e_{JK} \} \\
  e_{J} & \:\bot\:\{ e_{IA}, e_{JK} \} \\
  e_{IA} & \:\bot\:e_{JK}
\end{aligned}
\]

}

\subcaption{\label{fig-cj16_scm}SCM}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\[
\begin{aligned}
  & P( O_{R} \mid D_{R}, S, C ) \\
  & P( D_{R} \mid T_{IA}, B_{JK} ) \\
  & P( T_{IA} \mid T_{I}, X_{IA}, e_{IA} ) \\
  & P( T_{I} \mid X_{I}, e_{I} ) \\
  & P( B_{JK} \mid B_{J}, Z_{JK}, e_{JK} ) \\
  & P( B_{J} \mid Z_{J}, e_{J} ) \\ \\
  & P( e_{I} ) P( e_{IA} ) P( e_{J} ) P( e_{JK} ) \\ \\ \\
\end{aligned}
\]

}

\subcaption{\label{fig-cj16_prob}Probabilistic model}

\end{minipage}%
%
\begin{minipage}{0.33\linewidth}

\centering{

\[
\begin{aligned}
  O_{R} & \overset{iid}{\sim} \text{Bernoulli} \left[ \text{inv\_logit}( D_{R} ) \right] \\
  D_{R} & = \left( T_{IA}[i,a] - T_{IA}[h,b] \right) + B_{JK}[j,k] \\
  T_{IA} & = T_{I} + \beta_{XA} X_{IA} + e_{IA} \\
  T_{I} & = \beta_{XI} X_{I} + e_{I} \\
  B_{JK} & = B_{J} + \beta_{ZK} Z_{JK} + e_{JK} \\
  B_{J} & = \beta_{ZJ} Z_{J} + e_{J} \\ \\
  \boldsymbol{e} & \sim \text{Multi-Normal}( \boldsymbol{\mu}, \boldsymbol{\Sigma} )
  \\
  \boldsymbol{\Sigma} &= \boldsymbol{V} \boldsymbol{Q} \boldsymbol{V} \\ \\
\end{aligned}
\]

}

\subcaption{\label{fig-cj16_stat}Statistical model}

\end{minipage}%

\caption{\label{fig-cj16}Comparative judgment model, SCM, probabilistic
and statistical model assuming different discriminal dispersions for the
student's traits}

\end{figure}%

While the joint distribution in Equation (\ref{eq-mat51}) includes the
probability distributions of the sampling and comparison mechanisms,
\(P(S)\) and \(P(C)\), as well as those of the predictor
variables--\(P(X_{IA})\), \(P(X_{I})\), \(P(Z_{JK})\), and
\(P(Z_{J})\)--all of these probabilities are omitted from the
statistical model \ref{fig-cj16_stat}. This omission is justified
because, while these distributions contribute to the overall joint
distribution of the data, the variables \(S\), \(C\), \(X_{IA}\),
\(X_{I}\), \(Z_{JK}\), and \(Z_{J}\) are observed and independent of any
other variable in the model. As observed variables, they do not require
distributional assumptions in the same way the idiosyncratic errors do.
Their independence follows from the underlying random selection
procedures that govern the variables\footnote{Randomization ensures that
  data--and, by extension, an estimator--satisfies several key
  identification properties, such as common support, no interference,
  and consistency. The most critical property, however, is the
  elimination of confounding. \emph{Confounding} occurs when an external
  variable, such as \(X_{I}\), simultaneously influences both the
  outcome (e.g., \(O_{R}\)) and a variable of interest (e.g., \(S\)),
  resulting in spurious associations between the latter two
  \citep{Everitt_et_al_2010}. Randomization ensure the absence of
  confounding by effectively decoupling the association between the
  variable of interest and any other variable, except for the outcome
  itself. For a more detailed discussion on the benefits of
  randomization, see \citet{Pearl_2009}, \citet{Morgan_et_al_2014},
  \citet{Neal_2020}, and \citet{Hernan_et_al_2025}.}.

Next \(D_{R}\) is defined as the difference between the discriminal
processes \(T_{IA}[i, a]\) and \(T_{IA}[h, b]\), representing the
underlying written-quality trait of the compared texts, plus the
corresponding repeated judge bias \(B_{JK}[j, k]\). Note that if it is
assumed that \(B_{JK}[j,k]\) reflects the difference in
stimulus-specific biases, i.e.,
\(B_{JK}[j,k] = B_{JK}[i,a,j,k] - B_{JK}[h,b,j,k]\), the discriminal
difference can be re-written as:
\begin{equation}\phantomsection\label{eq-mat52}{
\begin{aligned}
D_{R} &= \left( T_{IA}[i,a] - T_{IA}[h,b] \right) + B_{JK}[j,k] \\
&= \left( T_{IA}[i, a] + B_{JK}[i, a, j, k] \right) - \left( T_{IA}[h, b] + B_{JK}[h, b, j, k] \right) \\
& = T^{*}_{IA}[i,a] - T^{*}_{IA}[h,b]
\end{aligned}
}\end{equation}

This formulation reveals that the discriminal difference captures a
\emph{pure interaction effect}, in which neither the texts' discriminal
processes nor the judges' biases alone determine the outcome, but their
interaction does \citep{Attia_et_al_2022}. Put simply, this mathematical
description captures the idea that the stimuli' discriminal processes
become an observable outcome only through the lens of judges'
perceptions (i.e., their biases). For clarity, the square brackets in
\(D_{R}\) indicate the relevant indices for each trait vector; they do
not imply any subsetting of the data.

Now the functional forms for \(T_{IA}\), \(T_{I}\), \(B_{JK}\), and
\(B_{J}\) are specified. \(T_{IA}\) is modeled as a linear combination
of the students' underlying writing-quality traits \(T_{I}\), the
effects of relevant text-related variables on quality assessment
\(\beta_{XA}X_{IA}\) (such as the influence of text length), and the
text-specific idiosyncratic errors \(e_{IA}\). Similarly, \(T_{I}\) is
expressed as a linear combination of relevant student-related variables
affecting the quality assessment \(\beta_{XI} X_{I}\), and
student-specific idiosyncratic errors \(e_{I}\). For the judge-specific
terms, \(B_{JK}\) is modeled as a linear combination of the judge's
individual bias \(B_{J}\), the influence of relevant judgment-related
variables on quality assessment \(\beta_{ZK}Z_{JK}\) (e.g., how the
number of judgments affect the evaluation), and judgment-specific
idiosyncratic errors \(e_{JK}\). Finally, \(B_{J}\) is defined as a
linear combination of relevant judge-level variables influencing the
quality assessment \(\beta_{ZJ}Z_{J}\) (such as judgment expertise) and
judge-specific idiosyncratic errors \(e_{J}\).

Next, the probabilistic assumptions for the idiosyncratic errors
\(e_{I}\), \(e_{IA}\), \(e_{J}\), and \(e_{JK}\) are specified. Unlike
other variables in the model, these error terms exhibit indeterminacies
in their \emph{location}, \emph{orientation}, and \emph{scale} due to
the lack of an inherent scale in the associated latent variables
\(T_{I}\), \(T_{IA}\), \(B_{J}\), and \(B_{JK}\). Thus, to identify the
latent variable model these indeterminacies must be resolved
\citep{Depaoli_2021, deAyala_2009}. Drawing on principles from SEM
\citep{Hoyle_et_al_2023}, the vector of idiosyncratic errors
\(\boldsymbol{e} = [e_{I}, e_{IA}, e_{J}, e_{JK}]^{T}\) are assumed to
follow a Multivariate Normal distribution with mean vector
\(\boldsymbol{\mu}\) and a covariance matrix
\(\boldsymbol{\Sigma} = \boldsymbol{V} \boldsymbol{Q} \boldsymbol{V}\),
with \(\boldsymbol{V}\) denoting a diagonal matrix of standard
deviations and \(\boldsymbol{Q}\) a correlation matrix. To address the
\emph{location} indeterminacy, the errors' mean vector is set to zero:
\begin{equation}\phantomsection\label{eq-mat53}{
\boldsymbol{\mu} = [0, 0, 0, 0]^{T}
}\end{equation}

Following SCM~\ref{fig-cj16_scm}, the \emph{orientation} indeterminacy
is solved by assuming that the errors are uncorrelated. This assumption
leads to the definition of the error's correlation matrix,
\(\boldsymbol{Q}\), as the identity matrix:
\begin{equation}\phantomsection\label{eq-mat54}{
\boldsymbol{Q} = \begin{bmatrix}
    1 & 0 & 0 & 0 \\
    0 & 1 & 0 & 0 \\
    0 & 0 & 1 & 0 \\
    0 & 0 & 0 & 1 
  \end{bmatrix}
}\end{equation}

To resolve the \emph{scale} indeterminacy, the diagonal matrix
\(\boldsymbol{V}\) is defined as follows:
\begin{equation}\phantomsection\label{eq-mat55}{
\boldsymbol{V} = \begin{bmatrix}
    s_{XI} & 0 & 0 & 0 \\
    0 & p_{IA} & 0 & 0 \\
    0 & 0 & s_{ZJ} & 0 \\
    0 & 0 & 0 & p_{JK} 
  \end{bmatrix}
}\end{equation}

Here, \(s_{XI}\) represents the standard deviation for the individuals,
\(p_{IA}\) for the stimuli, \(s_{ZJ}\) for the judges, and \(p_{JK}\)
for the judgments. It is assumed \(s_{XI}\) varies depending on the
teaching method group to which each student belongs. Using the example
from Section~\ref{sec-theoretical}, where the teaching method
\(X_{I} = \{1,2\}\), the model sets the constraint according to Equation
(\ref{eq-mat56}). This constraint anchors the scale of the individuals'
latent trait while relaxing the assumption of equal dispersion for the
stimuli, thereby addressing the concerns raised in
Section~\ref{sec-theory-issue1a}.
\begin{equation}\phantomsection\label{eq-mat56}{
\sum_{g=1}^{2} s_{XI}[g]/2 = 1
}\end{equation}

Because the error vector \(\boldsymbol{e}\) follows an uncorrelated
Multivariate Normal distribution, the marginal distribution of
\(e_{IA}\) is a univariate Normal distribution with mean zero and
standard deviation \(p_{IA}\). Thus, \(p_{IA}\) is set as a proportion
of \(1\) to establish the scale of the stimuli' latent trait relative to
the scale of the individuals' trait. Note that as a result, \(T_{IA}\)
is also normally distributed. This configuration effectively reinstates
Thurstone's original assumption of Normal discriminal processes for the
stimuli (see Table~\ref{tbl-thurstone_cases}).

Similarly, it is assumed that \(s_{ZJ}\) varies depending on the groups
to which each judge belongs. For instance, if \(Z_{J} = \{1,2,3\}\)
represents three groups of judges with varying expertise, the model sets
the constraint according to Equation (\ref{eq-mat57}). This constraint
anchors the scale of the judges' latent trait and relaxes the assumption
of equal dispersion for the judgments.
\begin{equation}\phantomsection\label{eq-mat57}{
\sum_{g=1}^{2} s_{ZJ}[g]/3 = 1
}\end{equation}

Conversely, \(p_{JK}\) is defined as a proportion of \(1\) to establish
the scale of the judgments' latent trait relative to the scale of the
judges' trait.

Finally, we use \emph{Bayesian inference methods} to convert the
statistical model \ref{fig-cj16_stat} into a practical statistical tool
for analyzing paired comparison data. Bayesian inference offers three
main advantages in this context. First, it handles complex and
overparameterized models, where the number of parameters exceeds the
number of observations \citep{Baker_1998, Kim_et_al_1999}. This feature
is essential for our implementation, as the proposed model is indeed
overparameterized. Second, it incorporates prior information to
constrain parameter estimates within plausible bounds, thereby
mitigating estimation issues like non-convergence or improper solutions
that often affect frequentist methods
\citep{Martin_et_al_1975, Seaman_et_al_2011}. Prior distributions are
used to define the error distribution and set the scale of latent
variables \citep{Depaoli_2014}. Third, Bayesian inference supports
robust inferences from small samples, where the asymptotic properties
underlying frequentist methods are less reliable
\citep{Baldwin_et_al_2013, Lambert_et_al_2006, Depaoli_2014}. This
feature is particularly relevant in CJ assessments, as researchers often
collect large volumes of paired comparisons but work with relatively
small samples of judges, stimuli, and individuals to test hypotheses.

The \textbf{Declarations} section of this document provides a link to
the model code, along with an alternative specification that assumes
equal discriminal dispersions. We tested both versions of the model with
success using \texttt{Stan} \citep[version 2.26.1]{Stan_2020}.

\section{Discussion}\label{sec-discussion}

Thurstone introduced the Law of Comparative Judgment to measure
psychological traits of stimuli through pairwise comparisons
\citep{Thurstone_1927a, Thurstone_1927b}. In its general form, the
theory models single-judge comparisons across multiple, potentially
correlated stimuli. Each comparison produces a dichotomous outcome
indicating which stimulus the judge perceives as having a higher trait
level. However, Thurstone identified one key challenge in this general
formulation: the measurement model required estimating more ``unknown''
parameters than the number of available pairwise comparisons
\citep{Thurstone_1927b}. To address this issue and to facilitate the
theory's practical applicability, he formulated five cases, each
progressively incorporating several simplifying assumptions.

Among these, Case V remains the most widely used model in empirical CJ
research, mainly due to the widespread adoption of the BTL model. The
BTL model incorporates the core assumptions of Case V--namely, equal
discriminal dispersions and zero correlation among stimuli' discriminal
processes--but replaces the processes' normal distribution with the more
mathematically tractable logistic distribution
\citep{Andrich_1978, Bramley_2008}. Although this substitution has
minimal impact on trait estimation or model interpretation
\citep{vanderLinden_et_al_2017_I, McElreath_2021}, the simplifying
assumptions of the BTL model--and by extension, of Case V--may fail to
capture the complexity of some traits or account for heterogeneous
stimuli
\citep{Thurstone_1927a, Andrich_1978, Bramley_2008, Kelly_et_al_2022},
potentially leading to unreliable and inaccurate trait estimates
\citep{Ackerman_1989, Zimmerman_1994, McElreath_2020, Hoyle_et_al_2023}.

Moreover, because Thurstone's original goal was to produce a ``coarse
scaling'' of traits and allocate stimuli along this continuum
\citep[p.~269]{Thurstone_1927a}, his theory offered no guidance on how
to use trait estimates for statistical inference. The CJ tradition has
attempted to address this gap by separating trait estimation from
hypothesis testing, relying on point estimates, such as BTL scores or
their transformations, for inference. While this approach simplifies
analysis, it can also introduce bias and compromise the reliability of
the resulting inferences
\citep{McElreath_2020, Kline_et_al_2023, Hoyle_et_al_2023}.

To address the limitations of Thurstone's Case V and the BTL model, this
study extended Thurstone's general form using a systematic, integrated
approach that combined causal and Bayesian inference methods. The
approach began with the development of a conceptual model, formalized as
a Structural Causal Model (SCM) and represented graphically by a
Directed Acyclic Graph (DAG)
\citep{Pearl_2009, Pearl_et_al_2016, Gross_et_al_2018, Neal_2020}. This
model integrated Thurstone's core theoretical principles, such as the
discriminal processes of stimuli, alongside key CJ assessment design
features, including judges' bias, sampling procedures, and comparison
mechanisms. Together, these components allowed the causal processes
underlying the CJ system to be disentangled.

The approach then translated the SCM into a bespoke statistical model
that enabled the analysis of CJ data in cases where the assumptions of
equal dispersion and zero correlation were violated, and where
statistical inference was required. In particular, the model accounted
for judge biases, captured the hierarchical structure of stimuli,
incorporated measurement error into the hypothesis testing process, and
accommodated heterogeneity in discriminal dispersions. By addressing
these issues, the methodological innovations aimed to enhance the
reliability and validity of trait measurement in CJ, while also
improving the accuracy of statistical inferences.

Beyond these potential benefits, the approach offered two additional
advantages. First, it clarified the roles and interactions of all actors
and processes involved in CJ assessments. Second, it shifted the
analytic paradigm from passively accepting the assumptions of Case V and
the BTL model to actively testing their fit with observed data.
Together, these advantages established a principled framework for
evaluating best practices in designing CJ assessments--one that better
aligns with the demands of contemporary CJ contexts
\citep{Kelly_et_al_2022}--providing new insights into existing research
and opening promising avenues for future inquiry.

\subsection{Future research directions using our
approach}\label{sec-discussion_RA}

Among the many potential directions for future research, three avenues
deserve particular attention due to their direct impact on the
reliability and validity of CJ trait estimates, as well as on the
accuracy of statistical inferences. The following sections outline these
avenues and explain how our approach facilitates their investigation.

\subsubsection{The impact of sampling and comparison
mechanisms}\label{sec-discussion_RA1}

Although sampling and comparison mechanisms are central to modern CJ
assessments, it is striking that most CJ literature has examined them
within a limited scope. Researchers have primarily investigated the
effects of adaptive comparative judgment (ACJ) designs on trait
reliability
\citep{Pollitt_2012a, Pollitt_2012b, Bramley_2015, Verhavert_et_al_2022, Mikhailiuk_et_al_2021, Gray_et_al_2024}
or proposed practical guidelines for the number of comparisons judges
should make \citep{Verhavert_et_al_2019, Crompvoets_et_al_2022}. While
these studies offer valuable insights, they also overlook the broader
role that these mechanisms play within the CJ system. As this oversight
likely stems from a more fundamental lack of conceptual clarity about
how these mechanisms function within the system, the present study
integrated these mechanisms into the conceptual model of CJ.

The explicit integration of the sampling and comparison mechanisms
offers a new perspective on how these mechanisms shape the CJ process.
Specifically, it clarified their role as sources of missing data in CJ's
data-generating process--that is, as mechanisms that determine which
observations are missing from the final data sample. This new
perspective invites the application of Little and Rubin's principled
missing data framework \citep{Little_et_al_2020}, allowing a more
rigorous evaluation of existing claims about these missing data
mechanisms, their influence on CJ outcomes, and their implications for
designing and evaluating more complex assessments setups.

This study circumvented the need to apply the missing data framework by
deliberately structuring the sampling and comparison mechanisms to be
independent of any observed or unobserved variables, including the
outcome. In other words, these mechanisms were intentionally designed to
yield data that are \emph{missing completely at random} (MCAR)
\citep{Little_et_al_2020}. This design choice offered one key advantage:
it generated simple random samples that satisfied the condition of
\emph{ignorability}, thereby allowing researchers to legitimately
\emph{ignore} missing data during analysis without introducing bias
\citep{Everitt_et_al_2010, Kohler_et_al_2019, Neal_2020}.

However, many modern CJ applications rely on more complex assessment
designs, in which the sampling and comparison mechanisms introduce more
intricate forms of missingness such as \emph{missing at random} (MAR) or
\emph{missing not at random} (MNAR) \citep{Little_et_al_2020}. One
prominent example is ACJ designs, where prior judgment outcomes inform
the selection of stimulus pairs for subsequent comparisons
\citep{Pollitt_2012a, Pollitt_2012b, Bramley_2015}. This pair selection
procedure suggests that ACJ's comparison mechanism is outcome-dependent,
potentially classifying the method as a generator of MNAR data. If this
classification holds, ACJ might violate the condition of ignorability,
making it an unsuitable pair selection procedure for reliable and valid
trait estimation and inference. Moreover, under this interpretation, the
mixed findings on ACJ's effectiveness become more comprehensible: while
some studies find that the method improves trait reliability
\citep{Pollitt_et_al_2003, Pollitt_2012a, Pollitt_2012b}, others argue
that that these gains may be artificially inflated
\citep{Bramley_2015, Bramley_et_al_2019, Crompvoets_et_al_2020, Crompvoets_et_al_2022}.

Regardless of the underlying missingness mechanisms, any CJ assessment
design would benefit from explicitly defining its assumptions--a
practice supported by this study's approach. This clarity enables a
thorough evaluation of how the sampling and comparison mechanisms affect
trait estimation and statistical inference in each design. Such
assessments are particularly relevant given the common misconception in
the CJ literature that Thurstone's model can naturally handle even
non-random missing data without compromising the reliability or validity
of trait estimates \citep{Bramley_2008}.

\subsubsection{The effects of judges' bias on the reliability of trait
estimates}\label{sec-discussion_RA2}

Despite the growing notion that various stimulus-related factors
influence judges' perceptions
\citep{vanDaal_et_al_2016, Lesterhuis_et_al_2018, Chambers_et_al_2022}
and that these influences may not always cancel each other out,
empirical evidence of judges' biases remains scarce in the CJ literature
\citep{Pollitt_et_al_2003, vanDaal_et_al_2016, Bartholomew_et_al_2020a}.
This gap likely persists not from a lack of interest or research but
because the identification of such biases often depends on ad-hoc
detection methods, like `misfit' statistics, that may not be well-suited
for the task \citep{Kelly_et_al_2022}. To overcome this limitation, the
present study treated judges' biases as an integral component of the CJ
system from the outset. This approach offers one key advantage: it
provides a more accurate representation of the data-generating process
behind pairwise comparisons, one that acknowledges that the discriminal
processes of stimuli becomes an observable outcome only through judges'
perceptions, which may exhibit bias.

The explicit integration of judges' bias into CJ's conceptual model then
paves the way for investigating several relevant research questions. One
key question is whether CJ data can be validly analyzed under the
assumption of ``sample-free'' trait calibration, specifically under the
hypothesis that judges exhibit no systematic bias. This question is
particularly relevant because ``sample-freeness'' is still regarded as
an inherent property of the BTL model \citep{Bramley_2008, Andrich_1978}
despite growing evidence of persistent biases. Another critical question
is whether training or expertise can help judges avoid focusing on
irrelevant stimulus features, such as handwriting, over more central
criteria like argumentative quality in writing assessments
\citep{Kelly_et_al_2022}. Exploring these questions may also provide
insights into what it truly means to be an ``expert'' within the CJ
context \citep{Kelly_et_al_2022}.

Moreover, since judges rely on these stimulus-related factors when
evaluating complex, multidimensional traits (e.g., structure, style)
\citep{vanDaal_et_al_2016, Lesterhuis_et_al_2018, Chambers_et_al_2022},
and these factors account for variation in judgment accuracy
\citep{Gill_et_al_2013, vanDaal_et_al_2017, vanDaal_2020, Gijsen_et_al_2021},
it is reasonable to expect that assessments also vary according to
judge-specific attributes such as gender, age, culture, income,
education, training, or expertise \citep{Kelly_et_al_2022}. Prior
studies support this view
\citep[e.g.,][]{Bartholomew_et_al_2020a, McMahon_et_al_2015}. Thus,
building on the discussion in Section~\ref{sec-discussion_RA1}, further
exploration is warranted into how judge selection influences the
formation of a ``shared consensus'' and whether certain judge attributes
introduce systematic biases or distortions in the stimuli trait
distribution \citep{Deffner_et_al_2022}. Furthermore, if such attributes
do in fact compromise the assumption of ``sample-freeness,'' it becomes
essential to consider strategies for mitigating their effects and to
determine how many judges (and how many judgments per judge) are
required to produce reliable trait estimates under these conditions.
Additionally, it is worth examining whether \emph{repeated measures
designs}, in which judges evaluate the same stimulus pairs multiple
times \citep{Lawson_2015}, can improve judgment consistency and
accuracy. As anticipated, the approach presented in this study provides
a structured foundation to rigorously investigate these questions.

\subsubsection{The identification of `misfitting' judges and
stimuli}\label{sec-discussion_RA3}

Although the CJ literature clearly defines \emph{misfit} judges and
stimuli, CJ researchers have rarely examined how these observations
relate to Thurstonian theory. In particular, they have not identified
which elements of Thurstone's theory account for the occurrence of
misfits. This disconnect likely stems from the fact that misfit
statistics are derived from residual analysis and outlier detection
methods rather than from Thurstonian principles. Specifically,
\emph{misfit judges} are typically defined as those whose assessments
diverge significantly from the ``shared consensus''
\citep{Pollitt_2012a, Pollitt_2012b, vanDaal_et_al_2016, Goossens_et_al_2018, Wu_et_al_2022},
while \emph{misfit stimuli} are those that elicit more judgment
discrepancies than others
\citep{Pollitt_2004, Pollitt_2012a, Pollitt_2012b, Goossens_et_al_2018}.
Both definitions closely mirror the statistical concept of outliers,
that is, observations that deviate markedly from the rest of the sample
in which they occur \citep{Grubbs_1969}. But this resemblance extends
beyond the definitions themselves, as misfits are often identified using
conventional outlier detection procedures, such as transforming BTL
model residuals into diagnostic statistics and comparing them against
predefined thresholds
\citep{Pollitt_2012a, Pollitt_2012b, Wu_et_al_2022}.

Nevertheless, is not the classification of misfits as outliers that
raises concerns for trait measurement and inference. Rather, the concern
lies in the prevailing CJ practice of identifying these observations
through ad-hoc procedures and excluding them from analysis
\citep{Pollitt_2012a, Pollitt_2012b}, often without empirical support
for the various hypotheses suggested to explain their occurrence. It is
essential to recognize that outliers are defined relative to a specific
model \citep{McElreath_2020}. Thus, detection procedures based on models
like the BTL--which rests on strong assumptions--should be approached
with caution, as they may not be appropriate for this purpose
\citep{Kelly_et_al_2022}. If the BTL model does not accurately represent
the actual data-generating process of the CJ system, the method may miss
classify valid observations as misfits. On top this, excluding these
observations carries additional risks. The statistical literature
cautions that removing outliers can discard valuable information
\citep{Miller_2023} and introduce bias into trait estimates. The
direction and magnitude of this bias are often unpredictable, as they
depend on which observations are excluded from the analysis
\citep{Zimmerman_1994, OHagan_2018, McElreath_2020}. Finally, even when
the model aligns well with the data, its rigid assumptions limit the
model's ability to adequately test many of the hypotheses proposed to
explain misfit behavior.

In contrast, the approach presented in this study provides a rigorous
framework for testing several relevant hypotheses. For instance, it
allows to investigate whether misfit judges are those who exhibit (an
outlying degree of) systematic bias or greater variability in their
judgments compared to their peers. Similarly, the approach makes
possible to explore whether misfit stimuli exhibit more variable
discriminal processes relative to other stimuli or if they are genuinely
outlying cases. Moreover, since outliers are not inherently ``bad data''
\citep{McElreath_2020}, the present approach offers a principled
alternative to exclusion, one that retains misfits in the analysis
without compromising trait estimation or inference. This is achieved by
adapting the proposed model into robust measurement models
\citep{McElreath_2020}, a broad class of procedures designed to reduce
the sensitivity of parameter estimates to mild or moderate departures
from model assumptions \citep{Everitt_et_al_2010}. This strategy also
speaks to a broader concern in the social sciences: when predictive
power is low, the solution may not lie solely in seeking new variables
or procedures, but also in adopting more sophisticated measurement
models \citep{Wainer_et_al_1978}.

\subsection{Study limitations and practical challenges for applied CJ
researchers}\label{sec-discussion_challenges}

Drawing conclusions from observed data always requires assumptions,
whether the data are observational or experimental
\citep{Kohler_et_al_2019, Deffner_et_al_2022}. The proposed approach is
not an exception to this fundamental principle. Like all approaches
grounded on causal inference, it relies on expert knowledge and
assumptions about the variables' causal structure that are often
untestable at the outset \citep{Hernan_et_al_2025}. However, its purpose
is not to deliver automatic answers when applied to a given CJ dataset.
Instead, it encourages the formulation of precise questions and the
explicit articulation of assumptions, fostering a generalizable
understanding of the CJ system under study
\citep{Rohrer_et_al_2022, Deffner_et_al_2022, Sterner_et_al_2024}. This
clarity is crucial because the accuracy of estimates and validity of
inferences depend on how well the data and inferential goals align with
a model's assumptions \citep{Kohler_et_al_2019}. Although this alignment
remains to be empirically tested for the proposed models, the
theory-driven nature of this approach provides a solid foundation for
future empirical evaluation of its causal assumptions
\citep{Deffner_et_al_2022}, grounded in both established theory and
existing evidence.

The theoretical commitment to causal inference also introduces several
practical challenges that applied CJ researchers must navigate. These
fall into two main categories: first, acquiring the foundational
knowledge necessary to apply the approach effectively; and second,
dedicating greater attention to both conceptual and statistical
modeling.

\subsubsection{Required foundational
knowledge}\label{sec-discussion_challenges1}

Applying the approach effectively requires foundational knowledge in two
areas. First, a solid understanding of causal inference principles.
Second, the ability to translate the functional and probabilistic
aspects of conceptual models into bespoke statistical models. A clear
example illustrating the importance of causal inference is the recurrent
assumption that predictor variables are ``relevant'' to the research
context--interpreting this relevance as their inclusion in a
\emph{sufficient adjustment set}
\citep{Pearl_2009, Pearl_et_al_2016, Morgan_et_al_2014}. However, this
study does not explore the full implications of that assumption for
model specification, estimation, and inference. Nevertheless, to assist
with the effort of engaging with this and other complex causal inference
concepts, key references are provided throughout the study to guide
applied CJ researchers toward a deeper understanding of these ideas
\footnote{Also refer to the detailed online document referenced in the
  Declarations section}.

Developing the skills to translate conceptual models into bespoke
statistical models also presents challenges. Although Bayesian inference
methods offer a more accessible path for many applied CJ researchers to
develop these skills--by reducing the need for specialized knowledge in
areas like optimization theory, which frequentist methods often
require--they still demand familiarity with technical concepts such as
probabilistic programming languages (PPLs), probability distributions,
and convergence diagnostics. Thus, to support this skill development,
this study provides a link to the statistical model code and alternative
model specifications in the \textbf{Declarations} section of this
document \footnote{Seminal texts on Bayesian inference methods, such as
  \citet{Gelman_et_al_2014} and \citet{McElreath_2020}, offer valuable
  support for developing a deeper understanding of these models.}.

\subsubsection{Attention to conceptual and statistical
modeling}\label{sec-discussion_challenges2}

Even after acquiring the necessary foundational knowledge, applying this
approach to a specific CJ context involves two additional challenges.
First, it is essential to verify whether the conceptual model offers a
\emph{faithful} \footnote{Avoid confusing this term with the
  \emph{faithfulness assumption} \citep{Neal_2020, Hernan_et_al_2025}
  described in the causal inference literature. For further details, see
  footnote 1.} representation of the CJ system under study. Second, the
statistical translation of the model must be assessed to determine if it
can accurately estimate the intended estimands (i.e., parameters) from
empirical data. To ensure fidelity, the conceptual and statistical
models presented here should be treated as starting points, not
universal solutions for all CJ designs or datasets. While these models
may be adequate in some contexts, assuming so without evaluation is
unwise. Instead, adaptation to the specific context is necessary, with
careful attention to assessment design features and causal assumptions.
As discussed in Section~\ref{sec-discussion_challenges}, this approach
supports such adaptation by providing a transparent framework for
articulating new assumptions and guiding CJ assessment design.

Conversely, evaluating the estimation capabilities of the statistical
model requires addressing the challenge of identification analysis. As
outlined in Section~\ref{sec-theoretical}, \emph{identification
analysis} determines whether a statistical model can accurately compute
a given estimand (e.g., a parameter) based solely on its (causal)
assumptions, independent of random variability
\citep{Schuessler_et_al_2023}. Identification is crucial because it is a
necessary condition for consistency. \emph{Consistency} is the property
of an estimator (e.g., a statistical model) whose estimates converge to
the ``true'' value of an estimand as the data size approaches infinity
\citep{Everitt_et_al_2010}. Without identification, consistency is
impossible--even with infinite, error-free data--and meaningful
inference from finite samples cannot be achieved
\citep{Schuessler_et_al_2023}. While formal derivations of
identification may seem a natural next step, the complexity of the CJ
system makes this approach impractical at the outset. Instead,
simulation-based methods like power analysis offer a more practical and
flexible alternative, enabling examination of estimate consistency
without relying on complex mathematical proofs. Notably, the approach
presented here supports both strategies by providing the probabilistic
foundation for formal derivations and the statistical structure needed
for simulation-based methods.

\section{Conclusion}\label{sec-conclusion}

The present study highlights the need to extend Thurstone's theory to
address the demands of contemporary empirical CJ research. It advocates
for developing bespoke CJ models tailored to the specific
data-generating processes of different CJ assessment designs. These
models aim to enhance the robustness of trait estimates and clarity of
inferences. Moreover, this work outlines a clear path for advancing both
theoretical and applied CJ research and lays the foundation for broader
adoption of more robust and interpretable methodologies across the
social sciences.

\newpage{}

\section*{Declarations}\label{declarations}
\addcontentsline{toc}{section}{Declarations}

\textbf{Funding:} The Research Fund (BOF) of the University of Antwerp
funded this project.

\textbf{Financial interests:} The authors declare no relevant financial
interests.

\textbf{Non-financial interests:} The authors declare no relevant
non-financial interests.

\textbf{Ethics approval:} The University of Antwerp Research Ethics
Committee confirmed that this study does not require ethical approval.

\textbf{Consent to participate:} Not applicable

\textbf{Consent for publication:} All authors have read and approved the
final version of the manuscript for publication.

\textbf{Data availability:} This study did not use any data.

\textbf{Materials and code availability:} The \texttt{CODE\ LINK}
section at the top of the digital document located at:
\url{https://jriveraespejo.github.io/paper2_manuscript/} provides access
to all materials and code.

\textbf{AI-assisted technologies in the writing process:} The authors
used various AI-based language tools to refine phrasing, optimize
wording, and enhance clarity and coherence throughout the manuscript.
They take full responsibility for the final content of the publication.

\textbf{CRediT authorship contribution statement:}
\emph{Conceptualization:} J.M.R.E, T.vD., S.DM., and S.G.;
\emph{Methodology:} J.M.R.E, T.vD., and S.DM.; \emph{Software:}
J.M.R.E.; \emph{Validation:} J.M.R.E.; \emph{Formal Analysis:} J.M.R.E.;
\emph{Investigation:} J.M.R.E; \emph{Resources:} T.vD. and S.DM.;
\emph{Data curation:} J.M.R.E.; \emph{Writing - original draft:}
J.M.R.E.; \emph{Writing - review and editing:} T.vD., S.DM., and S.G.;
\emph{Visualization:} J.M.R.E.; \emph{Supervision:} S.G. and S.DM.;
\emph{Project administration:} S.G. and S.DM.; \emph{Funding
acquisition:} S.G. and S.DM.

\newpage{}

\section{Appendix}\label{sec-appendix}

\subsection{Statistical and Causal inference}\label{sec-appendixB}

This section introduces fundamental statistical and causal inference
concepts necessary for understanding the core theoretical principles
described in this document. It does not, however, offer a comprehensive
overview of causal inference methods. Readers seeking more in-depth
understanding may wish to explore introductory papers such as
\citet{Pearl_2010}, \citet{Rohrer_2018}, \citet{Pearl_2019}, and
\citet{Cinelli_et_al_2020}. They may also find it helpful to consult
introductory books like \citet{Pearl_et_al_2018}, \citet{Neal_2020}, and
\citet{McElreath_2020}. For more advanced study, readers may refer to
seminal intermediate papers such as \citet{Neyman_et_al_1923},
\citet{Rubin_1974}, \citet{Spirtes_et_al_1991}, and \citet{Sekhon_2009},
as well as books such as \citet{Pearl_2009}, \citet{Morgan_et_al_2014},
and \citet{Hernan_et_al_2025}.

\subsubsection{Empirical research and randomized
experiments}\label{sec-appendixB1}

Empirical research uses evidence from observation and experimentation to
address real-world challenges. In this context, researchers typically
formulate their research questions as \emph{estimands} or \emph{targets
of inference}, i.e., the specific quantities they seek to determine
\citep{Everitt_et_al_2010}. For instance, researchers might be
interested in answering the following question: ``To what extent do
different teaching methods \((T)\) influence students' ability to
produce high-quality written texts \((Y)\)?'' To investigate this,
researchers could randomly assign students to two groups, each exposed
to a different teaching method \((T_{i} = \{1,2\})\). Then, they would
perform pairwise comparisons, generating a dichotomous outcome
\((Y_{i} = \{0,1\})\) showing which student exhibits more of the
ability. In this scenario, the research question can be rephrased as the
estimand, ``\emph{On average}, is there a difference in the ability to
produce high-quality written texts between the two groups of students?''
and this estimand can be mathematically represented by the random
associational quantity in Equation~\ref{eq-group_diff}, where
\(E[\cdot]\) denotes the expected value.

\begin{equation}\phantomsection\label{eq-group_diff}{
E[Y_{i} \mid T_{i}=1] - E[Y_{i} \mid T_{i}=2]
}\end{equation}

Researchers then proceed to identify the estimands.
\emph{Identification} determines whether an estimator can accurately
compute the estimand based solely on its assumptions, regardless of
random variability \citep[p.~4]{Schuessler_et_al_2023}. An
\emph{estimator} refers to a method or function that transforms data
into an estimate \citep{Neal_2020}. \emph{Estimates} are numerical
values that approximate the estimand derived through the process of
\emph{estimation}, which integrates data with an estimator
\citep{Everitt_et_al_2010}. The Identification-Estimation flowchart
\citep{McElreath_2020, Neal_2020}, shown in Figure~\ref{fig-IEflow},
visually represents the transition from estimands to estimates.

\begin{figure}

\centering{

\includegraphics[width=0.35\linewidth,height=\textheight,keepaspectratio]{images/png/IEflow.png}

}

\caption{\label{fig-IEflow}Identification-Estimation flowchart.
Extracted and slightly modified from \citet[p.~32]{Neal_2020}}

\end{figure}%

Identification is a necessary condition to ensure \emph{consistent}
estimators. An estimator achieves \emph{consistency} when it converges
to the ``true'' value of an estimand as the data size approaches
infinity \citep{Everitt_et_al_2010}. Without identification, researchers
cannot achieve consistency, even with ``infinite'' and error-free data.
As a result, deriving meaningful insights about an estimand from finite
data becomes impossible \citep[p.~5]{Schuessler_et_al_2023}. Therefore,
to ensure accurate and reliable estimates, researchers prioritize
estimators with desirable identification properties. For instance, the
Z-test is a widely used estimator for comparing group proportions,
yielding accurate estimates when its underlying assumptions are
satisfied \citep{Kanji_2006}. Furthermore, researchers can interpret
estimates from the Z-test as causal, provided the data is collected
through a randomized experiment.

Randomized experiments are widely recognized as the gold standard in
evidence-based science \citep{Hariton_et_al_2018, Hansson_2014}. This
recognition stems from their ability to enable researchers interpret
associational estimates as causal. They achieve this by ensuring data,
and by extension an estimator, satisfies several key identification
properties, such as common support, no interference, and consistency
\citep{Morgan_et_al_2014, Neal_2020}. The most critical property,
however, is the elimination of confounding. \emph{Confounding} occurs
when an external variable \(X\) simultaneously influences the outcome
\(Y\) and the variable of interest \(T\), resulting in spurious
associations \citep{Everitt_et_al_2010}. Randomization addresses this
issue by decoupling the association between the intervention allocation
\(T\) from any other variable \(X\)
\citep{Morgan_et_al_2014, Neal_2020}.

Nevertheless, researchers often face constraints that limit their
ability to conduct randomized experiments. These constraints include
ethical concerns, such as the assignment of individuals to potentially
harmful interventions, and practical limitations, such as the
infeasibility of, for example, assigning individuals to genetic
modifications or physical impairments \citep{Neal_2020}. In these cases,
causal inference offers a valuable alternative for generating causal
estimates and understanding the mechanisms underlying specific data. In
addition, the framework can provide significant theoretical insights
that can enhance the design of experimental and observational studies
\citep{McElreath_2020}.

\subsubsection{Identification under causal
inference}\label{sec-appendixB2}

Unlike classical statistical modeling, which focuses primarily on
summarizing data and inferring associations, the \emph{causal inference}
framework is designed to identify causes and estimate their effects
using data \citep{Shaughnessy_et_al_2010, Neal_2020}. The framework uses
rigorous mathematical techniques to address the \emph{fundamental
problem of causality}
\citep{Pearl_2009, Pearl_et_al_2016, Morgan_et_al_2014}. This problem
revolves around the question, ``What would have happened `in the world'
under different circumstances?'' This question introduces the concept of
counterfactuals, which are instrumental in defining and identifying
causal effects.

\emph{Counterfactuals} are hypothetical scenarios that are
\emph{contrary to fact}, where alternative outcomes resulting from a
given cause are neither observed nor observable
\citep{Neal_2020, Counterfactual_2024}. The structural approach to
causal inference \citep{Pearl_2009, Pearl_et_al_2016} provides a formal
framework for defining counterfactuals. For instance, in the scenario
described in Section~\ref{sec-appendixB1}, the approach begins by
defining the \emph{individual causal effect} (ICE) as the difference
between each student's potential outcomes, as in Equation~\ref{eq-ICE}.

\begin{equation}\phantomsection\label{eq-ICE}{
\tau_{i} = Y_{i} \mid do(T_{i}=1) - Y_{i} \mid do(T_{i}=2)
}\end{equation}

where \(do(T_{i}=t)\) represents the intervention operator,
\(Y_{i} \mid do(T_{i}=1)\) represents the potential outcome under
intervention \(T_{i}=1\), and \(Y_{i} \mid do(T_{i}=1)\) represents the
potential outcome under intervention \(T_{i}=2\). Here, an
\emph{intervention} involves assigning a constant value to the treatment
variable for each student's potential outcomes. Note that if a student
is assigned to intervention \(T_{i}=1\), the potential outcome under
\(T_{i}=2\) becomes a counterfactual, as it is no longer observed nor
observable. To address this challenge, the structural approach extends
the ICE to the \emph{average causal effect} (ACE,
Equation~\ref{eq-ACE}), representing the average difference between the
students' observed potential outcomes and their counterfactual
counterparts.

\begin{equation}\phantomsection\label{eq-ACE}{
\begin{aligned}
\tau & = E[\tau_{i}] \\
  & = E[Y_{i} \mid do(T_{i}=1)]- E[Y_{i} \mid do(T_{i}=2)]
\end{aligned}
}\end{equation}

Even though counterfactuals are unobservable, researchers can still
identify the ACE from associational estimates by leveraging the
structural approach. The approach identifies the ACE by statistically
conditioning data on a \emph{sufficient adjustment set} of variables
\(X\) \citep{Pearl_2009, Pearl_et_al_2016, Morgan_et_al_2014}. This
\emph{sufficient} set (potentially empty) must block all non-causal
paths between \(T\) to \(Y\) without opening new ones. When such a set
exists, then \(T\) and \(Y\) are \emph{d-separated} by \(X\)
(\(T \:\bot\:Y \mid X\)) \citep{Pearl_2009}, and \(X\) satisfies the
\emph{backdoor criterion} \citep[pp 37]{Neal_2020}. Here,
\emph{conditioning} describes the process of restricting the focus to
the subset of the population defined by the conditioning variable
\citep[p.~32]{Neal_2020} (see Equation~\ref{eq-CACE}).

Conditioning on a sufficient adjustment set enables researchers to
estimate the ACE, even when the data comes from an observational study.
This process is feasible because such conditioning ensures that the ACE
estimator satisfies several critical properties, including confounding
elimination \citep{Morgan_et_al_2014}. Naturally, the validity of claims
about the causal effects of \(T\) on \(Y\) now hinges on the assumption
that \(X\) serves as a sufficient adjustment set. However, as
\citet[p.~150]{Kohler_et_al_2019} noted, drawing conclusions about the
real world from observed data inevitably requires assumptions. This
requirement holds true for both observational and experimental data.

For instance, if researchers cannot conduct the randomized experiments
described in Section~\ref{sec-appendixB1} and must instead rely on
observational data, they can still identify the ACE as long as an
observed variable \(X\), such as the socio-economic status of the
school, satisfies the backdoor criterion. Under these circumstances,
researchers first identify the \emph{conditional average causal effect}
(CACE, Equation~\ref{eq-CACE})

\begin{equation}\phantomsection\label{eq-CACE}{
CACE_{t} = E[Y_{i} \mid T_{i}=t, X]
}\end{equation}

From the CACE, researchers can identify the ACE from associational
quantities as in Equation~\ref{eq-mACE1}. This identification process is
commonly known as the \emph{backdoor adjustment}. Here, \(E_{X}[\cdot]\)
represents the marginal expected value over \(X\)
\citep{Morgan_et_al_2014}.

\begin{equation}\phantomsection\label{eq-mACE1}{
\begin{aligned}
  \tau & = E[Y_{i} \mid do(T_{i}=1)]- E[Y_{i} \mid do(T_{i}=2)] \\
  & = E_{X}[CACE_{1} - CACE_{2}] \\
  & = E_{X}\left[ E[Y_{i} \mid T_{i}=1, X] - E[Y_{i} \mid T_{i}=2, X] \right]
\end{aligned}
}\end{equation}

Notably, the approach extends the ACE identification for a continuous
variable \(T\) as in Equation~\ref{eq-mACE_cont}, ensuring broad
applicability across different causal scenarios \citep[p.~45]{Neal_2020}

\begin{equation}\phantomsection\label{eq-mACE_cont}{
\begin{aligned}
  \tau &= E[Y_{i} \mid do(T_{i}=t)] \\
  & = d E_{X}\left[ E[Y_{i} \mid T_{i}=t, X]\right]/ dt
  \end{aligned}
}\end{equation}

\subsubsection{Diving into the specifics}\label{sec-appendixB3}

The structural approach to causal inference uses SCMs and DAGs to
formally and graphically represent the presumed causal structure
underlying the ACE
\citep{Pearl_2009, Pearl_et_al_2016, Gross_et_al_2018, Neal_2020}.
Essentially, these tools serve as \emph{conceptual (theoretical) models}
on which identification analysis rests
\citep[p.~4]{Schuessler_et_al_2023}. Thus, using these tools,
researchers can determine which statistical models can identify (ACE,
CACE, or other), assuming the depicted causal structure is correct
\citep{McElreath_2020}, thus enabling valid causal inference.
Figure~\ref{fig-IEflow} shows the role of theoretical models in the
inference process.

SCMs and DAGs support identification analysis through two key
advantages. First, regardless of complexity, they can represent various
causal structures using only five fundamental building blocks
\citep{Neal_2020, McElreath_2020}. This feature allows researchers to
decompose complex structures into manageable components, facilitating
their analysis \citep{McElreath_2020}. Second, they depict causal
relationships in a non-parametric and fully interactive way. This
flexibility enables feasible ACE identification strategies without
defining the variables' data types, the functional form between them, or
their parameters \citep[p.~35]{Pearl_et_al_2016}.

Thus, Section~\ref{sec-appendixB31} and Section~\ref{sec-appendixB32}
elaborate on the first advantage, while Section~\ref{sec-appendixB32}
and Section~\ref{sec-appendixB33} do so for the second. Finally,
Section~\ref{sec-appendixB34} explains how researchers use SCMs and DAGs
alongside Bayesian inference methods in the estimation process.

\paragraph{The five fundamental block for SCMs and
DAGs}\label{sec-appendixB31}

Figures \ref{fig-dags_scms1}, \ref{fig-dags_scms2},
\ref{fig-dags_scms3}, \ref{fig-dags_scms4}, and \ref{fig-dags_scms5}
display the five fundamental building blocks for SCMs and DAGs. The left
panels of the figures show the formal mathematical models, represented
by the SCMs, defined in terms of a set of \emph{endogenous} variables
\(V=\{X_{1},X_{2},X_{3}\}\), a set of \emph{exogenous} variables
\(E=\{e_{X1},e_{X2},e_{X3}\}\), and a set of functions
\(F=\{f_{X1},f_{X2},f_{X3}\}\) \citep{Pearl_2009, Cinelli_et_al_2020}.
Endogenous variables are those whose causal mechanisms a researcher
chooses to model \citep{Neal_2020}. In contrast, exogenous variables
represent \emph{errors} or \emph{disturbances} arising from omitted
factors that the investigator chooses not to model explicitly
\citep[p.~27,68]{Pearl_2009}. Lastly, the functions, referred to as
\emph{structural equations}, express the endogenous variables as
non-parametric functions of other variables. These functions use the
symbol `\(:=\)' to denote the asymmetrical causal dependence of the
variables and the symbol `\(\:\bot\:\)' to represent
\emph{d-separation}, a concept akin to (conditional) independence.

Notably, every SCM has an associated DAG
\citep{Pearl_et_al_2016, Cinelli_et_al_2020}. The right panels of the
figures display these DAGs. A DAG is a graph consisting of nodes
connected by edges, where the nodes represent random variables. The term
\emph{directed} means that the edges extend from one node to another,
with arrows indicating the direction of causal influence. The term
\emph{acyclic} implies that the causal influences do not form loops,
ensuring the influences do not cycle back on themselves
\citep{McElreath_2020}. DAGs represent observed variables as solid black
circles, while they use open circles for unobserved (latent) variables
\citep{Morgan_et_al_2014}. Although the \emph{standard representation}
of DAGs typically omits exogenous variables for simplicity, the
\emph{magnified representation} depicted in the figures offers one key
advantage: including exogenous variables can help researchers highlight
potential issues related to conditioning and confounding
\citep{Cinelli_et_al_2020}.

\begin{figure}[H]

\begin{minipage}{0.50\linewidth}

\centering{

\[
\begin{aligned}
  X_{1} & := f_{X1}(e_{X1}) \\
  X_{3} & := f_{X3}(e_{X3}) \\
  e_{X1} & \:\bot\:e_{X3}
\end{aligned}
\]

}

\subcaption{\label{fig-scm_bb1}SCM}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=0.54\linewidth,height=\textheight,keepaspectratio]{images/png/mdag_bb1.png}

}

\subcaption{\label{fig-mdag_bb1}DAG}

\end{minipage}%

\caption{\label{fig-dags_scms1}Two unconnected nodes}

\end{figure}%

\begin{figure}[H]

\begin{minipage}{0.50\linewidth}

\centering{

\[
\begin{aligned}
  X_{1} & := f_{X1}(e_{X1}) \\
  X_{3} & := f_{X3}(X_{1},e_{X3}) \\
  e_{X1} & \:\bot\:e_{X3}
\end{aligned}
\]

}

\subcaption{\label{fig-scm_bb2}SCM}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=0.54\linewidth,height=\textheight,keepaspectratio]{images/png/mdag_bb2.png}

}

\subcaption{\label{fig-mdag_bb2}DAG}

\end{minipage}%

\caption{\label{fig-dags_scms2}Two connected nodes or descendant}

\end{figure}%

\begin{figure}[H]

\begin{minipage}{0.50\linewidth}

\centering{

\[
\begin{aligned}
  X_{1} & := f_{X1}(e_{X1}) \\
  X_{2} & := f_{X2}(X_{1},e_{X2}) \\
  X_{3} & := f_{X3}(X_{2},e_{X3}) \\
  e_{X1} & \:\bot\:e_{X2} \\
  e_{X1} & \:\bot\:e_{X3} \\
  e_{X2} & \:\bot\:e_{X3}
\end{aligned}
\]

}

\subcaption{\label{fig-scm_bb3}SCM}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=0.65\linewidth,height=\textheight,keepaspectratio]{images/png/mdag_bb3.png}

}

\subcaption{\label{fig-mdag_bb3}DAG}

\end{minipage}%

\caption{\label{fig-dags_scms3}Chain or mediator}

\end{figure}%

\begin{figure}[H]

\begin{minipage}{0.50\linewidth}

\centering{

\[
\begin{aligned}
  X_{1} & := f_{X1}(X_{2},e_{X1}) \\
  X_{2} & := f_{X2}(e_{X2}) \\
  X_{3} & := f_{X3}(X_{2},e_{X3}) \\
  e_{X1} & \:\bot\:e_{X2} \\
  e_{X1} & \:\bot\:e_{X3} \\
  e_{X2} & \:\bot\:e_{X3}
\end{aligned}
\]

}

\subcaption{\label{fig-scm_bb4}SCM}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=0.65\linewidth,height=\textheight,keepaspectratio]{images/png/mdag_bb4.png}

}

\subcaption{\label{fig-mdag_bb4}DAG}

\end{minipage}%

\caption{\label{fig-dags_scms4}Fork or confounder}

\end{figure}%

\begin{figure}[H]

\begin{minipage}{0.50\linewidth}

\centering{

\[
\begin{aligned}
  X_{1} & := f_{X1}(e_{X1}) \\
  X_{2} & := f_{X2}(X_{1},X_{3},e_{X2}) \\
  X_{3} & := f_{X3}(e_{X3}) \\
  e_{X1} & \:\bot\:e_{X2} \\
  e_{X1} & \:\bot\:e_{X3} \\
  e_{X2} & \:\bot\:e_{X3}
\end{aligned}
\]

}

\subcaption{\label{fig-scm_bb5}SCM}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=0.65\linewidth,height=\textheight,keepaspectratio]{images/png/mdag_bb5.png}

}

\subcaption{\label{fig-mdag_bb5}DAG}

\end{minipage}%

\caption{\label{fig-dags_scms5}Collider or inmorality}

\end{figure}%

A careful examination of these building blocks highlights the
theoretical assumptions underlying their observed variables. SCM
\ref{fig-scm_bb1} and DAG \ref{fig-mdag_bb1} depict two unconnected
nodes, representing a scenario where variables \(X_{1}\) and \(X_{3}\)
are independent or not causally related. SCM \ref{fig-scm_bb2} and DAG
\ref{fig-mdag_bb2} illustrate two connected nodes, representing a
scenario where a \emph{parent} node \(X_{1}\) exerts a causal influence
on a \emph{child} node \(X_{3}\). In this setup, \(X_{3}\) is considered
a \emph{descendant} of \(X_{1}\). Additionally, \(X_{1}\) and \(X_{3}\)
are described as \emph{adjacent} because there is a \emph{direct path}
connecting them. SCM \ref{fig-scm_bb3} and DAG \ref{fig-mdag_bb3} depict
a \emph{chain}, where \(X_{1}\) influences \(X_{2}\), and \(X_{2}\)
influences \(X_{3}\). In this configuration, \(X_{1}\) is a parent node
of \(X_{2}\), which is a parent node of \(X_{3}\). This structure
creates a \emph{directed path} between \(X_{1}\) and \(X_{3}\).
Consequently, \(X_{1}\) is an \emph{ancestor} of \(X_{3}\), and
\(X_{2}\) fully \emph{mediates} the relationship between the two. SCM
\ref{fig-scm_bb4} and DAG \ref{fig-mdag_bb4} illustrate a \emph{fork},
where variables \(X_{1}\) and \(X_{3}\) are both influenced by
\(X_{2}\). Here, \(X_{2}\) is a parent node that \emph{confounds} the
relationship between \(X_{1}\) and \(X_{3}\). Finally, SCM
\ref{fig-scm_bb5} and DAG \ref{fig-mdag_bb5} show a \emph{collider},
where variables \(X_{1}\) and \(X_{3}\) are concurrent causes of
\(X_{2}\). In this configuration, \(X_{1}\) and \(X_{3}\) are not
causally related to each other but both influence \(X_{2}\) (an
\emph{inmorality}). Notably, all building blocks assume the errors are
independent of each other and from all other variables in the graph, as
evidenced by the pairwise relations \(e_{X1} \:\bot\:e_{X2}\),
\(e_{X1} \:\bot\:e_{X3}\), and \(e_{X2} \:\bot\:e_{X3}\).

Researchers can then use these building blocks to represent the scenario
outlined in Section~\ref{sec-appendixB2}. SCM \ref{fig-scm_example1} and
DAG \ref{fig-mdag_example1} depict the plausible causal structure for
this example. In this context, the variable \(X\) (socio-economic status
of the school) is thought to be a confounder in the relationship between
the teaching method \(T\) and the outcome \(Y\). The figures display
multiple descendant relationships such as \(X \rightarrow T\),
\(X \rightarrow Y\), and \(T \rightarrow Y\). They also highlight
unconnected node pairs, evident from the relationships
\(e_{T} \:\bot\:e_{X}\), \(e_{T} \:\bot\:e_{Y}\), and
\(e_{X} \:\bot\:e_{Y}\). Additional, the figures show one fork,
\(X \rightarrow \{T, Y\}\), and two colliders:
\(\{X, e_{T}\} \rightarrow T\) and \(\{X, T, e_{Y}\} \rightarrow Y\).

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\[
\begin{aligned}
  X & := f_{X}(e_{X}) \\
  T & := f_{T}(X,e_{T}) \\
  Y & := f_{Y}(T,X,e_{Y}) \\
  e_{T} & \:\bot\:e_{X} \\
  e_{T} & \:\bot\:e_{Y} \\
  e_{X} & \:\bot\:e_{Y}
\end{aligned}
\]

}

\subcaption{\label{fig-scm_example1}SCM}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=0.65\linewidth,height=\textheight,keepaspectratio]{images/png/mdag_example1.png}

}

\subcaption{\label{fig-mdag_example1}DAG}

\end{minipage}%

\caption{\label{fig-example1}Plausible causal structure the scenario
outlined in Section~\ref{sec-appendixB2}.}

\end{figure}%

\paragraph{The probabilistic implications of these
blocks}\label{sec-appendixB32}

Beyond their graphical capabilities, researchers can translate SCMs and
DAGs into statistical models because these conceptual models encode
functional and probabilistic information, which they can then replace
with suitable functions and probabilistic assumptions
\citep{Pearl_et_al_2016}. Notably, this encoding relies on two
fundamental assumptions: the causal Markov and the faithfulness
assumption. The \emph{causal Markov assumption} links the graph
structure to the joint probability distribution of variables by
asserting that any variable in a graph is independent of all its
non-descendants conditional on its parents
\citetext{\citealp[p.~20]{Neal_2020}; \citealp[p.~80]{Hernan_et_al_2025}}.
The \emph{faithfulness assumption}, in turn, states that the observed
statistical (in)dependencies in the data reflect the underlying causal
structure among variables--except in rare degenerate cases
\citetext{\citealp[p.~100]{Neal_2020}; \citealp[p.~81]{Hernan_et_al_2025}}.

A notable implication of the assumptions underlying the probabilistic
encoding is that any conceptual model described by an SCM and DAG can
represent the joint distribution of variables more efficiently
\citep[p.~29]{Pearl_et_al_2016}. This expression takes the form of a
product of conditional probability distributions (CPDs) of the type
\(P(child \mid parents)\). This property is formally known as the
\emph{Bayesian Network factorization} (BNF,
Equation~\ref{eq-net_factor})
\citetext{\citealp[p.~29]{Pearl_et_al_2016}; \citealp[p.~21]{Neal_2020}}.
In this expression, \(pa(X_{i})\) denotes the set of variables that are
the parents of \(X_{i}\).
\begin{equation}\phantomsection\label{eq-net_factor}{
\begin{aligned}
P(X_{1}, X_{2}, \dots, X_{P}) & = X_{1} \cdot \prod^{P}_{p=2} P(X_{i} \mid X_{i-1}, \dots, X_{1}) & (\small{\text{by chain rule})}\\
& = X_{1} \cdot \prod^{P}_{p=2} P(X_{i} \mid pa(X_{i}) ) & (\small{\text{by BNF}})
\end{aligned}
}\end{equation}

This encoding enables researchers with conceptual (theoretical)
knowledge in the form of an SCM and DAG to predict patterns of
(in)dependencies in the data. As highlighted by
\citet[p.~35]{Pearl_et_al_2016}, these predictions depend solely on the
structure of these conceptual models without requiring the quantitative
details of the equations or the distributions of the errors. Moreover,
once researchers observe empirical data, the patterns of
(in)dependencies in the data can provide significant insights into the
validity of the proposed conceptual model.

The five fundamental building blocks described in
Section~\ref{sec-appendixB31} clearly illustrate which (in)dependencies
can SMCs and DAGs predict. For instance, applying the BNF to the causal
structure shown in the SCM \ref{fig-scm_bb1} and DAG \ref{fig-mdag_bb1}
enables researchers to express the joint probability distribution of the
observed variables as \(P(X_{1}, X_{3}) = P(X_{1}) P(X_{3})\),
supporting the theoretical assumption that the observed variables
\(X_{1}\) and \(X_{3}\) are unconditionally independent
(\(X_{1} \:\bot\:X_{3}\)) \citep[p.~24]{Neal_2020}. Conversely, when
\(X_{3}\) is unconditionally dependent on \(X_{1}\)
(\(X_{1} \:\not\bot\:X_{3}\)), as depicted in the SCM \ref{fig-scm_bb2}
and DAG \ref{fig-mdag_bb2}, the BNF express their joint probability
distribution as \(P(X_{1}, X_{3}) = P(X_{3} \mid X_{1}) P(X_{1})\).
Notably, these descriptions demonstrate the clear correspondence between
the structural equations illustrated in Section~\ref{sec-appendixB31}
and the CPDs.

Beyond the insights gained from two-node structures, researchers can
uncover more nuanced patterns of(in)dependencies from chains, forks, and
colliders. These (in)dependencies apply to any data set generated by a
causal model with those structures, regardless of the specific functions
attached to the SCM \citep[p.~36]{Pearl_et_al_2016}. For instance,
applying the BNF to the chain structure depicted in the SCM
\ref{fig-scm_bb3} and DAG \ref{fig-mdag_bb3} allow researchers to
represent the joint distribution for the observed variables as
\(P(X_{1},X_{2},X_{3}) =\)
\(P(X_{1}) P(X_{2} \mid X_{1}) P(X_{3} \mid X_{2})\). This expression
implies that \(X_{1}\) and \(X_{3}\) are unconditionally dependent
\((X_{1} \:\not\bot\:X_{3})\), but conditionally independent when
controlling for \(X_{2}\) \((X_{1} \:\bot\:X_{3} \mid X_{2})\).
Moreover, in the fork structure shown in the SCM \ref{fig-scm_bb4} and
DAG \ref{fig-mdag_bb4}, researchers can express the joint distribution
of the observed variables as \(P(X_{1},X_{2},X_{3}) =\)
\(P(X_{1} \mid X_{2}) P(X_{2}) P(X_{3} \mid X_{2})\). Similar to the
chain structure, this expression allows researchers to further infer
that \(X_{1}\) and \(X_{3}\) are unconditionally dependent
\((X_{1} \:\not\bot\:X_{3})\), but conditionally independent when
controlling for \(X_{2}\) \((X_{1} \:\bot\:X_{3} \mid X_{2})\). Finally,
researchers analyzing the collider structure illustrated in the SCM
\ref{fig-scm_bb5} and DAG \ref{fig-mdag_bb5} can express the joint
distribution of the observed variables as \(P(X_{1},X_{2},X_{3}) =\)
\(P(X_{1}) P(X_{2} \mid X_{1}, X_{3}) P(X_{3})\). This representation
allows researchers to infer that \(X_{1}\) and \(X_{3}\) are
unconditionally independent \((X_{1} \:\bot\:X_{3})\), but conditionally
dependent when controlling for \(X_{2}\)
\((X_{1} \:\not\bot\:X_{3} \mid X_{2})\). The authors \citet[p.~37, 40,
41]{Pearl_et_al_2016} and \citet[p.~25--26]{Neal_2020} provide the
mathematical proofs for these conclusions.

Using these additional probabilistic insights, researchers can revisit
the scenario in Section~\ref{sec-appendixB2}. In this context, applying
the BNF to the SCM \ref{fig-scm_example2} structure, enables the
representation of the joint probability distribution of the observed
variables as \(P(Y, T, X) =\) \(P(Y \mid T, X) P(T \mid X) P(X)\). From
this expression, researchers can infer that the outcome \(Y\) is
unconditionally dependent on the teaching method \(T\)
\((Y \:\not\bot\:T)\). This dependency arises from two key structures: a
direct causal path from the teaching method \(T\) to the outcome \(Y\),
represented by the two-connected-nodes structure \(T \rightarrow Y\)
(black path in DAG \ref{fig-mdag_example2}), and a confounding
non-causal path from the teaching method \(T\) to the outcome \(Y\)
through the socio-economic status of the school \(X\), represented by
the fork structure \(T \leftarrow X \rightarrow Y\) (gray path in DAG
\ref{fig-mdag_example2}).

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\[
\begin{aligned}
  X & := x \\
  T & := f_{T}(x,e_{T}) \\
  Y & := f_{Y}(T,x,e_{Y}) \\
  e_{T} & \:\bot\:e_{X} \\
  e_{T} & \:\bot\:e_{Y} \\
  e_{X} & \:\bot\:e_{Y}
\end{aligned}
\]

}

\subcaption{\label{fig-scm_example2}SCM}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=0.65\linewidth,height=\textheight,keepaspectratio]{images/png/mdag_example1a.png}

}

\subcaption{\label{fig-mdag_example2}Conditioned DAG}

\end{minipage}%

\caption{\label{fig-example2}Plausible causal structure the scenario
outlined in Section~\ref{sec-appendixB2}.}

\end{figure}%

\paragraph{From probability to causality}\label{sec-appendixB33}

The structural approach to causal inference translates probabilistic
insights into actionable strategies seeking to identify the ACE from
associational quantities. The approach achieves this by relying on the
\emph{modularity assumption}, which posits that intervening on a node
alters only the causal mechanism of that node, leaving others unchanged
\citep[p.~34]{Neal_2020}.

The modularity assumption underpins the concepts of manipulated graphs
and Truncated Factorization, which are essential for representing
interventions \(P(Y_{i} \mid do(T_{i}=t))\) within SCMs and DAGs.
\emph{Manipulated graphs} simulate physical interventions by removing
specific edges from a DAG, while preserving the remaining structure
unchanged \citep[p.~34]{Neal_2020}. In parallel, \emph{Truncated
Factorization} (TF) achieves a similar simulation by removing specific
functions from the conceptual model and replacing them with constants,
while keeping the rest of the structure unchanged \citep{Pearl_2010}.
The probabilistic implications of this factorization are formalized in
Equation~\ref{eq-trunc_factor}, where \(S\) represents the subset of
variables \(X_{p}\) directly influenced by the intervention, while an
example illustrating these concepts follows below.

\begin{equation}\phantomsection\label{eq-trunc_factor}{
P(X_{1}, X_{2}, \dots, X_{P} \mid do(S)) =
\begin{cases}
  \prod P(X_{p} \mid pa(X_{p}) ) & \text{if} \: p \not\in S \\
  1 \quad & \text{otherwise}
\end{cases}
}\end{equation}

Using the TF, researchers can define the \emph{backdoor adjustment} to
identify the ACE. This adjustment states that if a variable
\(X_{p} \in S\) serves as a \emph{sufficient adjustment set} for the
effect of \(X_{a}\) on \(X_{b}\), then the ACE can be identified using
Equation~\ref{eq-backdoor}. The sufficient adjustment set (potentially
empty) must block all non-causal paths between \(X_{a}\) and \(X_{b}\)
without introducing new paths. If such a set exists, then \(X_{a}\) and
\(X_{b}\) are \emph{d-separated} by \(X_{p}\)
(\(X_{a} \:\bot\:X_{b} \mid X_{p}\)) \citep{Pearl_2009}, and \(X_{p}\)
satisfies the \emph{backdoor criterion} \citep[p.~37]{Neal_2020}.

\begin{equation}\phantomsection\label{eq-backdoor}{
P(X_{a} \mid do(X_{b}=x)) = \sum_{Xp} P(X_{a} \mid X_{b}=x, X_{p}) P(X_{p})
}\end{equation}

Ultimately, the backdoor adjustment enables researchers to express the
ACE as:

\begin{equation}\phantomsection\label{eq-backdoor_adjustment}{
\begin{aligned}
\tau & = E[X_{a} \mid do(X_{b}=1)]- E[X_{a} \mid do(X_{b}=2)] \\
  & = E_{Xp}\left[ E[X_{a} \mid do(X_{b}=1), X_{p}]- E[X_{a} \mid do(X_{b}=2), X_{p}] \right] \\
  & = \sum_{Xp} X_{a} \cdot P(X_{a} \mid X_{b}=1, X_{p}) \cdot P(X_{p}) - \sum_{Xp} X_{a} \cdot P(X_{a} \mid X_{b}=2, X_{p}) \cdot P(X_{p})
\end{aligned}
}\end{equation}

With these new insights, researchers revisiting the scenario in
Section~\ref{sec-appendixB32} can infer that the socio-economic status
of the school, \(X\), satisfies the backdoor criterion, assuming the
causal structure depicted by the SCM \ref{fig-scm_example2} and DAG
\ref{fig-mdag_example2} is correct. This means that \(X\) serves as a
sufficient adjustment set, as it effectively blocks all confounding
non-causal paths introduced by the fork structure. Nevertheless, since
\(Y\) remains dependent on \(T\) even after conditioning
\((Y \:\not\bot\:T \mid X)\), this dependency can only be attributed to
the direct causal effect \(T \rightarrow Y\). Notably, for the purpose
of identification, the conditioned DAG \ref{fig-mdag_example2} is
equivalent to the manipulated DAG \ref{fig-mdag_example3}, because \(X\)
satisfies the backdoor criterion.

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\[
\begin{aligned}
  X & := f_{X}(e_{X}) \\
  T & := t \\
  Y & := f_{Y}(t,X,e_{Y}) \\
  e_{T} & \:\bot\:e_{X} \\
  e_{T} & \:\bot\:e_{Y} \\
  e_{X} & \:\bot\:e_{Y}
\end{aligned}
\]

}

\subcaption{\label{fig-scm_example3}SCM}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics[width=0.65\linewidth,height=\textheight,keepaspectratio]{images/png/mdag_example1b.png}

}

\subcaption{\label{fig-mdag_example3}Manipulated DAG}

\end{minipage}%

\caption{\label{fig-example3}Plausible causal structure the scenario
outlined in Section~\ref{sec-appendixB32}.}

\end{figure}%

Researchers can then apply the \emph{backdoor adjustment} to identify
the ACE of \(T\) on \(Y\). They achieve this by first identifying the
CACE of \(T\) on \(Y\) by conditioning on \(X\), and then marginalizing
this effect over \(X\) to obtain the ACE. This process is expressed in
Equation~\ref{eq-mACE2} (see Section~\ref{sec-appendixB2}).

\begin{equation}\phantomsection\label{eq-mACE2}{
\begin{aligned}
  \tau & = E[Y_{i} \mid do(T_{i}=1)]- E[Y_{i} \mid do(T_{i}=2)] \\
  & = E_{X}\left[ E[Y_{i} \mid T_{i}=1, X] - E[Y_{i} \mid T_{i}=2, X] \right] \\
  & = \sum_{X} Y_{i} \cdot P( Y_{i} \mid T_{i}=1, X) \cdot P(X) - \sum_{X} Y_{i} \cdot P( Y_{i} \mid T_{i}=2, X) \cdot P(X)
\end{aligned}
}\end{equation}

\paragraph{The estimation process}\label{sec-appendixB34}

Ultimately, researchers can use Bayesian inference methods to estimate
the ACE. The approach begins by defining two probability distributions:
the likelihood of the data,
\(P(X_{1}, X_{2}, \dots, X_{P} \mid \theta)\), and the prior
distribution, \(P(\theta)\) \citep{Everitt_et_al_2010}, where \(X_{P}\)
represents a random variable, and \(\theta\) represents a
one-dimensional parameter space for simplicity. After observing
empirical data, researchers can update the priors to posterior
distributions using Bayes' rule in Equation~\ref{eq-bayes_rule}:

\begin{equation}\phantomsection\label{eq-bayes_rule}{
P(\theta \mid X_{1}, X_{2}, \dots, X_{P}) = \frac{P(X_{1}, X_{2}, \dots, X_{P} \mid \theta) \cdot P(\theta)}{P(X_{1}, X_{2}, \dots, X_{P})}
}\end{equation}

Given that the denominator on the right-hand side of
Equation~\ref{eq-bayes_rule} serves as a normalizing constant
independent of the parameter \(\theta\), researchers can simplify the
posterior updating process into three steps. First, they integrate new
empirical data through the likelihood. Second, they update the
parameters' priors to a posterior distribution according to
Equation~\ref{eq-prop_rule}. Ultimately, they normalize these results to
obtain a valid probability distribution.

\begin{equation}\phantomsection\label{eq-prop_rule}{
P(\theta \mid X_{1}, X_{2}, \dots, X_{P}) \propto P(X_{1}, X_{2}, \dots, X_{P}\mid \theta) \cdot P(\theta)
}\end{equation}

Temporarily setting aside the definition of prior distributions
\(P(\theta)\), note that the posterior updating process depends heavily
on the assumptions underlying the likelihood of the data. However, as
the number of random variables, \(P\), increases, this joint
distribution quickly becomes intractable \citep{Neal_2020}. This
intractability is evident from Equation~\ref{eq-like_chain}, where the
likelihood distribution is expressed by multiple chained CPDs.

\begin{equation}\phantomsection\label{eq-like_chain}{
P(X_{1}, X_{2}, \dots, X_{P} \mid \theta) = P(X_{1} \mid \theta) \prod^{P}_{p=2} P(X_{i} \mid X_{i-1}, \dots, X_{1}, \theta )
}\end{equation}

Nevertheless, researchers can manage the complexity of the likelihood by
assuming specific local (in)dependencies among variables. SCMs and DAGs
provide a formal framework to represent these assumptions, as detailed
in Section~\ref{sec-appendixB32}. These assumptions improve model
tractability and simplify the estimation process by enabling the
derivation of the BNF of the likelihood (Equation~\ref{eq-like_BNF}).
With this simplified structure, any probabilistic programming language
can model the system and compute the parameter's posterior distribution
using Equation~\ref{eq-bayes_rule}.

\begin{equation}\phantomsection\label{eq-like_BNF}{
P(X_{1}, X_{2}, \dots, X_{P} \mid \theta) = P(X_{1} \mid \theta) \prod^{P}_{p=2} P(X_{i} \mid pa(X_{i}), \theta )
}\end{equation}

\newpage{}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\renewcommand{\bibsection}{}
\bibliography{references.bib}

\end{document}